### 2024-05-29 article/main.tex
initial draft

\documentclass[runningheads,a4paper,11pt]{article}

\usepackage{algorithmic}
\usepackage{algorithm} 
\usepackage{array}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{cite}
\usepackage{comment} 
\usepackage{enumitem}
\usepackage{epsfig} 
\usepackage{fancyhdr}
\usepackage[T1]{fontenc}
\usepackage{geometry} 
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref} 
\usepackage[latin1]{inputenc}
\usepackage{interval}
\usepackage{multicol}
\usepackage{multirow} 
\usepackage{rotating}
\usepackage{setspace}
\usepackage{subfigure}
\usepackage{tabularx}
\usepackage{tikz}
\usepackage{url}
\usepackage{verbatim}
\usepackage{xcolor}
\usepackage{xstring}

\geometry{a4paper,top=2.5cm,left=2cm,right=2cm,bottom=2cm}

\begin{document}
\title{How Consistent are LLMs in Applying Handcrafted Language Features}
\author{Andrei Olar --- andrei.olar@ubbcluj.ro}
\maketitle
\begin{abstract}
    Large language models are sought for solving an increasing number of
    increasingly diverse tasks. Perhaps naturally one such task is text style
    transfer: the task of changing the style of a text while preserving its
    meaning. Besides the natural approach of fine-tuning a large language model,
    a simpler and cheaper alternative should also be investigated: using
    handcrafted language features to direct the large language model.
\end{abstract}

\section{Introduction}\label{introduction}

Large language models have taken one of the main spots in research for quite
some time.
Because language models such as BERT~\cite{devlin2018bert} or GPT~\cite{gpt-2018,gpt2-2019,
gpt3-2020} constantly advance the state of the art for many natural language
processing tasks, it has become natural to want to evaluate these models
performance on ever more tasks.
We know from multiple surveys~\cite{minaee2024llmsurvey,zhao2023survey} and
benchmarks~\cite{papcode2024hellaswag,chiang2024chatbot} that large language
models (especially the more popular ones) are very good at following
instructions.
We can also intuit that training LLMs on diverse data (for instance the
Pile~\cite{gao2020pile}) uniquely qualifies them to produce text in a wide
variety of styles.

Text style transfer is defined as the ``task of transforming the stylistic manner
in which a sentence is written, while preserving the meaning of the original
sentence''~\cite{tst-review-2021}.
This definition can be extended, to entire articles or corpuses containing
multiple articles without fear of text style transfer losing its meaning.
The task of transfering text style is an old preocupation for the natural
language processing and computational linguistics communities which has picked
up interest again in the context of the advancements in the field of deep
learning and particularly since the inception of the transformer architecture.

Handcrafted linguistic features (HLFs) are single numerical values produced by a
uniquely identifiable method on any natural language~\cite{lftk-2023}. More
often than not, these features are easy to compute and intuitively very telling
of the style in which a text is written.

For example, take the number of exclamations in a text.
This particular feature is easier to compute than most other metrics in
computational linguistics and high values of this particular HLF can pretty
reliably signal an inflamatory news writing style, spam or an alert narrative.
The hope that arises from this observation is to describe style sufficiently by
using only HLFs.
<CITATIONS NEEDED>
The appeal of this approach lies within the idempotent nature of most HLFs.
That makes them useful in identifying or verifying styles, too. <CITATIONS NEEDED>

<SOME MORE IDEAS AROUND STYLEPTB, lexical, syntax, semantics,
and thematics, that span the fundamental atomic
transfers that text can undergo (McDonald and
Pustejovsky, 1985; DiMarco and Hirst, 1993).Chomsky 2002,>

From the above ideas, a couple of questions arise naturally.
Firstly, is it possible to influence an LLMs writing style by instructing it to
follow certain HLFs?
If this is possible then it's very financially advantageous to use this approach
instead of fine tuning an LLM.
For one, fine tuning an LLM incurs some loss of generality~\cite{yang2024unveiling}.
And then there's the matter of cost.

Secondly, we ask if it's possible to tell just how well the most prominent LLMs
of today are able to generate text according to instructions concerning HLFs.
If LLMs respond well enough to prompt instructions, this might prove to be an
easy and affordable solution for the text style transfer problem.

\section{Related Work}\label{related}

There is abundent literature on the topic of text style transfer taken from
various perspectives.
For writing this paper, we have mostly consulted literature reviews and
syntheses~\cite{tst-review-2021,tst-survey-2022} on the subject in order to
inform the selection of HLFs that we would be using in our experiments.

The literature on HLFs is extensive and spans most of the field of computational
linguistics.
This work is well referenced and synthesised by Lee and Lee~\cite{lftk-2023}.
To our knowledge there is no work that connects HLFs with LLMs in the manner
described in this paper.
Most other literature focuses on achieving results on connected tasks, such as
assessing text readability~\cite{lee-etal-2021-pushing}.
This paper differs from other endeavours in at least two significant ways.
Firstly, this paper is simply about feeding well crafted user prompts to an LLM
instead of diving into deeper concepts such as the transformer architecture,
deep learning or even neural networks.
Secondly, it focuses squarely on manipulating the style attributes of the text
generated by an LLM using HLFs and nothing else.

That said, text style transfer has two components: transferring style and
retaining meaning.
While this paper focuses on transferring style, there are numerous benchmarks,
studies and reviews that atest to the ability of state of the art LLMs to retain
meaning in rewording exercises. <CITATIONS NEEDED>

- delineate from work in the StylePTB paper (https://aclanthology.org/2021.naacl-main.171.pdf)
- the styleptb paper evaluates style transfer quality which is the subject of
future work

\section{Experiment Design}\label{method}

Our experiment consists of analysing the behavior of an LLM when asked to
generate new text based on a given text and and some instructions derived from
HLF values computed on a benchmark dataset.
To perform the experiment we must know which LLMs we are going to evaluate and
what benchmark datasets we are going to use.
We must also select the HLFs for our experiment.

Before answering those questions, let us outline the experimental process.

The first task in the experiment is to establish a baseline.
We do so by instructing the target LLM to generate text retaining the meaning of
the input text without any further instruction.
We repeat the text generation step 10 times.
For each text generation, we compute the following statistics for each of our
chosen HLFs:

\begin{itemize}
    \item \textbf{min}: the minimum value of the HLF on the generated text;
    \item \textbf{max}: the maximum value of the HLF on the generated text;
    \item \textbf{avg}: the mean value of the HLF on the generated text;
    \item \textbf{var}: the variance of the HLF values on the generated text.
\end{itemize}

A second level baseline is then obtained by following the above process with a
prompt that builds upon the first one by adding the instruction to use example
text from our benchmark dataset of choice.
The examples are five texts chosen randomly once for each benchmark dataset.
These example texts are the same for all evaluated LLMs.

The next step is to devise HLF prompts derived from both of the baseline
prompts.
The HLF prompts contain additional instructions to generate text that would
comply with certain HLF values.
These HLF values are obtained by computing the statistics outlined above on
the entirety of the target benchmark dataset.

We then repeat the text generation and statistics computation with each of the
HLF prompts in the same way as we did for their corresponding baselines.

To answer our first question, we compare the first baseline with the results
obtained by using the corresponding modified HLF prompt.
Our reasoning is that if the LLM were influenced by the HLF prompt, the output
would show different HLF values compared to the baseline.
This assumption is rooted in the work on detecting LLM generated text.<CITATION NEEDED>

To answer the question of how much LLMs are influenced by being prompted with
instructions about HLFs, we turn to the second level baseline.
Given that the LLM now also has example articles, it can freely immitate those
articles when generating the text.
If there is a consistent difference between the HLF and non-HLF prompt, this
difference can only stem from the HLF instructions.
Therefore, how much the HLF values differ between the second level baseline and
the corresponding HLF generated texts is strictly due to the HLF instructions.

One final observation before diving into details is that we can use the variance
computed for HLFs on the benchmark datasets to our advantage.
This statistic shows us whether the LLM is at least roughly compliant with the
instructions provided in the HLF prompts.
To understand how let us denote an HLF computed on a text generated by using an
HLF prompt with $HLF_g$.
Similarly, we denote the average and the variance of the same HLF on the
benchmark dataset with $avg(HLF_b)$ and $var(HLF_b)$, respectively.
If we consistently notice that
    \[HLF_g - var(HLF_b) \leq avg(HLF_b) \leq HLF_g + var(HLF_b)\]
then we consider that the LLM consistently follows the HLF instructions in the
prompt.

\subsection{LLM Selection}\label{llm-selection}

In order to perform our experiment we must first choose proper candidates.
Since the interaction with the model happens only via prompts, we may use models
with any type of licensing.
Our choice will be the top model from each of the highest scoring five different
major vendors listed in the Chatbot Arena benchmark~\cite{chiang2024chatbot}.

<TODO - make a table with model specs: no params, context size, ELO, license>
OpenAI - gpt-4o
Google - Gemini-1.5-Pro-API-0514
Anthropic - Claude 3 Opus
Meta - Llama-3-70b-Instruct
Cohere - Command R+

\subsection{HLF Selection}\label{hlf-selection}

On the other hand, we need to look at which handcrafted linguistic features are
relevant for our experiment.
In our opinion, good handcrafted linguistic features for the task of text style
transfer are those features that can reasonably be thought to affect one of the
coordinates by which people judge the style of text.
These coordinates are well enumerated both in~\cite{tst-review-2021} and
in~\cite{tst-survey-2022}.

In order to build upon the previous work done regarding computing handcrafted
linguistic features, we will rely only on the features available in
LFTK~\cite{lftk-2023}.

Based on the above two selection criteria, we have selected the handcrafted
linguistic features in <INSERT TABLE HERE>.

\subsection{DataSet Selection}\label{ds-selection}

- Are there benchmark datasets for text style transfer? (STYLEPTB + Yelp Reviews)

\subsection{Input Prompt and Text}\label{input-text}

- link to text from within the corpus
- link to text from outside the corpus
- link to each system prompt (2 baselines, 2 HLF prompt templates)

\section{Experiment Results}

\subsection{Graph Interpretation}

Ox - the number of the generation 1..10
Oy - HLF value
Continuous grey line - baseline metric
continuous grenat line - hlf metric
grey around grenat line - corpus target zone
dotted black lines - corpus min, max
continuous black line - corpus avg

\subsection{Baseline}

Yelp Results

STYLEPTB Results

\subsection{Context Aware}

Yelp Results
StylePTB Results

\section{Discussion}

Discuss experiment results here.

\section{Conclusions and Future Work}

Yes, you can instruct at least some of the LLMs with HLFs.

compute BLEU, meteor, cider on styleptb using these LLMs and the prompt to
evaluate whether we actually achieve style transfer because of/in spite of HLF
instructions.
Human survey (amazon mechanical turk) on whether the text generations are good
or not. Likert scale (0 no good, 10 matches style perfectly)

\bibliographystyle{plain}
\bibliography{references}
\end{document}

### 2024-05-30 article/main.tex
tweak intro and related work. add llm table

@@
 
-Large language models have taken one of the main spots in research for quite
-some time.
+\textbf{Large language models (LLMs)} have are now a commonplace research topic.
 Because language models such as BERT~\cite{devlin2018bert} or GPT~\cite{gpt-2018,gpt2-2019,
@@
 
-Text style transfer is defined as the ``task of transforming the stylistic manner
-in which a sentence is written, while preserving the meaning of the original
-sentence''~\cite{tst-review-2021}.
+\textbf{Text style transfer (TST)} is defined as the ``task of transforming the
+stylistic manner in which a sentence is written, while preserving the meaning of
+the original sentence''~\cite{tst-review-2021}.
 This definition can be extended, to entire articles or corpuses containing
@@
 
-Handcrafted linguistic features (HLFs) are single numerical values produced by a
+\textbf{Handcrafted linguistic features (HLFs)} are single numerical values produced by a
 uniquely identifiable method on any natural language~\cite{lftk-2023}. More
@@
 
-For example, take the number of exclamations in a text.
-This particular feature is easier to compute than most other metrics in
-computational linguistics and high values of this particular HLF can pretty
-reliably signal an inflamatory news writing style, spam or an alert narrative.
-The hope that arises from this observation is to describe style sufficiently by
-using only HLFs.
-<CITATIONS NEEDED>
-The appeal of this approach lies within the idempotent nature of most HLFs.
-That makes them useful in identifying or verifying styles, too. <CITATIONS NEEDED>
-
-<SOME MORE IDEAS AROUND STYLEPTB, lexical, syntax, semantics,
-and thematics, that span the fundamental atomic
-transfers that text can undergo (McDonald and
-Pustejovsky, 1985; DiMarco and Hirst, 1993).Chomsky 2002,>
+Let's take the number of exclamations in a text as an example.
+This particular HLF is one of the easier computational linguistics features to
+compute.
+Yet high values of this particular HLF often provide a clear signal for
+inflamatory news writing style, spam or an alert narrative.
+Similar observations have been made about other characteristics of written text
+that might be computed as HLFs~\cite{hovy1987generating}.
+This gives rise to the hope that text style can be sufficiently, although not
+completely, described through the use of HLFs.
+What's most appealing about HLFs is their idempotent nature.
+They could potentially be used both for interpreting style and generating text
+with a given style.
 
@@
 Firstly, is it possible to influence an LLMs writing style by instructing it to
-follow certain HLFs?
-If this is possible then it's very financially advantageous to use this approach
-instead of fine tuning an LLM.
-For one, fine tuning an LLM incurs some loss of generality~\cite{yang2024unveiling}.
-And then there's the matter of cost.
+generate text in accordance to certain HLFs?
+This problem of controlled text generation has implications for text style
+transfer and authorship evaluation.
+It might also constitute a cheaper and less intrusive approach than fine tuning
+an LLM.
+Fine tuning an LLM incurs some loss of generality~\cite{yang2024unveiling}.
 
 Secondly, we ask if it's possible to tell just how well the most prominent LLMs
-of today are able to generate text according to instructions concerning HLFs.
-If LLMs respond well enough to prompt instructions, this might prove to be an
-easy and affordable solution for the text style transfer problem.
+of today are able to generate text according to instructions concerning HLFs?
+The answer to this question could hint at a relatively simple and accurate way
+of at least partially evaluating LLM performance on the text style transfer
+task.
 
@@
 
-There is abundent literature on the topic of text style transfer taken from
-various perspectives.
-For writing this paper, we have mostly consulted literature reviews and
-syntheses~\cite{tst-review-2021,tst-survey-2022} on the subject in order to
-inform the selection of HLFs that we would be using in our experiments.
+There is abundent literature on the topic of text style transfer.
+Reviews and surveys~\cite{tst-review-2021,tst-survey-2022} were helpful for
+informing our selection of HLFs.
+Studies on attaining fine-grained text style transfer using language models
+have served as inspiration for writing this paper~\cite{lyu-etal-2023-fine}.
+We should note though that we are not attempting to transfer text style, but
+merely to instruct an LLM to follow instructions based on HLFs.
 
-The literature on HLFs is extensive and spans most of the field of computational
-linguistics.
-This work is well referenced and synthesised by Lee and Lee~\cite{lftk-2023}.
+From a linguistic perspective, text style can be thought of as the form used for
+delivering meaning~\cite{tst_sigkdd_review_2022}.
+The ample literature on the subject seems to agree that style depends on context
+and author choice vis-a-vis a communication
+goal~\cite{mcdonald1985computational,hovy1987generating}.
+The author's choice is available at all levels: lexical, syntactic and
+semantic~\cite{dimarco1994model}.
+These findings provide the hope and evidence that the current paper is useful
+especially when it comes to the influence of syntactic structure over text
+style~\cite{chomsky2002syntactic}.
+
+The idea of constructing style from fine-grained aspects is not new and there
+are excellent studies on the subject. The StylePTB authors (StylePTB is one of
+the benchmark datasets used in this paper) explain how this dataset leverages
+lexical, syntactic, semantic and thematic aspects to allow verifying text style
+transfer more thoroughly than previous methods~\cite{lyu-etal-2021-styleptb}.
+In the current paper we use this dataset merely to instruct LLMs to rewrite text
+from it, without actually measuring style transfer quality.
+
+The body of work on HLFs is extensive.
+It is well referenced and synthesised by Lee and Lee~\cite{lftk-2023}.
 To our knowledge there is no work that connects HLFs with LLMs in the manner
@@
 assessing text readability~\cite{lee-etal-2021-pushing}.
-This paper differs from other endeavours in at least two significant ways.
+This paper differs from other tasks that aim to control the style of generated
+text in at least two significant ways.
 Firstly, this paper is simply about feeding well crafted user prompts to an LLM
@@
 deep learning or even neural networks.
-Secondly, it focuses squarely on manipulating the style attributes of the text
-generated by an LLM using HLFs and nothing else.
+Secondly, it focuses on manipulating the style attributes of the text generated
+by an LLM squarely using HLFs and measuring the change in output.
+There is no ulterior motive.
 
-That said, text style transfer has two components: transferring style and
-retaining meaning.
-While this paper focuses on transferring style, there are numerous benchmarks,
-studies and reviews that atest to the ability of state of the art LLMs to retain
-meaning in rewording exercises. <CITATIONS NEEDED>
-
-- delineate from work in the StylePTB paper (https://aclanthology.org/2021.naacl-main.171.pdf)
-- the styleptb paper evaluates style transfer quality which is the subject of
-future work
+Finally, this paper documents a form of controllable text generation (CTG)~\cite{zhang-ctg-2022}.
+Hightened recent interest in CTG and especially in benchmarking LLMs in the
+context of CTG~\cite{chen2024benchmarking} is particularly relevant for this
+paper.
+Such research alleviates the burden of proof regarding the level to which
+LLMs respond well to diversified instructions for controlled text generation in
+general.
+The novelty of this paper lies both in basing the control instructions on
+HLFs and in using those HLFs to evaluate the performance of the LLM.
 
@@
 Our experiment consists of analysing the behavior of an LLM when asked to
-generate new text based on a given text and and some instructions derived from
-HLF values computed on a benchmark dataset.
+generate new text based on a given text and and instructions derived from HLF
+statistics computed on a benchmark dataset.
 To perform the experiment we must know which LLMs we are going to evaluate and
@@
 
-A second level baseline is then obtained by following the above process with a
-prompt that builds upon the first one by adding the instruction to use example
+A second baseline is obtained by following the above process with a new prompt
+that builds upon the first one by adding the instruction to use example
 text from our benchmark dataset of choice.
@@
 
-The next step is to devise HLF prompts derived from both of the baseline
-prompts.
+The next step is to devise HLF prompts derived from both baseline prompts.
 The HLF prompts contain additional instructions to generate text that would
@@
 the entirety of the target benchmark dataset.
-
-We then repeat the text generation and statistics computation with each of the
-HLF prompts in the same way as we did for their corresponding baselines.
+We then repeat the process through which we created the baselines for each of
+the HLF prompts.
 
@@
 would show different HLF values compared to the baseline.
-This assumption is rooted in the work on detecting LLM generated text.<CITATION NEEDED>
 
 To answer the question of how much LLMs are influenced by being prompted with
-instructions about HLFs, we turn to the second level baseline.
+instructions about HLFs, we turn our attention to the second baseline.
 Given that the LLM now also has example articles, it can freely immitate those
@@
 difference can only stem from the HLF instructions.
-Therefore, how much the HLF values differ between the second level baseline and
-the corresponding HLF generated texts is strictly due to the HLF instructions.
+We perform this second experiment with input text from both within and outside
+the corpus.
 
-One final observation before diving into details is that we can use the variance
-computed for HLFs on the benchmark datasets to our advantage.
-This statistic shows us whether the LLM is at least roughly compliant with the
-instructions provided in the HLF prompts.
-To understand how let us denote an HLF computed on a text generated by using an
-HLF prompt with $HLF_g$.
-Similarly, we denote the average and the variance of the same HLF on the
-benchmark dataset with $avg(HLF_b)$ and $var(HLF_b)$, respectively.
+One final observation before diving into details is that we can use an HLF's
+variance on the benchmark dataset to our advantage.
+This statistic shows whether the LLM takes the HLF prompts into account enough
+for the generated text to fall within the parameters of the benchmark dataset.
+Let us denote an HLF computed on a generated text using an HLF prompt with
+$HLF_g$.
+Similarly denote the average and the variance of the same HLF on the benchmark
+dataset with $avg(HLF_b)$ and $var(HLF_b)$, respectively.
 If we consistently notice that
     \[HLF_g - var(HLF_b) \leq avg(HLF_b) \leq HLF_g + var(HLF_b)\]
-then we consider that the LLM consistently follows the HLF instructions in the
-prompt.
+then we consider that the LLM follows HLF prompts in a relevant way.
 
@@
 with any type of licensing.
-Our choice will be the top model from each of the highest scoring five different
-major vendors listed in the Chatbot Arena benchmark~\cite{chiang2024chatbot}.
+For our experiment we choose the top model from each of the highest scoring five
+different major vendors listed on the Chatbot Arena~\cite{chiang2024chatbot}
+leaderboard.
+The final choice is documented in Table~\ref{table-llm}.
 
-<TODO - make a table with model specs: no params, context size, ELO, license>
-OpenAI - gpt-4o
-Google - Gemini-1.5-Pro-API-0514
-Anthropic - Claude 3 Opus
-Meta - Llama-3-70b-Instruct
-Cohere - Command R+
+\begin{table}[ht]
+    \setlength\tabcolsep{6pt}
+    \centering
+\begin{tabular}{@{}lllll@{}}\toprule
+Name                      & Parameters           & Context Size        & Elo Score & License Type                                      \\ \toprule
+OpenAI GPT-4o             & undisclosed          & 128000              & 1287      & Proprietary                                       \\
+Google Gemini 1.5 Pro     & undisclosed          & 1048576             & 1268      & Proprietary                                       \\
+Anthropic Claude 3 Opus   & undisclosed          & 200000              & 1248      & Proprietary                                       \\
+Meta Llama 3-70B-Instruct & 70 Billion           & 8000                & 1208      & Llama 3 Community                                 \\
+Cohere Command R+         & 104 Billion          & 128000              & 1189      & CC-BY-NC 4.0 \\
+                          &                      &                     &           & with Acceptable Use\\
+                          &                      &                     &           & Addendum\\ \bottomrule
+\end{tabular}
+\caption{Experiment LLMs}\label{table-llm}
+\end{table}
 

### 2024-06-01 article/main.tex
fill out hlfs, datasets and correct existing text

@@
 
-\textbf{Large language models (LLMs)} have are now a commonplace research topic.
-Because language models such as BERT~\cite{devlin2018bert} or GPT~\cite{gpt-2018,gpt2-2019,
-gpt3-2020} constantly advance the state of the art for many natural language
-processing tasks, it has become natural to want to evaluate these models
-performance on ever more tasks.
+\textbf{Large language models (LLMs)} have are a commonplace research topic
+today.
+Because language models such as BERT~\cite{devlin2018bert} or
+GPT~\cite{gpt-2018,gpt2-2019,gpt3-2020} constantly advance the state of the art
+on many natural language processing tasks, it has become natural to want to
+evaluate the performance of these generalist models on ever more specialized
+tasks.
 We know from multiple surveys~\cite{minaee2024llmsurvey,zhao2023survey} and
 benchmarks~\cite{papcode2024hellaswag,chiang2024chatbot} that large language
-models (especially the more popular ones) are very good at following
-instructions.
+models (especially the more popular ones) are good at following instructions.
 We can also intuit that training LLMs on diverse data (for instance the
@@
 the original sentence''~\cite{tst-review-2021}.
-This definition can be extended, to entire articles or corpuses containing
-multiple articles without fear of text style transfer losing its meaning.
+This definition can be extended to entire articles or corpuses of text because
+these are also the object of linguistic style and
+stylistics~\cite{lugea2023stylistics}.
 The task of transfering text style is an old preocupation for the natural
@@
 
-\textbf{Handcrafted linguistic features (HLFs)} are single numerical values produced by a
-uniquely identifiable method on any natural language~\cite{lftk-2023}. More
-often than not, these features are easy to compute and intuitively very telling
-of the style in which a text is written.
+\textbf{Handcrafted linguistic features (HLFs)} are single numerical values
+produced by a uniquely identifiable method on any natural
+language~\cite{lftk-2023}.
+More often than not, these values are easy to compute while the methods used for
+obtaining them are usually idempotent.
+This makes them very attractive because of the dual purpose that they can serve.
+On the one hand, they can be used to instruct the generation of text (``write
+no more than 10 sentences'') while on the other they can be used to analyse text
+and, therefore, verify generated text.
 
-Let's take the number of exclamations in a text as an example.
-This particular HLF is one of the easier computational linguistics features to
-compute.
-Yet high values of this particular HLF often provide a clear signal for
-inflamatory news writing style, spam or an alert narrative.
+Some HLFs statistically describe text in terms of its basic syntactical and
+morphological components.
+Knowing this we could follow Noam Chomsky's ideas~\cite{chomsky2002syntactic}
+which imply that the style of a text is influenced by its underlying syntactic
+structures.
+Traditionally, linguistic style was thought to be the result of morphological
+and syntactic arrangements~\cite{lugea2023stylistics}.
+More recent schools of thought on linguistic style find semantics and pragmatics
+to be inextricably linked to author style~\cite{verma2019lexical}, but argue
+that a great deal can still be inferred from lexicon, syntax and a `surface'
+level which contains basic HLFs such as the number of words in a sentence.
+
+For example, a high number of exclamations in a text is often a revealing
+signal of an inflamatory news writing style, of spam or of alert narration or
+dialogue.
 Similar observations have been made about other characteristics of written text
-that might be computed as HLFs~\cite{hovy1987generating}.
-This gives rise to the hope that text style can be sufficiently, although not
-completely, described through the use of HLFs.
-What's most appealing about HLFs is their idempotent nature.
-They could potentially be used both for interpreting style and generating text
-with a given style.
+that might be computed as HLFs~\cite{hovy1987generating,lugea2023stylistics}.
+Combining multiple HLFs that are revelatory of linguistic style could provide us
+with an easily interpretable and verifiable stylistic profile.
 
-From the above ideas, a couple of questions arise naturally.
-Firstly, is it possible to influence an LLMs writing style by instructing it to
-generate text in accordance to certain HLFs?
+These apparently disparate strands of thought lead us to the two research
+questions we'd like to answer in this paper.
+
+Firstly, \textbf{is it possible to influence an LLM's writing style by
+    instructing it to generate text that complies to certain HLFs}?
 This problem of controlled text generation has implications for text style
-transfer and authorship evaluation.
-It might also constitute a cheaper and less intrusive approach than fine tuning
-an LLM.
-Fine tuning an LLM incurs some loss of generality~\cite{yang2024unveiling}.
+transfer and authorship verification.
+Answering this question might point out cheaper, less intrusive and more user
+friendly solutions than fine tuning large language models for these tasks.
+Fine tuning LLMs also incurs some loss of generality~\cite{yang2024unveiling},
+going directly against the purpose of having a generic language model.
 
-Secondly, we ask if it's possible to tell just how well the most prominent LLMs
-of today are able to generate text according to instructions concerning HLFs?
+Secondly, we ask \textbf{if it's possible to quantify how closely an LLM is able
+    to follow prompts concerning the compliance of the generated text to HLFs?}
 The answer to this question could hint at a relatively simple and accurate way
-of at least partially evaluating LLM performance on the text style transfer
-task.
+of evaluating the quality of (partial) text style transfer using a large
+language model.
 
@@
 
-From a linguistic perspective, text style can be thought of as the form used for
-delivering meaning~\cite{tst_sigkdd_review_2022}.
+Linguistic style, then, is the sum of identifiable language choices manifest in
+a text made from the language system by the text
+producer~\cite{lugea2023stylistics}.
+Another way to think about it is that style is the form used for delivering
+meaning~\cite{tst_sigkdd_review_2022}.
 The ample literature on the subject seems to agree that style depends on context
@@
 goal~\cite{mcdonald1985computational,hovy1987generating}.
-The author's choice is available at all levels: lexical, syntactic and
-semantic~\cite{dimarco1994model}.
-These findings provide the hope and evidence that the current paper is useful
-especially when it comes to the influence of syntactic structure over text
-style~\cite{chomsky2002syntactic}.
+The author's choice is available at all linguistic levels, including the
+morphological, lexical, syntactic, semantic and pragmatic
+levels~\cite{dimarco1994model,lugea2023stylistics}.
+The influence of syntactic structure over text style is apparent also from a
+formalist perspective~\cite{chomsky2002syntactic}.
+This work grounds our choice of handcrafted linguistic features related to
+linguistic style.
 
 The idea of constructing style from fine-grained aspects is not new and there
-are excellent studies on the subject. The StylePTB authors (StylePTB is one of
-the benchmark datasets used in this paper) explain how this dataset leverages
-lexical, syntactic, semantic and thematic aspects to allow verifying text style
-transfer more thoroughly than previous methods~\cite{lyu-etal-2021-styleptb}.
-In the current paper we use this dataset merely to instruct LLMs to rewrite text
-from it, without actually measuring style transfer quality.
+are excellent studies on the subject.
+The StylePTB authors explain how this dataset leverages lexical, syntactic,
+semantic and thematic aspects to allow verifying text style transfer more
+thoroughly than previous methods~\cite{lyu-etal-2021-styleptb}.
+Besides providing the understanding required to distinguish between semantic
+transfer and other forms of style transfer, the dataset itself is useful for
+evaluating HLFs that function at the sentence level.
 
@@
 described in this paper.
-Most other literature focuses on achieving results on connected tasks, such as
-assessing text readability~\cite{lee-etal-2021-pushing}.
-This paper differs from other tasks that aim to control the style of generated
-text in at least two significant ways.
-Firstly, this paper is simply about feeding well crafted user prompts to an LLM
-instead of diving into deeper concepts such as the transformer architecture,
-deep learning or even neural networks.
-Secondly, it focuses on manipulating the style attributes of the text generated
-by an LLM squarely using HLFs and measuring the change in output.
-There is no ulterior motive.
+HLFs are used in the context of other, connected tasks such as assessing text
+readability~\cite{lee-etal-2021-pushing}.
 
-Finally, this paper documents a form of controllable text generation (CTG)~\cite{zhang-ctg-2022}.
+In terms of controlling the style of the generated text, our approach differs
+from others in at least two significant ways: simplicity and focus.
+We are constructing well crafted user prompts for an LLM instead of diving into
+more complex artificial intelligence solutions.
+Our focus is on controlling text generation using HLFs and measuring the change
+relative to a baseline obtained without using HLFs in the generation process.
+We are not discussing the effectiveness of style transfer, author verification
+or other tasks (yet).
+
+Finally, this paper documents a form of controllable text generation
+(CTG)~\cite{zhang-ctg-2022}.
 Hightened recent interest in CTG and especially in benchmarking LLMs in the
@@
 general.
-The novelty of this paper lies both in basing the control instructions on
-HLFs and in using those HLFs to evaluate the performance of the LLM.
+Compared to the existing work, our prompts are guided by HLFs.
+Furthermore, we use those same HLFs to evaluate the performance of the LLM.
 
@@
 Our experiment consists of analysing the behavior of an LLM when asked to
-generate new text based on a given text and and instructions derived from HLF
-statistics computed on a benchmark dataset.
+generate new text based on a given input text and some instructions derived from
+HLF statistics computed on a benchmark dataset.
 To perform the experiment we must know which LLMs we are going to evaluate and
@@
 We must also select the HLFs for our experiment.
+Before going into details, let us outline the experiment process.
 
-Before answering those questions, let us outline the experimental process.
-
-The first task in the experiment is to establish a baseline.
+The first task is to establish a baseline.
 We do so by instructing the target LLM to generate text retaining the meaning of
 the input text without any further instruction.
-We repeat the text generation step 10 times.
+We repeat the text generation step 10 times using the same prompt and the same
+input text.
 For each text generation, we compute the following statistics for each of our
@@
 
-A second baseline is obtained by following the above process with a new prompt
-that builds upon the first one by adding the instruction to use example
+A second baseline is obtained by following the same process but with a new
+prompt that builds upon the first one by adding the instruction to use example
 text from our benchmark dataset of choice.
-The examples are five texts chosen randomly once for each benchmark dataset.
-These example texts are the same for all evaluated LLMs.
+The examples are three texts chosen randomly once for each benchmark dataset.
+The reason behind choosing three texts is to not overrun the context window of
+the LLM.
+This hints at the main drawback for our approach: it uses up the context window
+with system instructions.
+While this isn't an issue with all models, some only allow a context window of
+around 8000 tokens so we should be mindful.
+We use the same example texts for all evaluated LLMs.
+Same as before, we construct the baseline by conducting 10 trials using the same
+prompt and input text.
 
@@
 the entirety of the target benchmark dataset.
-We then repeat the process through which we created the baselines for each of
-the HLF prompts.
+The HLF prompts are constructed from templates that take into account the HLF
+values.
+The prompt templates are available in the code repository that accompanies this
+experiment~\cite{olar2024experimentcode}.
+The instructions concerning all choseon HLFs are given concomitently within the
+same prompt.
+In a similar manner to how we generated the baselines, we run ten trials for
+each case (with or without example articles from the target benchmark dataset).
 
-To answer our first question, we compare the first baseline with the results
-obtained by using the corresponding modified HLF prompt.
+We compare the first baseline with the results obtained by using the first
+modified HLF prompt to understand whether we can influence LLM text generation
+using HLFs.
 Our reasoning is that if the LLM were influenced by the HLF prompt, the output
@@
 
-To answer the question of how much LLMs are influenced by being prompted with
-instructions about HLFs, we turn our attention to the second baseline.
+We now turn our attention to the second baseline.
 Given that the LLM now also has example articles, it can freely immitate those
 articles when generating the text.
-If there is a consistent difference between the HLF and non-HLF prompt, this
-difference can only stem from the HLF instructions.
-We perform this second experiment with input text from both within and outside
-the corpus.
+If there is a consistent difference over the ten trials between generating text
+using the baseline or using the HLF prompt, this difference can only be a result
+of the LLM complying to HLF instructions.
+This difference is expressed by the change of value for each HLF between the
+baseline and HLF enhanced text generation.
+
+Because of the examples present in the prompt, we perform the second experiment
+with input text both from within and from outside the corpus.
 
@@
         Cohere Command R+         & 104 Billion & 128000       & 1189      & CC-BY-NC 4.0        \\
-                          &                      &                     &           & with Acceptable Use\\
-                          &                      &                     &           & Addendum\\ \bottomrule
+                                  &             &              &           & with Acceptable Use \\
+                                  &             &              &           & Addendum            \\ \bottomrule
     \end{tabular}
@@
 
-On the other hand, we need to look at which handcrafted linguistic features are
-relevant for our experiment.
-In our opinion, good handcrafted linguistic features for the task of text style
-transfer are those features that can reasonably be thought to affect one of the
-coordinates by which people judge the style of text.
-These coordinates are well enumerated both in~\cite{tst-review-2021} and
-in~\cite{tst-survey-2022}.
+Our main focus is to understand how susceptible an LLMs is to prompts telling it
+to conform to HLFs.
+LFTK~\cite{lftk-2023} authors categorize HLFs by domain and family.
+Furthermore, information is available on the suitability of each HLF for given
+tasks such as readability assessment or fake news detection.
+However, there is no direct mapping between HLFs and style.
+It falls onto us to choose features based on their influence over text style as
+recognized in literature~\cite{verma2019lexical,lugea2023stylistics}.
 
-In order to build upon the previous work done regarding computing handcrafted
-linguistic features, we will rely only on the features available in
-LFTK~\cite{lftk-2023}.
+Another HLF characteristic we have in mind is their level of abstraction.
+Features such as the word count or number of punctuation mark are very concrete.
+On the other side of the spectrum, features such as the Flesch-Kincaid grade
+level~\cite{kincaid1975derivation} are highly abstract.
+We divide HLFs in five categories ranging from very concrete to highly abstract
+and choose 2 features from each category for a grand total of 10 HLFs.
+Our final selection is available in Table~\ref{table-hlf}.
 
-Based on the above two selection criteria, we have selected the handcrafted
-linguistic features in <INSERT TABLE HERE>.
+\begin{table}[ht]
+    \setlength\tabcolsep{6pt}
+    \centering
+    \begin{tabular}{@{}lllll@{}}\toprule
+        Name                              & Family           & Domain           & Abstraction     & Style Level\footnotemark \\ \toprule
+        total unique words                & wordsent         & surface          & very concrete   & lexical                  \\
+        total sentences                   & wordsent         & surface          & very concrete   & syntactic                \\ \bottomrule
+        total number of named entities    & entity           & discourse        & concrete        & semantic                 \\
+        total number of unique adjectives & partofspeech     & syntax           & concrete        & syntactic                \\ \bottomrule
+        simple type token ratio           & typetokenratio   & lexico-semantics & regular         & pragmatic                \\
+        number of verbs per word          & partofspeech     & surface          & regular         & syntactic                \\ \bottomrule
+        corrected adjective variation     & lexicalvariation & lexico-semantics & abstract        & lexical                  \\
+        corrected verb variation          & lexicalvariation & lexico-semantics & abstract        & lexical                  \\ \bottomrule
+        Flesch-Kincaid grade level        & readformula      & surface          & highly abstract & pragmatic                \\
+        Kuperman age of aquisition        & worddiff         & lexico-semantics & highly abstract & lexical                  \\ \bottomrule
+    \end{tabular}
+    \caption{Handcrafted Linguistic Features}\label{table-hlf}
+\end{table}
+\footnotetext{as described in the figure in~\cite{lugea2023stylistics}, chapter 1, page 7}
 
@@
 
-- Are there benchmark datasets for text style transfer? (STYLEPTB + Yelp Reviews)
+We have run this experiment on the datasets listed in Table~\ref{table-ds}.
+
+\begin{table}[ht]
+    \setlength\tabcolsep{6pt}
+    \centering
+    \begin{tabular}{@{}lllll@{}}\toprule
+        Name              & Task                       & Row Count & Content Type & Reference                     \\ \toprule
+        Yelp Reviews Full & Sentiment Classification   & 700000    & Reviews      & \cite{yelp2015neurips}        \\
+                          & Text Style Transfer        & 700000    &              &                               \\
+                          & Text Classification        & 700000    &              &                               \\ \midrule
+        CNN / DailyMail   & Summarization              & 300000    & News         & \cite{cnndm2015}              \\
+                          & Question Answering         &           & Stories      & \cite{cnndm2017}              \\
+                          & Text Generation            &           &              &                               \\ \midrule
+        HANNA             & Automatic Story Generation & 1056      & Stories      & \cite{hanna2024}              \\ \midrule
+        StylePTB          & Text Style Transfer        & 60000     & Sentences    & \cite{lyu-etal-2021-styleptb} \\ \bottomrule
+    \end{tabular}
+    \caption{Experiment LLMs}\label{table-ds}
+\end{table}
+
+The Yelp review data set provides a diverse assortment of writing styles and
+thus allowing us to home in on an LLMs ability to follow HLFs instead of copying
+a writing style.
+The reviews are often relatively short texts written in a coloquial style with
+a fair amount of variation in sentiment.
+
+On the other hand the CNN and DailyMail stories are news stories and even though
+their style might vary from one author to another, they will definitely be more
+formal than Yelp reviews.
+The writing is likely to be more mature, too.
+
+The Human-ANnotated NArratives (HANNA) dataset is interesting because it
+contains stories generated by LLMs. Since LLM generated stories can be assessed
+just as easily as human generated stories by using HLFs, we thought it would be
+a welcome addition to run our experiment for non-human writers, too.
+
+Finally, StylePTB is a dataset containing fine grained stylistic changes at the
+sentence level. We use this dataset to see how well LLMs do in scenarios
+involving HLFs at the sentence level (for example, part of speech HLFs).
 
@@
 
-- link to text from within the corpus
-- link to text from outside the corpus
-- link to each system prompt (2 baselines, 2 HLF prompt templates)
+The input prompts are derived from the templates available in the Git repository
+that accompanies this article~\cite{olar2024experimentcode}.
+We provide the LLM with a system prompt that contains instructions concerning
+all HLFs.
 
@@
 
-Ox - the number of the generation 1..10
-Oy - HLF value
-Continuous grey line - baseline metric
-continuous grenat line - hlf metric
-grey around grenat line - corpus target zone
-dotted black lines - corpus min, max
-continuous black line - corpus avg
+Handcrafted linguistic features are integer or floating point values that we
+can compute on text.
+All our graphs show the HLF on the vertical axis, while the horizontal axis is
+reserved for the trial number.
+The plots lend themselves to easy interpretation based on some simple rules.
+We plot the HLF values for the text generated from baseline prompts as a
+continuous grey line.
+The similar values obtained by using HLF enhanced prompts are plotted using a
+continuous grenat line.
+To highlight the `tolerance' obtained by adding and subtracting the variance of
+the HLF on the benchmark dataset, we plot light gray areas around the baseline
+and HLF-enhanced plot lines.
+The minimum and maximum HLF on the benchmark dataset are plotted using dotted
+black lines.
+A continuous black line designates the average HLF on the benchmark dataset.
 
-\subsection{Baseline}
+Following this description of how to interpret the graphs, we present sample
+findings using the features for which we've observed the best and the worst
+compliance by the LLM.
+These findings were all made on the <DECIDE LATER> dataset.
 
-Yelp Results
+\subsection{Can an LLM Follow HLF Instructions?}
 
-STYLEPTB Results
+Yelp Results on the best and worst performing metric.
+The rest of the metrics in an Appendix.
 
@@
 
-Yelp Results
-StylePTB Results
+Yelp results on the best and worst performing metric.
+The rest of the metrics (including on all other datasets) in an Appendix.
 

### 2024-06-03 src/llm_tst_consistency/evaluator.py
create prompts for HLF instructions

@@ -1,0 +1,70 @@
+from dataclasses import asdict
+from pathlib import Path
+
+import lftk
+import spacy
+from datasets import load_dataset
+from jinja2 import Environment, FileSystemLoader, select_autoescape
+from spacy.tokens.doc import Doc
+
+from llm_tst_consistency.stats import Stats
+
+
+def load_template(name):
+    templates_dir = Path(__file__).parent / "templates"
+    env = Environment(
+        loader=FileSystemLoader(templates_dir),
+        autoescape=select_autoescape()
+    )
+    return env.get_template(name)
+
+
+def is_cnn(record: dict) -> bool:
+    return not record.get("article", "").upper().startswith("(CNN)")
+
+
+def load_cnn_daily_mail():
+    return load_dataset("abisee/cnn_dailymail", "3.0.0", split="test").filter(is_cnn)
+
+
+HLF_MAP = {
+    "t_uword": "total_unique_words.j2",
+    "t_sent": "total_sentences.j2",
+    "t_n_ent": "total_named_entities.j2",
+    "n_uadj": "total_unique_adjectives.j2",
+    "simp_ttr": "simple_type_token_ratio.j2",
+    "a_verb_pw": "avg_verbs_per_word.j2",
+    "corr_adj_var": "corrected_adjective_variation.j2",
+    "corr_verb_var": "corrected_verb_variation.j2",
+    "fkgl": "grade_level.j2",
+    "a_kup_ps": "kuperman_age.j2",
+}
+
+
+def render_hlf_instruction(template_name: str, stats: Stats) -> str:
+    template = load_template(template_name)
+    return template.render(**asdict(stats))
+
+
+class HandcraftedLinguisticFeature:
+    def __init__(self, name: str):
+        self._name = name
+
+    def __call__(self, doc: Doc) -> float:
+        extractor = lftk.Extractor(docs=doc)
+        return extractor.extract(features=[self._name])[self._name]
+
+
+if __name__ == "__main__":
+    tpl = load_template("prompt_1.j2")
+    nlp = spacy.load("en_core_web_sm")
+    ds = load_cnn_daily_mail()
+    corpus = [row["article"] for row in ds][:10]
+    docs = list(map(nlp, corpus))
+    ds_stats = {
+        HLF_MAP[key]: Stats.hlf(docs, HandcraftedLinguisticFeature(key))
+        for key in HLF_MAP
+    }
[…]

### 2024-06-03 src/llm_tst_consistency/stats/__init__.py

@@ -1,0 +1,1 @@
+from llm_tst_consistency.stats._stats import Stats

### 2024-06-03 src/llm_tst_consistency/stats/_stats.py

@@ -1,0 +1,79 @@
+from dataclasses import dataclass, field
+from sys import maxsize
+from typing import Callable, Any, Iterable
+
+
+@dataclass(unsafe_hash=True)
+class Stats:
+    """Stats container."""
+
+    mean: float = field(init=True, hash=True, compare=True)
+    """average over a container of numbers"""
+    low: float = field(init=True, hash=True, compare=True)
+    """minimum value from a container of numbers"""
+    high: float = field(init=True, hash=True, compare=True)
+    """maximum value from a container of numbers"""
+    variance: float = field(init=True, hash=True, compare=True)
+    """variance computed over a container of numbers"""
+
+    @classmethod
+    def empty(cls) -> "Stats":
+        """Return an object that has all stats initialized to 0."""
+        return Stats(0, 0, 0, 0)
+
+    @classmethod
+    def hlf(cls, docs: list, hlf: Callable[[Any], float]) -> "Stats":
+        computer = Welford(hlf)
+        return computer(docs)
+
+
+class Welford:
+    def __init__(self, extract_x_callback: Callable[[Any], float]) -> None:
+        self._extract_numeric_value = extract_x_callback
+
+    @classmethod
+    def _welford_pass(
+        cls, result: Stats, x: float, n: int, m2: float
+    ) -> tuple[Stats, int, float]:
+        """Implement a single pass of the Welford algorithm.
+
+        More details on the algorithm can be found here:
+        * https://natural-blogarithm.com/post/variance-welford-vs-numpy/
+        * https://en.wikipedia.org/wiki/Algorithms_for_calculating_variance
+        """
+        result.high = max(result.high, x)
+        result.low = min(result.low, x)
+
+        n += 1
+        delta = x - result.mean
+        result.mean += float(delta) / n
+        delta2 = x - result.mean
+        m2 += delta * delta2
+
+        return result, n, m2
+
+    @classmethod
+    def _compute_final_numeric_result(
+        cls, result: Stats, n: int, m2: float
+    ) -> Stats:
+        if n < 1:
+            return Stats.empty()
+        if n > 1:
+            result.variance = m2 / float(n - 1)
+        result.low = round(result.low, 2)
+        result.high = round(result.high, 2)
+        result.mean = round(result.mean, 2)
+        result.variance = round(result.variance, 2)
+        return result
+
+    def __call__(self, items: Iterable) -> Stats:
+        if not items or not any(item for item in items):
+            return Stats.empty()
[…]

### 2024-06-03 src/llm_tst_consistency/evaluator.py
small tweaks

@@ -37,7 +37,7 @@ HLF_MAP = {
     "corr_adj_var": "corrected_adjective_variation.j2",
     "corr_verb_var": "corrected_verb_variation.j2",
     "fkgl": "grade_level.j2",
-    "a_kup_ps": "kuperman_age.j2",
+    "a_kup_pw": "kuperman_age.j2",
 }
 
 
@@ -67,4 +67,5 @@ if __name__ == "__main__":
     }
     instr = list(map(lambda x: render_hlf_instruction(*x), ds_stats.items()))
     print(tpl.render())
+    print("----------------------------------------------------------------")
     print(tpl.render(hlf_instructions=instr))

### 2024-06-04 src/llm_tst_consistency/evaluator.py
add plots

@@ -1,12 +1,16 @@
+import os.path
 from dataclasses import asdict
 from pathlib import Path
 
-import lftk
+import orjson as json
+import pandas as pd
 import spacy
 from datasets import load_dataset
 from jinja2 import Environment, FileSystemLoader, select_autoescape
-from spacy.tokens.doc import Doc
 
+from llm_tst_consistency.hlf import HandcraftedLinguisticFeature
+from llm_tst_consistency.llm import Ollama
+from llm_tst_consistency.plot import draw_plots
 from llm_tst_consistency.stats import Stats
 
 
@@ -27,7 +31,7 @@ def load_cnn_daily_mail():
     return load_dataset("abisee/cnn_dailymail", "3.0.0", split="test").filter(is_cnn)
 
 
-HLF_MAP = {
+HLF_TEMPLATES = {
     "t_uword": "total_unique_words.j2",
     "t_sent": "total_sentences.j2",
     "t_n_ent": "total_named_entities.j2",
@@ -39,6 +43,7 @@ HLF_MAP = {
     "fkgl": "grade_level.j2",
     "a_kup_pw": "kuperman_age.j2",
 }
+HLFs = {name: HandcraftedLinguisticFeature(name) for name in HLF_TEMPLATES}
 
 
 def render_hlf_instruction(template_name: str, stats: Stats) -> str:
@@ -46,13 +51,7 @@ def render_hlf_instruction(template_name: str, stats: Stats) -> str:
     return template.render(**asdict(stats))
 
 
-class HandcraftedLinguisticFeature:
-    def __init__(self, name: str):
-        self._name = name
-
-    def __call__(self, doc: Doc) -> float:
-        extractor = lftk.Extractor(docs=doc)
-        return extractor.extract(features=[self._name])[self._name]
+INPUT = "Interesting fact: by the year 2020 all actors on american tv shows will be australian."
 
 
 if __name__ == "__main__":
@@ -62,10 +61,30 @@ if __name__ == "__main__":
     corpus = [row["article"] for row in ds][:10]
     docs = list(map(nlp, corpus))
     ds_stats = {
-        HLF_MAP[key]: Stats.hlf(docs, HandcraftedLinguisticFeature(key))
-        for key in HLF_MAP
+        key: Stats.hlf(docs, HandcraftedLinguisticFeature(key))
+        for key in HLF_TEMPLATES
     }
[…]

### 2024-06-04 src/llm_tst_consistency/hlf/__init__.py

@@ -1,0 +1,4 @@
+from llm_tst_consistency.hlf._hlf import HandcraftedLinguisticFeature
+
+
+__all__ = ["HandcraftedLinguisticFeature"]

### 2024-06-04 src/llm_tst_consistency/hlf/_hlf.py

@@ -1,0 +1,11 @@
+import lftk
+from spacy.tokens import Doc
+
+
+class HandcraftedLinguisticFeature:
+    def __init__(self, name: str):
+        self._name = name
+
+    def __call__(self, doc: Doc) -> float:
+        extractor = lftk.Extractor(docs=doc)
+        return extractor.extract(features=[self._name])[self._name]

### 2024-06-04 src/llm_tst_consistency/llm/__init__.py

@@ -1,0 +1,4 @@
+from llm_tst_consistency.llm._llama import Ollama
+
+
+__all__ = ["Ollama"]

### 2024-06-04 src/llm_tst_consistency/llm/_llama.py

@@ -1,0 +1,24 @@
+import ollama
+import spacy
+
+from llm_tst_consistency.hlf import HandcraftedLinguisticFeature
+
+
+class Ollama:
+    def __init__(
+        self, text: str, prompt: str, hlf_cfg: dict[str, HandcraftedLinguisticFeature]
+    ) -> None:
+        self._messages = [
+            {"role": "system", "content": prompt},
+            {"role": "user", "content": text},
+        ]
+        self._hlf_cfg = hlf_cfg
+        self._nlp = spacy.load("en_core_web_sm")
+
+    def __call__(self) -> dict[str, float]:
+        response = ollama.chat(model="llama3", messages=self._messages)
+        doc = self._nlp(response["message"]["content"])
+        return {
+            name: hlf(doc)
+            for name, hlf in self._hlf_cfg.items()
+        }

### 2024-06-04 src/llm_tst_consistency/plot.py

@@ -1,0 +1,79 @@
+import pandas as pd
+import plotly.graph_objects as go
+from plotly.subplots import make_subplots
+
+from llm_tst_consistency.stats import Stats
+
+
+def draw_plots(
+    title: str,
+    df: pd.DataFrame,
+    columns: list[str],
+    benchmark_stats: dict[str, Stats]
+):
+
+    # Assuming df is your DataFrame
+    for metric in columns:
+        df[f"{metric}_varlo"] = df[metric] - benchmark_stats[metric].variance
+        df[f"{metric}_varhi"] = df[metric] + benchmark_stats[metric].variance
+
+    COL_COUNT = 2
+    fig = make_subplots(rows=len(columns)//COL_COUNT+1, subplot_titles=columns, cols=COL_COUNT)
+
+    for i, metric in enumerate(columns):
+        fig.add_trace(
+            go.Scatter(
+                x=df.index,
+                y=df[metric],
+                mode='lines',
+                name=metric,
+                legendgroup=metric,
+                line={"color": "#6E0B14"}
+            ),
+            row=i//COL_COUNT+1,
+            col=i%COL_COUNT+1
+        )
+        fig.add_trace(
+            go.Scatter(
+                x=df.index,
+                y=[benchmark_stats[metric].low] * len(df),
+                mode='lines',
+                name=f'corpus {metric} min',
+                line=dict(dash='dot', color="#676767"),
+                showlegend=False,
+                legendgroup=metric
+            ),
+            row=i//COL_COUNT+1,
+            col=i%COL_COUNT+1
+        )
+
+        fig.add_trace(
+            go.Scatter(
+                x=df.index,
+                y=[benchmark_stats[metric].high] * len(df),
+                mode='lines',
+                name=f'corpus {metric} max',
+                line=dict(dash='dot', color="#676767"),
+                showlegend=False,
+                legendgroup=metric
+            ),
+            row=i//COL_COUNT+1,
+            col=i%COL_COUNT+1
+        )
+
+        fig.add_trace(
+            go.Scatter(
+                x=df.index,
+                y=[benchmark_stats[metric].mean] * len(df),
+                mode='lines',
+                name=f'corpus {metric} average',
+                line={"color": "#000000"},
+                showlegend=False,
+                legendgroup=metric,
+            ),
+            row=i//COL_COUNT+1,
+            col=i%COL_COUNT+1
+        )
+
+    fig.update_layout(height=200*len(columns), width=900, title_text=title)
+    fig.show()

### 2024-06-04 src/llm_tst_consistency/evaluator.py
tweak the prompt

@@ -2,7 +2,6 @@ import os.path
 from dataclasses import asdict
 from pathlib import Path
 
-import orjson as json
 import pandas as pd
 import spacy
 from datasets import load_dataset
@@ -51,7 +50,69 @@ def render_hlf_instruction(template_name: str, stats: Stats) -> str:
     return template.render(**asdict(stats))
 
 
-INPUT = "Interesting fact: by the year 2020 all actors on american tv shows will be australian."
+INPUT = """
+Young people with internet addiction experience changes in their brain chemistry
+which could lead to more addictive behaviours, research suggests.
+
+The study, published in PLOS Mental Health, reviewed previous research using
+functional magnetic resonance imaging (fMRI) to examine how regions of the brain
+interact in people with internet addiction.
+
+They found that the effects were evident throughout multiple neural networks in
+the brains of young people, and that there was increased activity in parts of
+the brain when participants were resting.
+
+At the same time, there was an overall decrease in the functional connectivity
+in parts of the brain involved in active thinking, which is the executive
+control network of the brain responsible for memory and decision-making.
+
+The research found that these changes resulted in addictive behaviours and
+tendencies in adolescents, as well as behavioural changes linked to mental
+health, development, intellectual ability and physical coordination.
+
+The researchers reviewed 12 previous studies involving 237 10- to 19-year-olds
+with a formal diagnosis of internet addiction between 2013 and 2023.
+
+Almost half of British teenagers have said they feel addicted to social media,
+according to a survey this year.
+
+Max Chang, the study’s lead author and an MSc student at the UCL Great Ormond
+Street Institute of Child Health (GOS ICH), said: “Adolescence is a crucial
+developmental stage during which people go through significant changes in their
+biology, cognition and personalities.
+
+“As a result, the brain is particularly vulnerable to internet addiction-related
+urges during this time, such as compulsive internet usage, cravings towards
+usage of the mouse or keyboard and consuming media.
+
+“The findings from our study show that this can lead to potentially negative
+behavioural and developmental changes that could impact the lives of
+adolescents. For example, they may struggle to maintain relationships and social
+activities, lie about online activity and experience irregular eating and
+disrupted sleep.”
+
+Chang added that he hoped the findings demonstrated “how internet addiction
+alters the connection between the brain networks in adolescence”, which would
+then allow early signs of internet addiction to be treated effectively.
+
+He added: “Clinicians could potentially prescribe treatment to aim at certain
[…]

### 2024-06-05 src/llm_tst_consistency/evaluator.py
update prompts. update plots

@@ -4,8 +4,9 @@ from pathlib import Path
 
 import pandas as pd
 import spacy
-from datasets import load_dataset
+from datasets import load_dataset, Dataset
 from jinja2 import Environment, FileSystemLoader, select_autoescape
+from spacy import Language
 
 from llm_tst_consistency.hlf import HandcraftedLinguisticFeature
 from llm_tst_consistency.llm import Ollama
@@ -43,11 +44,6 @@ HLF_TEMPLATES = {
     "a_kup_pw": "kuperman_age.j2",
 }
 HLFs = {name: HandcraftedLinguisticFeature(name) for name in HLF_TEMPLATES}
-
-
-def render_hlf_instruction(template_name: str, stats: Stats) -> str:
-    template = load_template(template_name)
-    return template.render(**asdict(stats))
 
 
 INPUT = """
@@ -115,37 +111,46 @@ implications of spending too much time online.”
 """
 
 
-if __name__ == "__main__":
-    tpl = load_template("prompt_1.j2")
-    nlp = spacy.load("en_core_web_sm")
-    ds = load_cnn_daily_mail()
+def _compute_dataset_features(nlp: Language, ds: Dataset) -> dict[str, Stats]:
     corpus = [row["article"] for row in ds][:10]
     docs = list(map(nlp, corpus))
-    ds_stats = {
+    return {
         key: Stats.hlf(docs, HandcraftedLinguisticFeature(key))
         for key in HLF_TEMPLATES
     }
-    instr = list(map(lambda x: render_hlf_instruction(HLF_TEMPLATES[x], ds_stats[x]), ds_stats))
-    column_names = list(HLFs)
 
-    baseline_csv = "baseline_llama_3.csv"
-    if not os.path.exists(baseline_csv):
-        baseline_prompt = tpl.render()
-        generate_text = Ollama(INPUT, baseline_prompt, HLFs)
-        stats_df = pd.DataFrame(data=[generate_text() for _ in range(10)], columns=column_names)
-        stats_df.to_csv(baseline_csv)
+
+def _write_hlf_instructions(ds_stats: dict[str, Stats]) -> list[str]:
+    result = []
+    for feature, stats in ds_stats.items():
+        template = load_template(HLF_TEMPLATES[feature])
+        result.append(template.render(**asdict(stats)))
+    return result
+
+
+generate = True
+if __name__ == "__main__":
+    file_name = Path(__file__).parent / "data.csv"
[…]

### 2024-06-05 src/llm_tst_consistency/llm/_llama.py

@@ -15,10 +15,16 @@ class Ollama:
         self._hlf_cfg = hlf_cfg
         self._nlp = spacy.load("en_core_web_sm")
 
-    def __call__(self) -> dict[str, float]:
+    @staticmethod
+    def _apply_prefix(name: str, prefix: str) -> str:
+        if prefix is None or len(prefix.strip()) < 1:
+            return name
+        return f"{prefix}_{name}"
+
+    def __call__(self, prefix: str = "") -> dict[str, float]:
         response = ollama.chat(model="llama3", messages=self._messages)
         doc = self._nlp(response["message"]["content"])
         return {
-            name: hlf(doc)
+            self._apply_prefix(name, prefix): hlf(doc)
             for name, hlf in self._hlf_cfg.items()
         }

### 2024-06-05 src/llm_tst_consistency/plot.py

@@ -1,79 +1,94 @@
+from typing import Callable
+
 import pandas as pd
 import plotly.graph_objects as go
+from plotly.basedatatypes import BaseTraceType
+from plotly.graph_objs import Figure
 from plotly.subplots import make_subplots
 
 from llm_tst_consistency.stats import Stats
 
 
+def _add_variance(
+        df: pd.DataFrame,
+        benchmark_stats: dict[str, Stats],
+        features: list[str],
+        name_extractor: Callable[[str], str]
+) -> pd.DataFrame:
+    for feature in features:
+        df[f"{name_extractor(feature)}_varlo"] = df[f"{name_extractor(feature)}"] - benchmark_stats[feature].variance
+        df[f"{name_extractor(feature)}_varhi"] = df[f"{name_extractor(feature)}"] + benchmark_stats[feature].variance
+    return df
+
+
+def _baseline_name(feature: str) -> str:
+    return f"baseline_{feature}"
+
+
+def _hlf_name(feature: str) -> str:
+    return f"hlf_{feature}"
+
+
+def _create_benchmark_trace(
+    df: pd.DataFrame, feature: str, title: str, benchmark_value: float, color: str = "#575757", dash: str = "dot"
+) -> BaseTraceType:
+    return go.Scatter(
+        x=df.index,
+        y=[benchmark_value] * len(df),
+        mode='lines',
+        name=title,
+        line={"color": color, "dash": dash},
+        showlegend=False,
+        legendgroup=feature,
+    )
+
+
+def _create_baseline(df: pd.DataFrame, feature: str) -> BaseTraceType:
+    return go.Scatter(
+        x=df.index,
+        y=df[_baseline_name(feature)],
+        mode='lines',
+        name=f"baseline {feature}",
+        legendgroup=feature,
+        line={"color": "#575757"}
+    )
+
+
+def _create_hlf(df: pd.DataFrame, feature: str) -> BaseTraceType:
+    return go.Scatter(
+        x=df.index,
+        y=df[_hlf_name(feature)],
+        mode='lines',
+        name=f"hlf-enhanced {feature}",
+        legendgroup=feature,
+        line={"color": "#6E0B14"}
+    )
+
+
+def _add_trace_to_plot(fig: Figure, trace: BaseTraceType, i: int, col_count: int) -> None:
+    fig.add_trace(trace, row=i // col_count + 1, col=i % col_count + 1)
+
+
 def draw_plots(
[…]

### 2024-06-05 src/llm_tst_consistency/evaluator.py
add benchmark ds stats. compute statistical significance

@@ -1,7 +1,9 @@
-import os.path
 from dataclasses import asdict
+from math import isnan, erf
 from pathlib import Path
 
+import numpy as np
+import orjson as json
 import pandas as pd
 import spacy
 from datasets import load_dataset, Dataset
@@ -128,12 +130,42 @@ def _write_hlf_instructions(ds_stats: dict[str, Stats]) -> list[str]:
     return result
 
 
-generate = True
+def areas_to_target(a: list[float], b: list[float], target: list[float]) -> tuple[float, float]:
+    area_a = np.trapz(np.abs(np.array(a) - np.array(target)))
+    area_b = np.trapz(np.abs(np.array(b) - np.array(target)))
+    return area_a, area_b
+
+
+def distance_norms_to_target(a: list[float], b: list[float], target: list[float]) -> tuple[float, float]:
+    distance_a = np.linalg.norm(np.array(a) - np.array(target))
+    distance_b = np.linalg.norm(np.array(b) - np.array(target))
+    return distance_a, distance_b
+
+
+def is_significant(x: float, y: float, confidence: float = 0.95) -> bool:
+    c = x - y
+    std_dev = np.std([x, y])
+    z_score = c / std_dev
+    p_value = 1 - 0.5 * (1 + erf(z_score / np.sqrt(2)))
+    threshold = 1 - confidence
+    if isnan(p_value):
+        return False
+    # Return the p-value and a message indicating whether c is statistically significant
+    return p_value < threshold
+
+
 if __name__ == "__main__":
     file_name = Path(__file__).parent / "data.csv"
     features = list(HLFs)
     tpl = load_template("prompt_1.j2")
-    ds_stats = _compute_dataset_features(spacy.load("en_core_web_sm"), load_cnn_daily_mail())
+    stats_file_name = Path(__file__).parent / "cnn_dailymail.json"
+    if not stats_file_name.exists():
+        ds_stats = _compute_dataset_features(spacy.load("en_core_web_sm"), load_cnn_daily_mail())
+        stats_file_name.write_bytes(json.dumps(ds_stats))
+    else:
+        ds_stats = json.loads(stats_file_name.read_text())
+        for key in ds_stats:
+            ds_stats[key] = Stats(**ds_stats[key])
     if not file_name.exists():
         # generate baseline
         baseline_generator = Ollama(INPUT, tpl.render(), HLFs)
@@ -152,5 +184,21 @@ if __name__ == "__main__":
[…]

### 2024-06-05 src/llm_tst_consistency/evaluator.py
refactor statistical significance code

@@ -142,15 +142,16 @@ def distance_norms_to_target(a: list[float], b: list[float], target: list[float]
     return distance_a, distance_b
 
 
-def is_significant(x: float, y: float, confidence: float = 0.95) -> bool:
-    c = x - y
+def is_diff_significant(x: float, y: float, confidence: float = 0.95) -> bool:
     std_dev = np.std([x, y])
-    z_score = c / std_dev
+    z_score = abs(x - y) / std_dev
+
+    # need to double-check this formula
     p_value = 1 - 0.5 * (1 + erf(z_score / np.sqrt(2)))
-    threshold = 1 - confidence
     if isnan(p_value):
         return False
-    # Return the p-value and a message indicating whether c is statistically significant
+
+    threshold = 1 - confidence
     return p_value < threshold
 
 
@@ -195,10 +196,10 @@ if __name__ == "__main__":
 
         dist_a, dist_b = distance_norms_to_target(a, b, target)
         euclid[feature] = dist_a - dist_b
-        euclid[f"is_{feature}_significant"] = is_significant(dist_a, dist_b)
+        euclid[f"is_{feature}_significant"] = is_diff_significant(dist_a, dist_b)
 
         area_a, area_b = areas_to_target(a, b, target)
         area[feature] = area_a - area_b
-        area[f"is_{feature}_significant"] = is_significant(area_a, area_b)
+        area[f"is_{feature}_significant"] = is_diff_significant(area_a, area_b)
 
     diff_df = pd.DataFrame(columns=list(euclid.keys()), data=[euclid, area])

### 2024-06-05 src/llm_tst_consistency/evaluator.py
refactor generation code

@@ -171,17 +171,18 @@ if __name__ == "__main__":
         # generate baseline
         baseline_generator = Ollama(INPUT, tpl.render(), HLFs)
         data = [baseline_generator("baseline") for _ in range(10)]
+
         # augment with hlf instructions
         hlf_generator = Ollama(INPUT, tpl.render(hlf_instructions=_write_hlf_instructions(ds_stats)), HLFs)
         for obj in data:
             obj.update(hlf_generator("hlf"))
+
         # create dataframe
-        column_names = [
+        df = pd.DataFrame(data=data, columns=[
             f"{metric_set}_{feature}"
             for metric_set in ["baseline", "hlf"]
             for feature in features
-        ]
-        df = pd.DataFrame(data=data, columns=column_names)
+        ])
         df.to_csv(file_name)
     else:
         df = pd.read_csv(file_name)

### 2024-06-05 src/llm_tst_consistency/evaluator.py
some more tweaks

@@ -144,7 +144,7 @@ def distance_norms_to_target(a: list[float], b: list[float], target: list[float]
 
 def is_diff_significant(x: float, y: float, confidence: float = 0.95) -> bool:
     std_dev = np.std([x, y])
-    z_score = abs(x - y) / std_dev
+    z_score = x - y / std_dev
 
     # need to double-check this formula
     p_value = 1 - 0.5 * (1 + erf(z_score / np.sqrt(2)))

### 2024-06-05 src/llm_tst_consistency/evaluator.py
implement kolmogorov-smirnoff test

@@ -8,6 +8,7 @@ import pandas as pd
 import spacy
 from datasets import load_dataset, Dataset
 from jinja2 import Environment, FileSystemLoader, select_autoescape
+from scipy.stats import ks_2samp
 from spacy import Language
 
 from llm_tst_consistency.hlf import HandcraftedLinguisticFeature
@@ -190,17 +191,29 @@ if __name__ == "__main__":
     draw_plots("llama 3", df, features, ds_stats)
     euclid = {}
     area = {}
+    ks = {}
     for feature in features:
+        feat_sig_name = f"is_{feature}_significant"
         target = [ds_stats[feature].mean]*len(df)
         a = df[f"baseline_{feature}"].values
         b = df[f"hlf_{feature}"].values
 
+        # Kolmogorov-Smirnov test
+        ks_ab = ks_2samp(a, b)
+        ks_a = ks_2samp(a, target)
+        ks_b = ks_2samp(b, target)
+        ks[feature] = (ks_a.statistic - ks_b.statistic) < 0
+        ks[feat_sig_name] = ks_ab.pvalue < 0.25
+
+        # Euclidian distance
         dist_a, dist_b = distance_norms_to_target(a, b, target)
-        euclid[feature] = dist_a - dist_b
-        euclid[f"is_{feature}_significant"] = is_diff_significant(dist_a, dist_b)
+        euclid[feature] = (dist_a - dist_b) > 0
+        euclid[feat_sig_name] = is_diff_significant(dist_a, dist_b)
 
+        # Area surface to target
         area_a, area_b = areas_to_target(a, b, target)
-        area[feature] = area_a - area_b
-        area[f"is_{feature}_significant"] = is_diff_significant(area_a, area_b)
+        area[feature] = (area_a - area_b) > 0
+        area[feat_sig_name] = is_diff_significant(area_a, area_b)
 
-    diff_df = pd.DataFrame(columns=list(euclid.keys()), data=[euclid, area])
+    diff_df = pd.DataFrame(columns=list(euclid.keys()), data=[euclid, area, ks])
+    pass

### 2024-06-05 src/llm_tst_consistency/evaluator.py
add more explicit columns to the stats dataframe

@@ -193,27 +193,32 @@ if __name__ == "__main__":
     area = {}
     ks = {}
     for feature in features:
-        feat_sig_name = f"is_{feature}_significant"
+        diff_col = f"{feature}_diff"
+        is_closer_col = f"is_{feature}_closer"
+        is_different_col = f"is_{feature}_different"
         target = [ds_stats[feature].mean]*len(df)
-        a = df[f"baseline_{feature}"].values
-        b = df[f"hlf_{feature}"].values
+        baseline = df[f"baseline_{feature}"].values
+        hlf = df[f"hlf_{feature}"].values
 
         # Kolmogorov-Smirnov test
-        ks_ab = ks_2samp(a, b)
-        ks_a = ks_2samp(a, target)
-        ks_b = ks_2samp(b, target)
-        ks[feature] = (ks_a.statistic - ks_b.statistic) < 0
-        ks[feat_sig_name] = ks_ab.pvalue < 0.25
+        ks_bh = ks_2samp(baseline, hlf)
+        ks_bt = ks_2samp(baseline, target)
+        ks_ht = ks_2samp(hlf, target)
+        ks[diff_col] = ks_bt.statistic - ks_ht.statistic
+        ks[is_closer_col] = ks_bt.statistic > ks_ht.statistic
+        ks[is_different_col] = ks_bh.pvalue < 0.25
 
         # Euclidian distance
-        dist_a, dist_b = distance_norms_to_target(a, b, target)
-        euclid[feature] = (dist_a - dist_b) > 0
-        euclid[feat_sig_name] = is_diff_significant(dist_a, dist_b)
+        dist_a, dist_b = distance_norms_to_target(baseline, hlf, target)
+        euclid[diff_col] = dist_a - dist_b
+        euclid[is_closer_col] = dist_a > dist_b
+        euclid[is_different_col] = is_diff_significant(dist_a, dist_b)
 
         # Area surface to target
-        area_a, area_b = areas_to_target(a, b, target)
-        area[feature] = (area_a - area_b) > 0
-        area[feat_sig_name] = is_diff_significant(area_a, area_b)
+        area_a, area_b = areas_to_target(baseline, hlf, target)
+        area[diff_col] = area_a - area_b
+        area[is_closer_col] = area_a > area_b
+        area[is_different_col] = is_diff_significant(area_a, area_b)
 
     diff_df = pd.DataFrame(columns=list(euclid.keys()), data=[euclid, area, ks])
     pass

### 2024-06-05 src/llm_tst_consistency/evaluator.py
try new settings

@@ -27,11 +27,13 @@ def load_template(name):
 
 
 def is_cnn(record: dict) -> bool:
-    return not record.get("article", "").upper().startswith("(CNN)")
+    return "(CNN)" in record.get("article", "")
 
 
 def load_cnn_daily_mail():
-    return load_dataset("abisee/cnn_dailymail", "3.0.0", split="test").filter(is_cnn)
+    return load_dataset(
+        "abisee/cnn_dailymail", "3.0.0", split="test"
+    ).filter(is_cnn)
 
 
 HLF_TEMPLATES = {
@@ -49,73 +51,20 @@ HLF_TEMPLATES = {
 HLFs = {name: HandcraftedLinguisticFeature(name) for name in HLF_TEMPLATES}
 
 
-INPUT = """
-Young people with internet addiction experience changes in their brain chemistry
-which could lead to more addictive behaviours, research suggests.
+INPUT = """I've been to this cafe a few times. The service is great and the menu
+is affordable. It has a large spacious space, reliable free wifi, and good food.
 
-The study, published in PLOS Mental Health, reviewed previous research using
-functional magnetic resonance imaging (fMRI) to examine how regions of the brain
-interact in people with internet addiction.
+It is quite new and seems a bit confused with which demographic of clienteles
+they are catering to. It has a great potential to be a hub for creative, hipster
+crowd.
 
-They found that the effects were evident throughout multiple neural networks in
-the brains of young people, and that there was increased activity in parts of
-the brain when participants were resting.
-
-At the same time, there was an overall decrease in the functional connectivity
-in parts of the brain involved in active thinking, which is the executive
-control network of the brain responsible for memory and decision-making.
-
-The research found that these changes resulted in addictive behaviours and
-tendencies in adolescents, as well as behavioural changes linked to mental
-health, development, intellectual ability and physical coordination.
-
-The researchers reviewed 12 previous studies involving 237 10- to 19-year-olds
-with a formal diagnosis of internet addiction between 2013 and 2023.
-
-Almost half of British teenagers have said they feel addicted to social media,
-according to a survey this year.
-
-Max Chang, the study’s lead author and an MSc student at the UCL Great Ormond
-Street Institute of Child Health (GOS ICH), said: “Adolescence is a crucial
-developmental stage during which people go through significant changes in their
-biology, cognition and personalities.
-
[…]

### 2024-06-05 src/llm_tst_consistency/stats/_stats.py

@@ -28,8 +28,8 @@ class Stats:
 
 
 class Welford:
-    def __init__(self, extract_x_callback: Callable[[Any], float]) -> None:
-        self._extract_numeric_value = extract_x_callback
+    def __init__(self, number_extractor: Callable[[Any], float]) -> None:
+        self._get_x = number_extractor
 
     @classmethod
     def _welford_pass(
@@ -72,8 +72,9 @@ class Welford:
         result = Stats(0, maxsize, -1, 0)
         count = 0
         m2: float = 0
-        for text in items:
+        for i, item in enumerate(items):
             result, count, m2 = self._welford_pass(
-                result, self._extract_numeric_value(text), count, m2
+                result, self._get_x(item), count, m2
             )
+            print("processing item", i)
         return self._compute_final_numeric_result(result, count, m2)

### 2024-06-07 src/llm_tst_consistency/evaluator.py
add openai support. tweak the prompts

@@ -2,6 +2,7 @@ from dataclasses import asdict
 from math import isnan, erf
 from pathlib import Path
 
+import dotenv
 import numpy as np
 import orjson as json
 import pandas as pd
@@ -11,10 +12,13 @@ from jinja2 import Environment, FileSystemLoader, select_autoescape
 from scipy.stats import ks_2samp
 from spacy import Language
 
-from llm_tst_consistency.hlf import HandcraftedLinguisticFeature
-from llm_tst_consistency.llm import Ollama
+from llm_tst_consistency.hlf import HandcraftedLinguisticFeature, KupermanAgeOfAcquisition
+from llm_tst_consistency.llm import OpenAI
 from llm_tst_consistency.plot import draw_plots
 from llm_tst_consistency.stats import Stats
+
+
+MAX_CORPUS_SIZE = 100
 
 
 def load_template(name):
@@ -37,10 +41,10 @@ def load_cnn_daily_mail():
 
 
 HLF_TEMPLATES = {
-    "t_uword": "total_unique_words.j2",
+    "t_word": "total_words.j2",
     "t_sent": "total_sentences.j2",
-    "t_n_ent": "total_named_entities.j2",
-    "n_uadj": "total_unique_adjectives.j2",
+    "n_uverb": "number_of_unique_verbs.j2",
+    "n_uadj": "number_of_unique_adjectives.j2",
     "simp_ttr": "simple_type_token_ratio.j2",
     "a_verb_pw": "avg_verbs_per_word.j2",
     "corr_adj_var": "corrected_adjective_variation.j2",
@@ -48,7 +52,10 @@ HLF_TEMPLATES = {
     "fkgl": "grade_level.j2",
     "a_kup_pw": "kuperman_age.j2",
 }
-HLFs = {name: HandcraftedLinguisticFeature(name) for name in HLF_TEMPLATES}
+HLFs = {
+    name: KupermanAgeOfAcquisition(name) if name == "a_kup_pw" else HandcraftedLinguisticFeature(name)
+    for name in HLF_TEMPLATES
+}
 
 
 INPUT = """I've been to this cafe a few times. The service is great and the menu
@@ -64,11 +71,11 @@ stars they deserve.
 
 
 def _compute_dataset_features(nlp: Language, ds: Dataset) -> dict[str, Stats]:
-    corpus = [row["article"] for row in ds]
+    corpus = [row["article"] for row in ds][:MAX_CORPUS_SIZE]
     docs = list(map(nlp, corpus))
     return {
[…]

### 2024-06-07 src/llm_tst_consistency/hlf/__init__.py

@@ -1,4 +1,4 @@
-from llm_tst_consistency.hlf._hlf import HandcraftedLinguisticFeature
+from llm_tst_consistency.hlf._hlf import HandcraftedLinguisticFeature, KupermanAgeOfAcquisition
 
 
-__all__ = ["HandcraftedLinguisticFeature"]
+__all__ = ["HandcraftedLinguisticFeature", "KupermanAgeOfAcquisition"]

### 2024-06-07 src/llm_tst_consistency/hlf/_hlf.py

@@ -1,4 +1,7 @@
+from pathlib import Path
+
 import lftk
+import pandas as pd
 from spacy.tokens import Doc
 
 
@@ -9,3 +12,22 @@ class HandcraftedLinguisticFeature:
     def __call__(self, doc: Doc) -> float:
         extractor = lftk.Extractor(docs=doc)
         return extractor.extract(features=[self._name])[self._name]
+
+
+class KupermanAgeOfAcquisition(HandcraftedLinguisticFeature):
+    TABLE_PATH = Path(__file__).parent / "kup-aoa-ratings.csv"
+
+    def __init__(self, name: str):
+        super().__init__(name)
+        self._aoa_ratings = pd.read_csv(self.TABLE_PATH)
+
+    def __call__(self, doc: Doc) -> float:
+        unique_words = set(token.lemma_.lower() for token in doc)
+        aoa_of_unique_words = [
+            round(rating)
+            for word in unique_words
+            for rating in self._aoa_ratings.loc[self._aoa_ratings["Word"] == word, "Rating.Mean"]
+        ]
+        if len(aoa_of_unique_words) == 0:
+            return 100
+        return round(sum(aoa_of_unique_words) / len(aoa_of_unique_words), 2)

### 2024-06-07 src/llm_tst_consistency/llm/__init__.py

@@ -1,4 +1,5 @@
 from llm_tst_consistency.llm._llama import Ollama
+from llm_tst_consistency.llm._openai import OpenAI
 
 
-__all__ = ["Ollama"]
+__all__ = ["Ollama", "OpenAI"]

### 2024-06-07 src/llm_tst_consistency/llm/_openai.py

@@ -1,10 +1,12 @@
-import ollama
+import os
+
+import openai
 import spacy
 
 from llm_tst_consistency.hlf import HandcraftedLinguisticFeature
 
 
-class Ollama:
+class OpenAI:
     def __init__(
         self, text: str, prompt: str, hlf_cfg: dict[str, HandcraftedLinguisticFeature]
     ) -> None:
@@ -14,7 +16,7 @@ class Ollama:
         ]
         self._hlf_cfg = hlf_cfg
         self._nlp = spacy.load("en_core_web_sm")
-        self._client = ollama.Client(host="172.30.212.3")
+        self._client = openai.Client(api_key=os.getenv("LTC_OPENAI_KEY"))
 
     @staticmethod
     def _apply_prefix(name: str, prefix: str) -> str:
@@ -23,8 +25,10 @@ class Ollama:
         return f"{prefix}_{name}"
 
     def __call__(self, prefix: str = "") -> dict[str, float]:
-        response = self._client.chat(model="llama3:70b-text", messages=self._messages)
-        doc = self._nlp(response["message"]["content"])
+        response = self._client.chat.completions.create(
+            model="gpt-3.5-turbo", messages=self._messages
+        )
+        doc = self._nlp(response.choices[0].message.content)
         return {
             self._apply_prefix(name, prefix): hlf(doc)
             for name, hlf in self._hlf_cfg.items()

### 2024-06-07 src/llm_tst_consistency/stats/_stats.py

@@ -76,5 +76,4 @@ class Welford:
             result, count, m2 = self._welford_pass(
                 result, self._get_x(item), count, m2
             )
-            print("processing item", i)
         return self._compute_final_numeric_result(result, count, m2)

### 2024-06-07 src/llm_tst_consistency/llm/_openai.py
use gpt-4o

@@ -26,7 +26,7 @@ class OpenAI:
 
     def __call__(self, prefix: str = "") -> dict[str, float]:
         response = self._client.chat.completions.create(
-            model="gpt-3.5-turbo", messages=self._messages
+            model="gpt-4o", messages=self._messages
         )
         doc = self._nlp(response.choices[0].message.content)
         return {

### 2024-06-07 src/llm_tst_consistency/evaluator.py
Add Gemini 1.5 Pro

@@ -13,7 +13,7 @@ from scipy.stats import ks_2samp
 from spacy import Language
 
 from llm_tst_consistency.hlf import HandcraftedLinguisticFeature, KupermanAgeOfAcquisition
-from llm_tst_consistency.llm import OpenAI
+from llm_tst_consistency.llm import OpenAI, Gemini
 from llm_tst_consistency.plot import draw_plots
 from llm_tst_consistency.stats import Stats
 
@@ -129,11 +129,11 @@ if __name__ == "__main__":
             ds_stats[key] = Stats(**ds_stats[key])
     if not file_name.exists():
         # generate baseline
-        baseline_generator = OpenAI(INPUT, tpl.render(), HLFs)
+        baseline_generator = Gemini(INPUT, tpl.render(), HLFs)
         data = [baseline_generator("baseline") for _ in range(10)]
 
         # augment with hlf instructions
-        hlf_generator = OpenAI(INPUT, tpl.render(hlf_instructions=_write_hlf_instructions(ds_stats)), HLFs)
+        hlf_generator = Gemini(INPUT, tpl.render(hlf_instructions=_write_hlf_instructions(ds_stats)), HLFs)
         for obj in data:
             obj.update(hlf_generator("hlf"))
 

### 2024-06-07 src/llm_tst_consistency/llm/__init__.py

@@ -1,5 +1,6 @@
 from llm_tst_consistency.llm._llama import Ollama
-from llm_tst_consistency.llm._openai import OpenAI
+from llm_tst_consistency.llm._gemini import Gemini
+from llm_tst_consistency.llm._gpt import OpenAI
 
 
-__all__ = ["Ollama", "OpenAI"]
+__all__ = ["Gemini", "Ollama", "OpenAI"]

### 2024-06-07 src/llm_tst_consistency/llm/_gemini.py

@@ -1,0 +1,39 @@
+import os
+import time
+
+import vertexai
+import vertexai.generative_models as gai
+import spacy
+
+from llm_tst_consistency.hlf import HandcraftedLinguisticFeature
+
+_API_KEY=os.getenv("LTC_GOOGLE_API_KEY")
+
+
+class Gemini:
+    def __init__(
+        self, text: str, prompt: str, hlf_cfg: dict[str, HandcraftedLinguisticFeature]
+    ) -> None:
+        self._input_text = text
+        self._hlf_cfg = hlf_cfg
+        self._nlp = spacy.load("en_core_web_sm")
+        vertexai.init(project="llm-tst-consistency", location="us-central1")
+        self._model = gai.GenerativeModel(
+            model_name="gemini-1.5-pro",
+            system_instruction=[prompt]
+        )
+
+    @staticmethod
+    def _apply_prefix(name: str, prefix: str) -> str:
+        if prefix is None or len(prefix.strip()) < 1:
+            return name
+        return f"{prefix}_{name}"
+
+    def __call__(self, prefix: str = "") -> dict[str, float]:
+        time.sleep(12)
+        response = self._model.generate_content(self._input_text)
+        doc = self._nlp(response.text)
+        return {
+            self._apply_prefix(name, prefix): hlf(doc)
+            for name, hlf in self._hlf_cfg.items()
+        }

### 2024-06-07 src/llm_tst_consistency/evaluator.py
Add Claude3 support

@@ -13,7 +13,7 @@ from scipy.stats import ks_2samp
 from spacy import Language
 
 from llm_tst_consistency.hlf import HandcraftedLinguisticFeature, KupermanAgeOfAcquisition
-from llm_tst_consistency.llm import OpenAI, Gemini
+from llm_tst_consistency.llm import OpenAI, Gemini, Claude3
 from llm_tst_consistency.plot import draw_plots
 from llm_tst_consistency.stats import Stats
 
@@ -129,11 +129,11 @@ if __name__ == "__main__":
             ds_stats[key] = Stats(**ds_stats[key])
     if not file_name.exists():
         # generate baseline
-        baseline_generator = Gemini(INPUT, tpl.render(), HLFs)
+        baseline_generator = Claude3(INPUT, tpl.render(), HLFs)
         data = [baseline_generator("baseline") for _ in range(10)]
 
         # augment with hlf instructions
-        hlf_generator = Gemini(INPUT, tpl.render(hlf_instructions=_write_hlf_instructions(ds_stats)), HLFs)
+        hlf_generator = Claude3(INPUT, tpl.render(hlf_instructions=_write_hlf_instructions(ds_stats)), HLFs)
         for obj in data:
             obj.update(hlf_generator("hlf"))
 

### 2024-06-07 src/llm_tst_consistency/llm/__init__.py

@@ -1,6 +1,7 @@
 from llm_tst_consistency.llm._llama import Ollama
 from llm_tst_consistency.llm._gemini import Gemini
 from llm_tst_consistency.llm._gpt import OpenAI
+from llm_tst_consistency.llm._claude3 import Claude3
 
 
-__all__ = ["Gemini", "Ollama", "OpenAI"]
+__all__ = ["Gemini", "Ollama", "OpenAI", "Claude3"]

### 2024-06-07 src/llm_tst_consistency/llm/_claude3.py

@@ -1,0 +1,38 @@
+import os
+import time
+
+import anthropic
+import spacy
+
+from llm_tst_consistency.hlf import HandcraftedLinguisticFeature
+
+
+class Claude3:
+    def __init__(
+        self, text: str, prompt: str, hlf_cfg: dict[str, HandcraftedLinguisticFeature]
+    ) -> None:
+        self._system_prompt = prompt
+        self._messages = [{"role": "user", "content": text}]
+        self._hlf_cfg = hlf_cfg
+        self._nlp = spacy.load("en_core_web_sm")
+        self._client = anthropic.Client(api_key=os.getenv("LTC_ANTHROPIC_KEY"))
+
+    @staticmethod
+    def _apply_prefix(name: str, prefix: str) -> str:
+        if prefix is None or len(prefix.strip()) < 1:
+            return name
+        return f"{prefix}_{name}"
+
+    def __call__(self, prefix: str = "") -> dict[str, float]:
+        time.sleep(5)
+        response = self._client.messages.create(
+            model="claude-3-opus-20240229",
+            system=self._system_prompt,
+            messages=self._messages,
+            max_tokens=1024,
+        )
+        doc = self._nlp(response.content[0].text)
+        return {
+            self._apply_prefix(name, prefix): hlf(doc)
+            for name, hlf in self._hlf_cfg.items()
+        }

### 2024-06-07 src/llm_tst_consistency/evaluator.py
parametrize everything (input text, dataset, llm)

@@ -13,7 +13,7 @@ from scipy.stats import ks_2samp
 from spacy import Language
 
 from llm_tst_consistency.hlf import HandcraftedLinguisticFeature, KupermanAgeOfAcquisition
-from llm_tst_consistency.llm import OpenAI, Gemini, Claude3
+from llm_tst_consistency.llm import OpenAI, Gemini, Claude3, Ollama
 from llm_tst_consistency.plot import draw_plots
 from llm_tst_consistency.stats import Stats
 
@@ -114,40 +114,65 @@ def is_diff_significant(x: float, y: float, confidence: float = 0.95) -> bool:
     return p_value < threshold
 
 
-if __name__ == "__main__":
-    dotenv.load_dotenv()
-    file_name = Path(__file__).parent / "data.csv"
-    features = list(HLFs)
-    tpl = load_template("prompt_1.j2")
-    stats_file_name = Path(__file__).parent / "cnn_dailymail.json"
+LLMs = {
+    "claude3": Claude3,
+    "gemini": Gemini,
+    "gpt": OpenAI,
+    "llama3": Ollama,
+}
+CURRENT_LLM = "llama3"
+
+DATASETS = {
+    "cnn_dailymail": load_cnn_daily_mail,
+}
+CURRENT_DATASET = "cnn_dailymail"
+
+
+def _load_dataset_stats(ds_name, ds_loader):
+    stats_file_name = Path(__file__).parent / f"{ds_name}.json"
     if not stats_file_name.exists():
-        ds_stats = _compute_dataset_features(spacy.load("en_core_web_sm"), load_cnn_daily_mail())
+        ds_stats = _compute_dataset_features(spacy.load("en_core_web_sm"), ds_loader())
         stats_file_name.write_bytes(json.dumps(ds_stats))
     else:
         ds_stats = json.loads(stats_file_name.read_text())
         for key in ds_stats:
             ds_stats[key] = Stats(**ds_stats[key])
+    return ds_stats
+
+
+def _generate_text_metrics(trials, prompt_template, hlf_instructions, input_text, model_cls):
+    features = list(HLFs)
+    tpl = load_template(prompt_template)
+
+    # generate baseline
+    baseline_prompt = tpl.render()
+    baseline_generator = model_cls(input_text, baseline_prompt, HLFs)
+    data = [baseline_generator("baseline") for _ in range(trials)]
+
+    # augment with hlf instructions
+    hlf_prompt = tpl.render(hlf_instructions=hlf_instructions)
[…]

### 2024-06-07 src/llm_tst_consistency/llm/_llama.py

@@ -1,3 +1,5 @@
+import os
+
 import ollama
 import spacy
 
@@ -14,7 +16,7 @@ class Ollama:
         ]
         self._hlf_cfg = hlf_cfg
         self._nlp = spacy.load("en_core_web_sm")
-        self._client = ollama.Client(host="172.30.212.3")
+        self._client = ollama.Client(host=os.getenv("LTC_OLLAMA_HOST"))
 
     @staticmethod
     def _apply_prefix(name: str, prefix: str) -> str:
@@ -23,7 +25,7 @@ class Ollama:
         return f"{prefix}_{name}"
 
     def __call__(self, prefix: str = "") -> dict[str, float]:
-        response = self._client.chat(model="llama3:70b-text", messages=self._messages)
+        response = self._client.chat(model="llama3", messages=self._messages)
         doc = self._nlp(response["message"]["content"])
         return {
             self._apply_prefix(name, prefix): hlf(doc)

### 2024-06-07 src/llm_tst_consistency/dataset_loaders.py
split code from evaluator.py to other modules

@@ -1,0 +1,11 @@
+from datasets import load_dataset
+
+
+def is_cnn(record: dict) -> bool:
+    return "(CNN)" in record.get("article", "")
+
+
+def load_cnn_daily_mail():
+    return load_dataset(
+        "abisee/cnn_dailymail", "3.0.0", split="test"
+    ).filter(is_cnn)

### 2024-06-07 src/llm_tst_consistency/evaluator.py

@@ -7,39 +7,20 @@ import numpy as np
 import orjson as json
 import pandas as pd
 import spacy
-from datasets import load_dataset, Dataset
+from datasets import Dataset
 from jinja2 import Environment, FileSystemLoader, select_autoescape
 from scipy.stats import ks_2samp
 from spacy import Language
 
+from llm_tst_consistency.dataset_loaders import load_cnn_daily_mail
 from llm_tst_consistency.hlf import HandcraftedLinguisticFeature, KupermanAgeOfAcquisition
 from llm_tst_consistency.llm import OpenAI, Gemini, Claude3, Ollama
 from llm_tst_consistency.plot import draw_plots
+from llm_tst_consistency.reporting import make_report
 from llm_tst_consistency.stats import Stats
 
 
 MAX_CORPUS_SIZE = 100
-
-
-def load_template(name):
-    templates_dir = Path(__file__).parent / "templates"
-    env = Environment(
-        loader=FileSystemLoader(templates_dir),
-        autoescape=select_autoescape()
-    )
-    return env.get_template(name)
-
-
-def is_cnn(record: dict) -> bool:
-    return "(CNN)" in record.get("article", "")
-
-
-def load_cnn_daily_mail():
-    return load_dataset(
-        "abisee/cnn_dailymail", "3.0.0", split="test"
-    ).filter(is_cnn)
-
-
 HLF_TEMPLATES = {
     "t_word": "total_words.j2",
     "t_sent": "total_sentences.j2",
@@ -57,63 +38,6 @@ HLFs = {
     for name in HLF_TEMPLATES
 }
 
-
-INPUT = """I've been to this cafe a few times. The service is great and the menu
-is affordable. It has a large spacious space, reliable free wifi, and good food.
-
-It is quite new and seems a bit confused with which demographic of clienteles
-they are catering to. It has a great potential to be a hub for creative, hipster
-crowd.
-
-At the moment, it is underrated. Thus, I give them an extra star on top of the 4
-stars they deserve.
-"""
-
-
-def _compute_dataset_features(nlp: Language, ds: Dataset) -> dict[str, Stats]:
-    corpus = [row["article"] for row in ds][:MAX_CORPUS_SIZE]
-    docs = list(map(nlp, corpus))
-    return {
-        name: Stats.hlf(docs, feature)
-        for name, feature in HLFs.items()
-    }
-
-
[…]

### 2024-06-07 src/llm_tst_consistency/reporting.py

@@ -1,0 +1,71 @@
+from math import isnan, erf
+
+import numpy as np
+import pandas as pd
+from scipy.stats import ks_2samp
+
+
+def _compute_areas_to_target(a: list[float], b: list[float], target: list[float]) -> tuple[float, float]:
+    area_a = np.trapz(np.abs(np.array(a) - np.array(target)))
+    area_b = np.trapz(np.abs(np.array(b) - np.array(target)))
+    return area_a, area_b
+
+
+def _compute_euclidian_norms_to_target(a: list[float], b: list[float], target: list[float]) -> tuple[float, float]:
+    distance_a = np.linalg.norm(np.array(a) - np.array(target))
+    distance_b = np.linalg.norm(np.array(b) - np.array(target))
+    return distance_a, distance_b
+
+
+def _is_diff_significant(x: float, y: float, confidence: float = 0.95) -> bool:
+    c = abs(x - y)
+    std_dev = np.std([x, y])
+    mu = np.mean([x, y])
+    z_score = c - mu / std_dev
+
+    # need to double-check this formula
+    p_value = 1 - 0.5 * (1 + erf(z_score / np.sqrt(2)))
+    if isnan(p_value):
+        return False
+
+    threshold = 1 - confidence
+    return p_value < threshold
+
+
+def make_report(features, ds_stats, df):
+    euclid = {}
+    area = {}
+    ks = {}
+    for feature in features:
+        diff_col = f"{feature}_diff"
+        is_closer_col = f"is_{feature}_closer"
+        is_different_col = f"is_{feature}_different"
+        target = [ds_stats[feature].mean]*len(df)
+        baseline = df[f"baseline_{feature}"].values
+        hlf = df[f"hlf_{feature}"].values
+
+        # Kolmogorov-Smirnov test
+        ks_bh = ks_2samp(baseline, hlf)
+        ks_bt = ks_2samp(baseline, target)
+        ks_ht = ks_2samp(hlf, target)
+        ks[diff_col] = ks_bt.statistic - ks_ht.statistic
+        ks[is_closer_col] = ks_bt.statistic > ks_ht.statistic
+        ks[is_different_col] = ks_bh.pvalue < 0.25
+
+        # Euclidian norm
+        dist_a, dist_b = _compute_euclidian_norms_to_target(baseline, hlf, target)
+        euclid[diff_col] = dist_a - dist_b
+        euclid[is_closer_col] = dist_a > dist_b
[…]

### 2024-06-07 src/llm_tst_consistency/evaluator.py
remove unused imports from evaluator.py

@@ -1,15 +1,12 @@
 from dataclasses import asdict
-from math import isnan, erf
 from pathlib import Path
 
 import dotenv
-import numpy as np
 import orjson as json
 import pandas as pd
 import spacy
 from datasets import Dataset
 from jinja2 import Environment, FileSystemLoader, select_autoescape
-from scipy.stats import ks_2samp
 from spacy import Language
 
 from llm_tst_consistency.dataset_loaders import load_cnn_daily_mail

### 2024-06-08 src/llm_tst_consistency/evaluator.py
refactor text generators one (hopefully) last time

@@ -1,4 +1,7 @@
+import os
 from dataclasses import asdict
+from enum import StrEnum
+from functools import partial
 from pathlib import Path
 
 import dotenv
@@ -11,13 +14,15 @@ from spacy import Language
 
 from llm_tst_consistency.dataset_loaders import load_cnn_daily_mail
 from llm_tst_consistency.hlf import HandcraftedLinguisticFeature, KupermanAgeOfAcquisition
-from llm_tst_consistency.llm import OpenAI, Gemini, Claude3, Ollama
+from llm_tst_consistency.llm import GPT, Gemini, Claude3, Ollama
 from llm_tst_consistency.plot import draw_plots
 from llm_tst_consistency.reporting import make_report
 from llm_tst_consistency.stats import Stats
 
 
-MAX_CORPUS_SIZE = 100
+dotenv.load_dotenv()
+MAX_CORPUS_SIZE = os.getenv("LTC_MAX_CORPUS_SIZE", 1)
+TRIAL_COUNT = os.getenv("LTC_TRIAL_COUNT", 3)
 HLF_TEMPLATES = {
     "t_word": "total_words.j2",
     "t_sent": "total_sentences.j2",
@@ -36,17 +41,49 @@ HLFs = {
 }
 
 LLMs = {
-    "claude3": Claude3,
-    "gemini": Gemini,
-    "gpt": OpenAI,
-    "llama3": Ollama,
+    "claude3": partial(
+        Claude3,
+        model_name="claude-3-opus-20240229",
+        api_key=os.getenv("LTC_ANTHROPIC_KEY"),
+        hlf_cfg=HLFs,
+    ),
+    "gemini": partial(
+        Gemini,
+        project_name=os.getenv("LTC_GOOGLE_PROJECT"),
+        location=os.getenv("LTC_GOOGLE_LOCATION"),
+        hlf_cfg=HLFs,
+        model_name="gemini-1.5-pro",
+    ),
+    "gpt": partial(
+        GPT,
+        model_name="gpt-4o",
+        api_key=os.getenv("LTC_OPENAI_KEY"),
+        hlf_cfg=HLFs,
+    ),
+    "llama3_8b": partial(
+        Ollama,
+        model_name="llama3",
+        host=os.getenv("LTC_OLLAMA_HOST"),
+        hlf_cfg=HLFs,
+    ),
+    "mixtral_8x7b": partial(
+        Ollama,
[…]

### 2024-06-08 src/llm_tst_consistency/llm/__init__.py

@@ -1,7 +1,7 @@
-from llm_tst_consistency.llm._llama import Ollama
-from llm_tst_consistency.llm._gemini import Gemini
-from llm_tst_consistency.llm._gpt import OpenAI
-from llm_tst_consistency.llm._claude3 import Claude3
+from llm_tst_consistency.llm._ollama import Ollama
+from llm_tst_consistency.llm._google import Gemini
+from llm_tst_consistency.llm._openai import GPT
+from llm_tst_consistency.llm._anthropic import Claude3
 
 
-__all__ = ["Gemini", "Ollama", "OpenAI", "Claude3"]
+__all__ = ["Gemini", "Ollama", "GPT", "Claude3"]

### 2024-06-08 src/llm_tst_consistency/llm/_base.py

@@ -1,0 +1,38 @@
+from abc import abstractmethod
+
+import spacy
+
+from llm_tst_consistency.hlf import HandcraftedLinguisticFeature
+
+
+DEFAULT_NLP_MODEL = "en_core_web_sm"
+
+
+class BaseModel:
+    def __init__(
+        self,
+        model_name: str,
+        hlf_cfg: dict[str, HandcraftedLinguisticFeature],
+        metric_level: str = None,
+        nlp_model_name: str = None,
+    ) -> None:
+        self._hlf_cfg = hlf_cfg
+        self._nlp = spacy.load(nlp_model_name or DEFAULT_NLP_MODEL)
+        self._model_name = model_name
+        self._prefix = metric_level
+
+    def _apply_prefix(self, name: str) -> str:
+        if self._prefix is None or len(self._prefix.strip()) < 1:
+            return name
+        return f"{self._prefix}_{name}"
+
+    @abstractmethod
+    def _generate_text(self) -> str:
+        pass
+
+    def __call__(self) -> dict[str, float]:
+        doc = self._nlp(self._generate_text())
+        return {
+            self._apply_prefix(name): hlf(doc)
+            for name, hlf in self._hlf_cfg.items()
+        }

### 2024-06-08 src/llm_tst_consistency/llm/_claude3.py

@@ -1,38 +1,0 @@
-import os
-import time
-
-import anthropic
-import spacy
-
-from llm_tst_consistency.hlf import HandcraftedLinguisticFeature
-
-
-class Claude3:
-    def __init__(
-        self, text: str, prompt: str, hlf_cfg: dict[str, HandcraftedLinguisticFeature]
-    ) -> None:
-        self._system_prompt = prompt
-        self._messages = [{"role": "user", "content": text}]
-        self._hlf_cfg = hlf_cfg
-        self._nlp = spacy.load("en_core_web_sm")
-        self._client = anthropic.Client(api_key=os.getenv("LTC_ANTHROPIC_KEY"))
-
-    @staticmethod
-    def _apply_prefix(name: str, prefix: str) -> str:
-        if prefix is None or len(prefix.strip()) < 1:
-            return name
-        return f"{prefix}_{name}"
-
-    def __call__(self, prefix: str = "") -> dict[str, float]:
-        time.sleep(5)
-        response = self._client.messages.create(
-            model="claude-3-opus-20240229",
-            system=self._system_prompt,
-            messages=self._messages,
-            max_tokens=1024,
-        )
-        doc = self._nlp(response.content[0].text)
-        return {
-            self._apply_prefix(name, prefix): hlf(doc)
-            for name, hlf in self._hlf_cfg.items()
-        }

### 2024-06-08 src/llm_tst_consistency/llm/_gemini.py

@@ -1,39 +1,0 @@
-import os
-import time
-
-import vertexai
-import vertexai.generative_models as gai
-import spacy
-
-from llm_tst_consistency.hlf import HandcraftedLinguisticFeature
-
-_API_KEY=os.getenv("LTC_GOOGLE_API_KEY")
-
-
-class Gemini:
-    def __init__(
-        self, text: str, prompt: str, hlf_cfg: dict[str, HandcraftedLinguisticFeature]
-    ) -> None:
-        self._input_text = text
-        self._hlf_cfg = hlf_cfg
-        self._nlp = spacy.load("en_core_web_sm")
-        vertexai.init(project="llm-tst-consistency", location="us-central1")
-        self._model = gai.GenerativeModel(
-            model_name="gemini-1.5-pro",
-            system_instruction=[prompt]
-        )
-
-    @staticmethod
-    def _apply_prefix(name: str, prefix: str) -> str:
-        if prefix is None or len(prefix.strip()) < 1:
-            return name
-        return f"{prefix}_{name}"
-
-    def __call__(self, prefix: str = "") -> dict[str, float]:
-        time.sleep(12)
-        response = self._model.generate_content(self._input_text)
-        doc = self._nlp(response.text)
-        return {
-            self._apply_prefix(name, prefix): hlf(doc)
-            for name, hlf in self._hlf_cfg.items()
-        }

### 2024-06-08 src/llm_tst_consistency/llm/_google.py

@@ -1,0 +1,37 @@
+import time
+
+import vertexai
+import vertexai.generative_models as gai
+
+from llm_tst_consistency.hlf import HandcraftedLinguisticFeature
+from llm_tst_consistency.llm._base import BaseModel, DEFAULT_NLP_MODEL
+
+
+class Gemini(BaseModel):
+    def __init__(
+        self,
+        project_name: str,
+        location: str,
+        model_name: str,
+        text: str,
+        prompt: str,
+        hlf_cfg: dict[str, HandcraftedLinguisticFeature],
+        **kwargs: str
+    ) -> None:
+        super().__init__(
+            model_name,
+            hlf_cfg,
+            metric_level=kwargs.get("metric_level"),
+            nlp_model_name=kwargs.get("nlp_model_name")
+        )
+        self._input_text = text
+        vertexai.init(project=project_name, location=location)
+        self._model = gai.GenerativeModel(
+            model_name=self._model_name,
+            system_instruction=[prompt]
+        )
+
+    def _generate_text(self) -> str:
+        time.sleep(12)
+        response = self._model.generate_content(self._input_text)
+        return response.text

### 2024-06-08 src/llm_tst_consistency/llm/_gpt.py

@@ -1,35 +1,0 @@
-import os
-
-import openai
-import spacy
-
-from llm_tst_consistency.hlf import HandcraftedLinguisticFeature
-
-
-class OpenAI:
-    def __init__(
-        self, text: str, prompt: str, hlf_cfg: dict[str, HandcraftedLinguisticFeature]
-    ) -> None:
-        self._messages = [
-            {"role": "system", "content": prompt},
-            {"role": "user", "content": text},
-        ]
-        self._hlf_cfg = hlf_cfg
-        self._nlp = spacy.load("en_core_web_sm")
-        self._client = openai.Client(api_key=os.getenv("LTC_OPENAI_KEY"))
-
-    @staticmethod
-    def _apply_prefix(name: str, prefix: str) -> str:
-        if prefix is None or len(prefix.strip()) < 1:
-            return name
-        return f"{prefix}_{name}"
-
-    def __call__(self, prefix: str = "") -> dict[str, float]:
-        response = self._client.chat.completions.create(
-            model="gpt-4o", messages=self._messages
-        )
-        doc = self._nlp(response.choices[0].message.content)
-        return {
-            self._apply_prefix(name, prefix): hlf(doc)
-            for name, hlf in self._hlf_cfg.items()
-        }

### 2024-06-08 src/llm_tst_consistency/llm/_llama.py

@@ -1,33 +1,0 @@
-import os
-
-import ollama
-import spacy
-
-from llm_tst_consistency.hlf import HandcraftedLinguisticFeature
-
-
-class Ollama:
-    def __init__(
-        self, text: str, prompt: str, hlf_cfg: dict[str, HandcraftedLinguisticFeature]
-    ) -> None:
-        self._messages = [
-            {"role": "system", "content": prompt},
-            {"role": "user", "content": text},
-        ]
-        self._hlf_cfg = hlf_cfg
-        self._nlp = spacy.load("en_core_web_sm")
-        self._client = ollama.Client(host=os.getenv("LTC_OLLAMA_HOST"))
-
-    @staticmethod
-    def _apply_prefix(name: str, prefix: str) -> str:
-        if prefix is None or len(prefix.strip()) < 1:
-            return name
-        return f"{prefix}_{name}"
-
-    def __call__(self, prefix: str = "") -> dict[str, float]:
-        response = self._client.chat(model="llama3", messages=self._messages)
-        doc = self._nlp(response["message"]["content"])
-        return {
-            self._apply_prefix(name, prefix): hlf(doc)
-            for name, hlf in self._hlf_cfg.items()
-        }

### 2024-06-08 src/llm_tst_consistency/llm/_ollama.py

@@ -1,0 +1,30 @@
+import ollama
+from llm_tst_consistency.hlf import HandcraftedLinguisticFeature
+from llm_tst_consistency.llm._base import BaseModel
+
+
+class Ollama(BaseModel):
+    def __init__(
+        self,
+        host: str,
+        model_name: str,
+        text: str,
+        prompt: str,
+        hlf_cfg: dict[str, HandcraftedLinguisticFeature],
+        **kwargs: str
+    ) -> None:
+        super().__init__(
+            model_name,
+            hlf_cfg,
+            metric_level=kwargs.get("metric_level"),
+            nlp_model_name=kwargs.get("nlp_model_name")
+        )
+        self._messages = [
+            {"role": "system", "content": prompt},
+            {"role": "user", "content": text},
+        ]
+        self._client = ollama.Client(host=host)
+
+    def _generate_text(self) -> str:
+        response = self._client.chat(model=self._model_name, messages=self._messages)
+        return response["message"]["content"]

### 2024-06-08 src/llm_tst_consistency/llm/_openai.py

@@ -1,0 +1,36 @@
+import openai
+import spacy
+
+from llm_tst_consistency.hlf import HandcraftedLinguisticFeature
+from llm_tst_consistency.llm._base import BaseModel
+
+
+class GPT(BaseModel):
+    def __init__(
+        self,
+        api_key: str,
+        model_name: str,
+        text: str,
+        prompt: str,
+        hlf_cfg: dict[str, HandcraftedLinguisticFeature],
+        **kwargs: str
+    ) -> None:
+        super().__init__(
+            model_name,
+            hlf_cfg,
+            metric_level=kwargs.get("metric_level"),
+            nlp_model_name=kwargs.get("nlp_model_name")
+        )
+        self._messages = [
+            {"role": "system", "content": prompt},
+            {"role": "user", "content": text},
+        ]
+        self._hlf_cfg = hlf_cfg
+        self._nlp = spacy.load("en_core_web_sm")
+        self._client = openai.Client(api_key=api_key)
+
+    def _generate_text(self) -> str:
+        response = self._client.chat.completions.create(
+            model=self._model_name, messages=self._messages
+        )
+        return response.choices[0].message.content

### 2024-06-08 src/llm_tst_consistency/dataset_loaders.py
reformat code using black

@@ -6,6 +6,4 @@ def is_cnn(record: dict) -> bool:
 
 
 def load_cnn_daily_mail():
-    return load_dataset(
-        "abisee/cnn_dailymail", "3.0.0", split="test"
-    ).filter(is_cnn)
+    return load_dataset("abisee/cnn_dailymail", "3.0.0", split="test").filter(is_cnn)

### 2024-06-08 src/llm_tst_consistency/evaluator.py

@@ -13,7 +13,10 @@ from jinja2 import Environment, FileSystemLoader, select_autoescape
 from spacy import Language
 
 from llm_tst_consistency.dataset_loaders import load_cnn_daily_mail
-from llm_tst_consistency.hlf import HandcraftedLinguisticFeature, KupermanAgeOfAcquisition
+from llm_tst_consistency.hlf import (
+    HandcraftedLinguisticFeature,
+    KupermanAgeOfAcquisition,
+)
 from llm_tst_consistency.llm import GPT, Gemini, Claude3, Ollama
 from llm_tst_consistency.plot import draw_plots
 from llm_tst_consistency.reporting import make_report
@@ -36,7 +39,11 @@ HLF_TEMPLATES = {
     "a_kup_pw": "kuperman_age.j2",
 }
 HLFs = {
-    name: KupermanAgeOfAcquisition(name) if name == "a_kup_pw" else HandcraftedLinguisticFeature(name)
+    name: (
+        KupermanAgeOfAcquisition(name)
+        if name == "a_kup_pw"
+        else HandcraftedLinguisticFeature(name)
+    )
     for name in HLF_TEMPLATES
 }
 
@@ -71,7 +78,7 @@ LLMs = {
         model_name="mixtral:8x7b",
         host=os.getenv("LTC_OLLAMA_HOST"),
         hlf_cfg=HLFs,
-    )
+    ),
 }
 CURRENT_LLM = "mixtral"
 
@@ -96,8 +103,7 @@ def _load_input_text(ds_name):
 def _load_j2(template_name):
     templates_dir = Path(__file__).parent / "templates"
     env = Environment(
-        loader=FileSystemLoader(templates_dir),
-        autoescape=select_autoescape()
+        loader=FileSystemLoader(templates_dir), autoescape=select_autoescape()
     )
     return env.get_template(template_name)
 
@@ -105,10 +111,7 @@ def _load_j2(template_name):
 def _compute_dataset_features(nlp: Language, ds: Dataset) -> dict[str, Stats]:
     corpus = [row["article"] for row in ds][:MAX_CORPUS_SIZE]
     docs = list(map(nlp, corpus))
-    return {
-        name: Stats.hlf(docs, feature)
-        for name, feature in HLFs.items()
-    }
+    return {name: Stats.hlf(docs, feature) for name, feature in HLFs.items()}
 
 
 def _load_dataset_stats(ds_name, ds_loader):
[…]

### 2024-06-08 src/llm_tst_consistency/hlf/__init__.py

@@ -1,4 +1,7 @@
-from llm_tst_consistency.hlf._hlf import HandcraftedLinguisticFeature, KupermanAgeOfAcquisition
+from llm_tst_consistency.hlf._hlf import (
+    HandcraftedLinguisticFeature,
+    KupermanAgeOfAcquisition,
+)
 
 
 __all__ = ["HandcraftedLinguisticFeature", "KupermanAgeOfAcquisition"]

### 2024-06-08 src/llm_tst_consistency/hlf/_hlf.py

@@ -26,7 +26,9 @@ class KupermanAgeOfAcquisition(HandcraftedLinguisticFeature):
         aoa_of_unique_words = [
             round(rating)
             for word in unique_words
-            for rating in self._aoa_ratings.loc[self._aoa_ratings["Word"] == word, "Rating.Mean"]
+            for rating in self._aoa_ratings.loc[
+                self._aoa_ratings["Word"] == word, "Rating.Mean"
+            ]
         ]
         if len(aoa_of_unique_words) == 0:
             return 100

### 2024-06-08 src/llm_tst_consistency/llm/_base.py

@@ -33,6 +33,5 @@ class BaseModel:
     def __call__(self) -> dict[str, float]:
         doc = self._nlp(self._generate_text())
         return {
-            self._apply_prefix(name): hlf(doc)
-            for name, hlf in self._hlf_cfg.items()
+            self._apply_prefix(name): hlf(doc) for name, hlf in self._hlf_cfg.items()
         }

### 2024-06-08 src/llm_tst_consistency/llm/_google.py

@@ -22,13 +22,12 @@ class Gemini(BaseModel):
             model_name,
             hlf_cfg,
             metric_level=kwargs.get("metric_level"),
-            nlp_model_name=kwargs.get("nlp_model_name")
+            nlp_model_name=kwargs.get("nlp_model_name"),
         )
         self._input_text = text
         vertexai.init(project=project_name, location=location)
         self._model = gai.GenerativeModel(
-            model_name=self._model_name,
-            system_instruction=[prompt]
+            model_name=self._model_name, system_instruction=[prompt]
         )
 
     def _generate_text(self) -> str:

### 2024-06-08 src/llm_tst_consistency/llm/_ollama.py

@@ -17,7 +17,7 @@ class Ollama(BaseModel):
             model_name,
             hlf_cfg,
             metric_level=kwargs.get("metric_level"),
-            nlp_model_name=kwargs.get("nlp_model_name")
+            nlp_model_name=kwargs.get("nlp_model_name"),
         )
         self._messages = [
             {"role": "system", "content": prompt},

### 2024-06-08 src/llm_tst_consistency/llm/_openai.py

@@ -19,7 +19,7 @@ class GPT(BaseModel):
             model_name,
             hlf_cfg,
             metric_level=kwargs.get("metric_level"),
-            nlp_model_name=kwargs.get("nlp_model_name")
+            nlp_model_name=kwargs.get("nlp_model_name"),
         )
         self._messages = [
             {"role": "system", "content": prompt},

### 2024-06-08 src/llm_tst_consistency/plot.py

@@ -10,14 +10,18 @@ from llm_tst_consistency.stats import Stats
 
 
 def _add_variance(
-        df: pd.DataFrame,
-        benchmark_stats: dict[str, Stats],
-        features: list[str],
-        name_extractor: Callable[[str], str]
+    df: pd.DataFrame,
+    benchmark_stats: dict[str, Stats],
+    features: list[str],
+    name_extractor: Callable[[str], str],
 ) -> pd.DataFrame:
     for feature in features:
-        df[f"{name_extractor(feature)}_varlo"] = df[f"{name_extractor(feature)}"] - benchmark_stats[feature].variance
-        df[f"{name_extractor(feature)}_varhi"] = df[f"{name_extractor(feature)}"] + benchmark_stats[feature].variance
+        df[f"{name_extractor(feature)}_varlo"] = (
+            df[f"{name_extractor(feature)}"] - benchmark_stats[feature].variance
+        )
+        df[f"{name_extractor(feature)}_varhi"] = (
+            df[f"{name_extractor(feature)}"] + benchmark_stats[feature].variance
+        )
     return df
 
 
@@ -30,12 +34,17 @@ def _hlf_name(feature: str) -> str:
 
 
 def _create_benchmark_trace(
-    df: pd.DataFrame, feature: str, title: str, benchmark_value: float, color: str = "#575757", dash: str = "dot"
+    df: pd.DataFrame,
+    feature: str,
+    title: str,
+    benchmark_value: float,
+    color: str = "#575757",
+    dash: str = "dot",
 ) -> BaseTraceType:
     return go.Scatter(
         x=df.index,
         y=[benchmark_value] * len(df),
-        mode='lines',
+        mode="lines",
         name=title,
         line={"color": color, "dash": dash},
         showlegend=False,
@@ -47,10 +56,10 @@ def _create_baseline(df: pd.DataFrame, feature: str) -> BaseTraceType:
     return go.Scatter(
         x=df.index,
         y=df[_baseline_name(feature)],
-        mode='lines',
+        mode="lines",
         name=f"baseline {feature}",
         legendgroup=feature,
-        line={"color": "#575757"}
+        line={"color": "#575757"},
     )
 
 
@@ -58,35 +67,50 @@ def _create_hlf(df: pd.DataFrame, feature: str) -> BaseTraceType:
     return go.Scatter(
         x=df.index,
         y=df[_hlf_name(feature)],
-        mode='lines',
+        mode="lines",
         name=f"hlf-enhanced {feature}",
[…]

### 2024-06-08 src/llm_tst_consistency/reporting.py

@@ -5,13 +5,17 @@ import pandas as pd
 from scipy.stats import ks_2samp
 
 
-def _compute_areas_to_target(a: list[float], b: list[float], target: list[float]) -> tuple[float, float]:
+def _compute_areas_to_target(
+    a: list[float], b: list[float], target: list[float]
+) -> tuple[float, float]:
     area_a = np.trapz(np.abs(np.array(a) - np.array(target)))
     area_b = np.trapz(np.abs(np.array(b) - np.array(target)))
     return area_a, area_b
 
 
-def _compute_euclidian_norms_to_target(a: list[float], b: list[float], target: list[float]) -> tuple[float, float]:
+def _compute_euclidian_norms_to_target(
+    a: list[float], b: list[float], target: list[float]
+) -> tuple[float, float]:
     distance_a = np.linalg.norm(np.array(a) - np.array(target))
     distance_b = np.linalg.norm(np.array(b) - np.array(target))
     return distance_a, distance_b
@@ -40,7 +44,7 @@ def make_report(features, ds_stats, df):
         diff_col = f"{feature}_diff"
         is_closer_col = f"is_{feature}_closer"
         is_different_col = f"is_{feature}_different"
-        target = [ds_stats[feature].mean]*len(df)
+        target = [ds_stats[feature].mean] * len(df)
         baseline = df[f"baseline_{feature}"].values
         hlf = df[f"hlf_{feature}"].values
 
@@ -67,5 +71,5 @@ def make_report(features, ds_stats, df):
     return pd.DataFrame(
         columns=list(euclid.keys()),
         data=[euclid, area, ks],
-        index=["by euclidian norm", "by area", "k-s test"]
+        index=["by euclidian norm", "by area", "k-s test"],
     )

### 2024-06-08 src/llm_tst_consistency/stats/_stats.py

@@ -53,9 +53,7 @@ class Welford:
         return result, n, m2
 
     @classmethod
-    def _compute_final_numeric_result(
-        cls, result: Stats, n: int, m2: float
-    ) -> Stats:
+    def _compute_final_numeric_result(cls, result: Stats, n: int, m2: float) -> Stats:
         if n < 1:
             return Stats.empty()
         if n > 1:
@@ -73,7 +71,5 @@ class Welford:
         count = 0
         m2: float = 0
         for i, item in enumerate(items):
-            result, count, m2 = self._welford_pass(
-                result, self._get_x(item), count, m2
-            )
+            result, count, m2 = self._welford_pass(result, self._get_x(item), count, m2)
         return self._compute_final_numeric_result(result, count, m2)

### 2024-06-08 src/llm_tst_consistency/evaluator.py
allow saving one plot per file

@@ -80,7 +80,7 @@ LLMs = {
         hlf_cfg=HLFs,
     ),
 }
-CURRENT_LLM = "mixtral"
+CURRENT_LLM = "llama3_8b"
 
 DATASETS = {
     "cnn_dailymail": load_cnn_daily_mail,

### 2024-06-08 src/llm_tst_consistency/plot.py

@@ -1,4 +1,5 @@
-from typing import Callable
+from pathlib import Path
+from typing import Callable, Iterable
 
 import pandas as pd
 import plotly.graph_objects as go
@@ -74,10 +75,55 @@ def _create_hlf(df: pd.DataFrame, feature: str) -> BaseTraceType:
     )
 
 
-def _add_trace_to_plot(
-    fig: Figure, trace: BaseTraceType, i: int, col_count: int
-) -> None:
-    fig.add_trace(trace, row=i // col_count + 1, col=i % col_count + 1)
+def draw_one_figure(title, features, df, ds_stats, plot_col_count):
+    fig = make_subplots(
+        rows=len(features) // plot_col_count + 1,
+        subplot_titles=features,
+        cols=plot_col_count,
+    )
+    for i, f in enumerate(features):
+        for trace in _make_feature_traces(f, df, ds_stats):
+            fig.add_trace(
+                trace, row=i // plot_col_count + 1, col=i % plot_col_count + 1
+            )
+    fig.update_layout(height=200 * len(features), width=900, title_text=title)
+    return fig
+
+
+def draw_multiple_figures(title, features, df, ds_stats):
+    for i, f in enumerate(features):
+        fig = go.Figure()
+        for trace in _make_feature_traces(f, df, ds_stats):
+            fig.add_trace(trace)
+        fig.update_layout(height=480, width=640, title_text=f"{title} - {f}")
+        yield f, fig
+
+
+def _make_feature_traces(feature_name, df, ds_stats):
+    return [
+        _create_hlf(df, feature_name),
+        _create_baseline(df, feature_name),
+        _create_benchmark_trace(
+            df,
+            feature_name,
+            f"{feature_name} benchmark min",
+            ds_stats[feature_name].low,
+        ),
+        _create_benchmark_trace(
+            df,
+            feature_name,
+            f"{feature_name} benchmark max",
+            ds_stats[feature_name].high,
+        ),
+        _create_benchmark_trace(
+            df,
+            feature_name,
+            f"{feature_name} benchmark average",
+            ds_stats[feature_name].mean,
+            "#000000",
+            "solid",
+        ),
+    ]
 
 
 def draw_plots(
@@ -89,30 +135,11 @@ def draw_plots(
 ):
     df = _add_variance(df, benchmark_stats, features, _baseline_name)
     df = _add_variance(df, benchmark_stats, features, _hlf_name)
-    fig = make_subplots(
[…]

### 2024-06-08 src/llm_tst_consistency/dataset_loaders.py
improve reporting and the experiment's parametrization

@@ -1,3 +1,5 @@
+from itertools import islice
+
 from datasets import load_dataset
 
 
@@ -5,5 +7,9 @@ def is_cnn(record: dict) -> bool:
     return "(CNN)" in record.get("article", "")
 
 
-def load_cnn_daily_mail():
-    return load_dataset("abisee/cnn_dailymail", "3.0.0", split="test").filter(is_cnn)
+def load_cnn_daily_mail(max_size: int = 0):
+    ds = load_dataset("abisee/cnn_dailymail", "3.0.0", split="test").filter(is_cnn)
+    articles = map(lambda row: row["article"], ds)
+    if max_size > 0:
+        articles = islice(articles, max_size)
+    return list(articles)

### 2024-06-08 src/llm_tst_consistency/evaluator.py

@@ -1,14 +1,14 @@
 import os
 from dataclasses import asdict
-from enum import StrEnum
 from functools import partial
 from pathlib import Path
+from random import sample
 
 import dotenv
 import orjson as json
 import pandas as pd
 import spacy
-from datasets import Dataset
+import typer
 from jinja2 import Environment, FileSystemLoader, select_autoescape
 from spacy import Language
 
@@ -18,15 +18,16 @@ from llm_tst_consistency.hlf import (
     KupermanAgeOfAcquisition,
 )
 from llm_tst_consistency.llm import GPT, Gemini, Claude3, Ollama
+from llm_tst_consistency.parameters import LLMName, MetricLevel, DatasetName, PromptName
 from llm_tst_consistency.plot import draw_plots
 from llm_tst_consistency.reporting import make_report
 from llm_tst_consistency.stats import Stats
 
 
 dotenv.load_dotenv()
-MAX_CORPUS_SIZE = os.getenv("LTC_MAX_CORPUS_SIZE", 1)
-TRIAL_COUNT = os.getenv("LTC_TRIAL_COUNT", 3)
-HLF_TEMPLATES = {
+cli = typer.Typer(name="evaluator")
+
+HLF_TEMPLATE_CONFIG = {
     "t_word": "total_words.j2",
     "t_sent": "total_sentences.j2",
     "n_uverb": "number_of_unique_verbs.j2",
@@ -44,53 +45,38 @@ HLFs = {
         if name == "a_kup_pw"
         else HandcraftedLinguisticFeature(name)
     )
-    for name in HLF_TEMPLATES
+    for name in HLF_TEMPLATE_CONFIG
 }
-
-LLMs = {
-    "claude3": partial(
+LLM_CONFIG = {
+    LLMName.CLAUDE: partial(
         Claude3,
         model_name="claude-3-opus-20240229",
         api_key=os.getenv("LTC_ANTHROPIC_KEY"),
         hlf_cfg=HLFs,
     ),
-    "gemini": partial(
+    LLMName.GEMINI: partial(
         Gemini,
         project_name=os.getenv("LTC_GOOGLE_PROJECT"),
         location=os.getenv("LTC_GOOGLE_LOCATION"),
         hlf_cfg=HLFs,
         model_name="gemini-1.5-pro",
     ),
-    "gpt": partial(
+    LLMName.GPT: partial(
         GPT,
         model_name="gpt-4o",
[…]

### 2024-06-08 src/llm_tst_consistency/parameters.py

@@ -1,0 +1,24 @@
+from enum import StrEnum
+
+
+class MetricLevel(StrEnum):
+    BASELINE = "baseline"
+    HLF = "hlf"
+
+
+class LLMName(StrEnum):
+    CLAUDE = "claude3"
+    GEMINI = "gemini"
+    GPT = "gpt"
+    LLAMA3 = "llama3_8b"
+    MIXTRAL = "mixtral_8x7b"
+
+
+class DatasetName(StrEnum):
+    CNN_DAILY_MAIL = "cnn_dailymail"
+    YELP_REVIEWS = "yelp"
+
+
+class PromptName(StrEnum):
+    PROMPT1 = "prompt_1"
+    PROMPT2 = "prompt_2"

### 2024-06-08 src/llm_tst_consistency/plot.py

@@ -127,6 +127,7 @@ def _make_feature_traces(feature_name, df, ds_stats):
 
 
 def draw_plots(
+    prompt_name: str,
     title: str,
     df: pd.DataFrame,
     features: list[str],
@@ -135,7 +136,7 @@ def draw_plots(
 ):
     df = _add_variance(df, benchmark_stats, features, _baseline_name)
     df = _add_variance(df, benchmark_stats, features, _hlf_name)
-    img_path = Path(__file__).parent.parent.parent / "article" / "plots" / title
+    img_path = Path(__file__).parent.parent.parent / "article" / "plots" / prompt_name / title
     if not (img_path.exists() and img_path.is_dir()):
         img_path.mkdir(exist_ok=True, parents=True)
 

### 2024-06-08 src/llm_tst_consistency/reporting.py

@@ -36,7 +36,7 @@ def _is_diff_significant(x: float, y: float, confidence: float = 0.95) -> bool:
     return p_value < threshold
 
 
-def make_report(features, ds_stats, df):
+def make_report(llm_name, ds_name, features, ds_stats, df):
     euclid = {}
     area = {}
     ks = {}
@@ -44,6 +44,7 @@ def make_report(features, ds_stats, df):
         diff_col = f"{feature}_diff"
         is_closer_col = f"is_{feature}_closer"
         is_different_col = f"is_{feature}_different"
+
         target = [ds_stats[feature].mean] * len(df)
         baseline = df[f"baseline_{feature}"].values
         hlf = df[f"hlf_{feature}"].values
@@ -71,5 +72,9 @@ def make_report(features, ds_stats, df):
     return pd.DataFrame(
         columns=list(euclid.keys()),
         data=[euclid, area, ks],
-        index=["by euclidian norm", "by area", "k-s test"],
+        index=[
+            f"({llm_name},{ds_name}) euclidian norm",
+            f"({llm_name},{ds_name}) area",
+            f"({llm_name},{ds_name}) k-s test",
+        ],
     )

### 2024-06-08 src/llm_tst_consistency/evaluator.py
add cnn_dailymail settings to git

@@ -100,7 +100,7 @@ def _compute_dataset_features(nlp: Language, corpus: list[str]) -> dict[str, Sta
 
 
 def _process_dataset(ds_name, ds_loader, max_corpus_size, max_example_count):
-    stats_file_name = Path(__file__).parent / f"{ds_name}.json"
+    stats_file_name = Path(__file__).parent.parent.parent / "data" / f"{ds_name}_settings.json"
     if not stats_file_name.exists():
         corpus = ds_loader(max_corpus_size)
         obj = {"stats": _compute_dataset_features(spacy.load("en_core_web_sm"), corpus), "examples": list(sample(corpus, max_example_count))}

### 2024-06-08 src/llm_tst_consistency/evaluator.py
implement 2nd prompt (with examples)

@@ -126,14 +126,15 @@ def _generate_text_metrics(
     tpl = _load_j2(prompt_template)
 
     # generate baseline
-    baseline_prompt = tpl.render(examples=examples)
+    example_text = f"{os.linesep}---{os.linesep}".join(examples)
+    baseline_prompt = tpl.render(examples=example_text)
     baseline_generator = model_cls(
         text=input_text, prompt=baseline_prompt, metric_level=MetricLevel.BASELINE
     )
     data = [baseline_generator() for _ in range(trial_count)]
 
     # augment with hlf instructions
-    hlf_prompt = tpl.render(hlf_instructions=hlf_instructions, examples=examples)
+    hlf_prompt = tpl.render(hlf_instructions=hlf_instructions, examples=example_text)
     hlf_generator = model_cls(
         text=input_text, prompt=hlf_prompt, metric_level=MetricLevel.HLF
     )

### 2024-06-08 src/llm_tst_consistency/reporting.py

@@ -24,8 +24,10 @@ def _compute_euclidian_norms_to_target(
 def _is_diff_significant(x: float, y: float, confidence: float = 0.95) -> bool:
     c = abs(x - y)
     std_dev = np.std([x, y])
+    if std_dev == 0:
+        return False
     mu = np.mean([x, y])
-    z_score = c - mu / std_dev
+    z_score = (c - mu) / std_dev
 
     # need to double-check this formula
     p_value = 1 - 0.5 * (1 + erf(z_score / np.sqrt(2)))

### 2024-06-08 src/llm_tst_consistency/evaluator.py
add support for command-r+

@@ -17,7 +17,7 @@ from llm_tst_consistency.hlf import (
     HandcraftedLinguisticFeature,
     KupermanAgeOfAcquisition,
 )
-from llm_tst_consistency.llm import GPT, Gemini, Claude3, Ollama
+from llm_tst_consistency.llm import GPT, Gemini, Claude3, Ollama, CommandR
 from llm_tst_consistency.parameters import LLMName, MetricLevel, DatasetName, PromptName
 from llm_tst_consistency.plot import draw_plots
 from llm_tst_consistency.reporting import make_report
@@ -73,6 +73,12 @@ LLM_CONFIG = {
         host=os.getenv("LTC_OLLAMA_HOST"),
         hlf_cfg=HLFs,
     ),
+    LLMName.COMMAND_R: partial(
+        CommandR,
+        model_name="command-r-plus",
+        api_key=os.getenv("LTC_COHERE_API_KEY"),
+        hlf_cfg=HLFs,
+    ),
 }
 DATASET_CONFIG = {
     DatasetName.CNN_DAILY_MAIL: load_cnn_daily_mail,
@@ -100,10 +106,15 @@ def _compute_dataset_features(nlp: Language, corpus: list[str]) -> dict[str, Sta
 
 
 def _process_dataset(ds_name, ds_loader, max_corpus_size, max_example_count):
-    stats_file_name = Path(__file__).parent.parent.parent / "data" / f"{ds_name}_settings.json"
+    stats_file_name = (
+        Path(__file__).parent.parent.parent / "data" / f"{ds_name}_settings.json"
+    )
     if not stats_file_name.exists():
         corpus = ds_loader(max_corpus_size)
-        obj = {"stats": _compute_dataset_features(spacy.load("en_core_web_sm"), corpus), "examples": list(sample(corpus, max_example_count))}
+        obj = {
+            "stats": _compute_dataset_features(spacy.load("en_core_web_sm"), corpus),
+            "examples": list(sample(corpus, max_example_count)),
+        }
         stats_file_name.write_bytes(json.dumps(obj))
     else:
         obj = json.loads(stats_file_name.read_text())
@@ -122,7 +133,14 @@ def _write_hlf_instructions(ds_stats: dict[str, Stats]) -> list[str]:
 
 
 def _generate_text_metrics(
-    trial_count, prompt_template, hlf_instructions, input_text, model_cls, features, examples=None):
+    trial_count,
+    prompt_template,
[…]

### 2024-06-08 src/llm_tst_consistency/llm/__init__.py

@@ -2,6 +2,7 @@ from llm_tst_consistency.llm._ollama import Ollama
 from llm_tst_consistency.llm._google import Gemini
 from llm_tst_consistency.llm._openai import GPT
 from llm_tst_consistency.llm._anthropic import Claude3
+from llm_tst_consistency.llm._cohere import CommandR
 
 
-__all__ = ["Gemini", "Ollama", "GPT", "Claude3"]
+__all__ = ["Gemini", "Ollama", "GPT", "Claude3", "CommandR"]

### 2024-06-08 src/llm_tst_consistency/llm/_cohere.py

@@ -1,0 +1,37 @@
+import time
+
+import cohere
+import spacy
+
+from llm_tst_consistency.hlf import HandcraftedLinguisticFeature
+from llm_tst_consistency.llm._base import BaseModel
+
+
+class CommandR(BaseModel):
+    def __init__(
+        self,
+        api_key: str,
+        model_name: str,
+        text: str,
+        prompt: str,
+        hlf_cfg: dict[str, HandcraftedLinguisticFeature],
+        **kwargs: str
+    ) -> None:
+        super().__init__(
+            model_name,
+            hlf_cfg,
+            metric_level=kwargs.get("metric_level"),
+            nlp_model_name=kwargs.get("nlp_model_name"),
+        )
+        self._message = text
+        self._preamble = prompt
+        self._hlf_cfg = hlf_cfg
+        self._nlp = spacy.load("en_core_web_sm")
+        self._client = cohere.Client(api_key=api_key)
+
+    def _generate_text(self) -> str:
+        time.sleep(12)
+        response = self._client.chat(
+            model=self._model_name, message=self._message, preamble=self._preamble
+        )
+        return response.text

### 2024-06-08 src/llm_tst_consistency/parameters.py

@@ -12,6 +12,7 @@ class LLMName(StrEnum):
     GPT = "gpt"
     LLAMA3 = "llama3_8b"
     MIXTRAL = "mixtral_8x7b"
+    COMMAND_R = "command_r"
 
 
 class DatasetName(StrEnum):

### 2024-06-08 src/llm_tst_consistency/plot.py

@@ -136,7 +136,9 @@ def draw_plots(
 ):
     df = _add_variance(df, benchmark_stats, features, _baseline_name)
     df = _add_variance(df, benchmark_stats, features, _hlf_name)
-    img_path = Path(__file__).parent.parent.parent / "article" / "plots" / prompt_name / title
+    img_path = (
+        Path(__file__).parent.parent.parent / "article" / "plots" / prompt_name / title
+    )
     if not (img_path.exists() and img_path.is_dir()):
         img_path.mkdir(exist_ok=True, parents=True)
 

### 2024-06-08 src/llm_tst_consistency/dataset_loaders.py
add support for the yelp dataset

@@ -13,3 +13,11 @@ def load_cnn_daily_mail(max_size: int = 0):
     if max_size > 0:
         articles = islice(articles, max_size)
     return list(articles)
+
+
+def load_yelp(max_size: int = 0):
+    ds = load_dataset("Yelp/yelp_review_full", split="test")
+    reviews = map(lambda row: row["text"], ds)
+    if max_size > 0:
+        reviews = islice(reviews, max_size)
+    return list(reviews)

### 2024-06-08 src/llm_tst_consistency/evaluator.py

@@ -12,7 +12,7 @@ import typer
 from jinja2 import Environment, FileSystemLoader, select_autoescape
 from spacy import Language
 
-from llm_tst_consistency.dataset_loaders import load_cnn_daily_mail
+from llm_tst_consistency.dataset_loaders import load_cnn_daily_mail, load_yelp
 from llm_tst_consistency.hlf import (
     HandcraftedLinguisticFeature,
     KupermanAgeOfAcquisition,
@@ -82,11 +82,15 @@ LLM_CONFIG = {
 }
 DATASET_CONFIG = {
     DatasetName.CNN_DAILY_MAIL: load_cnn_daily_mail,
+    DatasetName.YELP_REVIEWS: load_yelp,
 }
 
 
-def _load_input_text(ds_name):
-    fpath = Path(__file__).parent.parent.parent / "data" / f"{ds_name}_input.txt"
+def _load_input_text(ds_name, from_dataset=False):
+    filename = f"{ds_name}_input.txt"
+    if from_dataset:
+        filename = f"{ds_name}_input_from_dataset.txt"
+    fpath = Path(__file__).parent.parent.parent / "data" / filename
     if fpath.exists() and fpath.is_file():
         return fpath.read_text()
     raise ValueError(f"could not find input file associated with dataset", ds_name)
@@ -214,24 +218,31 @@ def main(
         help="the maximum number of examples to use in the prompts that make use of them",
         envvar="LTC_EXAMPLE_COUNT",
     ),
+    input_from_dataset: bool = typer.Option(
+        False,
+        "--input-from-dataset",
+        is_flag=True,
+        help="use this flag to load input text from the chosen dataset"
+    )
 ):
+    features = list(HLFs)
     for prompt_name in prompt_names:
         for ds_name in ds_names:
             report = pd.DataFrame()
+            # read input text corresponding to dataset
+            input_text = _load_input_text(ds_name, from_dataset=input_from_dataset)
+            # compute/load dataset stats
+            ds_stats, examples = _process_dataset(
+                ds_name,
+                DATASET_CONFIG[ds_name],
+                max_corpus_size=max_size,
+                max_example_count=example_count,
+            )
+            # create prompt based on dataset stats
+            hlf_instructions = _write_hlf_instructions(ds_stats)
+
             for llm in llms:
-                # read input text corresponding to dataset
-                input_text = _load_input_text(ds_name)
-                # compute/load dataset stats
[…]

### 2024-06-08 run_experiment.sh
add a shell script to run the entire experiment

@@ -1,0 +1,6 @@
+#!/bin/sh
+
+# run all LLMs on all datasets with all prompts
+poetry run evaluator -n 10 -s 1000 -e 3
+# then run all LLMs on all datasets with input from the dataset, but only for prompt 2
+poetry run evaluator -n 10 -s 1000 -e 3 -p prompt_2 --input-from-dataset

### 2024-06-08 src/llm_tst_consistency/evaluator.py

@@ -227,6 +227,7 @@ def main(
 ):
     features = list(HLFs)
     for prompt_name in prompt_names:
+        prompt_discriminator = prompt_name if not input_from_dataset else f"{prompt_name}_ifd"
         for ds_name in ds_names:
             report = pd.DataFrame()
             # read input text corresponding to dataset
@@ -243,13 +244,14 @@ def main(
 
             for llm in llms:
                 # load text generation results / run text generation and compute results
+                template_name = f"{prompt_name}.j2"
                 df = _load_generated_text_metrics(
-                    prompt_name,
+                    prompt_discriminator,
                     llm,
                     ds_name,
                     lambda: _generate_text_metrics(
                         trial_count,
-                        f"{prompt_name}.j2",
+                        template_name,
                         hlf_instructions,
                         input_text,
                         LLM_CONFIG[llm],
@@ -257,11 +259,8 @@ def main(
                         examples,
                     ),
                 )
-                plot_dir = prompt_name
-                if input_from_dataset:
-                    plot_dir += "-ifd"
                 draw_plots(
-                    plot_dir,
+                    prompt_discriminator,
                     f"{prompt_name}-{llm}-{ds_name}",
                     df,
                     features,
@@ -269,10 +268,7 @@ def main(
                 )
                 individual_report = make_report(llm, ds_name, features, ds_stats, df)
                 report = pd.concat([report, individual_report], axis=0)
-            report_name = f"report_{prompt_name}_{ds_name}.csv"
-            if input_from_dataset:
-                report_name = f"report_{prompt_name}_ifd_{ds_name}.csv"
-            report.to_csv(report_name)
+            report.to_csv(f"report_{prompt_discriminator}_{ds_name}.csv")
 
 
 if __name__ == "__main__":

### 2024-06-08 src/llm_tst_consistency/evaluator.py
tweak the gemini safety settings

@@ -222,12 +222,14 @@ def main(
         False,
         "--input-from-dataset",
         is_flag=True,
-        help="use this flag to load input text from the chosen dataset"
-    )
+        help="use this flag to load input text from the chosen dataset",
+    ),
 ):
     features = list(HLFs)
     for prompt_name in prompt_names:
-        prompt_discriminator = prompt_name if not input_from_dataset else f"{prompt_name}_ifd"
+        prompt_discriminator = (
+            prompt_name if not input_from_dataset else f"{prompt_name}_ifd"
+        )
         for ds_name in ds_names:
             report = pd.DataFrame()
             # read input text corresponding to dataset

### 2024-06-08 src/llm_tst_consistency/llm/_google.py

@@ -4,7 +4,31 @@ import vertexai
 import vertexai.generative_models as gai
 
 from llm_tst_consistency.hlf import HandcraftedLinguisticFeature
-from llm_tst_consistency.llm._base import BaseModel, DEFAULT_NLP_MODEL
+from llm_tst_consistency.llm._base import BaseModel
+
+
+safety_config = [
+    gai.SafetySetting(
+        category=gai.HarmCategory.HARM_CATEGORY_DANGEROUS_CONTENT,
+        threshold=gai.HarmBlockThreshold.BLOCK_MEDIUM_AND_ABOVE,
+    ),
+    gai.SafetySetting(
+        category=gai.HarmCategory.HARM_CATEGORY_HARASSMENT,
+        threshold=gai.HarmBlockThreshold.BLOCK_MEDIUM_AND_ABOVE,
+    ),
+    gai.SafetySetting(
+        category=gai.HarmCategory.HARM_CATEGORY_HATE_SPEECH,
+        threshold=gai.HarmBlockThreshold.BLOCK_MEDIUM_AND_ABOVE,
+    ),
+    gai.SafetySetting(
+        category=gai.HarmCategory.HARM_CATEGORY_SEXUALLY_EXPLICIT,
+        threshold=gai.HarmBlockThreshold.BLOCK_MEDIUM_AND_ABOVE,
+    ),
+    gai.SafetySetting(
+        category=gai.HarmCategory.HARM_CATEGORY_UNSPECIFIED,
+        threshold=gai.HarmBlockThreshold.BLOCK_MEDIUM_AND_ABOVE,
+    ),
+]
 
 
 class Gemini(BaseModel):
@@ -32,5 +56,7 @@ class Gemini(BaseModel):
 
     def _generate_text(self) -> str:
         time.sleep(12)
-        response = self._model.generate_content(self._input_text)
+        response = self._model.generate_content(
+            self._input_text, safety_settings=safety_config
+        )
         return response.text

### 2024-06-09 src/llm_tst_consistency/evaluator.py
use k-s test for significance and geometry for relative closeness

@@ -226,51 +226,69 @@ def main(
     ),
 ):
     features = list(HLFs)
-    for prompt_name in prompt_names:
-        prompt_discriminator = (
-            prompt_name if not input_from_dataset else f"{prompt_name}_ifd"
-        )
-        for ds_name in ds_names:
-            report = pd.DataFrame()
-            # read input text corresponding to dataset
-            input_text = _load_input_text(ds_name, from_dataset=input_from_dataset)
-            # compute/load dataset stats
-            ds_stats, examples = _process_dataset(
-                ds_name,
-                DATASET_CONFIG[ds_name],
-                max_corpus_size=max_size,
-                max_example_count=example_count,
+    count = len(prompt_names)*len(ds_names)*len(llms)
+    with typer.progressbar(range(count), count) as progress:
+        for prompt_name in prompt_names:
+            prompt_discriminator = (
+                prompt_name if not input_from_dataset else f"{prompt_name}_ifd"
             )
-            # create prompt based on dataset stats
-            hlf_instructions = _write_hlf_instructions(ds_stats)
-
-            for llm in llms:
-                # load text generation results / run text generation and compute results
-                template_name = f"{prompt_name}.j2"
-                df = _load_generated_text_metrics(
-                    prompt_discriminator,
-                    llm,
+            for ds_name in ds_names:
+                report = pd.DataFrame()
+                tex_lines = []
+                # read input text corresponding to dataset
+                input_text = _load_input_text(ds_name, from_dataset=input_from_dataset)
+                # compute/load dataset stats
+                ds_stats, examples = _process_dataset(
                     ds_name,
-                    lambda: _generate_text_metrics(
-                        trial_count,
-                        template_name,
-                        hlf_instructions,
-                        input_text,
-                        LLM_CONFIG[llm],
+                    DATASET_CONFIG[ds_name],
+                    max_corpus_size=max_size,
+                    max_example_count=example_count,
+                )
+                # create prompt based on dataset stats
+                hlf_instructions = _write_hlf_instructions(ds_stats)
+
+                for llm in llms:
+                    # load text generation results / run text generation and compute results
+                    template_name = f"{prompt_name}.j2"
+                    df = _load_generated_text_metrics(
+                        prompt_discriminator,
+                        llm,
+                        ds_name,
+                        lambda: _generate_text_metrics(
+                            trial_count,
+                            template_name,
+                            hlf_instructions,
+                            input_text,
+                            LLM_CONFIG[llm],
[…]

### 2024-06-09 src/llm_tst_consistency/llm/_google.py

@@ -10,23 +10,23 @@ from llm_tst_consistency.llm._base import BaseModel
 safety_config = [
     gai.SafetySetting(
         category=gai.HarmCategory.HARM_CATEGORY_DANGEROUS_CONTENT,
-        threshold=gai.HarmBlockThreshold.BLOCK_MEDIUM_AND_ABOVE,
+        threshold=gai.HarmBlockThreshold.BLOCK_ONLY_HIGH,
     ),
     gai.SafetySetting(
         category=gai.HarmCategory.HARM_CATEGORY_HARASSMENT,
-        threshold=gai.HarmBlockThreshold.BLOCK_MEDIUM_AND_ABOVE,
+        threshold=gai.HarmBlockThreshold.BLOCK_ONLY_HIGH,
     ),
     gai.SafetySetting(
         category=gai.HarmCategory.HARM_CATEGORY_HATE_SPEECH,
-        threshold=gai.HarmBlockThreshold.BLOCK_MEDIUM_AND_ABOVE,
+        threshold=gai.HarmBlockThreshold.BLOCK_ONLY_HIGH,
     ),
     gai.SafetySetting(
         category=gai.HarmCategory.HARM_CATEGORY_SEXUALLY_EXPLICIT,
-        threshold=gai.HarmBlockThreshold.BLOCK_MEDIUM_AND_ABOVE,
+        threshold=gai.HarmBlockThreshold.BLOCK_ONLY_HIGH,
     ),
     gai.SafetySetting(
         category=gai.HarmCategory.HARM_CATEGORY_UNSPECIFIED,
-        threshold=gai.HarmBlockThreshold.BLOCK_MEDIUM_AND_ABOVE,
+        threshold=gai.HarmBlockThreshold.BLOCK_ONLY_HIGH,
     ),
 ]
 

### 2024-06-09 src/llm_tst_consistency/reporting.py

@@ -38,13 +38,11 @@ def _is_diff_significant(x: float, y: float, confidence: float = 0.95) -> bool:
     return p_value < threshold
 
 
-def make_report(llm_name, ds_name, features, ds_stats, df):
+def make_report(llm_name, features, ds_stats, df):
     euclid = {}
     area = {}
-    ks = {}
     for feature in features:
         diff_col = f"{feature}_diff"
-        is_closer_col = f"is_{feature}_closer"
         is_different_col = f"is_{feature}_different"
 
         target = [ds_stats[feature].mean] * len(df)
@@ -53,30 +51,23 @@ def make_report(llm_name, ds_name, features, ds_stats, df):
 
         # Kolmogorov-Smirnov test
         ks_bh = ks_2samp(baseline, hlf)
-        ks_bt = ks_2samp(baseline, target)
-        ks_ht = ks_2samp(hlf, target)
-        ks[diff_col] = ks_bt.statistic - ks_ht.statistic
-        ks[is_closer_col] = ks_bt.statistic > ks_ht.statistic
-        ks[is_different_col] = ks_bh.pvalue < 0.25
+        is_diff = ks_bh.pvalue < 0.05
 
         # Euclidian norm
         dist_a, dist_b = _compute_euclidian_norms_to_target(baseline, hlf, target)
-        euclid[diff_col] = dist_a - dist_b
-        euclid[is_closer_col] = dist_a > dist_b
-        euclid[is_different_col] = _is_diff_significant(dist_a, dist_b)
+        euclid[diff_col] = round(dist_a - dist_b, 2)
+        euclid[is_different_col] = is_diff
 
         # Area between curves
         area_a, area_b = _compute_areas_to_target(baseline, hlf, target)
-        area[diff_col] = area_a - area_b
-        area[is_closer_col] = area_a > area_b
-        area[is_different_col] = _is_diff_significant(area_a, area_b)
+        area[diff_col] = round(area_a - area_b, 2)
+        area[is_different_col] = is_diff
 
     return pd.DataFrame(
         columns=list(euclid.keys()),
-        data=[euclid, area, ks],
+        data=[euclid, area],
         index=[
-            f"({llm_name},{ds_name}) euclidian norm",
-            f"({llm_name},{ds_name}) area",
-            f"({llm_name},{ds_name}) k-s test",
+            f"{llm_name} $Diff_N$",
+            f"{llm_name} $Diff_A$",
         ],
     )

### 2024-06-09 article/main.tex
update prompts. update plots
incorporate first results in article text

@@
 \usepackage{booktabs}
-\usepackage{caption}
+\usepackage[justification=centering,font=scriptsize]{caption}
 \usepackage{cite}
@@
 \usepackage[colorlinks]{hyperref}
+\usepackage{inconsolata}
 \usepackage[latin1]{inputenc}
@@
 
-In order to perform our experiment we must first choose proper candidates.
-Since the interaction with the model happens only via prompts, we may use models
-with any type of licensing.
-For our experiment we choose the top model from each of the highest scoring five
-different major vendors listed on the Chatbot Arena~\cite{chiang2024chatbot}
-leaderboard.
-The final choice is documented in Table~\ref{table-llm}.
+For our experiment we will use some of the best large language models available.
+We trust the competence judgements issued in the the Chatbot Arena~\cite{chiang2024chatbot}
+for each LLM.
+In the interest of covering more ground, we choose large language models from
+different vendors.
+Our choice includes four top performing LLMs and one model that anyone would be
+able to run on their own hardware.
+Our final choice is documented in Table~\ref{table-llm}.
 
@@
     \centering
-    \begin{tabular}{@{}lllll@{}}\toprule
+    \begin{tabular}{@{}lllll@{}}
+        \toprule
         Name                    & Parameters  & Context Size & Elo Score & License Type        \\ \toprule
@@
         Anthropic Claude 3 Opus & undisclosed & 200000       & 1248      & Proprietary         \\
-        Meta Llama 3-70B-Instruct & 70 Billion  & 8000         & 1208      & Llama 3 Community   \\
+        Meta Llama 3-8B         & 8 Billion   & 8000         & 1208      & Llama 3 Community   \\
         Cohere Command R+       & 104 Billion & 128000       & 1189      & CC-BY-NC 4.0        \\
@@
 
-Another HLF characteristic we have in mind is their level of abstraction.
-Features such as the word count or number of punctuation mark are very concrete.
-On the other side of the spectrum, features such as the Flesch-Kincaid grade
-level~\cite{kincaid1975derivation} are highly abstract.
+We can also describe HLFs in terms of how abstract the instruction to conform to
+them is.
+For example, `write 100 words' is a more concrete instruction than `write as if
+you were in junior highschool'~\cite{kincaid1975derivation}.
 We divide HLFs in five categories ranging from very concrete to highly abstract
 and choose 2 features from each category for a grand total of 10 HLFs.
-Our final selection is available in Table~\ref{table-hlf}.
+Table~\ref{table-hlf} showcases our selection of features.
 
@@
     \centering
-    \begin{tabular}{@{}lllll@{}}\toprule
+    \begin{tabular}{@{}lllll@{}}
+        \toprule
         Name                              & Family           & Domain           & Abstraction     & Style Level\footnotemark \\ \toprule
-        total unique words                & wordsent         & surface          & very concrete   & lexical                  \\
+        total words                       & wordsent         & surface          & very concrete   & lexical                  \\
         total sentences                   & wordsent         & surface          & very concrete   & syntactic                \\ \bottomrule
-        total number of named entities    & entity           & discourse        & concrete        & semantic                 \\
+        total number of unique adjectives & partofspeech     & syntax           & concrete        & semantic                 \\
         total number of unique adjectives & partofspeech     & syntax           & concrete        & syntactic                \\ \bottomrule
@@
 
+Each of these features is well documented in literature~\cite{lftk-2023}.
+
 \subsection{DataSet Selection}\label{ds-selection}
@@
     \centering
-    \begin{tabular}{@{}lllll@{}}\toprule
+    \begin{tabular}{@{}lllll@{}}
+        \toprule
         Name              & Task                     & Row Count & Content Type & Reference              \\ \toprule
@@
                           & Text Generation          &           &              &                        \\ \midrule
-        HANNA             & Automatic Story Generation & 1056      & Stories      & \cite{hanna2024}              \\ \midrule
-        StylePTB          & Text Style Transfer        & 60000     & Sentences    & \cite{lyu-etal-2021-styleptb} \\ \bottomrule
     \end{tabular}
@@
 
-The Human-ANnotated NArratives (HANNA) dataset is interesting because it
-contains stories generated by LLMs. Since LLM generated stories can be assessed
-just as easily as human generated stories by using HLFs, we thought it would be
-a welcome addition to run our experiment for non-human writers, too.
+We use the Huggingface ecosystem to load these
+datasets~\cite{lhoest-etal-2021-datasets}.
+Because of the statistical nature of the experiment and in the interest of
+expediency, we only use the first thousand texts from the test split of each
+dataset.
 
-Finally, StylePTB is a dataset containing fine grained stylistic changes at the
-sentence level. We use this dataset to see how well LLMs do in scenarios
-involving HLFs at the sentence level (for example, part of speech HLFs).
+\subsection{User Prompt and Text}\label{input-text}
 
-\subsection{Input Prompt and Text}\label{input-text}
+This experiment consists of instructing various large language models to reword
+an input text so that it conforms to a bunch of handcrafted linguistic features.
+The input text is provided via a user prompt.
+We provide two user prompts for each dataset.
+These user prompts can be freely examined online in the
+\texttt{\href{https://github.com/koosie0507/llm-tst-consistency/tree/main/data}{data}}
+folder of the experiment's Git repository~\cite{olar2024experimentcode}.
 
-The input prompts are derived from the templates available in the Git repository
-that accompanies this article~\cite{olar2024experimentcode}.
-We provide the LLM with a system prompt that contains instructions concerning
-all HLFs.
+For the CNN/DailyMail dataset, the user prompt from outside the dataset is an
+extract from the classic text`The Life and Opinions of Tristram Shandy,
+Gentleman', by Laurence Sterne~\cite{sterne2003life}.
+We judge this text to be far apart in both style and exhibited linguistic
+features from the CNN/DailyMail dataset contents.
+For the Yelp dataset, our chosen user prompt is Oscar Wilde's `Sonnet to
+Liberty'~\cite{wilde1909poems}.
+This text is vastly different from anything in the Yelp review dataset.
+
+The instructions for the LLM are compiled from the Jinja templates available in
+the \texttt{\href{https://github.com/koosie0507/llm-tst-consistency/tree/main/data}{templates}}
+folder of the experiment's Git repository~\cite{olar2024experimentcode}.
+This folder includes both the main instruction prompts for the two scenarios
+that we cover in this experiment (`\texttt{prompt\_1}' and `\texttt{prompt\_2}')
+and the templates for generating the instructions for each HLF.
 
@@
 
-\subsection{Graph Interpretation}
+\subsection{Interpretation Conventions}
 
@@
 reserved for the trial number.
-The plots lend themselves to easy interpretation based on some simple rules.
 We plot the HLF values for the text generated from baseline prompts as a
@@
 
-Following this description of how to interpret the graphs, we present sample
-findings using the features for which we've observed the best and the worst
-compliance by the LLM.
-These findings were all made on the <DECIDE LATER> dataset.
+Our results tables highlight differences between the generated text's HLF values
+obtained by using the baseline and the HLF prompts.
+We are concerned with how different the plots of each HLF are from the
+continuous black line representing the average HLF on the benchmark dataset and
+how different they are from each other.
+
+We use the two-sampled Kolmogorov-Smirnov test to determine whether the baseline
+results are different from the HLF-enhanced results.
+We consider that any $p-value \le 0.05$ means that the distributions are
+different.
+
+To determine which of the two is closer to the target, we compute the area
+between each result and the target.
+Another way to measure the geometric closeness to the target is by computing the
+Euclidian norm of the vectorial difference between result and target.
+Let us denote with $A_b$ and $A_h$ the areas between baseline and target and HLF
+instruction result and target, respectively.
+Let us similarly denote with $N_b$ and $N_h$ the Euclidian norms.
+Finally, we define $Diff_A = A_b - A_h$ and $Diff_N = N_b - N_h$
+
+The tables that sum up the experiment results show $Diff_A$ and $Diff_N$ for
+each model and we provide a result table for each dataset and prompt variant in
+the online repository of the experiment.
+For each value we specify whether it is significant statistically according to
+the Kolmogorov-Smirnov test.
+
+In the next sections we only present the most relevant results of the
+experiment.
+To save space, we illustrate significant differences between the baseline and
+the HLF instruction results in \textbf{bold} typeface.
+Positive results are \underline{underlined} and negative results are
+\textit{italicized}.
 
@@
 
-Yelp Results on the best and worst performing metric.
-The rest of the metrics in an Appendix.
+The results on the CNN-DailyMail dataset showcased in
+Table~\ref{table-cnn-dailymail-results} provide the most eloquent answer to this
+question.
+The results on the Yelp dataset are positive, albeit less convincing.
+\begin{table}[!ht]
+    \centering
+    \resizebox{\textwidth}{!}{%
+        \begin{tabular}{lllllllllll}
+            model               & t\_word                      & t\_sent                    & n\_uverb                    & n\_uadj                    & simp\_ttr                 & a\_verb\_pw               & corr\_adj\_var            & corr\_verb\_var            & fkgl                    & a\_kup\_pw                \\
+            \toprule
+            claude3 $Diff_N$    & \underline{\textbf{623.16}}  & \underline{\textbf{24.74}} & \underline{\textbf{51.18}}  & \underline{\textbf{26.68}} & 0.06                      & -0.0                      & \underline{\textbf{0.46}} & \underline{\textbf{1.4}}   & \underline{1.44}        & \textit{\textbf{-1.66}}   \\
+            claude3 $Diff_A$    & \underline{\textbf{1774.0}}  & \underline{\textbf{74.0}}  & \underline{\textbf{158.84}} & \underline{\textbf{81.2}}  & 0.2                       & -0.01                     & \underline{\textbf{1.88}} & \underline{\textbf{4.49}}  & \underline{5.01}        & \textit{\textbf{-4.09}}   \\
+            gemini $Diff_N$     & \underline{\textbf{985.86}}  & \underline{\textbf{34.52}} & \underline{\textbf{100.36}} & \underline{\textbf{53.39}} & \underline{\textbf{0.52}} & 0.05                      & \underline{\textbf{2.62}} & \underline{\textbf{4.68}}  & \textit{-4.17}          & \textit{\textbf{-2.43}}   \\
+            gemini $Diff_A$     & \underline{\textbf{2990.5}}  & \underline{\textbf{106.5}} & \underline{\textbf{320.34}} & \underline{\textbf{159.1}} & \underline{\textbf{1.5}}  & 0.09                      & \underline{\textbf{7.24}} & \underline{\textbf{13.15}} & \textit{-12.65}         & \textit{\textbf{-7.13}}   \\
+            gpt $Diff_N$        & \underline{\textbf{125.87}}  & 1.46                       & 10.95                       & 10.77                      & 0.02                      & -0.01                     & 0.71                      & 0.81                       & \textit{\textbf{-1.72}} & 0.43                      \\
+            gpt $Diff_A$        & \underline{\textbf{327.0}}   & 3.0                        & 29.0                        & 27.0                       & 0.08                      & -0.01                     & 1.82                      & 2.24                       & \textit{\textbf{-5.72}} & 0.44                      \\
+            llama3\_8b $Diff_N$ & 171.31                       & 6.63                       & -9.8                        & 13.8                       & 0.04                      & -0.04                     & 1.03                      & -0.84                      & -0.11                   & -0.76                     \\
+            llama3\_8b $Diff_A$ & 524.0                        & 20.0                       & -32.5                       & 43.5                       & 0.1                       & -0.08                     & 3.64                      & -2.85                      & 1.57                    & -2.29                     \\
+            command\_r $Diff_N$ & \underline{\textbf{1024.78}} & \underline{\textbf{39.67}} & \underline{\textbf{81.58}}  & \underline{\textbf{41.46}} & \underline{\textbf{0.46}} & \underline{\textbf{0.03}} & \underline{\textbf{2.67}} & \underline{\textbf{3.78}}  & \underline{1.4}         & \underline{\textbf{0.14}} \\
+            command\_r $Diff_A$ & \underline{\textbf{3132.5}}  & \underline{\textbf{120.5}} & \underline{\textbf{246.84}} & \underline{\textbf{132.2}} & \underline{\textbf{1.36}} & \underline{\textbf{0.06}} & \underline{\textbf{7.88}} & \underline{\textbf{11.21}} & \underline{3.9}         & \underline{\textbf{1.03}} \\
+            \bottomrule
+        \end{tabular}%
+    }
+    \caption{CNN-DailyMail Results}
+    \label{table-cnn-dailymail-results}
+\end{table}
+
+Command-R+ was the best performer by far, while Llama3 (8B) was predictibly the
+worst at following HLF instructions.
+Figure~\ref{fig-command-r-t-word} shows a positive result, where the result
+obtained with the aid of HLF instructions is closer to the target.
+Figure~\ref{fig-gpt-fkgl} shows a negative result, where the baseline is closer
+to the target.
+Figure~\ref{fig-llama-simp-ttr} shows an inconclusive result.
+
+\begin{figure}[ht!]
+    \centering
+    \begin{minipage}{0.32\textwidth}
+        \includegraphics[width=\linewidth]{plots/prompt_1/prompt_1-command_r-cnn_dailymail/prompt_1-command_r-cnn_dailymail_t_word.png}
+        \caption{Command-R+\\Total Words}
+        \label{fig-command-r-t-word}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.32\textwidth}
+        \includegraphics[width=\linewidth]{plots/prompt_1/prompt_1-gpt-cnn_dailymail/prompt_1-gpt-cnn_dailymail_fkgl.png}
+        \caption{GPT\\Flesch-Kincaid Grade Level}
+        \label{fig-gpt-fkgl}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.32\textwidth}
+        \includegraphics[width=\linewidth]{plots/prompt_1/prompt_1-llama3_8b-cnn_dailymail/prompt_1-llama3_8b-cnn_dailymail_simp_ttr.png}
+        \caption[center]{Llama3-8B\\Simple TTR}
+        \label{fig-llama-simp-ttr}
+    \end{minipage}
+\end{figure}
 

### 2024-06-11 article/main.tex
review follow-up

@@
 generate new text based on a given input text and some instructions derived from
-HLF statistics computed on a benchmark dataset.
-To perform the experiment we must know which LLMs we are going to evaluate and
-what benchmark datasets we are going to use.
-We must also select the HLFs for our experiment.
-Before going into details, let us outline the experiment process.
+HLF statistics computed on a corpus containing multiple texts.
+The experiment relies on choosing LLMs to evaluate and finding datasets from
+which we can derive corpuses.
+Then there's the matter of selecting the HLFs for our experiment.
+Let us outline the experiment process before going into further detail.
 
@@
 We do so by instructing the target LLM to generate text retaining the meaning of
-the input text without any further instruction.
+the input text without any further instruction related to style or the
+linguistic features of the generated text.
 We repeat the text generation step 10 times using the same prompt and the same
 input text.
+
 For each text generation, we compute the following statistics for each of our
@@
     \item \textbf{avg}: the mean value of the HLF on the generated text;
-    \item \textbf{var}: the variance of the HLF values on the generated text.
 \end{itemize}
 
-A second baseline is obtained by following the same process but with a new
-prompt that builds upon the first one by adding the instruction to use example
-text from our benchmark dataset of choice.
-The examples are three texts chosen randomly once for each benchmark dataset.
-The reason behind choosing three texts is to not overrun the context window of
-the LLM.
+A second baseline is obtained by following the same process and same input text,
+but with a new prompt that builds upon the first one by adding the instruction
+to use example text from our corpus of choice.
+The examples are three texts chosen randomly once at the beginning of the
+experiment from the used corpus.
+We use the same example texts for all evaluated LLMs.
+
+We choose three texts to not overrun the context window of the LLM.
 This hints at the main drawback for our approach: it uses up the context window
@@
 around 8000 tokens so we should be mindful.
-We use the same example texts for all evaluated LLMs.
-Same as before, we construct the baseline by conducting 10 trials using the same
-prompt and input text.
+
+We construct the second baseline by conducting 10 trials using the same
+prompt and input text and otherwise in the same manner in which we created the
+first baseline.
 
@@
 
-Because of the examples present in the prompt, we perform the second experiment
-with input text both from within and from outside the corpus.
+Because we provide the LLM with examples from the corpus, it is more likely to
+intuit the linguistic features of the corpus.
+To understand whether HLF instructions actually make a difference, we perform
+the experiment using the second type of prompt using input text both from within
+and from outside the corpus.
 
@@
 
-For our experiment we will use some of the best large language models available.
-We trust the competence judgements issued in the the Chatbot Arena~\cite{chiang2024chatbot}
-for each LLM.
+For our experiment we will use state of the art large language models.
+We used the Chatbot Arena~\cite{chiang2024chatbot} leaderboard for constructing
+our selection.
 In the interest of covering more ground, we choose large language models from
@@
 able to run on their own hardware.
-Our final choice is documented in Table~\ref{table-llm}.
+The final selection is available in Table~\ref{table-llm}.
 
@@
 
-Our main focus is to understand how susceptible an LLMs is to prompts telling it
-to conform to HLFs.
-LFTK~\cite{lftk-2023} authors categorize HLFs by domain and family.
-Furthermore, information is available on the suitability of each HLF for given
-tasks such as readability assessment or fake news detection.
-However, there is no direct mapping between HLFs and style.
-It falls onto us to choose features based on their influence over text style as
-recognized in literature~\cite{verma2019lexical,lugea2023stylistics}.
+Our main focus is to understand how susceptible LLMs are to prompts containing
+instructions to conform to HLFs.
+LFTK authors~\cite{lftk-2023} categorize HLFs by domain and family and make
+available information on the suitability of each HLF for given tasks such as
+readability assessment or fake news detection.
+Besides the domain and family, we also choose features based on their influence
+over text style as recognized in literature~\cite{verma2019lexical,lugea2023stylistics}.
 
-We can also describe HLFs in terms of how abstract the instruction to conform to
-them is.
+In this experiment we create specific instructions for each HLF.
+The resulting instructions can be catalogued on the concrete-abstract spectrum.
 For example, `write 100 words' is a more concrete instruction than `write as if
@@
 
-Each of these features is well documented in literature~\cite{lftk-2023}.
+Each of these features is well documented~\cite{lftk-2023} barring the need for
+a deeper dive here into the subject matter.
 
@@
 
-We have run this experiment on the datasets listed in Table~\ref{table-ds}.
+We must choose source datasets based upon which we create corpuses for our
+experiment.
+Our final selection of datasets is listed in Table~\ref{table-ds}.
 
@@
         Name              & Task                     & Row Count & Content Type & Reference              \\ \toprule
-        Yelp Reviews Full & Sentiment Classification & 700000    & Reviews      & \cite{yelp2015neurips} \\
-                          & Text Style Transfer      & 700000    &              &                        \\
-                          & Text Classification      & 700000    &              &                        \\ \midrule
-        CNN / DailyMail   & Summarization            & 300000    & News         & \cite{cnndm2015}       \\
-                          & Question Answering       &           & Stories      & \cite{cnndm2017}       \\
+        Yelp Reviews      & Sentiment Classification & 50000     & Reviews      & \cite{yelp2015neurips} \\
+        (test split)      & Text Style Transfer      &           &              &                        \\
+                          & Text Classification      &           &              &                        \\ \midrule
+        CNN / DailyMail   & Summarization            & 11490     & News         & \cite{cnndm2015}       \\
+        (test split)      & Question Answering       &           & Stories      & \cite{cnndm2017}       \\
                           & Text Generation          &           &              &                        \\ \midrule
     \end{tabular}
-    \caption{Experiment LLMs}\label{table-ds}
+    \caption{Source Datasets}\label{table-ds}
 \end{table}
 
-The Yelp review data set provides a diverse assortment of writing styles and
-thus allowing us to home in on an LLMs ability to follow HLFs instead of copying
-a writing style.
-The reviews are often relatively short texts written in a coloquial style with
-a fair amount of variation in sentiment.
+We use the Huggingface ecosystem~\cite{lhoest-etal-2021-datasets} to load the
+datasets.
+Yelp reviews are loaded from \texttt{Yelp/yelp\_review\_full}, while CNN stories
+are loaded from \texttt{abisee/cnn\_dailymail}.
 
-On the other hand the CNN and DailyMail stories are news stories and even though
-their style might vary from one author to another, they will definitely be more
-formal than Yelp reviews.
-The writing is likely to be more mature, too.
+The Yelp review data set provides a diverse assortment of reviews from online
+users.
+The authors employ a casual writing style, very similar to what a average
+writing style of an average person would be like.
+The dataset as a whole is useful for many tasks.
+The corpus we construct based on this dataset contains the first 1000 reviews
+from its test split.
 
-We use the Huggingface ecosystem to load these
-datasets~\cite{lhoest-etal-2021-datasets}.
-Because of the statistical nature of the experiment and in the interest of
-expediency, we only use the first thousand texts from the test split of each
-dataset.
+CNN/DailyMail contains stories and news with diverse styles.
+The style varies from one author to another, but also based on the publishing
+entity.
+The articles are more formal, longer and more complex in structure than Yelp
+reviews which makes this dataset a good alternative choice.
+The corpus that is based on this dataset contains the first 1000 CNN articles
+from the test split of the dataset.
 
@@
 
-This experiment consists of instructing various large language models to reword
+This experiment presupposes instructing various large language models to reword
 an input text so that it conforms to a bunch of handcrafted linguistic features.
 The input text is provided via a user prompt.
-We provide two user prompts for each dataset.
-These user prompts can be freely examined online in the
+
+We provide two input texts that are external to our corpuses, one input text per
+corpus.
+They can be examined online in the
 \texttt{\href{https://github.com/koosie0507/llm-tst-consistency/tree/main/data}{data}}
 folder of the experiment's Git repository~\cite{olar2024experimentcode}.
+These user prompts are used when we're not using an input text from the
+corpus itself.
 
 For the CNN/DailyMail dataset, the user prompt from outside the dataset is an
-extract from the classic text`The Life and Opinions of Tristram Shandy,
-Gentleman', by Laurence Sterne~\cite{sterne2003life}.
-We judge this text to be far apart in both style and exhibited linguistic
-features from the CNN/DailyMail dataset contents.
-For the Yelp dataset, our chosen user prompt is Oscar Wilde's `Sonnet to
-Liberty'~\cite{wilde1909poems}.
-This text is vastly different from anything in the Yelp review dataset.
+extract from the classic text \textit{The Life and Opinions of Tristram Shandy,
+Gentleman}, by Laurence Sterne~\cite{sterne2003life}.
+This text is far from a news article in terms of both style and exhibited
+linguistic features from the CNN/DailyMail dataset contents.
+We see a text that is similarly divorced from the style and linguistic
+features of most Yelp reviews in Oscar Wilde's \textit{Sonnet to
+Liberty}~\cite{wilde1909poems}.
+Having input texts that are vastly different from a dataset should inform an LLM
+to generate a baseline that does not conform to the linguistic features of the
+dataset.
 
-The instructions for the LLM are compiled from the Jinja templates available in
-the \texttt{\href{https://github.com/koosie0507/llm-tst-consistency/tree/main/data}{templates}}
+We use Jinja templates to compile the text generation instructions for the LLM.
+Both the model prompts and the HLF instructions are constructed using these
+templates.
+The Jinja templates are available in the 
+\texttt{\href{https://github.com/koosie0507/llm-tst-consistency/tree/main/data}{templates}}
 folder of the experiment's Git repository~\cite{olar2024experimentcode}.
-This folder includes both the main instruction prompts for the two scenarios
-that we cover in this experiment (`\texttt{prompt\_1}' and `\texttt{prompt\_2}')
-and the templates for generating the instructions for each HLF.
+
+There are two main text generation instruction templates: one without example
+content~(called `\texttt{prompt\_1}') and one with examples from the
+corpus~(`\texttt{prompt\_2}').
+Each of the two instruction templates allow the presence of HLF instructions.
 
@@
 reserved for the trial number.
-We plot the HLF values for the text generated from baseline prompts as a
-continuous grey line.
-The similar values obtained by using HLF enhanced prompts are plotted using a
-continuous grenat line.
-To highlight the `tolerance' obtained by adding and subtracting the variance of
-the HLF on the benchmark dataset, we plot light gray areas around the baseline
-and HLF-enhanced plot lines.
+HLF values for the text generated using baseline prompts are plotted as a
+continuous grey line, while the values obtained by using HLF enhanced prompts
+are plotted using a continuous grenat line.
 The minimum and maximum HLF on the benchmark dataset are plotted using dotted
 black lines.
-A continuous black line designates the average HLF on the benchmark dataset.
+A continuous black line designates the average HLF on the corpus.
+We sometimes refer to this average as the \textit{target} value because the LLM
+is primarily instructed to conform to this value.
 
-Our results tables highlight differences between the generated text's HLF values
-obtained by using the baseline and the HLF prompts.
-We are concerned with how different the plots of each HLF are from the
-continuous black line representing the average HLF on the benchmark dataset and
-how different they are from each other.
+Result tables highlight the difference between plots obtained using a
+baseline prompt versus the ones obtained using HLF-enhanced prompts.
+Firstly, we care about the difference between each plot (baseline and HLF) and
+the target (the corpus average).
+Secondly, we measure how different baselines are from HLF-enhanced plots.
 
-We use the two-sampled Kolmogorov-Smirnov test to determine whether the baseline
-results are different from the HLF-enhanced results.
-We consider that any $p-value \le 0.05$ means that the distributions are
-different.
+We use the two-sampled Kolmogorov-Smirnov~\cite{kolmogorov1933,smirnov1939} test
+to determine whether the baseline results are different from the HLF-enhanced
+results.
+We consider that the results are different for a resulting $p\_value \le 0.05$.
 
-To determine which of the two is closer to the target, we compute the area
-between each result and the target.
+To determine which of the two results is closer to the target, we compute the
+area between each result and the target.
 Another way to measure the geometric closeness to the target is by computing the
@@
 Let us similarly denote with $N_b$ and $N_h$ the Euclidian norms.
-Finally, we define $Diff_A = A_b - A_h$ and $Diff_N = N_b - N_h$
+Finally, we define $Diff_A = A_b - A_h$ and $Diff_N = N_b - N_h$.
 
-The tables that sum up the experiment results show $Diff_A$ and $Diff_N$ for
-each model and we provide a result table for each dataset and prompt variant in
-the online repository of the experiment.
-For each value we specify whether it is significant statistically according to
-the Kolmogorov-Smirnov test.
+The result tables show $Diff_A$ and $Diff_N$ for each LLM in the context of a
+prompt and a dataset.
+All result tables for each dataset and prompt variant are available online, in
+the experiment's Git repository~\cite{olar2024experimentcode}.
 
@@
 the HLF instruction results in \textbf{bold} typeface.
-Positive results are \underline{underlined} and negative results are
-\textit{italicized}.
+The \underline{underlined} values represent situations when the HLF-enhanced
+prompt lead to results that are closer to the corpus average than the baseline.
+\textit{Italicized} values represent situations when the baseline was closer
+to the target than the HLF-enhanced result.
 
@@
 
-Yelp results on the best and worst performing metric.
-The rest of the metrics (including on all other datasets) in an Appendix.
-
+In this variant of the experiment, we feed the LLM three examples from the
+corpus.
 \section{Discussion}

### 2024-06-11 src/llm_tst_consistency/evaluator.py
improve latex reporting and add llama3:70b results

@@ -17,12 +17,12 @@ from llm_tst_consistency.hlf import (
     HandcraftedLinguisticFeature,
     KupermanAgeOfAcquisition,
 )
+from llm_tst_consistency.latex import make_latex_table
 from llm_tst_consistency.llm import GPT, Gemini, Claude3, Ollama, CommandR
 from llm_tst_consistency.parameters import LLMName, MetricLevel, DatasetName, PromptName
 from llm_tst_consistency.plot import draw_plots
 from llm_tst_consistency.reporting import make_report
 from llm_tst_consistency.stats import Stats
-
 
 dotenv.load_dotenv()
 cli = typer.Typer(name="evaluator")
@@ -67,16 +67,16 @@ LLM_CONFIG = {
         api_key=os.getenv("LTC_OPENAI_KEY"),
         hlf_cfg=HLFs,
     ),
-    LLMName.LLAMA3: partial(
-        Ollama,
-        model_name="llama3",
-        host=os.getenv("LTC_OLLAMA_HOST"),
-        hlf_cfg=HLFs,
-    ),
     LLMName.COMMAND_R: partial(
         CommandR,
         model_name="command-r-plus",
         api_key=os.getenv("LTC_COHERE_API_KEY"),
+        hlf_cfg=HLFs,
+    ),
+    LLMName.LLAMA3_70B: partial(
+        Ollama,
+        model_name="llama3:70b",
+        host=os.getenv("LTC_OLLAMA_HOST"),
         hlf_cfg=HLFs,
     ),
 }
@@ -111,7 +111,7 @@ def _compute_dataset_features(nlp: Language, corpus: list[str]) -> dict[str, Sta
 
 def _process_dataset(ds_name, ds_loader, max_corpus_size, max_example_count):
     stats_file_name = (
-        Path(__file__).parent.parent.parent / "data" / f"{ds_name}_settings.json"
+            Path(__file__).parent.parent.parent / "data" / f"{ds_name}_settings.json"
     )
     if not stats_file_name.exists():
         corpus = ds_loader(max_corpus_size)
@@ -137,13 +137,13 @@ def _write_hlf_instructions(ds_stats: dict[str, Stats]) -> list[str]:
 
 
 def _generate_text_metrics(
-    trial_count,
-    prompt_template,
-    hlf_instructions,
-    input_text,
-    model_cls,
-    features,
-    examples=None,
+        trial_count,
+        prompt_template,
+        hlf_instructions,
+        input_text,
[…]

### 2024-06-11 src/llm_tst_consistency/latex.py

@@ -1,0 +1,37 @@
+def format_value(value, is_different):
+    if is_different:
+        if value > 0:
+            return f"\\textbf{{\\underline{{{value}}}}}"
+        elif value < 0:
+            return f"\\textbf{{\\textit{{{value}}}}}"
+    return str(value)
+
+
+def make_latex_table(df, caption, label, features, alignment):
+    # Start the table and specify the alignment
+    latex_table = "\\begin{table}[!ht]\n\\centering\n"
+    latex_table += f"\\resizebox{{\\textwidth}}{{!}}{{%\n\\begin{{tabular}}{{{''.join(alignment)}}}\n"
+
+    # Add the column headers
+    headers = [f"{feat}_diff" for feat in features]
+    latex_table += " & ".join(map(lambda x: x.replace("_", "\\_"), ["model"] + features)) + " \\\\\n\\toprule\n"
+
+    # Process each row
+    table_rows = []
+    for index, row in df.iterrows():
+        row_values = [index.replace("_", "\\_")]
+        for col in headers:
+            # Get the corresponding significance column name
+            significance_col = f"is_{col}erent"
+            # Format the value based on the significance
+            formatted_value = format_value(row[col], row[significance_col])
+            row_values.append(formatted_value)
+        table_rows.append(" & ".join(row_values) + " \\\\ ")
+    latex_table += "\\midrule\n".join(table_rows)
+
+    # End the table
+    latex_table += "\\bottomrule\n\\end{tabular}}\n"
+    latex_table += f"\\caption{{{caption}}}\n\\label{{{label}}}\n"
+    latex_table += "\\end{table}"
+
+    return latex_table

### 2024-06-11 src/llm_tst_consistency/parameters.py

@@ -11,6 +11,7 @@ class LLMName(StrEnum):
     GEMINI = "gemini"
     GPT = "gpt"
     LLAMA3 = "llama3_8b"
+    LLAMA3_70B = "llama3_70b"
     MIXTRAL = "mixtral_8x7b"
     COMMAND_R = "command_r"
 

### 2024-06-11 src/llm_tst_consistency/latex.py
fix bug in latex exporter

@@ -30,7 +30,7 @@ def make_latex_table(df, caption, label, features, alignment):
     latex_table += "\\midrule\n".join(table_rows)
 
     # End the table
-    latex_table += "\\bottomrule\n\\end{tabular}}\n"
+    latex_table += "\\bottomrule\n\\end{tabular}%\n}\n"
     latex_table += f"\\caption{{{caption}}}\n\\label{{{label}}}\n"
     latex_table += "\\end{table}"
 

### 2024-06-11 src/llm_tst_consistency/latex.py
fix another bug in the latex table generator

@@ -31,7 +31,7 @@ def make_latex_table(df, caption, label, features, alignment):
 
     # End the table
     latex_table += "\\bottomrule\n\\end{tabular}%\n}\n"
-    latex_table += f"\\caption{{{caption}}}\n\\label{{{label}}}\n"
+    latex_table += f"\\caption{{{caption.replace('_', ' ')}}}\n\\label{{{label.replace('_', '-')}}}\n"
     latex_table += "\\end{table}"
 
     return latex_table

### 2024-06-11 article/main.tex
finish up first draft

@@
 our selection.
-In the interest of covering more ground, we choose large language models from
-different vendors.
-Our choice includes four top performing LLMs and one model that anyone would be
-able to run on their own hardware.
+In the interest of covering more ground, we chose five top performers in the
+Chatbot Arena from different vendors.
 The final selection is available in Table~\ref{table-llm}.
@@
         Anthropic Claude 3 Opus & undisclosed & 200000       & 1248      & Proprietary         \\
-        Meta Llama 3-8B         & 8 Billion   & 8000         & 1208      & Llama 3 Community   \\
+        Meta Llama 3-70B        & 70 Billion  & 8000         & 1208      & Llama 3 Community   \\
         Cohere Command R+       & 104 Billion & 128000       & 1189      & CC-BY-NC 4.0        \\
@@
     \end{tabular}
-    \caption{Experiment LLMs}\label{table-llm}
+    \caption{Large Language Model Selection}\label{table-llm}
 \end{table}
@@
 
-In the next sections we only present the most relevant results of the
-experiment.
 To save space, we illustrate significant differences between the baseline and
@@
 The \underline{underlined} values represent situations when the HLF-enhanced
-prompt lead to results that are closer to the corpus average than the baseline.
-\textit{Italicized} values represent situations when the baseline was closer
-to the target than the HLF-enhanced result.
+prompt lead to results that are significantly closer to the corpus average than
+the baseline.
+\textit{Italicized} values represent situations when the baseline was
+significantly closer to the target than the HLF-enhanced result.
+
+In the next sections we only present the most relevant results of the
+experiment.
 
@@
 The results on the CNN-DailyMail dataset showcased in
-Table~\ref{table-cnn-dailymail-results} provide the most eloquent answer to this
-question.
+Table~\ref{table-prompt-1-cnn-dailymail} provide the most eloquent answer to
+this question.
 The results on the Yelp dataset are positive, albeit less convincing.
@@
             \toprule
-            claude3 $Diff_N$    & \underline{\textbf{623.16}}  & \underline{\textbf{24.74}} & \underline{\textbf{51.18}}  & \underline{\textbf{26.68}} & 0.06                      & -0.0                      & \underline{\textbf{0.46}} & \underline{\textbf{1.4}}   & \underline{1.44}        & \textit{\textbf{-1.66}}   \\
-            claude3 $Diff_A$    & \underline{\textbf{1774.0}}  & \underline{\textbf{74.0}}  & \underline{\textbf{158.84}} & \underline{\textbf{81.2}}  & 0.2                       & -0.01                     & \underline{\textbf{1.88}} & \underline{\textbf{4.49}}  & \underline{5.01}        & \textit{\textbf{-4.09}}   \\
-            gemini $Diff_N$     & \underline{\textbf{985.86}}  & \underline{\textbf{34.52}} & \underline{\textbf{100.36}} & \underline{\textbf{53.39}} & \underline{\textbf{0.52}} & 0.05                      & \underline{\textbf{2.62}} & \underline{\textbf{4.68}}  & \textit{-4.17}          & \textit{\textbf{-2.43}}   \\
-            gemini $Diff_A$     & \underline{\textbf{2990.5}}  & \underline{\textbf{106.5}} & \underline{\textbf{320.34}} & \underline{\textbf{159.1}} & \underline{\textbf{1.5}}  & 0.09                      & \underline{\textbf{7.24}} & \underline{\textbf{13.15}} & \textit{-12.65}         & \textit{\textbf{-7.13}}   \\
-            gpt $Diff_N$        & \underline{\textbf{125.87}}  & 1.46                       & 10.95                       & 10.77                      & 0.02                      & -0.01                     & 0.71                      & 0.81                       & \textit{\textbf{-1.72}} & 0.43                      \\
-            gpt $Diff_A$        & \underline{\textbf{327.0}}   & 3.0                        & 29.0                        & 27.0                       & 0.08                      & -0.01                     & 1.82                      & 2.24                       & \textit{\textbf{-5.72}} & 0.44                      \\
-            llama3\_8b $Diff_N$ & 171.31                       & 6.63                       & -9.8                        & 13.8                       & 0.04                      & -0.04                     & 1.03                      & -0.84                      & -0.11                   & -0.76                     \\
-            llama3\_8b $Diff_A$ & 524.0                        & 20.0                       & -32.5                       & 43.5                       & 0.1                       & -0.08                     & 3.64                      & -2.85                      & 1.57                    & -2.29                     \\
-            command\_r $Diff_N$ & \underline{\textbf{1024.78}} & \underline{\textbf{39.67}} & \underline{\textbf{81.58}}  & \underline{\textbf{41.46}} & \underline{\textbf{0.46}} & \underline{\textbf{0.03}} & \underline{\textbf{2.67}} & \underline{\textbf{3.78}}  & \underline{1.4}         & \underline{\textbf{0.14}} \\
-            command\_r $Diff_A$ & \underline{\textbf{3132.5}}  & \underline{\textbf{120.5}} & \underline{\textbf{246.84}} & \underline{\textbf{132.2}} & \underline{\textbf{1.36}} & \underline{\textbf{0.06}} & \underline{\textbf{7.88}} & \underline{\textbf{11.21}} & \underline{3.9}         & \underline{\textbf{1.03}} \\
-            \bottomrule
+            claude3 $Diff\_N$     & \textbf{\underline{623.16}}  & \textbf{\underline{24.74}} & \textbf{\underline{51.18}}  & \textbf{\underline{26.68}} & 0.06                      & -0.0                      & \textbf{\underline{0.46}} & \textbf{\underline{1.4}}   & 1.44                    & \textbf{\textit{-1.66}}   \\ \midrule
+            claude3 $Diff\_A$     & \textbf{\underline{1774.0}}  & \textbf{\underline{74.0}}  & \textbf{\underline{158.84}} & \textbf{\underline{81.2}}  & 0.2                       & -0.01                     & \textbf{\underline{1.88}} & \textbf{\underline{4.49}}  & 5.01                    & \textbf{\textit{-4.09}}   \\ \midrule
+            gemini $Diff\_N$      & \textbf{\underline{985.86}}  & \textbf{\underline{34.52}} & \textbf{\underline{100.36}} & \textbf{\underline{53.39}} & \textbf{\underline{0.52}} & 0.05                      & \textbf{\underline{2.62}} & \textbf{\underline{4.68}}  & -4.17                   & \textbf{\textit{-2.43}}   \\ \midrule
+            gemini $Diff\_A$      & \textbf{\underline{2990.5}}  & \textbf{\underline{106.5}} & \textbf{\underline{320.34}} & \textbf{\underline{159.1}} & \textbf{\underline{1.5}}  & 0.09                      & \textbf{\underline{7.24}} & \textbf{\underline{13.15}} & -12.65                  & \textbf{\textit{-7.13}}   \\ \midrule
+            gpt $Diff\_N$         & \textbf{\underline{125.87}}  & 1.46                       & 10.95                       & 10.77                      & 0.02                      & -0.01                     & 0.71                      & 0.81                       & \textbf{\textit{-1.72}} & 0.43                      \\ \midrule
+            gpt $Diff\_A$         & \textbf{\underline{327.0}}   & 3.0                        & 29.0                        & 27.0                       & 0.08                      & -0.01                     & 1.82                      & 2.24                       & \textbf{\textit{-5.72}} & 0.44                      \\ \midrule
+            command\_r $Diff\_N$  & \textbf{\underline{1024.78}} & \textbf{\underline{39.67}} & \textbf{\underline{81.58}}  & \textbf{\underline{41.46}} & \textbf{\underline{0.46}} & \textbf{\underline{0.03}} & \textbf{\underline{2.67}} & \textbf{\underline{3.78}}  & 1.4                     & \textbf{\underline{0.14}} \\ \midrule
+            command\_r $Diff\_A$  & \textbf{\underline{3132.5}}  & \textbf{\underline{120.5}} & \textbf{\underline{246.84}} & \textbf{\underline{132.2}} & \textbf{\underline{1.36}} & \textbf{\underline{0.06}} & \textbf{\underline{7.88}} & \textbf{\underline{11.21}} & 3.9                     & \textbf{\underline{1.03}} \\ \midrule
+            llama3\_70b $Diff\_N$ & \textbf{\underline{459.45}}  & \textbf{\underline{17.78}} & \textbf{\underline{36.5}}   & \textbf{\underline{19.75}} & \textbf{\underline{0.12}} & \textbf{\underline{0.03}} & \textbf{\underline{1.43}} & \textbf{\underline{2.18}}  & 1.5                     & \textbf{\textit{-0.06}}   \\ \midrule
+            llama3\_70b $Diff\_A$ & \textbf{\underline{1363.5}}  & \textbf{\underline{54.0}}  & \textbf{\underline{106.5}}  & \textbf{\underline{60.5}}  & \textbf{\underline{0.36}} & \textbf{\underline{0.09}} & \textbf{\underline{4.33}} & \textbf{\underline{6.38}}  & 7.46                    & \textbf{\textit{-0.11}}   \\ \bottomrule
         \end{tabular}%
-    }
-    \caption{CNN-DailyMail Results}
-    \label{table-cnn-dailymail-results}
+    }\caption{CNN Results}\label{table-prompt-1-cnn-dailymail}
 \end{table}
 
-Command-R+ was the best performer by far, while Llama3 (8B) was predictibly the
-worst at following HLF instructions.
+For our selection of features, on the CNN corpus of articles, Command-R+ was the
+best performer.
+GPT scored surprisingly low.
+
 Figure~\ref{fig-command-r-t-word} shows a positive result, where the result
@@
     \begin{minipage}{0.32\textwidth}
-        \includegraphics[width=\linewidth]{plots/prompt_1/prompt_1-llama3_8b-cnn_dailymail/prompt_1-llama3_8b-cnn_dailymail_simp_ttr.png}
-        \caption[center]{Llama3-8B\\Simple TTR}
+        \includegraphics[width=\linewidth]{plots/prompt_1/prompt_1-claude3-cnn_dailymail/prompt_1-claude3-cnn_dailymail_a_verb_pw.png}
+        \caption[center]{Claude3\\Average Verbs per Word}
         \label{fig-llama-simp-ttr}
@@
 
-In this variant of the experiment, we feed the LLM three examples from the
-corpus.
+The second type of instructional prompt we feed to the LLM contains three
+examples from the corpus.
+The hypothesis is that this will influence the baseline result to be closer to
+the corpus average.
+
+In this case, it is interesting to see how the LLM generates text based on input
+text from both within the corpus that contains the example articles and from
+outside the corpus.
+
+In Table~\ref{table-prompt-2-cnn-dailymail} we see the results when an input from
+outside the corpus is used.
+The outlook is similar for what we saw in the first experiment with GPT
+being the least responsive to HLF instructions.
+
+\begin{table}[!ht]
+    \centering
+    \resizebox{\textwidth}{!}{%
+        \begin{tabular}{lllllllllll}
+            model                 & t\_word                      & t\_sent                    & n\_uverb                    & n\_uadj                    & simp\_ttr                 & a\_verb\_pw               & corr\_adj\_var            & corr\_verb\_var           & fkgl                       & a\_kup\_pw                \\
+            \toprule
+            claude3 $Diff\_N$     & \textbf{\underline{718.49}}  & \textbf{\underline{24.13}} & \textbf{\underline{44.43}}  & \textbf{\underline{23.43}} & \textbf{\underline{0.4}}  & 0.03                      & 0.94                      & \textbf{\underline{0.77}} & 1.2                        & 0.07                      \\ \midrule
+            claude3 $Diff\_A$     & \textbf{\underline{2345.54}} & \textbf{\underline{90.18}} & \textbf{\underline{148.36}} & \textbf{\underline{63.6}}  & \textbf{\underline{1.18}} & 0.07                      & 1.36                      & \textbf{\underline{2.55}} & 3.33                       & 0.52                      \\ \midrule
+            gemini $Diff\_N$      & \textbf{\underline{785.6}}   & \textbf{\underline{25.8}}  & \textbf{\underline{60.5}}   & \textbf{\underline{41.6}}  & \textbf{\underline{0.41}} & \textbf{\underline{0.05}} & \textbf{\underline{1.95}} & \textbf{\underline{2.67}} & -4.62                      & \textbf{\textit{-1.94}}   \\ \midrule
+            gemini $Diff\_A$      & \textbf{\underline{2512.0}}  & \textbf{\underline{81.5}}  & \textbf{\underline{210.34}} & \textbf{\underline{125.7}} & \textbf{\underline{1.25}} & \textbf{\underline{0.09}} & \textbf{\underline{5.84}} & \textbf{\underline{8.05}} & -9.46                      & \textbf{\textit{-5.82}}   \\ \midrule
+            gpt $Diff\_N$         & 47.8                         & 5.18                       & 6.13                        & -0.13                      & 0.02                      & 0.0                       & 0.05                      & 0.3                       & 3.06                       & -0.27                     \\ \midrule
+            gpt $Diff\_A$         & 77.0                         & 12.5                       & 13.0                        & -4.0                       & 0.04                      & 0.0                       & -0.17                     & 0.77                      & 8.45                       & -0.16                     \\ \midrule
+            command\_r $Diff\_N$  & \textbf{\underline{263.51}}  & 5.35                       & \textbf{\underline{25.08}}  & \textbf{\underline{17.37}} & \textbf{\textit{-0.09}}   & -0.01                     & \textbf{\underline{1.74}} & \textbf{\underline{1.42}} & \textbf{\underline{10.46}} & \textbf{\underline{1.56}} \\ \midrule
+            command\_r $Diff\_A$  & \textbf{\underline{896.0}}   & 20.5                       & \textbf{\underline{86.0}}   & \textbf{\underline{60.6}}  & \textbf{\textit{-0.21}}   & -0.03                     & \textbf{\underline{5.45}} & \textbf{\underline{4.95}} & \textbf{\underline{31.29}} & \textbf{\underline{5.67}} \\ \midrule
+            llama3\_70b $Diff\_N$ & \textbf{\underline{376.0}}   & \textbf{\underline{15.63}} & \textbf{\underline{38.63}}  & 15.74                      & \textbf{\underline{0.24}} & \textbf{\underline{0.06}} & 1.69                      & \textbf{\underline{2.91}} & 0.31                       & -0.3                      \\ \midrule
+            llama3\_70b $Diff\_A$ & \textbf{\underline{1088.5}}  & \textbf{\underline{45.0}}  & \textbf{\underline{113.0}}  & 41.0                       & \textbf{\underline{0.65}} & \textbf{\underline{0.18}} & 3.53                      & \textbf{\underline{8.56}} & 0.27                       & -0.68                     \\ \bottomrule
+        \end{tabular}%
+    }
+    \caption{prompt 2 - CNN Results\\Input Outside Corpus}
+    \label{table-prompt-2-cnn-dailymail}
+\end{table}
+
+GPT's results would be good if the baseline were very close to the target value.
+As we can see in Figure~\ref{fig-p2-gpt-twords}, this is not always the case.
+We also notice that Command-R+(fig.~\ref{fig-p2-command-r-simpttr}) and Gemini
+(fig.~\ref{fig-p2-gemini-a-kup-pw}) both registered significantly negative
+results, in contrast to their performance in our first scenario.
+
+\begin{figure}[ht!]
+    \centering
+    \begin{minipage}{0.32\textwidth}
+        \includegraphics[width=\linewidth]{plots/prompt_2/prompt_2-gpt-cnn_dailymail/prompt_2-gpt-cnn_dailymail_t_word.png}
+        \caption{GPT\\Total Words}
+        \label{fig-p2-gpt-twords}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.32\textwidth}
+        \includegraphics[width=\linewidth]{plots/prompt_2/prompt_2-command_r-cnn_dailymail/prompt_2-command_r-cnn_dailymail_simp_ttr.png}
+        \caption{Command-R+\\Simple TTR}
+        \label{fig-p2-command-r-simpttr}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.32\textwidth}
+        \includegraphics[width=\linewidth]{plots/prompt_2/prompt_2-gemini-cnn_dailymail/prompt_2-gemini-cnn_dailymail_a_kup_pw.png}
+        \caption[center]{Gemini\\Kuperman Age of Aquisition}
+        \label{fig-p2-gemini-a-kup-pw}
+    \end{minipage}
+\end{figure}
+
+The results obtained in our first scenario are largely confirmed by our second
+scenario.
+So now we turn to see how using a text from within the corpus affects the
+performance of the large language models.
+The hypothesis is that if the text is `familiar', both the baseline and the HLF
+instruction prompt will produce results that are much closer to the target
+value.
+The expectation is to see less significant results.
+
+\begin{table}[!ht]
+    \centering
+    \resizebox{\textwidth}{!}{%
+        \begin{tabular}{lllllllllll}
+            model                 & t\_word                      & t\_sent                     & n\_uverb                    & n\_uadj                    & simp\_ttr                 & a\_verb\_pw               & corr\_adj\_var             & corr\_verb\_var            & fkgl                       & a\_kup\_pw              \\
+            \toprule
+            claude3 $Diff\_N$     & \textbf{\underline{871.02}}  & \textbf{\underline{34.2}}   & \textbf{\underline{34.48}}  & \textbf{\textit{-7.38}}    & \textbf{\underline{0.27}} & 0.01                      & \textbf{\textit{-0.93}}    & \textbf{\textit{-0.72}}    & 1.16                       & -0.14                   \\ \midrule
+            claude3 $Diff\_A$     & \textbf{\underline{2756.12}} & \textbf{\underline{117.18}} & \textbf{\underline{126.36}} & \textbf{\textit{-11.2}}    & \textbf{\underline{0.81}} & -0.01                     & \textbf{\textit{-2.52}}    & \textbf{\textit{-0.74}}    & 3.69                       & -0.4                    \\ \midrule
+            gemini $Diff\_N$      & \textbf{\underline{1315.41}} & \textbf{\underline{46.19}}  & \textbf{\underline{114.47}} & \textbf{\underline{72.38}} & \textbf{\underline{1.17}} & 0.03                      & \textbf{\underline{5.25}}  & \textbf{\underline{6.64}}  & -1.6                       & \textbf{\textit{-2.1}}  \\ \midrule
+            gemini $Diff\_A$      & \textbf{\underline{3996.0}}  & \textbf{\underline{138.0}}  & \textbf{\underline{357.36}} & \textbf{\underline{212.7}} & \textbf{\underline{3.3}}  & 0.06                      & \textbf{\underline{15.49}} & \textbf{\underline{19.4}}  & -13.22                     & \textbf{\textit{-6.19}} \\ \midrule
+            gpt $Diff\_N$         & \textbf{\underline{521.29}}  & \textbf{\underline{23.3}}   & \textbf{\underline{43.13}}  & 19.85                      & \textbf{\underline{0.2}}  & 0.02                      & 1.08                       & \textbf{\underline{1.04}}  & 5.09                       & 0.67                    \\ \midrule
+            gpt $Diff\_A$         & \textbf{\underline{1464.0}}  & \textbf{\underline{65.5}}   & \textbf{\underline{121.84}} & 58.1                       & \textbf{\underline{0.58}} & 0.05                      & 2.93                       & \textbf{\underline{2.59}}  & 14.23                      & 1.89                    \\ \midrule
+            command\_r $Diff\_N$  & \textbf{\underline{371.08}}  & \textbf{\underline{13.31}}  & 22.06                       & \textbf{\underline{17.19}} & \textbf{\underline{0.17}} & \textbf{\underline{0.02}} & 0.88                       & 1.09                       & \textbf{\underline{4.05}}  & -0.15                   \\ \midrule
+            command\_r $Diff\_A$  & \textbf{\underline{1198.0}}  & \textbf{\underline{42.5}}   & 73.0                        & \textbf{\underline{56.5}}  & \textbf{\underline{0.69}} & \textbf{\underline{0.07}} & 2.85                       & 3.99                       & \textbf{\underline{13.89}} & -0.11                   \\ \midrule
+            llama3\_70b $Diff\_N$ & \textbf{\underline{1040.85}} & \textbf{\underline{41.99}}  & \textbf{\underline{80.58}}  & \textbf{\underline{49.12}} & \textbf{\underline{0.55}} & \textbf{\underline{0.04}} & \textbf{\underline{3.99}}  & \textbf{\underline{5.05}}  & 7.18                       & \textbf{\textit{-1.94}} \\ \midrule
+            llama3\_70b $Diff\_A$ & \textbf{\underline{2928.0}}  & \textbf{\underline{118.5}}  & \textbf{\underline{222.0}}  & \textbf{\underline{133.5}} & \textbf{\underline{1.43}} & \textbf{\underline{0.05}} & \textbf{\underline{10.77}} & \textbf{\underline{13.42}} & 17.21                      & \textbf{\textit{-6.2}}  \\ \bottomrule
+        \end{tabular}%
+    }
+    \caption{CNN Results\\Input Inside Corpus}
+    \label{table-prompt-2-ifd-cnn-dailymail}
+\end{table}
+
+The results shown in Table~\ref{table-prompt-2-ifd-cnn-dailymail} contradict our
+expectations outright.
+The most surprising behaviour is exhibited in the context of more concrete HLFs
+such as the total number of words or the number of unique adjectives.
+Claude3 shows regressive behaviour for the number of unique adjectives in
+Figure~\ref{fig-p2-ifd-claude3-nuadj}, while Gemini constructs texts using a
+very different amount of words in Figure~\ref{fig-p2-ifd-gemini-twords}.
+Less surprisingly, Llama3 fails to generate text that conforms to the correct
+Kuperman age of acquisition per word in Figure~\ref{fig-p2-ifd-llama3-a-kup-pw}.
+
+\begin{figure}[ht!]
+    \centering
+    \begin{minipage}{0.32\textwidth}
+        \includegraphics[width=\linewidth]{plots/prompt_2_ifd/prompt_2-claude3-cnn_dailymail/prompt_2-claude3-cnn_dailymail_n_uadj.png}
+        \caption{Claude3\\Number of Unique Adjectives}
+        \label{fig-p2-ifd-claude3-nuadj}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.32\textwidth}
+        \includegraphics[width=\linewidth]{plots/prompt_2_ifd/prompt_2-gemini-cnn_dailymail/prompt_2-gemini-cnn_dailymail_t_word.png}
+        \caption{Gemini\\Total Words}
+        \label{fig-p2-ifd-gemini-twords}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.32\textwidth}
+        \includegraphics[width=\linewidth]{plots/prompt_2_ifd/prompt_2-llama3_70b-cnn_dailymail/prompt_2-llama3_70b-cnn_dailymail_a_kup_pw.png}
+        \caption[center]{Llama3\\Kuperman Age of Aquisition}
+        \label{fig-p2-ifd-llama3-a-kup-pw}
+    \end{minipage}
+\end{figure}
+
+The results in this last scenario show that whether LLMs take examples listed in
+the prompt into account or not with respect to conforming to certain linguistic
+features is debatable, providing them with specific instructions derived from
+HLFs makes a significant impact.
+
 \section{Discussion}
 
-Discuss experiment results here.
+In terms of our first question --- whether LLMs are able to follow instructions
+concerning handcrafted linguistic features --- the response is in our opinion
+affirmative.
+Certainly, not all LLMs are created equal in this regard.
+It is debatable whether this experiment is well designed because of the choice
+to derive target HLF values for the LLM from the corpus.
+This is evident in the results obtained for the Yelp dataset.
+While these results also show a significant influence when using HLFs in the
+prompt, the influence is mostly negative.
+This can be attributed to the choice of input text in relation to the corpus,
+the wording in the instruction prompt or to each model's specific training.
+
+This leads us to our second question: how much do handcrafted linguistic feature
+instructions affect the way LLMs write text.
+The results in this regards show us a much less conclusive picture.
+
+Firstly, we cannot assume that the LLM will follow HLF instructions the way we
+intend.
+Whether we obtain the expected outcome or not depends largely on the prompt,
+the chosen HLFs, the model, the target values for the HLFs, a.s.o
+We do, however, notice a few patterns.
+The more abstract the instruction, the less it seems to make a positve impact.
+More importantly, by inspecting the result data, a more abstract instruction
+seems more likely to produce a negative impact on the desired outcome.
+
+Secondly, the choice of input text in relation to the target HLF values seems
+very important.
+Choosing to reword a poem so that it has the same reading ease as a user review
+seems like a setup for poor results.
+This leads us to intuit that LLMs are clearly able to use HLF instructions to
+generate text differently, but there is a limit to their `creativity'.
+
+Finally, the choice of HLFs is very important.
+The text generated by an LLM is less likely to conform to HLFs that result in
+more abstract instructions.
+Furthermore, the combination of chosen HLFs is important.
+Choosing HLFs that measure the uniqueness of a linguistic element (words,
+structures, a.s.o) limits the LLMs ability to conform to other HLFs (for example
+total number of sentences or reading ease).
+This is highlighted in the experiment specifically in the tension between the
+age of acquisition and the adjective or verb variation.
+Predictably, the more concrete instructions take precedence over the more
+abstract ones.
 
@@
 
-Yes, you can instruct at least some of the LLMs with HLFs.
+We started out by trying to understand whether it is possible to conform the
+output of a large language model to exhibit specific linguistic features.
+To our surprise, it turns out that instructing language models to write text
+that conforms to a selected assortment of linguistic features almost always
+yields significant changes in how the text is being generated.
 
-compute BLEU, meteor, cider on styleptb using these LLMs and the prompt to
-evaluate whether we actually achieve style transfer because of/in spite of HLF
-instructions.
-Human survey (amazon mechanical turk) on whether the text generations are good
-or not. Likert scale (0 no good, 10 matches style perfectly)
+This doesn't mean that the results are always satisfying.
+What the model writes about and the chosen target values for the linguistic
+features are major factors in how successful the overall experience is.
+Unfortunately, at this stage, finding out good target values and good writing
+instructions is a question of trial and error.
+
+Other experiment designs might have been more opportune in examining the
+questions layed out in the introduction.
+For example, using set target values for HLFs (without deriving them from a
+corpus of text) and choosing input text where the HLFs vary greatly would have
+been an alternate design.
+Following through on those approaches is one of the possible expansions on the
+current work.
+
+Another way of expanding on the current work is finding the settings in which
+large language models are most responsive to linguistic features.
+The current experiment was inconclusive in this regard and a legitimate further
+question is whether settings in which text generated by language models
+consistently conforms to certain linguistic features actually exist.
+
+Lastly, a simpler question that is left unanswered relates to the number of
+linguistic features that were used.
+What if we would have measured models' behaviour for single linguistic features
+independently?
 

### 2024-06-12 src/llm_tst_consistency/plot.py
tweak article plots to look nicer for print

@@ -91,11 +91,20 @@ def draw_one_figure(title, features, df, ds_stats, plot_col_count):
 
 
 def draw_multiple_figures(title, features, df, ds_stats):
-    for i, f in enumerate(features):
+    for f in features:
         fig = go.Figure()
         for trace in _make_feature_traces(f, df, ds_stats):
             fig.add_trace(trace)
-        fig.update_layout(height=480, width=640, title_text=f"{title} - {f}")
+            
+        fig.update_layout(
+            height=480,
+            width=640,
+            font_color="black",
+            #title_text=f"{title} - {f}",
+            showlegend=False,
+            paper_bgcolor='rgba(0,0,0,0)',
+            plot_bgcolor='rgba(0,0,0,0)',
+        )
         yield f, fig
 
 

### 2024-06-12 article/main.tex
apply ACL styling

@@
-\documentclass[runningheads,a4paper,11pt]{article}
+\pdfoutput=1
+
+\documentclass[11pt]{article}
+\usepackage[review]{acl}
 
@@
 \usepackage{amsthm}
+\usepackage{balance}
 \usepackage{booktabs}
-\usepackage[justification=centering,font=scriptsize]{caption}
-\usepackage{cite}
+\usepackage[justification=centering]{caption}
 \usepackage{comment}
@@
 \usepackage{graphicx}
-\usepackage[colorlinks]{hyperref}
 \usepackage{inconsolata}
-\usepackage[latin1]{inputenc}
+\usepackage[utf8]{inputenc}
 \usepackage{interval}
+\usepackage{microtype}
 \usepackage{multicol}
@@
 
-\geometry{a4paper,top=2.5cm,left=2cm,right=2cm,bottom=2cm}
+\title{LLM Output Compliance with Handcrafted Linguistic Features --- An Experiment}
+\author{Andrei Olar\\
+Universitatea Babeș-Bolyai, Cluj-Napoca, România\\
+\texttt{andrei.olar@ubbcluj.ro}\\}
 
 \begin{document}
-\title{How Consistent are LLMs in Applying Handcrafted Language Features}
-\author{Andrei Olar --- andrei.olar@ubbcluj.ro}
 \maketitle
@@
 
-One final observation before diving into details is that we can use an HLF's
-variance on the benchmark dataset to our advantage.
-This statistic shows whether the LLM takes the HLF prompts into account enough
-for the generated text to fall within the parameters of the benchmark dataset.
-Let us denote an HLF computed on a generated text using an HLF prompt with
-$HLF_g$.
-Similarly denote the average and the variance of the same HLF on the benchmark
-dataset with $avg(HLF_b)$ and $var(HLF_b)$, respectively.
-If we consistently notice that
-\[HLF_g - var(HLF_b) \leq avg(HLF_b) \leq HLF_g + var(HLF_b)\]
-then we consider that the LLM follows HLF prompts in a relevant way.
-
 \subsection{LLM Selection}\label{llm-selection}
@@
 Chatbot Arena from different vendors.
-The final selection is available in Table~\ref{table-llm}.
+The final selection is available in Table~\ref{table-llm}. The identifier
 
-\begin{table}[ht]
+\begin{table*}[ht]
     \setlength\tabcolsep{6pt}
     \centering
-    \begin{tabular}{@{}lllll@{}}
+    \resizebox{\textwidth}{!}{%
+        \begin{tabular}{@{}llllll@{}}
             \toprule
-        Name                    & Parameters  & Context Size & Elo Score & License Type        \\ \toprule
-        OpenAI GPT-4o           & undisclosed & 128000       & 1287      & Proprietary         \\
-        Google Gemini 1.5 Pro   & undisclosed & 1048576      & 1268      & Proprietary         \\
-        Anthropic Claude 3 Opus & undisclosed & 200000       & 1248      & Proprietary         \\
-        Meta Llama 3-70B        & 70 Billion  & 8000         & 1208      & Llama 3 Community   \\
-        Cohere Command R+       & 104 Billion & 128000       & 1189      & CC-BY-NC 4.0        \\
-                                &             &              &           & with Acceptable Use \\
-                                &             &              &           & Addendum            \\ \bottomrule
-    \end{tabular}
+            ID          & Name                    & Parameters  & Context Size & Elo Score & License Type        \\ \toprule
+            gpt         & OpenAI GPT-4o           & undisclosed & 128000       & 1287      & Proprietary         \\
+            gemini      & Google Gemini 1.5 Pro   & undisclosed & 1048576      & 1268      & Proprietary         \\
+            claude3     & Anthropic Claude 3 Opus & undisclosed & 200000       & 1248      & Proprietary         \\
+            llama3\_70b & Meta Llama 3-70B        & 70 Billion  & 8000         & 1208      & Llama 3 Community   \\
+            command\_r  & Cohere Command R+       & 104 Billion & 128000       & 1189      & CC-BY-NC 4.0        \\
+                        &                         &             &              &           & with Acceptable Use \\
+                        &                         &             &              &           & Addendum            \\ \bottomrule
+        \end{tabular}%
+    }
     \caption{Large Language Model Selection}\label{table-llm}
-\end{table}
+\end{table*}
+
+The ID of the LLM is used as the reference throughout figures and tables.
 
@@
 
-\begin{table}[ht]
+\begin{table*}[ht]
     \setlength\tabcolsep{6pt}
     \centering
-    \begin{tabular}{@{}lllll@{}}
+    \resizebox{\textwidth}{!}{%
+        \begin{tabular}{@{}llllll@{}}
             \toprule
-        Name                              & Family           & Domain           & Abstraction     & Style Level\footnotemark \\ \toprule
-        total words                       & wordsent         & surface          & very concrete   & lexical                  \\
-        total sentences                   & wordsent         & surface          & very concrete   & syntactic                \\ \bottomrule
-        total number of unique adjectives & partofspeech     & syntax           & concrete        & semantic                 \\
-        total number of unique adjectives & partofspeech     & syntax           & concrete        & syntactic                \\ \bottomrule
-        simple type token ratio           & typetokenratio   & lexico-semantics & regular         & pragmatic                \\
-        number of verbs per word          & partofspeech     & surface          & regular         & syntactic                \\ \bottomrule
-        corrected adjective variation     & lexicalvariation & lexico-semantics & abstract        & lexical                  \\
-        corrected verb variation          & lexicalvariation & lexico-semantics & abstract        & lexical                  \\ \bottomrule
-        Flesch-Kincaid grade level        & readformula      & surface          & highly abstract & pragmatic                \\
-        Kuperman age of aquisition        & worddiff         & lexico-semantics & highly abstract & lexical                  \\ \bottomrule
-    \end{tabular}
+            LFTK ID         & Name                              & Family           & Domain           & Abstraction     & Style Level\footnotemark \\ \toprule
+            t\_word         & total words                       & wordsent         & surface          & very concrete   & lexical                  \\
+            t\_sent         & total sentences                   & wordsent         & surface          & very concrete   & syntactic                \\ \bottomrule
+            n\_uverb        & total number of unique adjectives & partofspeech     & syntax           & concrete        & semantic                 \\
+            n\_uadj         & total number of unique adjectives & partofspeech     & syntax           & concrete        & syntactic                \\ \bottomrule
+            simp\_ttr       & simple type token ratio           & typetokenratio   & lexico-semantics & regular         & pragmatic                \\
+            a\_verb\_pw     & number of verbs per word          & partofspeech     & surface          & regular         & syntactic                \\ \bottomrule
+            corr\_adj\_var  & corrected adjective variation     & lexicalvariation & lexico-semantics & abstract        & lexical                  \\
+            corr\_verb\_var & corrected verb variation          & lexicalvariation & lexico-semantics & abstract        & lexical                  \\ \bottomrule
+            fkgl            & Flesch-Kincaid grade level        & readformula      & surface          & highly abstract & pragmatic                \\
+            a\_kup\_pw      & Kuperman age of aquisition        & worddiff         & lexico-semantics & highly abstract & lexical                  \\ \bottomrule
+        \end{tabular}%
+    }
     \caption{Handcrafted Linguistic Features}\label{table-hlf}
-\end{table}
+\end{table*}
 \footnotetext{as described in the figure in~\cite{lugea2023stylistics}, chapter 1, page 7}
 
-Each of these features is well documented~\cite{lftk-2023} barring the need for
-a deeper dive here into the subject matter.
+Each of these features is well documented~\cite{lftk-2023}, barring the need for
+an extended description here.
+The LFTK ID is used throughout tables and figures to refer to a specific HLF.
 
@@
 
-We must choose source datasets based upon which we create corpuses for our
-experiment.
+The experiment design rests upon measuring HLFs on corpuses of texts and
+computing statistics.
+LLMs also make use of examples extracted from a corpus to guide text generation.
+We obtain a corpus from a dataset.
 Our final selection of datasets is listed in Table~\ref{table-ds}.
 
-\begin{table}[ht]
+\begin{table*}[ht]
     \setlength\tabcolsep{6pt}
     \centering
+    \resizebox{\textwidth}{!}{%
         \begin{tabular}{@{}lllll@{}}
             \toprule
-        Name            & Task                     & Row Count & Content Type & Reference              \\ \toprule
+            Name            & Task                     & Number of Samples & Content Type & Reference              \\ \toprule
             Yelp Reviews    & Sentiment Classification & 50000             & Reviews      & \cite{yelp2015neurips} \\
@@
                             & Text Generation          &                   &              &                        \\ \midrule
-    \end{tabular}
+        \end{tabular}%
+    }
     \caption{Source Datasets}\label{table-ds}
-\end{table}
+\end{table*}
 
@@
 
+Lastly, it should be noted that each HLF has an independent value range.
+The number of total words ranges from small numbers into the hundreds or more,
+while the corrected verb variation is always a small number.
+The area and the Euclidian norm are calculated using the raw values of each HLF.
+That makes the values in each table column incomparable to the values in another
+table column.
+Because we don't show the measure of the difference between the baseline and the
+result, but the relative closeness to a target based on that measure, comparing
+values from the same column is also ill-advised.
+
 In the next sections we only present the most relevant results of the
@@
 
-\subsection{Can an LLM Follow HLF Instructions?}
+\subsection{HLF Instruction Impact}
 
-The results on the CNN-DailyMail dataset showcased in
-Table~\ref{table-prompt-1-cnn-dailymail} provide the most eloquent answer to
-this question.
-The results on the Yelp dataset are positive, albeit less convincing.
-\begin{table}[!ht]
+The results on the corpus containing CNN articles showcased in
+Table~\ref{table-prompt-1-cnn-dailymail} shows that most LLMs take into account
+instructions concerning the linguistic features of their output.
+
+The results on the review corpus generated from the Yelp dataset are similarly
+significant, but the HLF instructions determine outcomes that are further away
+from the target.
+
+\begin{table*}[htp]
     \centering
@@
     }\caption{CNN Results}\label{table-prompt-1-cnn-dailymail}
-\end{table}
-
-For our selection of features, on the CNN corpus of articles, Command-R+ was the
-best performer.
-GPT scored surprisingly low.
+\end{table*}
 
@@
 
-\begin{figure}[ht!]
+\begin{figure*}[htp]
     \centering
@@
     \end{minipage}
-\end{figure}
+\end{figure*}
 
-\subsection{Context Aware}
+\subsection{HLF Instructions and Context Awareness}
 
@@
 
-\begin{table}[!ht]
+\begin{table*}[hpb]
     \centering
@@
     \label{table-prompt-2-cnn-dailymail}
-\end{table}
+\end{table*}
 
@@
 
-\begin{figure}[ht!]
+\begin{figure*}[hpb]
     \centering
@@
     \end{minipage}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{table}[!ht]
+\begin{table*}[ht]
     \centering
@@
     \label{table-prompt-2-ifd-cnn-dailymail}
-\end{table}
+\end{table*}
 
@@
 
-\begin{figure}[ht!]
+\begin{figure*}[ht]
     \centering
@@
 
-\bibliographystyle{plain}
 \bibliography{references}
+
 \end{document}

### 2024-06-12 article/main.tex
Prepare article for review

@@
 \usepackage[justification=centering]{caption}
+\usepackage{cleveref}
 \usepackage{comment}
@@
 
+\usepackage[true]{anonymous-acm}
+
 \title{LLM Output Compliance with Handcrafted Linguistic Features --- An Experiment}
+\authoranon{
 \author{Andrei Olar\\
@@
 \texttt{andrei.olar@ubbcluj.ro}\\}
+}
 
@@
 values.
-The prompt templates are available in the code repository that accompanies this
-experiment~\cite{olar2024experimentcode}.
+The prompt templates part of the experiment's source code.
 The instructions concerning all choseon HLFs are given concomitently within the
@@
 corpus.
-They can be examined online in the
-\texttt{\href{https://github.com/koosie0507/llm-tst-consistency/tree/main/data}{data}}
-folder of the experiment's Git repository~\cite{olar2024experimentcode}.
+The input texts are located in the
+\texttt{\textlinkanon{https://github.com/koosie0507/llm-tst-consistency/tree/main/data}{data}}
+folder of the experiment's source code.
 These user prompts are used when we're not using an input text from the
@@
 templates.
-The Jinja templates are available in the
-\texttt{\href{https://github.com/koosie0507/llm-tst-consistency/tree/main/data}{templates}}
-folder of the experiment's Git repository~\cite{olar2024experimentcode}.
+The Jinja templates are located in the
+\texttt{\textlinkanon{https://github.com/koosie0507/llm-tst-consistency/tree/main/data}{templates}}
+folder of the experiment's source code.
 
@@
 prompt and a dataset.
-All result tables for each dataset and prompt variant are available online, in
-the experiment's Git repository~\cite{olar2024experimentcode}.
+All result tables for each dataset and prompt variant are available in the
+appendix.
 
@@
 
+\appendix
+
+\section{Additional Result Tables}\label{sec:yelp-tables}
+\subsection{Yelp}
+\Cref{table-prompt-1-yelp,table-prompt-2-yelp,table-prompt-2-ifd-yelp} show the
+experiment results on the corpus of reviews extracted from the Yelp dataset.
+
+\begin{table*}[ht]
+    \centering
+    \resizebox{\textwidth}{!}{%
+        \begin{tabular}{lllllllllll}
+            model                 & t\_word                   & t\_sent                    & n\_uverb                 & n\_uadj                  & simp\_ttr & a\_verb\_pw             & corr\_adj\_var          & corr\_verb\_var         & fkgl                       & a\_kup\_pw               \\
+            \toprule
+            claude3 $Diff\_N$     & \textbf{\textit{-281.9}}  & \textbf{\textit{-13.22}}   & \textbf{\textit{-25.79}} & -14.56                   & -0.06     & 0.0                     & -1.27                   & -1.14                   & \textbf{\underline{6.89}}  & \textbf{\underline{1.0}} \\ \midrule
+            claude3 $Diff\_A$     & \textbf{\textit{-774.0}}  & \textbf{\textit{-27.76}}   & \textbf{\textit{-70.0}}  & -38.0                    & -0.19     & 0.02                    & -3.25                   & -3.39                   & \textbf{\underline{21.68}} & \textbf{\underline{2.9}} \\ \midrule
+            gemini $Diff\_N$      & -109.27                   & 0.05                       & -7.25                    & -13.18                   & -0.14     & 0.01                    & -1.01                   & -0.4                    & 1.62                       & -0.4                     \\ \midrule
+            gemini $Diff\_A$      & -252.82                   & -0.78                      & -17.5                    & -23.14                   & -0.52     & 0.01                    & -1.85                   & -1.07                   & 7.14                       & -0.93                    \\ \midrule
+            gpt $Diff\_N$         & -21.62                    & 2.42                       & -3.8                     & -8.84                    & 0.0       & -0.01                   & -0.8                    & -0.53                   & 8.94                       & -0.37                    \\ \midrule
+            gpt $Diff\_A$         & -34.72                    & 9.68                       & -10.54                   & -24.57                   & -0.03     & -0.02                   & -2.37                   & -1.58                   & 19.26                      & -1.03                    \\ \midrule
+            command\_r $Diff\_N$  & -40.1                     & 1.22                       & 2.43                     & \textbf{\textit{-13.64}} & 0.05      & \textbf{\textit{-0.04}} & \textbf{\textit{-1.17}} & 0.59                    & \textbf{\textit{-7.25}}    & \textbf{\textit{-1.67}}  \\ \midrule
+            command\_r $Diff\_A$  & -75.5                     & 1.5                        & 12.94                    & \textbf{\textit{-34.0}}  & 0.19      & \textbf{\textit{-0.12}} & \textbf{\textit{-2.76}} & 2.83                    & \textbf{\textit{-20.49}}   & \textbf{\textit{-4.76}}  \\ \midrule
+            llama3\_70b $Diff\_N$ & \textbf{\textit{-210.45}} & \textbf{\underline{6.87}}  & \textbf{\textit{-13.46}} & \textbf{\textit{-15.17}} & -0.16     & \textbf{\textit{-0.02}} & \textbf{\textit{-1.78}} & \textbf{\textit{-1.31}} & 3.7                        & \textbf{\textit{-1.58}}  \\ \midrule
+            llama3\_70b $Diff\_A$ & \textbf{\textit{-607.52}} & \textbf{\underline{19.42}} & \textbf{\textit{-41.22}} & \textbf{\textit{-44.77}} & -0.42     & \textbf{\textit{-0.05}} & \textbf{\textit{-5.48}} & \textbf{\textit{-4.34}} & 9.79                       & \textbf{\textit{-4.98}}  \\ \bottomrule
+        \end{tabular}%
+    }
+    \caption{Yelp Results - Prompt Without Examples}
+    \label{table-prompt-1-yelp}
+\end{table*}
+
+\begin{table*}[!ht]
+    \centering
+    \resizebox{\textwidth}{!}{%
+        \begin{tabular}{lllllllllll}
+            model                 & t\_word & t\_sent & n\_uverb & n\_uadj & simp\_ttr & a\_verb\_pw & corr\_adj\_var            & corr\_verb\_var & fkgl                       & a\_kup\_pw \\
+            \toprule
+            claude3 $Diff\_N$     & -121.77 & -4.61   & -9.9     & 18.41   & -0.11     & 0.03        & \textbf{\underline{1.4}}  & -0.33           & 1.81                       & 0.95       \\ \midrule
+            claude3 $Diff\_A$     & -394.5  & -11.5   & -37.5    & 60.0    & -0.37     & 0.09        & \textbf{\underline{4.66}} & -1.49           & 5.38                       & 2.62       \\ \midrule
+            gemini $Diff\_N$      & -23.49  & 1.52    & -0.55    & -1.29   & -0.02     & -0.01       & 0.09                      & -0.05           & \textbf{\underline{6.41}}  & 0.2        \\ \midrule
+            gemini $Diff\_A$      & -83.0   & 3.0     & -2.5     & -6.36   & -0.0      & -0.03       & 0.26                      & -0.08           & \textbf{\underline{19.63}} & 0.53       \\ \midrule
+            gpt $Diff\_N$         & 13.66   & 2.83    & -3.65    & -3.19   & 0.04      & -0.04       & -0.24                     & -0.22           & 7.5                        & 0.0        \\ \midrule
+            gpt $Diff\_A$         & 55.5    & 9.5     & -9.86    & -12.86  & 0.14      & -0.11       & -0.84                     & -0.22           & 18.67                      & -0.38      \\ \midrule
+            command\_r $Diff\_N$  & -188.71 & 0.69    & -10.73   & 1.17    & -0.23     & -0.01       & \textbf{\underline{0.74}} & -0.83           & \textbf{\textit{-5.02}}    & -0.08      \\ \midrule
+            command\_r $Diff\_A$  & -516.82 & 2.02    & -29.22   & 8.28    & -0.66     & -0.0        & \textbf{\underline{2.98}} & -2.48           & \textbf{\textit{-14.52}}   & -0.48      \\ \midrule
+            llama3\_70b $Diff\_N$ & -202.95 & -7.84   & -13.73   & 1.43    & -0.15     & -0.04       & 0.38                      & -1.39           & 10.67                      & 1.37       \\ \midrule
+            llama3\_70b $Diff\_A$ & -561.24 & -18.78  & -49.94   & 5.07    & -0.38     & -0.14       & 1.62                      & -4.81           & 29.26                      & 3.11       \\ \bottomrule
+        \end{tabular}%
+    }
+    \caption{Yelp Results - Prompt With Examples\\Input External To Corpus}
+    \label{table-prompt-2-yelp}
+\end{table*}
+\begin{table*}[!ht]
+    \centering
+    \resizebox{\textwidth}{!}{%
+        \begin{tabular}{lllllllllll}
+            model                 & t\_word                   & t\_sent                 & n\_uverb                  & n\_uadj                  & simp\_ttr               & a\_verb\_pw & corr\_adj\_var          & corr\_verb\_var         & fkgl                    & a\_kup\_pw              \\
+            \toprule
+            claude3 $Diff\_N$     & -148.13                   & -1.63                   & -3.93                     & \textbf{\textit{-46.03}} & -0.11                   & -0.0        & \textbf{\textit{-3.08}} & -0.31                   & -4.3                    & \textbf{\textit{-1.53}} \\ \midrule
+            claude3 $Diff\_A$     & -526.14                   & -7.98                   & -23.86                    & \textbf{\textit{-133.5}} & -0.41                   & -0.0        & \textbf{\textit{-8.98}} & -1.75                   & -21.03                  & \textbf{\textit{-5.34}} \\ \midrule
+            gemini $Diff\_N$      & 74.27                     & 0.65                    & 4.67                      & 2.98                     & 0.04                    & -0.0        & -0.22                   & 1.05                    & 1.44                    & -0.65                   \\ \midrule
+            gemini $Diff\_A$      & 221.0                     & 2.0                     & 10.5                      & 7.44                     & 0.12                    & -0.03       & -0.85                   & 2.73                    & 3.28                    & -1.79                   \\ \midrule
+            gpt $Diff\_N$         & 0.65                      & 4.03                    & -1.19                     & -2.66                    & -0.03                   & -0.01       & -0.17                   & -0.13                   & 1.05                    & -0.67                   \\ \midrule
+            gpt $Diff\_A$         & 17.0                      & 11.68                   & -3.68                     & -7.0                     & -0.1                    & -0.04       & -0.43                   & -0.05                   & 0.57                    & -1.75                   \\ \midrule
+            command\_r $Diff\_N$  & \textbf{\textit{-236.11}} & \textbf{\textit{-3.6}}  & \textbf{\textit{-0.97}}   & \textbf{\textit{-27.04}} & \textbf{\textit{-0.23}} & 0.0         & -1.24                   & \textbf{\textit{-1.08}} & \textbf{\textit{-2.13}} & \textbf{\textit{-1.09}} \\ \midrule
+            command\_r $Diff\_A$  & \textbf{\textit{-727.0}}  & \textbf{\textit{-9.24}} & \textbf{\underline{0.76}} & \textbf{\textit{-68.5}}  & \textbf{\textit{-0.64}} & 0.02        & -2.89                   & \textbf{\textit{-3.08}} & \textbf{\textit{-5.5}}  & \textbf{\textit{-3.0}}  \\ \midrule
+            llama3\_70b $Diff\_N$ & \textbf{\textit{-333.03}} & \textbf{\textit{-5.19}} & 4.43                      & \textbf{\textit{-21.38}} & -0.16                   & -0.03       & \textbf{\textit{-1.77}} & -0.03                   & 1.87                    & -0.44                   \\ \midrule
+            llama3\_70b $Diff\_A$ & \textbf{\textit{-687.34}} & \textbf{\textit{-3.42}} & 10.7                      & \textbf{\textit{-61.64}} & -0.31                   & -0.04       & \textbf{\textit{-5.6}}  & -0.89                   & 3.0                     & -1.92                   \\ \bottomrule
+        \end{tabular}%
+    }
+    \caption{Yelp Results - Prompt With Examples\\Input From Corpus}
+    \label{table-prompt-2-ifd-yelp}
+\end{table*}
+
+\section{Additional Plots}
+
+\subsection{GPT}
+
+\Cref{fig:gpt-prompt1-cnn,fig:gpt-prompt1-yelp,fig:gpt-prompt2-cnn,fig:gpt-prompt2-yelp,fig:gpt-prompt2-cnn-ifd,fig:gpt-prompt2-yelp-ifd}
+show the experiment results for the Claude 3 Opus large language model.
+
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-gpt-cnn_dailymail/prompt_1-gpt-cnn_dailymail.png}
+    \caption{GPT on CNN Corpus\\Prompt Without Examples}
+    \label{fig:gpt-prompt1-cnn}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-gpt-yelp/prompt_1-gpt-yelp.png}
+    \caption{GPT on Yelp Corpus\\Prompt Without Examples}
+    \label{fig:gpt-prompt1-yelp}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-gpt-cnn_dailymail/prompt_2-gpt-cnn_dailymail.png}
+    \caption{GPT on CNN Corpus\\Prompt With Examples\\Extraneous Input}
+    \label{fig:gpt-prompt2-cnn}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-gpt-yelp/prompt_2-gpt-yelp.png}
+    \caption{GPT on Yelp Corpus\\Prompt With Examples\\Extraneous Input}
+    \label{fig:gpt-prompt2-yelp}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-gpt-cnn_dailymail/prompt_2-gpt-cnn_dailymail.png}
+    \caption{GPT on CNN Corpus\\Prompt With Examples\\Input from Corpus}
+    \label{fig:gpt-prompt2-cnn-ifd}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-gpt-yelp/prompt_2-gpt-yelp.png}
+    \caption{GPT on Yelp Corpus\\Prompt With Examples\\Input from Corpus}
+    \label{fig:gpt-prompt2-yelp-ifd}
+\end{figure*}
+
+\subsection{Gemini}
+
+\Cref{fig:gemini-prompt1-cnn,fig:gemini-prompt1-yelp,fig:gemini-prompt2-cnn,fig:gemini-prompt2-yelp,fig:gemini-prompt2-cnn-ifd,fig:gemini-prompt2-yelp-ifd}
+show the experiment results for the Claude 3 Opus large language model.
+
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-gemini-cnn_dailymail/prompt_1-gemini-cnn_dailymail.png}
+    \caption{Gemini on CNN Corpus\\Prompt Without Examples}
+    \label{fig:gemini-prompt1-cnn}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-gemini-yelp/prompt_1-gemini-yelp.png}
+    \caption{Gemini on Yelp Corpus\\Prompt Without Examples}
+    \label{fig:gemini-prompt1-yelp}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-gemini-cnn_dailymail/prompt_2-gemini-cnn_dailymail.png}
+    \caption{Gemini on CNN Corpus\\Prompt With Examples\\Extraneous Input}
+    \label{fig:gemini-prompt2-cnn}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-gemini-yelp/prompt_2-gemini-yelp.png}
+    \caption{Gemini on Yelp Corpus\\Prompt With Examples\\Extraneous Input}
+    \label{fig:gemini-prompt2-yelp}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-gemini-cnn_dailymail/prompt_2-gemini-cnn_dailymail.png}
+    \caption{Gemini on CNN Corpus\\Prompt With Examples\\Input from Corpus}
+    \label{fig:gemini-prompt2-cnn-ifd}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-gemini-yelp/prompt_2-gemini-yelp.png}
+    \caption{Gemini on Yelp Corpus\\Prompt With Examples\\Input from Corpus}
+    \label{fig:gemini-prompt2-yelp-ifd}
+\end{figure*}
+
+\subsection{Claude3}
+
+\Cref{fig:claude3-prompt1-cnn,fig:claude3-prompt1-yelp,fig:claude3-prompt2-cnn,fig:claude3-prompt2-yelp,fig:claude3-prompt2-cnn-ifd,fig:claude3-prompt2-yelp-ifd}
+show the experiment results for the Claude 3 Opus large language model.
+
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-claude3-cnn_dailymail/prompt_1-claude3-cnn_dailymail.png}
+    \caption{Claude3 on CNN Corpus\\Prompt Without Examples}
+    \label{fig:claude3-prompt1-cnn}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-claude3-yelp/prompt_1-claude3-yelp.png}
+    \caption{Claude3 on Yelp Corpus\\Prompt Without Examples}
+    \label{fig:claude3-prompt1-yelp}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-claude3-cnn_dailymail/prompt_2-claude3-cnn_dailymail.png}
+    \caption{Claude3 on CNN Corpus\\Prompt With Examples\\Extraneous Input}
+    \label{fig:claude3-prompt2-cnn}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-claude3-yelp/prompt_2-claude3-yelp.png}
+    \caption{Claude3 on Yelp Corpus\\Prompt With Examples\\Extraneous Input}
+    \label{fig:claude3-prompt2-yelp}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-claude3-cnn_dailymail/prompt_2-claude3-cnn_dailymail.png}
+    \caption{Claude3 on CNN Corpus\\Prompt With Examples\\Input from Corpus}
+    \label{fig:claude3-prompt2-cnn-ifd}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-claude3-yelp/prompt_2-claude3-yelp.png}
+    \caption{Claude3 on Yelp Corpus\\Prompt With Examples\\Input from Corpus}
+    \label{fig:claude3-prompt2-yelp-ifd}
+\end{figure*}
+
+\subsection{Llama3}
+
+\Cref{fig:llama3_70b-prompt1-cnn,fig:llama3_70b-prompt1-yelp,fig:llama3_70b-prompt2-cnn,fig:llama3_70b-prompt2-yelp,fig:llama3_70b-prompt2-cnn-ifd,fig:llama3_70b-prompt2-yelp-ifd}
+show the experiment results for the Claude 3 Opus large language model.
+
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-llama3_70b-cnn_dailymail/prompt_1-llama3_70b-cnn_dailymail.png}
+    \caption{Llama3 on CNN Corpus\\Prompt Without Examples}
+    \label{fig:llama3_70b-prompt1-cnn}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-llama3_70b-yelp/prompt_1-llama3_70b-yelp.png}
+    \caption{Llama3 on Yelp Corpus\\Prompt Without Examples}
+    \label{fig:llama3_70b-prompt1-yelp}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-llama3_70b-cnn_dailymail/prompt_2-llama3_70b-cnn_dailymail.png}
+    \caption{Llama3 on CNN Corpus\\Prompt With Examples\\Extraneous Input}
+    \label{fig:llama3_70b-prompt2-cnn}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-llama3_70b-yelp/prompt_2-llama3_70b-yelp.png}
+    \caption{Llama3 on Yelp Corpus\\Prompt With Examples\\Extraneous Input}
+    \label{fig:llama3_70b-prompt2-yelp}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-llama3_70b-cnn_dailymail/prompt_2-llama3_70b-cnn_dailymail.png}
+    \caption{Llama3 on CNN Corpus\\Prompt With Examples\\Input from Corpus}
+    \label{fig:llama3_70b-prompt2-cnn-ifd}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-llama3_70b-yelp/prompt_2-llama3_70b-yelp.png}
+    \caption{Llama3 on Yelp Corpus\\Prompt With Examples\\Input from Corpus}
+    \label{fig:llama3_70b-prompt2-yelp-ifd}
+\end{figure*}
+
+\subsection{Command R+}
+
+\Cref{fig:command_r-prompt1-cnn,fig:command_r-prompt1-yelp,fig:command_r-prompt2-cnn,fig:command_r-prompt2-yelp,fig:command_r-prompt2-cnn-ifd,fig:command_r-prompt2-yelp-ifd}
+show the experiment results for the Claude 3 Opus large language model.
+
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-command_r-cnn_dailymail/prompt_1-command_r-cnn_dailymail.png}
+    \caption{Command R on CNN Corpus\\Prompt Without Examples}
+    \label{fig:command_r-prompt1-cnn}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-command_r-yelp/prompt_1-command_r-yelp.png}
+    \caption{Command R on Yelp Corpus\\Prompt Without Examples}
+    \label{fig:command_r-prompt1-yelp}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-command_r-cnn_dailymail/prompt_2-command_r-cnn_dailymail.png}
+    \caption{Command R on CNN Corpus\\Prompt With Examples\\Extraneous Input}
+    \label{fig:command_r-prompt2-cnn}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-command_r-yelp/prompt_2-command_r-yelp.png}
+    \caption{Command R on Yelp Corpus\\Prompt With Examples\\Extraneous Input}
+    \label{fig:command_r-prompt2-yelp}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-command_r-cnn_dailymail/prompt_2-command_r-cnn_dailymail.png}
+    \caption{Command R on CNN Corpus\\Prompt With Examples\\Input from Corpus}
+    \label{fig:command_r-prompt2-cnn-ifd}
+\end{figure*}
+\begin{figure*}[ht]
+    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-command_r-yelp/prompt_2-command_r-yelp.png}
+    \caption{Command R on Yelp Corpus\\Prompt With Examples\\Input from Corpus}
+    \label{fig:command_r-prompt2-yelp-ifd}
+\end{figure*}
+
 \end{document}

### 2024-06-13 article/main.tex
final draft

@@
 
-\textbf{Large language models (LLMs)} have are a commonplace research topic
+\textbf{Large language models (LLMs)} have become a commonplace research topic
 today.
@@
 GPT~\cite{gpt-2018,gpt2-2019,gpt3-2020} constantly advance the state of the art
-on many natural language processing tasks, it has become natural to want to
-evaluate the performance of these generalist models on ever more specialized
-tasks.
-We know from multiple surveys~\cite{minaee2024llmsurvey,zhao2023survey} and
-benchmarks~\cite{papcode2024hellaswag,chiang2024chatbot} that large language
-models (especially the more popular ones) are good at following instructions.
+on many natural language processing tasks, it is interesting to evaluate them on
+more specialized tasks.
+Multiple surveys~\cite{minaee2024llmsurvey,zhao2023survey} and
+benchmarks~\cite{papcode2024hellaswag,chiang2024chatbot} show that large
+language models are good at following instructions.
 We can also intuit that training LLMs on diverse data (for instance the
@@
 stylistics~\cite{lugea2023stylistics}.
-The task of transfering text style is an old preocupation for the natural
-language processing and computational linguistics communities which has picked
-up interest again in the context of the advancements in the field of deep
-learning and particularly since the inception of the transformer architecture.
+The task of transferring text style has a certain maturity.
+The interest in this task is renewed by the advancements made with LLMs.
 
@@
 obtaining them are usually idempotent.
-This makes them very attractive because of the dual purpose that they can serve.
-On the one hand, they can be used to instruct the generation of text (``write
-no more than 10 sentences'') while on the other they can be used to analyse text
-and, therefore, verify generated text.
-
-Some HLFs statistically describe text in terms of its basic syntactical and
-morphological components.
-Knowing this we could follow Noam Chomsky's ideas~\cite{chomsky2002syntactic}
-which imply that the style of a text is influenced by its underlying syntactic
-structures.
+This makes them very attractive because of their dual potential purpose.
+On one hand, they can be used to instruct the generation of text (``write no
+more than 10 sentences'').
+On the other hand they can be used to analyse text.
+This ability allows verifying whether a text has the expected value for an HLF.
+
+Taking Noam Chomsky's ideas~\cite{chomsky2002syntactic} that syntax and meaning
+are different things, we intuit that we can define writing style only in terms
+of both syntax and meaning.
+We shy away from discussing meaning, but syntax and orderly constructs like it
+seem to be within our grasp.
+
 Traditionally, linguistic style was thought to be the result of morphological
@@
 
-For example, a high number of exclamations in a text is often a revealing
-signal of an inflamatory news writing style, of spam or of alert narration or
-dialogue.
-Similar observations have been made about other characteristics of written text
-that might be computed as HLFs~\cite{hovy1987generating,lugea2023stylistics}.
+Some have observed that many characteristics of written text can be computed as
+HLFs~\cite{hovy1987generating,lugea2023stylistics}.
 Combining multiple HLFs that are revelatory of linguistic style could provide us
@@
 
-These apparently disparate strands of thought lead us to the two research
-questions we'd like to answer in this paper.
+These are the strands of thought that have inspired us to pursue two research
+questions in the age of LLMs.
 
@@
 
-Linguistic style, then, is the sum of identifiable language choices manifest in
-a text made from the language system by the text
-producer~\cite{lugea2023stylistics}.
+Linguistic style is the sum of identifiable language choices manifest in a text
+made from the language system by the text producer~\cite{lugea2023stylistics}.
 Another way to think about it is that style is the form used for delivering
@@
 
-Finally, this paper documents a form of controllable text generation
+Our paper describes an experiment in controllable text generation
 (CTG)~\cite{zhang-ctg-2022}.
@@
 general.
-Compared to the existing work, our prompts are guided by HLFs.
-Furthermore, we use those same HLFs to evaluate the performance of the LLM.
+This paper differs from the existing CTG work with its use of HLFs for writing
+the text generation instructions.
+It additionally uses HLFs to measure the performance of LLMs in terms of their
+ability to generate text that conforms to some chosen target HLF values.
 
@@
 
-Our experiment consists of analysing the behavior of an LLM when asked to
-generate new text based on a given input text and some instructions derived from
-HLF statistics computed on a corpus containing multiple texts.
-The experiment relies on choosing LLMs to evaluate and finding datasets from
-which we can derive corpuses.
+Our experiment analyses LLM output when asked to generate new text based on
+given input text and instructions derived from HLFs.
+We compute statistics of HLFs on a corpus to derive the text generation
+instructions.
+
+To run the experiment we must select the LLMs to evaluate and finding datasets
+from which we can derive corpuses.
 Then there's the matter of selecting the HLFs for our experiment.
-Let us outline the experiment process before going into further detail.
-
-The first task is to establish a baseline.
-We do so by instructing the target LLM to generate text retaining the meaning of
-the input text without any further instruction related to style or the
-linguistic features of the generated text.
-We repeat the text generation step 10 times using the same prompt and the same
-input text.
-
-For each text generation, we compute the following statistics for each of our
-chosen HLFs:
-
+
+We think it's best to describe the experimental process before going over our
+selection for the experiment's parameters.
+
+\subsection{Process}\label{subsec:experiment-process}
+
+Assume we have a target LLM, a corpus containing texts, some examples extracted
+from the corpus, a selection of HLFs and input text both from within and from
+outside the corpus.
+This description of the experiment process assumes those parameters have stable
+values.
+Changing the parameters requires restarting the process.
+
+The first activity involves establishing baselines for each experiment scenario.
+
+In the first scenario the target LLM is instructed to generate text retaining
+the meaning of  the input text without any further instruction related to style
+or the linguistic features of the generated text.
+This step is repeated 10 times, computing each HLF in our selection for the
+generated text.
+The computed values make up the first scenario baselines for each HLF.
+
+The second scenario consists in generating text from the same input text, but
+with the aid of some examples from the corpus.
+In this scenario, we believe that the choice of input text affects the outcome.
+Therefore we propose two variants for this scenario: one with input text
+extraneous to the corpus and one with input text from the corpus.
+
+We add the examples to the initial LLM prompt and follow the same process like
+in the first scenario.
+We do this for each variant of the second scenario to obtain two more baselines
+for each HLF for the corpus.
+
+The number of examples is chosen so that we do not overrun the context window of
+the LLM.
+Some context windows are smaller than others, settling the number of examples at
+three.
+
+The next step requires computing statistics on the corpus for each HLF:
 \begin{itemize}
-    \item \textbf{min}: the minimum value of the HLF on the generated text;
-    \item \textbf{max}: the maximum value of the HLF on the generated text;
-    \item \textbf{avg}: the mean value of the HLF on the generated text;
+    \item \textbf{min}: the minimum value of the HLF on the corpus;
+    \item \textbf{max}: the maximum value of the HLF on the corpus;
+    \item \textbf{avg}: the mean value of the HLF on the corpus.
 \end{itemize}
 
-A second baseline is obtained by following the same process and same input text,
-but with a new prompt that builds upon the first one by adding the instruction
-to use example text from our corpus of choice.
-The examples are three texts chosen randomly once at the beginning of the
-experiment from the used corpus.
-We use the same example texts for all evaluated LLMs.
-
-We choose three texts to not overrun the context window of the LLM.
-This hints at the main drawback for our approach: it uses up the context window
-with system instructions.
-While this isn't an issue with all models, some only allow a context window of
-around 8000 tokens so we should be mindful.
-
-We construct the second baseline by conducting 10 trials using the same
-prompt and input text and otherwise in the same manner in which we created the
-first baseline.
-
-The next step is to devise HLF prompts derived from both baseline prompts.
-The HLF prompts contain additional instructions to generate text that would
-comply with certain HLF values.
-These HLF values are obtained by computing the statistics outlined above on
-the entirety of the target benchmark dataset.
-The HLF prompts are constructed from templates that take into account the HLF
-values.
-The prompt templates part of the experiment's source code.
-The instructions concerning all choseon HLFs are given concomitently within the
-same prompt.
-In a similar manner to how we generated the baselines, we run ten trials for
-each case (with or without example articles from the target benchmark dataset).
-
-We compare the first baseline with the results obtained by using the first
-modified HLF prompt to understand whether we can influence LLM text generation
-using HLFs.
-Our reasoning is that if the LLM were influenced by the HLF prompt, the output
-would show different HLF values compared to the baseline.
-
-We now turn our attention to the second baseline.
-Given that the LLM now also has example articles, it can freely immitate those
-articles when generating the text.
-If there is a consistent difference over the ten trials between generating text
-using the baseline or using the HLF prompt, this difference can only be a result
-of the LLM complying to HLF instructions.
-This difference is expressed by the change of value for each HLF between the
-baseline and HLF enhanced text generation.
-
-Because we provide the LLM with examples from the corpus, it is more likely to
-intuit the linguistic features of the corpus.
-To understand whether HLF instructions actually make a difference, we perform
-the experiment using the second type of prompt using input text both from within
-and from outside the corpus.
+Doing so allows writing an HLF prompt derived from each of our baseline prompts.
+The HLF prompts instruct the LLM to generate text whose HLFs comply with the
+previously computed statistics.
+The individual (per-HLF) instructions are all are part of the same HLF prompt.
+The prompts are constructed using Jinja templates which are located in the
+\texttt{\textlinkanon{https://github.com/koosie0507/llm-tst-consistency/tree/main/src/llm_tst_consistency/templates}{templates}}
+directory of the source code used to run the experiment.
+The instruction prompts corresponding to our two scenarios are
+`\texttt{prompt\_1}' and `\texttt{prompt\_2}', respectively.
+The rest of the templates in the directory are used to construct the individual
+HLF instructions.
+
+In a similar manner to how we generated the baselines, we run 10 trials for each
+scenario and variant.
+We compare each baseline with the results obtained by using the corresponding
+HLF prompt.
+
+The first scenario is indicative of whether LLMs acknowledge HLF instructions.
+If they do then the results obtained from HLF instructions should be
+significantly different from the corresponding baseline results.
+
+Both variants of the second scenario quantify how similar the HLFs of the texts
+generated by the LLM are to the corpus average \textit{because of} the HLF
+instructions.
+The rationale here is that the LLM can generate text very similar to text from
+the corpus by using the supplied examples from the corpus.
+That would leave the HLF instructions as the remaining possible cause for
+generating text with HLFs that are consistently different from the corresponding
+baseline.
 
@@
 
-For our experiment we will use state of the art large language models.
-We used the Chatbot Arena~\cite{chiang2024chatbot} leaderboard for constructing
-our selection.
-In the interest of covering more ground, we chose five top performers in the
-Chatbot Arena from different vendors.
-The final selection is available in Table~\ref{table-llm}. The identifier
+We only use state of the art large language models for the experiment.
+The Chatbot Arena~\cite{chiang2024chatbot} leaderboard is essential in our
+selection.
+To cover more ground we find top performers from different vendors interesting.
+The final selection is available in Table~\ref{table-llm}.
 
@@
 
-The ID of the LLM is used as the reference throughout figures and tables.
+The ID of the LLM is referenced throughout figures, tables and generally
+throughout the rest of the paper.
 
@@
 
-In this experiment we create specific instructions for each HLF.
+The instructions for each HLF are specifically crafted using natural language.
 The resulting instructions can be catalogued on the concrete-abstract spectrum.
 For example, `write 100 words' is a more concrete instruction than `write as if
-you were in junior highschool'~\cite{kincaid1975derivation}.
+you were in junior highschool'.
 We divide HLFs in five categories ranging from very concrete to highly abstract
-and choose 2 features from each category for a grand total of 10 HLFs.
+and choose 2 features from each category.
 Table~\ref{table-hlf} showcases our selection of features.
@@
 
-Each of these features is well documented~\cite{lftk-2023}, barring the need for
-an extended description here.
 The LFTK ID is used throughout tables and figures to refer to a specific HLF.
 
-\subsection{DataSet Selection}\label{ds-selection}
-
-The experiment design rests upon measuring HLFs on corpuses of texts and
+\subsection{Corpus Selection}\label{ds-selection}
+
+The experiment design rests upon measuring HLFs on a corpus of texts and
 computing statistics.
 LLMs also make use of examples extracted from a corpus to guide text generation.
-We obtain a corpus from a dataset.
-Our final selection of datasets is listed in Table~\ref{table-ds}.
+Our approach idempotently derives a corpus of texts from a well known dataset.
+Our selection of datasets is listed in Table~\ref{table-ds}.
 
@@
 users.
-The authors employ a casual writing style, very similar to what a average
-writing style of an average person would be like.
-The dataset as a whole is useful for many tasks.
-The corpus we construct based on this dataset contains the first 1000 reviews
-from its test split.
+The authors employ a casual writing style, very similar to the writing style of
+an average person.
+We construct a corpus containing the first 1000 reviews from the test split of
+this dataset.
 
 CNN/DailyMail contains stories and news with diverse styles.
-The style varies from one author to another, but also based on the publishing
-entity.
+The style varies from one author to another and it also follows the publisher's
+editorial style.
 The articles are more formal, longer and more complex in structure than Yelp
-reviews which makes this dataset a good alternative choice.
+reviews which makes this dataset a good alternative.
 The corpus that is based on this dataset contains the first 1000 CNN articles
@@
 This experiment presupposes instructing various large language models to reword
-an input text so that it conforms to a bunch of handcrafted linguistic features.
-The input text is provided via a user prompt.
-
-We provide two input texts that are external to our corpuses, one input text per
+an input text so that it conforms to some HLFs.
+
+There is an input text corresponding to each corpus that is external to that
 corpus.
-The input texts are located in the
+These texts are located in the
 \texttt{\textlinkanon{https://github.com/koosie0507/llm-tst-consistency/tree/main/data}{data}}
 folder of the experiment's source code.
-These user prompts are used when we're not using an input text from the
-corpus itself.
-
-For the CNN/DailyMail dataset, the user prompt from outside the dataset is an
-extract from the classic text \textit{The Life and Opinions of Tristram Shandy,
-    Gentleman}, by Laurence Sterne~\cite{sterne2003life}.
-This text is far from a news article in terms of both style and exhibited
-linguistic features from the CNN/DailyMail dataset contents.
-We see a text that is similarly divorced from the style and linguistic
-features of most Yelp reviews in Oscar Wilde's \textit{Sonnet to
-    Liberty}~\cite{wilde1909poems}.
-Having input texts that are vastly different from a dataset should inform an LLM
-to generate a baseline that does not conform to the linguistic features of the
-dataset.
-
-We use Jinja templates to compile the text generation instructions for the LLM.
-Both the model prompts and the HLF instructions are constructed using these
-templates.
-The Jinja templates are located in the
-\texttt{\textlinkanon{https://github.com/koosie0507/llm-tst-consistency/tree/main/data}{templates}}
-folder of the experiment's source code.
-
-There are two main text generation instruction templates: one without example
-content~(called `\texttt{prompt\_1}') and one with examples from the
-corpus~(`\texttt{prompt\_2}').
-Each of the two instruction templates allow the presence of HLF instructions.
+
+For the variant that uses an input text from the corpus, we draw the text
+randomly from the corpus once, at the start of the experiment.
+
+The input text that is external to the CNN article corpus is an extract from the
+classic \textit{The Life and Opinions of Tristram Shandy, Gentleman}, by
+Laurence Sterne~\cite{sterne2003life}.
+The text style is very dissimilar compared to news articles.
+Its linguistic features are also very unlike the average on the CNN article
+corpus.
+A text that is similarly divorced from the style and linguistic features of most
+Yelp reviews is Oscar Wilde's \textit{Sonnet toLiberty}~\cite{wilde1909poems}.
+
+Choosing input texts that are vastly different from their corresponding corpus
+is based in the assumption that LLMs would generate baseline results that do not
+conform to the HLF statistics on that corpus.
 
@@
 can compute on text.
-All our graphs show the HLF on the vertical axis, while the horizontal axis is
+Plot figures show the HLF on the vertical axis, while the horizontal axis is
 reserved for the trial number.
@@
 We sometimes refer to this average as the \textit{target} value because the LLM
-is primarily instructed to conform to this value.
+is instructed to generate text that primarily conforms to this value.
 
 Result tables highlight the difference between plots obtained using a
-baseline prompt versus the ones obtained using HLF-enhanced prompts.
-Firstly, we care about the difference between each plot (baseline and HLF) and
-the target (the corpus average).
-Secondly, we measure how different baselines are from HLF-enhanced plots.
+baseline prompt versus the ones obtained using HLF instruction prompts.
+Firstly, we measure how different baselines are from HLF-enhanced plots.
+Secondly, we compare the closeness of the baseline and HLF-enhanced plots to the
+target average.
+Table values where the text is \textbf{bold} mark significant differences
+between baseline and HLF instruction results.
+\underline{Underlined} values mark situations where the result of using HLF
+instructions is closer to the target than the baseline.
+\textit{Italic} values mark situations where the baseline is closer to the
+target.
+
+\subsection{Measurements}\label{subsec:measurements}
 
@@
 
-To determine which of the two results is closer to the target, we compute the
-area between each result and the target.
-Another way to measure the geometric closeness to the target is by computing the
-Euclidian norm of the vectorial difference between result and target.
+To determine which of two results is closer to the target, we compute the area
+between each result and the target.
+Another way in which we measure the closeness of the result to the target is by
+computing the Euclidian norm of the vectorial difference between the result and
+target.
+
+Note that each HLF can have a different range of values than other HLFs.
+When we compute the area and the Euclidian norm, we use the raw values of each
+HLF, without preprocessing them.
+Therefore the results for one HLF are not comparable with the results for
+another HLF.
+
+Remember that we don't quantify how different the baseline is from the
+HLF-enhanced result.
+Instead we calculate the relative closeness to a common target.
+Because we took this approach, comparing values for the same HLF, but different
+LLMs is also ill-advised.
+
 Let us denote with $A_b$ and $A_h$ the areas between baseline and target and HLF
@@
 Let us similarly denote with $N_b$ and $N_h$ the Euclidian norms.
-Finally, we define $Diff_A = A_b - A_h$ and $Diff_N = N_b - N_h$.
+We define $Diff_A = A_b - A_h$ and $Diff_N = N_b - N_h$.
 
@@
 
-To save space, we illustrate significant differences between the baseline and
-the HLF instruction results in \textbf{bold} typeface.
-The \underline{underlined} values represent situations when the HLF-enhanced
-prompt lead to results that are significantly closer to the corpus average than
-the baseline.
-\textit{Italicized} values represent situations when the baseline was
-significantly closer to the target than the HLF-enhanced result.
-
-Lastly, it should be noted that each HLF has an independent value range.
-The number of total words ranges from small numbers into the hundreds or more,
-while the corrected verb variation is always a small number.
-The area and the Euclidian norm are calculated using the raw values of each HLF.
-That makes the values in each table column incomparable to the values in another
-table column.
-Because we don't show the measure of the difference between the baseline and the
-result, but the relative closeness to a target based on that measure, comparing
-values from the same column is also ill-advised.
-
-In the next sections we only present the most relevant results of the
-experiment.
+In the next sections we present our perspective on the most relevant results of
+the experiment.
 
@@
 
-The results on the corpus containing CNN articles showcased in
-Table~\ref{table-prompt-1-cnn-dailymail} shows that most LLMs take into account
-instructions concerning the linguistic features of their output.
+The results on the CNN article corpus showcased in Table~\ref{table-prompt-1-cnn-dailymail}
+shows that LLMs take into account instructions concerning the HLFs of their
+output.
 
@@
 
-The second type of instructional prompt we feed to the LLM contains three
-examples from the corpus.
-The hypothesis is that this will influence the baseline result to be closer to
-the corpus average.
-
-In this case, it is interesting to see how the LLM generates text based on input
-text from both within the corpus that contains the example articles and from
-outside the corpus.
-
-In Table~\ref{table-prompt-2-cnn-dailymail} we see the results when an input from
-outside the corpus is used.
-The outlook is similar for what we saw in the first experiment with GPT
-being the least responsive to HLF instructions.
+In the second scenario of the experiment we try to quantify the impact of using
+HLF instructions on the output of the LLM.
+There are two variants of this scenario: input text from within or from outside
+the corpus.
+
+Table~\ref{table-prompt-2-cnn-dailymail} contains results for the first variant.
+The outlook is similar to the first experiment scenario.
 
@@
 
-GPT's results would be good if the baseline were very close to the target value.
+A possible explanation for having fewer significant differences in results might
+lie in the closeness to the target.
 As we can see in Figure~\ref{fig-p2-gpt-twords}, this is not always the case.
-We also notice that Command-R+(fig.~\ref{fig-p2-command-r-simpttr}) and Gemini
-(fig.~\ref{fig-p2-gemini-a-kup-pw}) both registered significantly negative
-results, in contrast to their performance in our first scenario.
+In~\cref{fig-p2-command-r-simpttr,fig-p2-gemini-a-kup-pw} we observe regressions
+in performance with significant differences between baseline and result.
 
@@
 
-The results obtained in our first scenario are largely confirmed by our second
-scenario.
-So now we turn to see how using a text from within the corpus affects the
-performance of the large language models.
-The hypothesis is that if the text is `familiar', both the baseline and the HLF
-instruction prompt will produce results that are much closer to the target
-value.
+Even so, the results obtained in our first scenario are largely confirmed in the
+first variant of our second scenario.
+
+We finally look at the second variant of the second scenario of our experiment.
+The hypothesis in this variant is that LLMs will generate text that exhibits
+HLFs much closer to the target corpus average if the input text is from that
+corpus.
 The expectation is to see less significant results.
@@
 expectations outright.
-The most surprising behaviour is exhibited in the context of more concrete HLFs
-such as the total number of words or the number of unique adjectives.
-Claude3 shows regressive behaviour for the number of unique adjectives in
-Figure~\ref{fig-p2-ifd-claude3-nuadj}, while Gemini constructs texts using a
-very different amount of words in Figure~\ref{fig-p2-ifd-gemini-twords}.
-Less surprisingly, Llama3 fails to generate text that conforms to the correct
-Kuperman age of acquisition per word in Figure~\ref{fig-p2-ifd-llama3-a-kup-pw}.
+The most surprising behaviour is exhibited for HLFs which are derived into more
+concrete generation instructions, such as the total number of words.
+This behaviour is highlighted in figure~\ref{fig-p2-ifd-gemini-twords}.
+Additionally, we notice regressive behaviour for some models compared to the
+previously explored results in figure~\ref{fig-p2-ifd-claude3-nuadj}.
+Finally, the abstractness of the text generation instructions for certain HLFs
+leads to the generated text not being compliant as exemplified in
+figure~\ref{fig-p2-ifd-llama3-a-kup-pw}.
 
@@
 The results in this last scenario show that whether LLMs take examples listed in
-the prompt into account or not with respect to conforming to certain linguistic
-features is debatable, providing them with specific instructions derived from
-HLFs makes a significant impact.
+the prompt into account is debatable when it comes to generating text with
+similar linguistic features as the examples'.
 
@@
 Certainly, not all LLMs are created equal in this regard.
-It is debatable whether this experiment is well designed because of the choice
-to derive target HLF values for the LLM from the corpus.
-This is evident in the results obtained for the Yelp dataset.
+
+The design of the experiment leaves room for improvement because of the choices
+regarding computing HLF statistics on a corpus, input text, the corpus
+itself or the instruction prompt.
+The importance of these factors is evidentiated by the difference in results
+obtained using the Yelp reviews corpus when compared to the results obtained
+on the CNN article corpus.
 While these results also show a significant influence when using HLFs in the
 prompt, the influence is mostly negative.
-This can be attributed to the choice of input text in relation to the corpus,
-the wording in the instruction prompt or to each model's specific training.
 
@@
 The results in this regards show us a much less conclusive picture.
-
-Firstly, we cannot assume that the LLM will follow HLF instructions the way we
-intend.
-Whether we obtain the expected outcome or not depends largely on the prompt,
-the chosen HLFs, the model, the target values for the HLFs, a.s.o
-We do, however, notice a few patterns.
-The more abstract the instruction, the less it seems to make a positve impact.
-More importantly, by inspecting the result data, a more abstract instruction
-seems more likely to produce a negative impact on the desired outcome.
-
-Secondly, the choice of input text in relation to the target HLF values seems
-very important.
-Choosing to reword a poem so that it has the same reading ease as a user review
-seems like a setup for poor results.
-This leads us to intuit that LLMs are clearly able to use HLF instructions to
-generate text differently, but there is a limit to their `creativity'.
-
-Finally, the choice of HLFs is very important.
-The text generated by an LLM is less likely to conform to HLFs that result in
-more abstract instructions.
-Furthermore, the combination of chosen HLFs is important.
+It is not apparent the LLM will follow HLF instructions as intended.
+In fact, using the same prompt on different data lead to wildly different
+outcomes.
+Whether this is because of circumstances such as the selection of HLFs, the LLM,
+the HLF target values, the input text, a.s.o remains to be determined.
+We can observe some idioms, though.
+
+The more abstract the instruction to comply to an HLF, the fewer the chances of
+a significant impact on the outcome.
+When the outcome is significant, it tends to be negative.
+This tells us that large language models are likely good at following
+instructions about HLFs when those instructions are more concrete.
+
+Then, the choice of input text in relation to the target HLF values seems very
+important.
+LLMs don't seem able yet to cover the vast gulf between the linguistic features
+of an Oscar Wilde poem and the average linguistic features of text written by
+Yelp reviewers.
+This leads us to propose that even if LLMs are able to generate text that
+complies to certain target HLFs, there is a limit to this ability.
+
+Finally, the overall selection of HLFs is important.
 Choosing HLFs that measure the uniqueness of a linguistic element (words,
@@
 age of acquisition and the adjective or verb variation.
-Predictably, the more concrete instructions take precedence over the more
-abstract ones.
 
@@
 
-We started out by trying to understand whether it is possible to conform the
-output of a large language model to exhibit specific linguistic features.
+We started out by trying to understand whether it is possible to generate text
+that exhibits certain linguistic features using a large language model.
 To our surprise, it turns out that instructing language models to write text
 that conforms to a selected assortment of linguistic features almost always
-yields significant changes in how the text is being generated.
-
-This doesn't mean that the results are always satisfying.
-What the model writes about and the chosen target values for the linguistic
-features are major factors in how successful the overall experience is.
-Unfortunately, at this stage, finding out good target values and good writing
-instructions is a question of trial and error.
-
-Other experiment designs might have been more opportune in examining the
-questions layed out in the introduction.
-For example, using set target values for HLFs (without deriving them from a
-corpus of text) and choosing input text where the HLFs vary greatly would have
-been an alternate design.
-Following through on those approaches is one of the possible expansions on the
-current work.
-
-Another way of expanding on the current work is finding the settings in which
-large language models are most responsive to linguistic features.
-The current experiment was inconclusive in this regard and a legitimate further
-question is whether settings in which text generated by language models
-consistently conforms to certain linguistic features actually exist.
-
-Lastly, a simpler question that is left unanswered relates to the number of
-linguistic features that were used.
-What if we would have measured models' behaviour for single linguistic features
-independently?
+leads to significantly different linguistic featues of generated text.
+
+The results are not always satisfying, though.
+At this stage, the more pragmatic answer seems to involve prompt engineering and
+a careful choice of language features.
+Providing examples does not seem to influence the model's ability to comply with
+the user's desires.
+
+Other experiments are required to better understand the capabilities of large
+language models with regard to handcrafted linguistic features.
+For example, testing the response of language models to features in isolation
+or not using a corpus or input text are different setups that could eliminate
+some of the unknowns left uncovered by the current experiment.
 

### 2024-06-14 article/main.tex
incorporate review

@@
 seem to be within our grasp.
-
 Traditionally, linguistic style was thought to be the result of morphological
@@
 level which contains basic HLFs such as the number of words in a sentence.
-
 Some have observed that many characteristics of written text can be computed as
@@
 instructions.
-
 To run the experiment we must select the LLMs to evaluate and finding datasets
@@
 Then there's the matter of selecting the HLFs for our experiment.
-
 We think it's best to describe the experimental process before going over our
@@
 In the first scenario the target LLM is instructed to generate text retaining
-the meaning of  the input text without any further instruction related to style
-or the linguistic features of the generated text.
+the meaning of the input text from outside the corpus without any further
+instruction related to style or the linguistic features of the generated text.
 This step is repeated 10 times, computing each HLF in our selection for the
 generated text.
-The computed values make up the first scenario baselines for each HLF.
-
-The second scenario consists in generating text from the same input text, but
-with the aid of some examples from the corpus.
-In this scenario, we believe that the choice of input text affects the outcome.
-Therefore we propose two variants for this scenario: one with input text
-extraneous to the corpus and one with input text from the corpus.
-
-We add the examples to the initial LLM prompt and follow the same process like
-in the first scenario.
-We do this for each variant of the second scenario to obtain two more baselines
-for each HLF for the corpus.
-
+The computed values represent the baseline HLF values for the first scenario.
+
+The second scenario consists in generating text based on input text, but with
+the aid of some examples from the corpus.
+The examples are appended to the instruction prompt used in the first scenario.
 The number of examples is chosen so that we do not overrun the context window of
@@
 three.
+Adding examples to the LLM's instruction prompt also makes the provenance of the
+input text more important.
+Therefore we experiment with two variants of the second scenario: one with input
+text extraneous to the corpus and one with input text from the corpus.
+The text from outside the corpus is the same text we used in the first scenario.
+The text from within the corpus is chosen randomly once at the start of the
+experiment.
+Following the process from the first scenario results in obtaining baseline HLF
+values for each variant of the second scenario.
+
+From here on we refer to `baseline HLF values' as `baseline'.
+We now have three baselines: one for the first scenario and one for each variant
+of the second scenario.
 
@@
 
-Doing so allows writing an HLF prompt derived from each of our baseline prompts.
-The HLF prompts instruct the LLM to generate text whose HLFs comply with the
+These statistics allow writing an HLF prompt derived from each of our baseline
+prompts.
+The HLF prompts instruct the LLM to generate text with HLFs that comply with the
 previously computed statistics.
@@
 HLF prompt.
+From here on we refer to the results obtained by using prompts with HLF
+instructions as `HLF-enhanced results', `HLF results' or even simply `results'.
 
@@
 
-\subsection{User Prompt and Text}\label{input-text}
-
-This experiment presupposes instructing various large language models to reword
-an input text so that it conforms to some HLFs.
-
-There is an input text corresponding to each corpus that is external to that
-corpus.
+\subsection{External Input Texts}\label{input-text}
+
+The experiment involves using input text from both outside the corpus and from
+within it.
+This section describes how we use text external to the corpus, having already
+described our usage of text from within the corpus.
+
+We use a different external input text for each corpus.
 These texts are located in the
@@
 
-For the variant that uses an input text from the corpus, we draw the text
-randomly from the corpus once, at the start of the experiment.
-
-The input text that is external to the CNN article corpus is an extract from the
-classic \textit{The Life and Opinions of Tristram Shandy, Gentleman}, by
-Laurence Sterne~\cite{sterne2003life}.
+The text external to the CNN article corpus is an extract from the classic
+\textit{The Life and Opinions of Tristram Shandy, Gentleman}, by Laurence
+Sterne~\cite{sterne2003life}.
 The text style is very dissimilar compared to news articles.
-Its linguistic features are also very unlike the average on the CNN article
-corpus.
+Its linguistic features are also very unlike the average values computed for the
+CNN article corpus.
 A text that is similarly divorced from the style and linguistic features of most
@@
 
-Result tables highlight the difference between plots obtained using a
+Result tables highlight the difference between the results obtained using a
 baseline prompt versus the ones obtained using HLF instruction prompts.
@@
 
-In the second scenario of the experiment we try to quantify the impact of using
-HLF instructions on the output of the LLM.
+We attempt quantifying the impact of using HLF instructions on the output of the
+LLM by eliminating as many factors as possible that would lead to a different
+text being generated when compared to a baseline.
+In the second scenario of our experiment we use examples to allow the LLM to
+generate better baselines, thus hopefully narrowing the gap between baseline
+and HLF-enhanced result.
+
 There are two variants of this scenario: input text from within or from outside
 the corpus.
-
-Table~\ref{table-prompt-2-cnn-dailymail} contains results for the first variant.
-The outlook is similar to the first experiment scenario.
+Table~\ref{table-prompt-2-cnn-dailymail} contains results for the first variant:
+input from the corpus.
+We notice these results are similar to those in the first experiment scenario.
 

### 2024-06-14 article/main.tex
make article compliant to submission guidelines

@@
 \begin{abstract}
-    Large language models are sought for solving an increasing number of
-    increasingly diverse tasks. Perhaps naturally one such task is text style
-    transfer: the task of changing the style of a text while preserving its
-    meaning. Besides the natural approach of fine-tuning a large language model,
-    a simpler and cheaper alternative should also be investigated: using
-    handcrafted language features to direct the large language model.
+    Can we control the writing style of large language models (LLMs) by specifying
+    desired linguistic features?
+    We address this question by investigating the impact of handcrafted linguistic
+    feature (HLF) instructions on LLM-generated text.
+    Our experiment evaluates various state-of-the-art LLMs using prompts
+    incorporating HLF statistics derived from corpora of CNN articles and Yelp
+    reviews.
+    We find that LLMs demonstrate sensitivity to these instructions, particularly
+    when tasked with conforming to concrete features like word count.
+    However, compliance with abstract features, such as lexical variation, proves
+    more challenging, often resulting in negative impacts on compliance.
+    Our findings highlight the potential and limitations of utilizing HLFs for
+    guiding LLM text generation and underscore the need for further research into
+    optimizing prompt design and feature selection.
 \end{abstract}
@@
 language~\cite{lftk-2023}.
-More often than not, these values are easy to compute while the methods used for
-obtaining them are usually idempotent.
-This makes them very attractive because of their dual potential purpose.
-On one hand, they can be used to instruct the generation of text (``write no
-more than 10 sentences'').
-On the other hand they can be used to analyse text.
-This ability allows verifying whether a text has the expected value for an HLF.
+These values are usually easy to compute with methods that are idempotent.
+This makes HLFs very attractive because of their dual potential.
+They can be used to instruct the generation of text (``write no more than 10
+sentences''), but they can also be used to analyse text.
+This allows us to verify if a text has the \textit{expected} value for an HLF.
 
 Taking Noam Chomsky's ideas~\cite{chomsky2002syntactic} that syntax and meaning
-are different things, we intuit that we can define writing style only in terms
-of both syntax and meaning.
+are different things, we intuit that writing style is defined in terms of both.
 We shy away from discussing meaning, but syntax and orderly constructs like it
 seem to be within our grasp.
-Traditionally, linguistic style was thought to be the result of morphological
-and syntactic arrangements~\cite{lugea2023stylistics}.
-More recent schools of thought on linguistic style find semantics and pragmatics
-to be inextricably linked to author style~\cite{verma2019lexical}, but argue
-that a great deal can still be inferred from lexicon, syntax and a `surface'
-level which contains basic HLFs such as the number of words in a sentence.
+Traditionally, linguistic style revolved around morphological and syntactic
+arrangements~\cite{lugea2023stylistics}.
+More recent descriptions find semantics and pragmatics to be inextricably linked
+to author style~\cite{verma2019lexical}, but argue that a great deal can still
+be inferred from lexicon, syntax and a `surface' level which contains basic HLFs
+such as the number of words in a sentence.
 Some have observed that many characteristics of written text can be computed as
@@
 transfer and authorship verification.
-Answering this question might point out cheaper, less intrusive and more user
+Answering this question might point to cheaper, less intrusive and more user
 friendly solutions than fine tuning large language models for these tasks.
@@
 
-There is abundent literature on the topic of text style transfer.
-Reviews and surveys~\cite{tst-review-2021,tst-survey-2022} were helpful for
-informing our selection of HLFs.
+Text style transfer reviews and surveys~\cite{tst-review-2021,tst-survey-2022}
+were helpful for informing our selection of HLFs.
 Studies on attaining fine-grained text style transfer using language models
-have served as inspiration for writing this paper~\cite{lyu-etal-2023-fine}.
-We should note though that we are not attempting to transfer text style, but
-merely to instruct an LLM to follow instructions based on HLFs.
+have served as inspiration~\cite{lyu-etal-2023-fine}.
 
@@
 meaning~\cite{tst_sigkdd_review_2022}.
-The ample literature on the subject seems to agree that style depends on context
-and author choice vis-a-vis a communication
-goal~\cite{mcdonald1985computational,hovy1987generating}.
+Opinions that style depends on context and author choice vis-a-vis a
+communication goal seem to converge~\cite{mcdonald1985computational,
+    hovy1987generating}.
 The author's choice is available at all linguistic levels, including the
@@
 formalist perspective~\cite{chomsky2002syntactic}.
-This work grounds our choice of handcrafted linguistic features related to
-linguistic style.
+Our selection of HLFs is informed by aspects of linguistic style.
 
-The idea of constructing style from fine-grained aspects is not new and there
-are excellent studies on the subject.
-The StylePTB authors explain how this dataset leverages lexical, syntactic,
-semantic and thematic aspects to allow verifying text style transfer more
-thoroughly than previous methods~\cite{lyu-etal-2021-styleptb}.
-Besides providing the understanding required to distinguish between semantic
-transfer and other forms of style transfer, the dataset itself is useful for
-evaluating HLFs that function at the sentence level.
+Constructing style from fine-grained aspects is not new.
+The StylePTB authors explain this in detail with respect to lexical,
+syntactic, semantic and thematic aspects~\cite{lyu-etal-2021-styleptb}.
+The dataset was useful for understanding HLFs that function at the sentence
+level and the interplay of HLFs.
 
-The body of work on HLFs is extensive.
-It is well referenced and synthesised by Lee and Lee~\cite{lftk-2023}.
+The body of work on HLFs is extensive and well referenced and synthesised by
+Lee and Lee~\cite{lftk-2023}.
 To our knowledge there is no work that connects HLFs with LLMs in the manner
@@
 
-In terms of controlling the style of the generated text, our approach differs
-from others in at least two significant ways: simplicity and focus.
-We are constructing well crafted user prompts for an LLM instead of diving into
-more complex artificial intelligence solutions.
-Our focus is on controlling text generation using HLFs and measuring the change
-relative to a baseline obtained without using HLFs in the generation process.
-We are not discussing the effectiveness of style transfer, author verification
-or other tasks (yet).
+Our approach stands out for its simplicity and focus.
+We propose engineering instruction prompts for LLMs, a straightforward strategy
+distinct from complex alternatives.
+With discussions on style transfer or author verification beyond this paper's
+scope, we present a method controlling LLM text generation through the use of
+HLFs.
 
-Our paper describes an experiment in controllable text generation
-(CTG)~\cite{zhang-ctg-2022}.
+The presented experiment is one of controllable text
+generation (CTG)~\cite{zhang-ctg-2022}.
 Hightened recent interest in CTG and especially in benchmarking LLMs in the
@@
 paper.
-Such research alleviates the burden of proof regarding the level to which
-LLMs respond well to diversified instructions for controlled text generation in
-general.
-This paper differs from the existing CTG work with its use of HLFs for writing
+This lessens the burden of demonstrating how effectively LLMs respond to varied
+instructions for general controlled text generation.
+This paper differs from the existing CTG work in its use of HLFs for writing
 the text generation instructions.
@@
 instructions.
-To run the experiment we must select the LLMs to evaluate and finding datasets
-from which we can derive corpuses.
-Then there's the matter of selecting the HLFs for our experiment.
-We think it's best to describe the experimental process before going over our
-selection for the experiment's parameters.
+To run the experiment we must select LLMs to evaluate and find datasets from
+which we can derive corpuses.
+Then we must also select the HLFs for our experiment.
+Before going over our selection for the experiment's parameters, let us go
+through the experimental process.
 
@@
 The first activity involves establishing baselines for each experiment scenario.
-
 In the first scenario the target LLM is instructed to generate text retaining
@@
 The next step requires computing statistics on the corpus for each HLF:
-\begin{itemize}
-    \item \textbf{min}: the minimum value of the HLF on the corpus;
-    \item \textbf{max}: the maximum value of the HLF on the corpus;
-    \item \textbf{avg}: the mean value of the HLF on the corpus.
-\end{itemize}
-
+\textbf{min}(the minimum value), \textbf{max}(the maximum value) and
+\textbf{avg}(the mean).
 These statistics allow writing an HLF prompt derived from each of our baseline
 prompts.
+
 The HLF prompts instruct the LLM to generate text with HLFs that comply with the
@@
 The individual (per-HLF) instructions are all are part of the same HLF prompt.
-The prompts are constructed using Jinja templates which are located in the
-\texttt{\textlinkanon{https://github.com/koosie0507/llm-tst-consistency/tree/main/src/llm_tst_consistency/templates}{templates}}
+The prompts are constructed using the Jinja2~\cite{jinja2} templates which are
+located in the \texttt{\textlinkanon{https://github.com/koosie0507/llm-tst-consistency/tree/main/src/llm_tst_consistency/templates}{templates}}
 directory of the source code used to run the experiment.
@@
 
-The first scenario is indicative of whether LLMs acknowledge HLF instructions.
-If they do then the results obtained from HLF instructions should be
-significantly different from the corresponding baseline results.
+The first scenario indicates whether LLMs acknowledge HLF instructions.
+If they do, the HLF results should be significantly different from their
+baseline.
 
@@
 instructions.
-The rationale here is that the LLM can generate text very similar to text from
-the corpus by using the supplied examples from the corpus.
-That would leave the HLF instructions as the remaining possible cause for
-generating text with HLFs that are consistently different from the corresponding
-baseline.
+We assume that the LLM can generate text similar to text from the corpus when
+instructed it with examples from the corpus.
+The HLF instructions would remain as the possible cause for obtaining different
+results from baseline.
+The ID of the LLM is referenced throughout figures, tables and generally
+throughout the rest of the paper.
 
@@
 
-The ID of the LLM is referenced throughout figures, tables and generally
-throughout the rest of the paper.
-
 \subsection{HLF Selection}\label{hlf-selection}
@@
 instructions to conform to HLFs.
-LFTK authors~\cite{lftk-2023} categorize HLFs by domain and family and make
-available information on the suitability of each HLF for given tasks such as
-readability assessment or fake news detection.
+LFTK~\cite{lftk-2023} is a framework built on top of spaCy~\cite{spacy}
+that provides implementations for computing many HLFs.
+We use these two libraries to perform our experiment.
+They categorize HLFs by domain and family and make available information on the
+suitability of each HLF for given tasks such as readability assessment or fake
+news detection.
 Besides the domain and family, we also choose features based on their influence
@@
 Table~\ref{table-hlf} showcases our selection of features.
+The LFTK ID is used throughout tables and figures to refer to a specific HLF.
 
@@
 \footnotetext{as described in the figure in~\cite{lugea2023stylistics}, chapter 1, page 7}
-
-The LFTK ID is used throughout tables and figures to refer to a specific HLF.
 
@@
 Yelp reviews is Oscar Wilde's \textit{Sonnet toLiberty}~\cite{wilde1909poems}.
+Both texts are in the public domain and were obtained from Project
+Gutenberg~\cite{gutenberg}.
 
@@
 can compute on text.
-Plot figures show the HLF on the vertical axis, while the horizontal axis is
-reserved for the trial number.
-HLF values for the text generated using baseline prompts are plotted as a
-continuous grey line, while the values obtained by using HLF enhanced prompts
-are plotted using a continuous grenat line.
-The minimum and maximum HLF on the benchmark dataset are plotted using dotted
-black lines.
+Plot figures show the HLF on the vertical axis and the trial number on the horizontal
+axis.
+We plot the baseline using a continuous grey line and the HLF result using a
+continuous grenat line.
+The minimum and maximum HLF on the corpus are plotted using dotted black lines.
 A continuous black line designates the average HLF on the corpus.
-We sometimes refer to this average as the \textit{target} value because the LLM
-is instructed to generate text that primarily conforms to this value.
+We refer to the corpus average as the \textit{target} because the LLM is
+instructed to generate text that primarily conforms to this value.
 
-Result tables highlight the difference between the results obtained using a
-baseline prompt versus the ones obtained using HLF instruction prompts.
-Firstly, we measure how different baselines are from HLF-enhanced plots.
-Secondly, we compare the closeness of the baseline and HLF-enhanced plots to the
-target average.
-Table values where the text is \textbf{bold} mark significant differences
-between baseline and HLF instruction results.
-\underline{Underlined} values mark situations where the result of using HLF
-instructions is closer to the target than the baseline.
-\textit{Italic} values mark situations where the baseline is closer to the
-target.
+Result tables present an overview over an experiment scenario variant.
+Firstly, they display significant differences between baselines and HLF results.
+Secondly, they reveal the relative closeness of the baseline and HLF results to
+the target average.
+\textbf{Bold} text represents significant differences between the baseline and
+HLF results.
+\underline{Underlined} text marks HLF results that are closer to the target than
+the baseline.
+\textit{Italic} text marks HLF results that are further away than the baseline
+to the target.
 
@@
 to determine whether the baseline results are different from the HLF-enhanced
-results.
+results, as implemented in Scipy~\cite{2020SciPy-NMeth}.
 We consider that the results are different for a resulting $p\_value \le 0.05$.
@@
 Note that each HLF can have a different range of values than other HLFs.
-When we compute the area and the Euclidian norm, we use the raw values of each
-HLF, without preprocessing them.
+When we compute the area and the Euclidian norm using Numpy~\cite{harris2020array},
+we use the raw values of each HLF, without preprocessing them.
 Therefore the results for one HLF are not comparable with the results for
 another HLF.
-
-Remember that we don't quantify how different the baseline is from the
-HLF-enhanced result.
-Instead we calculate the relative closeness to a common target.
-Because we took this approach, comparing values for the same HLF, but different
-LLMs is also ill-advised.
+Since we don't quantify how different the baseline is from the result and we
+compute the relative closeness to a common target, comparing values for the same
+HLF but different LLMs is also ill-advised.
 
@@
 in performance with significant differences between baseline and result.
+Even so, the results obtained in our first scenario are largely confirmed in the
+first variant of our second scenario.
 
@@
 
-Even so, the results obtained in our first scenario are largely confirmed in the
-first variant of our second scenario.
-
 We finally look at the second variant of the second scenario of our experiment.
@@
 The expectation is to see less significant results.
+The results shown in Table~\ref{table-prompt-2-ifd-cnn-dailymail} contradict our
+expectations outright.
+The most surprising behaviour is exhibited for HLFs which are derived into more
+concrete generation instructions, such as the total number of words.
+This behaviour is highlighted in figure~\ref{fig-p2-ifd-gemini-twords}.
+Additionally, we notice regressive behaviour for some models compared to the
+previously explored results in figure~\ref{fig-p2-ifd-claude3-nuadj}.
+Finally, the abstractness of the text generation instructions for certain HLFs
+leads to the generated text not being compliant as exemplified in
+figure~\ref{fig-p2-ifd-llama3-a-kup-pw}.
 
@@
 \end{table*}
-
-The results shown in Table~\ref{table-prompt-2-ifd-cnn-dailymail} contradict our
-expectations outright.
-The most surprising behaviour is exhibited for HLFs which are derived into more
-concrete generation instructions, such as the total number of words.
-This behaviour is highlighted in figure~\ref{fig-p2-ifd-gemini-twords}.
-Additionally, we notice regressive behaviour for some models compared to the
-previously explored results in figure~\ref{fig-p2-ifd-claude3-nuadj}.
-Finally, the abstractness of the text generation instructions for certain HLFs
-leads to the generated text not being compliant as exemplified in
-figure~\ref{fig-p2-ifd-llama3-a-kup-pw}.
 
@@
 
-In terms of our first question --- whether LLMs are able to follow instructions
-concerning handcrafted linguistic features --- the response is in our opinion
-affirmative.
+In terms of our first question --- whether HLF instructions impact an LLM's
+writing, the response is affirmative.
 Certainly, not all LLMs are created equal in this regard.
+The experiment design involves choices that rely on the presence of a corpus and
+the usage of a fixed external input text.
+This is relevant as evidentiated by the difference in results obtained using the
+Yelp reviews corpus when compared to the results obtained on the CNN article
+corpus.
+While similarly significant, the HLF results on Yelp reviews are mostly worse
+than their corresponding baselines failing to confirm the positive impact of HLF
+instructions.
+The importance of the choice of the external text is highlighted by the better
+results obtained using an input text from the corpus.
 
-The design of the experiment leaves room for improvement because of the choices
-regarding computing HLF statistics on a corpus, input text, the corpus
-itself or the instruction prompt.
-The importance of these factors is evidentiated by the difference in results
-obtained using the Yelp reviews corpus when compared to the results obtained
-on the CNN article corpus.
-While these results also show a significant influence when using HLFs in the
-prompt, the influence is mostly negative.
+This leads us to our second question: how much do HLF instructions impact an
+LLM's writing?
+The metrics we compute dismiss the idea that HLF instructions have a consistent
+desired impact on the generated text.
+There isn't conclusive evidence that improving the LLM's odds of producing a
+better baseline results in a smaller difference between baseline and result
+relative to the target.
+More importantly, there is evidence that HLF instructions might have an effect
+that is opposite to the desired effect when comparing the results on the Yelp
+reviews dataset.
 
-This leads us to our second question: how much do handcrafted linguistic feature
-instructions affect the way LLMs write text.
-The results in this regards show us a much less conclusive picture.
-It is not apparent the LLM will follow HLF instructions as intended.
-In fact, using the same prompt on different data lead to wildly different
-outcomes.
-Whether this is because of circumstances such as the selection of HLFs, the LLM,
-the HLF target values, the input text, a.s.o remains to be determined.
-We can observe some idioms, though.
+Highly abstract HLF instructions are unlikely to have the desired effect.
+Furthermore, the impact tends to be adverse.
+This suggests that LLMs are particularly adept at following instructions
+regarding HLFs when those instructions are more concrete.
 
-The more abstract the instruction to comply to an HLF, the fewer the chances of
-a significant impact on the outcome.
-When the outcome is significant, it tends to be negative.
-This tells us that large language models are likely good at following
-instructions about HLFs when those instructions are more concrete.
+The choice of input text in relation to the target HLF values also seems
+important.
+LLMs don't yet seem able to cover the gulf between Oscar Wilde poems and the
+average Yelp review in terms of linguistic features.
+We surmise that even if LLMs are able to generate text that complies to certain
+target HLFs, there is a limit to this ability.
 
-Then, the choice of input text in relation to the target HLF values seems very
-important.
-LLMs don't seem able yet to cover the vast gulf between the linguistic features
-of an Oscar Wilde poem and the average linguistic features of text written by
-Yelp reviewers.
-This leads us to propose that even if LLMs are able to generate text that
-complies to certain target HLFs, there is a limit to this ability.
+Finally, selecting HLFs that are in tension with one another will affect the
+LLMs ability to comply.
+This is highlighted in the experiment with age of acquisition versus the
+adjective variation.
 
-Finally, the overall selection of HLFs is important.
-Choosing HLFs that measure the uniqueness of a linguistic element (words,
-structures, a.s.o) limits the LLMs ability to conform to other HLFs (for example
-total number of sentences or reading ease).
-This is highlighted in the experiment specifically in the tension between the
-age of acquisition and the adjective or verb variation.
-
-\section{Conclusions and Future Work}
-
-We started out by trying to understand whether it is possible to generate text
-that exhibits certain linguistic features using a large language model.
-To our surprise, it turns out that instructing language models to write text
-that conforms to a selected assortment of linguistic features almost always
-leads to significantly different linguistic featues of generated text.
-
-The results are not always satisfying, though.
-At this stage, the more pragmatic answer seems to involve prompt engineering and
-a careful choice of language features.
-Providing examples does not seem to influence the model's ability to comply with
-the user's desires.
+\section{Limitations and Future Work}
 
@@
 language models with regard to handcrafted linguistic features.
-For example, testing the response of language models to features in isolation
-or not using a corpus or input text are different setups that could eliminate
-some of the unknowns left uncovered by the current experiment.
+The selection of language models, the reliance on a corpus for computing target
+linguistic features, the fixed choice of input text are all limitations of the
+current approach.
+Using a selection of HLFs and not individually analysing the behaviour of the
+LLM under the influence of each individual HLF is another limitation of our
+approach.
+
+\section{Conclusions}
+
+We began by trying to understand whether it is possible to generate text that
+exhibits certain linguistic features by instructing a large language model.
+To our surprise, it turns out that instructing large language models to write
+text that conforms to a selected assortment of linguistic features leads to
+significantly different linguistic featues of the generated text.
+This is especially true for concrete instructions.
+
+However, the outcomes are not always desirable.
+From a pragmatic standpoint, prompt engineering and a careful choice of language
+features and input text seem like a good idea to alleviate results.
+Providing examples in the input prompt does not seem as beneficial.
 
 \bibliography{references}
-
+\newpage
 \appendix

### 2024-06-14 article/main.tex
final corrections

@@
 language models are good at following instructions.
-We can also intuit that training LLMs on diverse data (for instance the
+We intuit that training LLMs on diverse data (for instance the
 Pile~\cite{gao2020pile}) uniquely qualifies them to produce text in a wide
@@
 
-Taking Noam Chomsky's ideas~\cite{chomsky2002syntactic} that syntax and meaning
-are different things, we intuit that writing style is defined in terms of both.
-We shy away from discussing meaning, but syntax and orderly constructs like it
+Taking Noam Chomsky's ideas that syntax and meaning are different
+things~\cite{chomsky2002syntactic}, we intuit that writing style is defined in
+terms of both.
+We shy away from discussing meaning, but syntax and similarly orderly constructs
 seem to be within our grasp.
@@
 meaning~\cite{tst_sigkdd_review_2022}.
-Opinions that style depends on context and author choice vis-a-vis a
-communication goal seem to converge~\cite{mcdonald1985computational,
-    hovy1987generating}.
+Opinions seem to converge that style depends on context and author choice
+vis-a-vis a communication goal~\cite{mcdonald1985computational,hovy1987generating}.
 The author's choice is available at all linguistic levels, including the
@@
 
-Our approach stands out for its simplicity and focus.
+Our approach stands out through its simplicity and focus.
 We propose engineering instruction prompts for LLMs, a straightforward strategy
@@
 We use these two libraries to perform our experiment.
-They categorize HLFs by domain and family and make available information on the
+LFTK categorizes HLFs by domain and family and make available information on the
 suitability of each HLF for given tasks such as readability assessment or fake
@@
 Let us denote with $A_b$ and $A_h$ the areas between baseline and target and HLF
-instruction result and target, respectively.
+result and target, respectively.
 Let us similarly denote with $N_b$ and $N_h$ the Euclidian norms.
@@
 
-In the next sections we present our perspective on the most relevant results of
-the experiment.
+Next, we present our perspective on the most relevant results of the experiment.
 
@@
 Furthermore, the impact tends to be adverse.
-This suggests that LLMs are particularly adept at following instructions
-regarding HLFs when those instructions are more concrete.
+This suggests that LLMs are more adept at following instructions regarding HLFs
+when those instructions are more concrete.
 
@@
 Other experiments are required to better understand the capabilities of large
-language models with regard to handcrafted linguistic features.
+language models with regard to HLFs.
 The selection of language models, the reliance on a corpus for computing target
@@
 current approach.
+Making different experiment design choices in all these respects might shed more
+light on this matter.
 Using a selection of HLFs and not individually analysing the behaviour of the
-LLM under the influence of each individual HLF is another limitation of our
-approach.
+LLM under the influence of each individual HLF is another limiting choice that
+invites to future work on the alternative.
 
@@
 exhibits certain linguistic features by instructing a large language model.
-To our surprise, it turns out that instructing large language models to write
-text that conforms to a selected assortment of linguistic features leads to
-significantly different linguistic featues of the generated text.
+To our surprise, it turns out state of the art large language models are
+receptive to instructions regarding the linguistic features of the output.
 This is especially true for concrete instructions.
 
-However, the outcomes are not always desirable.
+However, the outcomes are not always good.
 From a pragmatic standpoint, prompt engineering and a careful choice of language
@@
 
+\newpage
+
 \bibliography{references}
+
 \newpage
+
 \appendix

### 2024-06-14 article/main.tex
more tweaks.

@@
             \toprule
-            ID          & Name                    & Parameters  & Context Size & Elo Score & License Type        \\ \toprule
+            ID          & Name                    & Parameters  & Context Size & Elo Score & License             \\ \toprule
             gpt         & OpenAI GPT-4o           & undisclosed & 128000       & 1287      & Proprietary         \\
@@
     \resizebox{\textwidth}{!}{%
-        \begin{tabular}{@{}lllll@{}}
+        \begin{tabular}{@{}llllll@{}}
             \toprule
-            Name            & Task                     & Number of Samples & Content Type & Reference              \\ \toprule
-            Yelp Reviews    & Sentiment Classification & 50000             & Reviews      & \cite{yelp2015neurips} \\
-            (test split)    & Text Style Transfer      &                   &              &                        \\
-                            & Text Classification      &                   &              &                        \\ \midrule
-            CNN / DailyMail & Summarization            & 11490             & News         & \cite{cnndm2015}       \\
-            (test split)    & Question Answering       &                   & Stories      & \cite{cnndm2017}       \\
-                            & Text Generation          &                   &              &                        \\ \midrule
+            Name            & Task                     & Number of Samples & Content Type & Reference              & License           \\ \toprule
+            Yelp Reviews    & Sentiment Classification & 50000             & Reviews      &~\cite{yelp2015neurips} & Yelp Dataset      \\
+            (test split)    & Text Style Transfer      &                   &              &                        & License Agreement \\
+                            & Text Classification      &                   &              &                        &                   \\ \midrule
+            CNN / DailyMail & Summarization            & 11490             & News         &~\cite{cnndm2015}       & Apache 2.0        \\
+            (test split)    & Question Answering       &                   & Stories      &~\cite{cnndm2017}       &                   \\
+                            & Text Generation          &                   &              &                        &                   \\ \bottomrule
         \end{tabular}%

### 2024-07-29 article/main.tex
fix main file
use the review option
fixup some wording

@@
+\documentclass[11pt]{article}
 \pdfoutput=1
 
-\documentclass[11pt]{article}
-\usepackage[review]{acl}
+\usepackage[review]{coling}
+
+\usepackage{times}
+\usepackage{latexsym}
 
@@
 Fine tuning LLMs also incurs some loss of generality~\cite{yang2024unveiling},
-going directly against the purpose of having a generic language model.
+going directly against the purpose of generic language models.
 
@@
 formalist perspective~\cite{chomsky2002syntactic}.
-Our selection of HLFs is informed by aspects of linguistic style.
+Our selection of HLFs is informed by the various aspects of linguistic style.
 
@@
 writing, the response is affirmative.
-Certainly, not all LLMs are created equal in this regard.
+Not all LLMs are equally susceptible to HLF instructions, though.
 The experiment design involves choices that rely on the presence of a corpus and
@@
 relative to the target.
-More importantly, there is evidence that HLF instructions might have an effect
-that is opposite to the desired effect when comparing the results on the Yelp
-reviews dataset.
+More importantly, there is evidence that HLF instructions have an adverse effect
+when comparing the results on the Yelp reviews dataset.
 
@@
 From a pragmatic standpoint, prompt engineering and a careful choice of language
-features and input text seem like a good idea to alleviate results.
+features and input text seem like the way to obtain desirable results.
 Providing examples in the input prompt does not seem as beneficial.
 
-\newpage
-
 \bibliography{references}
-
-\newpage
 

### 2024-07-29 article/main.tex
fix chktex errors

@@
 sentences''), but they can also be used to analyse text.
-This allows us to verify if a text has the \textit{expected} value for an HLF.
+This allows us to verify if a text has the \textit{expected} value for an HLF.\@
 
@@
 The number of examples is chosen so that we do not overrun the context window of
-the LLM.
+the LLM.\@
 Some context windows are smaller than others, settling the number of examples at
@@
 
-The next step requires computing statistics on the corpus for each HLF:
-\textbf{min}(the minimum value), \textbf{max}(the maximum value) and
-\textbf{avg}(the mean).
+The next step requires computing statistics on the corpus for each HLF:\@
+\textbf{min}\~(the minimum value), \textbf{max}\~(the maximum value) and
+\textbf{avg}\~(the mean).
 These statistics allow writing an HLF prompt derived from each of our baseline
@@
             claude3     & Anthropic Claude 3 Opus & undisclosed & 200000       & 1248      & Proprietary         \\
-            llama3\_70b & Meta Llama 3-70B        & 70 Billion  & 8000         & 1208      & Llama 3 Community   \\
+            llama3\_70b & Meta Llama 3-{}70B      & 70 Billion  & 8000         & 1208      & Llama 3 Community   \\
             command\_r  & Cohere Command R+       & 104 Billion & 128000       & 1189      & CC-BY-NC 4.0        \\
@@
 Table~\ref{table-hlf} showcases our selection of features.
-The LFTK ID is used throughout tables and figures to refer to a specific HLF.
+The LFTK ID is used throughout tables and figures to refer to a specific HLF.\@
 
@@
             \toprule
-            LFTK ID         & Name                              & Family           & Domain           & Abstraction     & Style Level\footnotemark \\ \toprule
+            LFTK ID         & Name                              & Family           & Domain           & Abstraction     & Style Level\footnotemark\\ \toprule
             t\_word         & total words                       & wordsent         & surface          & very concrete   & lexical                 \\
@@
 
-We use the Huggingface ecosystem~\cite{lhoest-etal-2021-datasets} to load the
-datasets.
+The Huggingface ecosystem~\cite{lhoest-etal-2021-datasets} makes the selected
+datasets easily accessible.
 Yelp reviews are loaded from \texttt{Yelp/yelp\_review\_full}, while CNN stories
@@
 
-We use the two-sampled Kolmogorov-Smirnov~\cite{kolmogorov1933,smirnov1939} test
-to determine whether the baseline results are different from the HLF-enhanced
-results, as implemented in Scipy~\cite{2020SciPy-NMeth}.
+Determining whether the baseline results differ from the HLF-enhanced results is
+a classic use case for the Kolmogorov-Smirnov~\cite{kolmogorov1933,smirnov1939}
+test, as implemented in Scipy~\cite{2020SciPy-NMeth}.
 We consider that the results are different for a resulting $p\_value \le 0.05$.
@@
 Therefore the results for one HLF are not comparable with the results for
-another HLF.
+another HLF.\@
 Since we don't quantify how different the baseline is from the result and we
@@
         \includegraphics[width=\linewidth]{plots/prompt_1/prompt_1-command_r-cnn_dailymail/prompt_1-command_r-cnn_dailymail_t_word.png}
-        \caption{Command-R+\\Total Words}
-        \label{fig-command-r-t-word}
+        \caption{Command-R+\\Total Words}\label{fig-command-r-t-word}
     \end{minipage}
@@
         \includegraphics[width=\linewidth]{plots/prompt_1/prompt_1-gpt-cnn_dailymail/prompt_1-gpt-cnn_dailymail_fkgl.png}
-        \caption{GPT\\Flesch-Kincaid Grade Level}
-        \label{fig-gpt-fkgl}
+        \caption{GPT\\Flesch-Kincaid Grade Level}\label{fig-gpt-fkgl}
     \end{minipage}
@@
         \includegraphics[width=\linewidth]{plots/prompt_1/prompt_1-claude3-cnn_dailymail/prompt_1-claude3-cnn_dailymail_a_verb_pw.png}
-        \caption[center]{Claude3\\Average Verbs per Word}
-        \label{fig-llama-simp-ttr}
+        \caption[center]{Claude3\\Average Verbs per Word}\label{fig-llama-simp-ttr}
     \end{minipage}
@@
     }
-    \caption{prompt 2 - CNN Results\\Input Outside Corpus}
-    \label{table-prompt-2-cnn-dailymail}
+    \caption{prompt 2 - CNN Results\\Input Outside Corpus}\label{table-prompt-2-cnn-dailymail} % chktex-file 8
 \end{table*}
@@
         \includegraphics[width=\linewidth]{plots/prompt_2/prompt_2-gpt-cnn_dailymail/prompt_2-gpt-cnn_dailymail_t_word.png}
-        \caption{GPT\\Total Words}
-        \label{fig-p2-gpt-twords}
+        \caption{GPT\\Total Words}\label{fig-p2-gpt-twords}
     \end{minipage}
@@
         \includegraphics[width=\linewidth]{plots/prompt_2/prompt_2-command_r-cnn_dailymail/prompt_2-command_r-cnn_dailymail_simp_ttr.png}
-        \caption{Command-R+\\Simple TTR}
-        \label{fig-p2-command-r-simpttr}
+        \caption{Command-R+\\Simple TTR}\label{fig-p2-command-r-simpttr}
     \end{minipage}
@@
         \includegraphics[width=\linewidth]{plots/prompt_2/prompt_2-gemini-cnn_dailymail/prompt_2-gemini-cnn_dailymail_a_kup_pw.png}
-        \caption[center]{Gemini\\Kuperman Age of Aquisition}
-        \label{fig-p2-gemini-a-kup-pw}
+        \caption[center]{Gemini\\Kuperman Age of Aquisition}\label{fig-p2-gemini-a-kup-pw}
     \end{minipage}
@@
     }
-    \caption{CNN Results\\Input Inside Corpus}
-    \label{table-prompt-2-ifd-cnn-dailymail}
+    \caption{CNN Results\\Input Inside Corpus}\label{table-prompt-2-ifd-cnn-dailymail}
 \end{table*}
@@
         \includegraphics[width=\linewidth]{plots/prompt_2_ifd/prompt_2-claude3-cnn_dailymail/prompt_2-claude3-cnn_dailymail_n_uadj.png}
-        \caption{Claude3\\Number of Unique Adjectives}
-        \label{fig-p2-ifd-claude3-nuadj}
+        \caption{Claude3\\Number of Unique Adjectives}\label{fig-p2-ifd-claude3-nuadj}
     \end{minipage}
@@
         \includegraphics[width=\linewidth]{plots/prompt_2_ifd/prompt_2-gemini-cnn_dailymail/prompt_2-gemini-cnn_dailymail_t_word.png}
-        \caption{Gemini\\Total Words}
-        \label{fig-p2-ifd-gemini-twords}
+        \caption{Gemini\\Total Words}\label{fig-p2-ifd-gemini-twords}
     \end{minipage}
@@
         \includegraphics[width=\linewidth]{plots/prompt_2_ifd/prompt_2-llama3_70b-cnn_dailymail/prompt_2-llama3_70b-cnn_dailymail_a_kup_pw.png}
-        \caption[center]{Llama3\\Kuperman Age of Aquisition}
-        \label{fig-p2-ifd-llama3-a-kup-pw}
+        \caption[center]{Llama3\\Kuperman Age of Aquisition}\label{fig-p2-ifd-llama3-a-kup-pw}
     \end{minipage}
@@
 
-The results in this last scenario show that whether LLMs take examples listed in
-the prompt into account is debatable when it comes to generating text with
-similar linguistic features as the examples'.
+The outcomes from the latest experiment are ambiguous. The LLMs do not seem to
+utilize the examples given in the prompt to their advantage when they generate
+text that should mirror the linguistic features of those examples.
 
@@
     }
-    \caption{Yelp Results - Prompt Without Examples}
-    \label{table-prompt-1-yelp}
+    \caption{Yelp Results - Prompt Without Examples}\label{table-prompt-1-yelp} % chktex-file 8
 \end{table*}
@@
     }
-    \caption{Yelp Results - Prompt With Examples\\Input External To Corpus}
-    \label{table-prompt-2-yelp}
+    \caption{Yelp Results - Prompt With Examples\\Input External To Corpus}\label{table-prompt-2-yelp}
 \end{table*}
@@
     }
-    \caption{Yelp Results - Prompt With Examples\\Input From Corpus}
-    \label{table-prompt-2-ifd-yelp}
+    \caption{Yelp Results - Prompt With Examples\\Input From Corpus}\label{table-prompt-2-ifd-yelp}
 \end{table*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-gpt-cnn_dailymail/prompt_1-gpt-cnn_dailymail.png}
-    \caption{GPT on CNN Corpus\\Prompt Without Examples}
-    \label{fig:gpt-prompt1-cnn}
+    \caption{GPT on CNN Corpus\\Prompt Without Examples}\label{fig:gpt-prompt1-cnn}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-gpt-yelp/prompt_1-gpt-yelp.png}
-    \caption{GPT on Yelp Corpus\\Prompt Without Examples}
-    \label{fig:gpt-prompt1-yelp}
+    \caption{GPT on Yelp Corpus\\Prompt Without Examples}\label{fig:gpt-prompt1-yelp}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-gpt-cnn_dailymail/prompt_2-gpt-cnn_dailymail.png}
-    \caption{GPT on CNN Corpus\\Prompt With Examples\\Extraneous Input}
-    \label{fig:gpt-prompt2-cnn}
+    \caption{GPT on CNN Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:gpt-prompt2-cnn}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-gpt-yelp/prompt_2-gpt-yelp.png}
-    \caption{GPT on Yelp Corpus\\Prompt With Examples\\Extraneous Input}
-    \label{fig:gpt-prompt2-yelp}
+    \caption{GPT on Yelp Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:gpt-prompt2-yelp}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-gpt-cnn_dailymail/prompt_2-gpt-cnn_dailymail.png}
-    \caption{GPT on CNN Corpus\\Prompt With Examples\\Input from Corpus}
-    \label{fig:gpt-prompt2-cnn-ifd}
+    \caption{GPT on CNN Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:gpt-prompt2-cnn-ifd}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-gpt-yelp/prompt_2-gpt-yelp.png}
-    \caption{GPT on Yelp Corpus\\Prompt With Examples\\Input from Corpus}
-    \label{fig:gpt-prompt2-yelp-ifd}
+    \caption{GPT on Yelp Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:gpt-prompt2-yelp-ifd}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-gemini-cnn_dailymail/prompt_1-gemini-cnn_dailymail.png}
-    \caption{Gemini on CNN Corpus\\Prompt Without Examples}
-    \label{fig:gemini-prompt1-cnn}
+    \caption{Gemini on CNN Corpus\\Prompt Without Examples}\label{fig:gemini-prompt1-cnn}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-gemini-yelp/prompt_1-gemini-yelp.png}
-    \caption{Gemini on Yelp Corpus\\Prompt Without Examples}
-    \label{fig:gemini-prompt1-yelp}
+    \caption{Gemini on Yelp Corpus\\Prompt Without Examples}\label{fig:gemini-prompt1-yelp}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-gemini-cnn_dailymail/prompt_2-gemini-cnn_dailymail.png}
-    \caption{Gemini on CNN Corpus\\Prompt With Examples\\Extraneous Input}
-    \label{fig:gemini-prompt2-cnn}
+    \caption{Gemini on CNN Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:gemini-prompt2-cnn}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-gemini-yelp/prompt_2-gemini-yelp.png}
-    \caption{Gemini on Yelp Corpus\\Prompt With Examples\\Extraneous Input}
-    \label{fig:gemini-prompt2-yelp}
+    \caption{Gemini on Yelp Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:gemini-prompt2-yelp}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-gemini-cnn_dailymail/prompt_2-gemini-cnn_dailymail.png}
-    \caption{Gemini on CNN Corpus\\Prompt With Examples\\Input from Corpus}
-    \label{fig:gemini-prompt2-cnn-ifd}
+    \caption{Gemini on CNN Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:gemini-prompt2-cnn-ifd}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-gemini-yelp/prompt_2-gemini-yelp.png}
-    \caption{Gemini on Yelp Corpus\\Prompt With Examples\\Input from Corpus}
-    \label{fig:gemini-prompt2-yelp-ifd}
+    \caption{Gemini on Yelp Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:gemini-prompt2-yelp-ifd}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-claude3-cnn_dailymail/prompt_1-claude3-cnn_dailymail.png}
-    \caption{Claude3 on CNN Corpus\\Prompt Without Examples}
-    \label{fig:claude3-prompt1-cnn}
+    \caption{Claude3 on CNN Corpus\\Prompt Without Examples}\label{fig:claude3-prompt1-cnn}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-claude3-yelp/prompt_1-claude3-yelp.png}
-    \caption{Claude3 on Yelp Corpus\\Prompt Without Examples}
-    \label{fig:claude3-prompt1-yelp}
+    \caption{Claude3 on Yelp Corpus\\Prompt Without Examples}\label{fig:claude3-prompt1-yelp}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-claude3-cnn_dailymail/prompt_2-claude3-cnn_dailymail.png}
-    \caption{Claude3 on CNN Corpus\\Prompt With Examples\\Extraneous Input}
-    \label{fig:claude3-prompt2-cnn}
+    \caption{Claude3 on CNN Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:claude3-prompt2-cnn}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-claude3-yelp/prompt_2-claude3-yelp.png}
-    \caption{Claude3 on Yelp Corpus\\Prompt With Examples\\Extraneous Input}
-    \label{fig:claude3-prompt2-yelp}
+    \caption{Claude3 on Yelp Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:claude3-prompt2-yelp}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-claude3-cnn_dailymail/prompt_2-claude3-cnn_dailymail.png}
-    \caption{Claude3 on CNN Corpus\\Prompt With Examples\\Input from Corpus}
-    \label{fig:claude3-prompt2-cnn-ifd}
+    \caption{Claude3 on CNN Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:claude3-prompt2-cnn-ifd}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-claude3-yelp/prompt_2-claude3-yelp.png}
-    \caption{Claude3 on Yelp Corpus\\Prompt With Examples\\Input from Corpus}
-    \label{fig:claude3-prompt2-yelp-ifd}
+    \caption{Claude3 on Yelp Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:claude3-prompt2-yelp-ifd}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-llama3_70b-cnn_dailymail/prompt_1-llama3_70b-cnn_dailymail.png}
-    \caption{Llama3 on CNN Corpus\\Prompt Without Examples}
-    \label{fig:llama3_70b-prompt1-cnn}
+    \caption{Llama3 on CNN Corpus\\Prompt Without Examples}\label{fig:llama3_70b-prompt1-cnn}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-llama3_70b-yelp/prompt_1-llama3_70b-yelp.png}
-    \caption{Llama3 on Yelp Corpus\\Prompt Without Examples}
-    \label{fig:llama3_70b-prompt1-yelp}
+    \caption{Llama3 on Yelp Corpus\\Prompt Without Examples}\label{fig:llama3_70b-prompt1-yelp}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-llama3_70b-cnn_dailymail/prompt_2-llama3_70b-cnn_dailymail.png}
-    \caption{Llama3 on CNN Corpus\\Prompt With Examples\\Extraneous Input}
-    \label{fig:llama3_70b-prompt2-cnn}
+    \caption{Llama3 on CNN Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:llama3_70b-prompt2-cnn}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-llama3_70b-yelp/prompt_2-llama3_70b-yelp.png}
-    \caption{Llama3 on Yelp Corpus\\Prompt With Examples\\Extraneous Input}
-    \label{fig:llama3_70b-prompt2-yelp}
+    \caption{Llama3 on Yelp Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:llama3_70b-prompt2-yelp}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-llama3_70b-cnn_dailymail/prompt_2-llama3_70b-cnn_dailymail.png}
-    \caption{Llama3 on CNN Corpus\\Prompt With Examples\\Input from Corpus}
-    \label{fig:llama3_70b-prompt2-cnn-ifd}
+    \caption{Llama3 on CNN Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:llama3_70b-prompt2-cnn-ifd}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-llama3_70b-yelp/prompt_2-llama3_70b-yelp.png}
-    \caption{Llama3 on Yelp Corpus\\Prompt With Examples\\Input from Corpus}
-    \label{fig:llama3_70b-prompt2-yelp-ifd}
+    \caption{Llama3 on Yelp Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:llama3_70b-prompt2-yelp-ifd}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-command_r-cnn_dailymail/prompt_1-command_r-cnn_dailymail.png}
-    \caption{Command R on CNN Corpus\\Prompt Without Examples}
-    \label{fig:command_r-prompt1-cnn}
+    \caption{Command R on CNN Corpus\\Prompt Without Examples}\label{fig:command_r-prompt1-cnn}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-command_r-yelp/prompt_1-command_r-yelp.png}
-    \caption{Command R on Yelp Corpus\\Prompt Without Examples}
-    \label{fig:command_r-prompt1-yelp}
+    \caption{Command R on Yelp Corpus\\Prompt Without Examples}\label{fig:command_r-prompt1-yelp}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-command_r-cnn_dailymail/prompt_2-command_r-cnn_dailymail.png}
-    \caption{Command R on CNN Corpus\\Prompt With Examples\\Extraneous Input}
-    \label{fig:command_r-prompt2-cnn}
+    \caption{Command R on CNN Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:command_r-prompt2-cnn}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-command_r-yelp/prompt_2-command_r-yelp.png}
-    \caption{Command R on Yelp Corpus\\Prompt With Examples\\Extraneous Input}
-    \label{fig:command_r-prompt2-yelp}
+    \caption{Command R on Yelp Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:command_r-prompt2-yelp}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-command_r-cnn_dailymail/prompt_2-command_r-cnn_dailymail.png}
-    \caption{Command R on CNN Corpus\\Prompt With Examples\\Input from Corpus}
-    \label{fig:command_r-prompt2-cnn-ifd}
+    \caption{Command R on CNN Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:command_r-prompt2-cnn-ifd}
 \end{figure*}
@@
     \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-command_r-yelp/prompt_2-command_r-yelp.png}
-    \caption{Command R on Yelp Corpus\\Prompt With Examples\\Input from Corpus}
-    \label{fig:command_r-prompt2-yelp-ifd}
+    \caption{Command R on Yelp Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:command_r-prompt2-yelp-ifd}
 \end{figure*}

### 2024-07-30 article/main.tex
place text after figures to ensure their availability in text links
fix figure layout
fix typos. rephrase unreadable paragraphs

@@
 The next step requires computing statistics on the corpus for each HLF:\@
-\textbf{min}\~(the minimum value), \textbf{max}\~(the maximum value) and
-\textbf{avg}\~(the mean).
+\textbf{min}~(the minimum value), \textbf{max}~(the maximum value) and
+\textbf{avg}~(the mean).
 These statistics allow writing an HLF prompt derived from each of our baseline
@@
 previously computed statistics.
-The individual (per-HLF) instructions are all are part of the same HLF prompt.
+The individual (per-HLF) instructions are all part of the same HLF prompt.
 The prompts are constructed using the Jinja2~\cite{jinja2} templates which are
@@
 instructions.
-We assume that the LLM can generate text similar to text from the corpus when
-instructed it with examples from the corpus.
-The HLF instructions would remain as the possible cause for obtaining different
-results from baseline.
-The ID of the LLM is referenced throughout figures, tables and generally
-throughout the rest of the paper.
+Working under the assumption that an LLM can generate text that is similar to an
+example from the corpus, the presence of HLF instructions in the prompt would
+remain among the few possible explanations were we to observe deviations from
+the baseline.
 
@@
 
-\begin{table*}[ht]
+\begin{table*}[ht!]
     \setlength\tabcolsep{6pt}
@@
 \end{table*}
+
+Throughout the remainder of this paper, we will reference individual LLMs by the
+ID assigned to each one in Table~\ref{table-llm}.
 
@@
 A text that is similarly divorced from the style and linguistic features of most
-Yelp reviews is Oscar Wilde's \textit{Sonnet toLiberty}~\cite{wilde1909poems}.
+Yelp reviews is Oscar Wilde's \textit{Sonnet to Liberty}~\cite{wilde1909poems}.
 Both texts are in the public domain and were obtained from Project
@@
 Determining whether the baseline results differ from the HLF-enhanced results is
-a classic use case for the Kolmogorov-Smirnov~\cite{kolmogorov1933,smirnov1939}
-test, as implemented in Scipy~\cite{2020SciPy-NMeth}.
+a classic use case for the Kolmogorov-Smirnov
+test~\cite{kolmogorov1933,smirnov1939}, as implemented in the
+Scipy library~\cite{2020SciPy-NMeth}.
 We consider that the results are different for a resulting $p\_value \le 0.05$.
@@
 
-\begin{table*}[htp]
+\begin{table*}[ht!]
     \centering
@@
             \toprule
-            claude3 $Diff\_N$     & \textbf{\underline{623.16}}  & \textbf{\underline{24.74}} & \textbf{\underline{51.18}}  & \textbf{\underline{26.68}} & 0.06                      & -0.0                      & \textbf{\underline{0.46}} & \textbf{\underline{1.4}}   & 1.44                    & \textbf{\textit{-1.66}}   \\ \midrule
-            claude3 $Diff\_A$     & \textbf{\underline{1774.0}}  & \textbf{\underline{74.0}}  & \textbf{\underline{158.84}} & \textbf{\underline{81.2}}  & 0.2                       & -0.01                     & \textbf{\underline{1.88}} & \textbf{\underline{4.49}}  & 5.01                    & \textbf{\textit{-4.09}}   \\ \midrule
-            gemini $Diff\_N$      & \textbf{\underline{985.86}}  & \textbf{\underline{34.52}} & \textbf{\underline{100.36}} & \textbf{\underline{53.39}} & \textbf{\underline{0.52}} & 0.05                      & \textbf{\underline{2.62}} & \textbf{\underline{4.68}}  & -4.17                   & \textbf{\textit{-2.43}}   \\ \midrule
-            gemini $Diff\_A$      & \textbf{\underline{2990.5}}  & \textbf{\underline{106.5}} & \textbf{\underline{320.34}} & \textbf{\underline{159.1}} & \textbf{\underline{1.5}}  & 0.09                      & \textbf{\underline{7.24}} & \textbf{\underline{13.15}} & -12.65                  & \textbf{\textit{-7.13}}   \\ \midrule
-            gpt $Diff\_N$         & \textbf{\underline{125.87}}  & 1.46                       & 10.95                       & 10.77                      & 0.02                      & -0.01                     & 0.71                      & 0.81                       & \textbf{\textit{-1.72}} & 0.43                      \\ \midrule
-            gpt $Diff\_A$         & \textbf{\underline{327.0}}   & 3.0                        & 29.0                        & 27.0                       & 0.08                      & -0.01                     & 1.82                      & 2.24                       & \textbf{\textit{-5.72}} & 0.44                      \\ \midrule
-            command\_r $Diff\_N$  & \textbf{\underline{1024.78}} & \textbf{\underline{39.67}} & \textbf{\underline{81.58}}  & \textbf{\underline{41.46}} & \textbf{\underline{0.46}} & \textbf{\underline{0.03}} & \textbf{\underline{2.67}} & \textbf{\underline{3.78}}  & 1.4                     & \textbf{\underline{0.14}} \\ \midrule
-            command\_r $Diff\_A$  & \textbf{\underline{3132.5}}  & \textbf{\underline{120.5}} & \textbf{\underline{246.84}} & \textbf{\underline{132.2}} & \textbf{\underline{1.36}} & \textbf{\underline{0.06}} & \textbf{\underline{7.88}} & \textbf{\underline{11.21}} & 3.9                     & \textbf{\underline{1.03}} \\ \midrule
-            llama3\_70b $Diff\_N$ & \textbf{\underline{459.45}}  & \textbf{\underline{17.78}} & \textbf{\underline{36.5}}   & \textbf{\underline{19.75}} & \textbf{\underline{0.12}} & \textbf{\underline{0.03}} & \textbf{\underline{1.43}} & \textbf{\underline{2.18}}  & 1.5                     & \textbf{\textit{-0.06}}   \\ \midrule
-            llama3\_70b $Diff\_A$ & \textbf{\underline{1363.5}}  & \textbf{\underline{54.0}}  & \textbf{\underline{106.5}}  & \textbf{\underline{60.5}}  & \textbf{\underline{0.36}} & \textbf{\underline{0.09}} & \textbf{\underline{4.33}} & \textbf{\underline{6.38}}  & 7.46                    & \textbf{\textit{-0.11}}   \\ \bottomrule
+            claude3 $Diff_N$     & \textbf{\underline{623.16}}  & \textbf{\underline{24.74}} & \textbf{\underline{51.18}}  & \textbf{\underline{26.68}} & 0.06                      & -0.0                      & \textbf{\underline{0.46}} & \textbf{\underline{1.4}}   & 1.44                    & \textbf{\textit{-1.66}}   \\ \midrule
+            claude3 $Diff_A$     & \textbf{\underline{1774.0}}  & \textbf{\underline{74.0}}  & \textbf{\underline{158.84}} & \textbf{\underline{81.2}}  & 0.2                       & -0.01                     & \textbf{\underline{1.88}} & \textbf{\underline{4.49}}  & 5.01                    & \textbf{\textit{-4.09}}   \\ \midrule
+            gemini $Diff_N$      & \textbf{\underline{985.86}}  & \textbf{\underline{34.52}} & \textbf{\underline{100.36}} & \textbf{\underline{53.39}} & \textbf{\underline{0.52}} & 0.05                      & \textbf{\underline{2.62}} & \textbf{\underline{4.68}}  & -4.17                   & \textbf{\textit{-2.43}}   \\ \midrule
+            gemini $Diff_A$      & \textbf{\underline{2990.5}}  & \textbf{\underline{106.5}} & \textbf{\underline{320.34}} & \textbf{\underline{159.1}} & \textbf{\underline{1.5}}  & 0.09                      & \textbf{\underline{7.24}} & \textbf{\underline{13.15}} & -12.65                  & \textbf{\textit{-7.13}}   \\ \midrule
+            gpt $Diff_N$         & \textbf{\underline{125.87}}  & 1.46                       & 10.95                       & 10.77                      & 0.02                      & -0.01                     & 0.71                      & 0.81                       & \textbf{\textit{-1.72}} & 0.43                      \\ \midrule
+            gpt $Diff_A$         & \textbf{\underline{327.0}}   & 3.0                        & 29.0                        & 27.0                       & 0.08                      & -0.01                     & 1.82                      & 2.24                       & \textbf{\textit{-5.72}} & 0.44                      \\ \midrule
+            command\_r $Diff_N$  & \textbf{\underline{1024.78}} & \textbf{\underline{39.67}} & \textbf{\underline{81.58}}  & \textbf{\underline{41.46}} & \textbf{\underline{0.46}} & \textbf{\underline{0.03}} & \textbf{\underline{2.67}} & \textbf{\underline{3.78}}  & 1.4                     & \textbf{\underline{0.14}} \\ \midrule
+            command\_r $Diff_A$  & \textbf{\underline{3132.5}}  & \textbf{\underline{120.5}} & \textbf{\underline{246.84}} & \textbf{\underline{132.2}} & \textbf{\underline{1.36}} & \textbf{\underline{0.06}} & \textbf{\underline{7.88}} & \textbf{\underline{11.21}} & 3.9                     & \textbf{\underline{1.03}} \\ \midrule
+            llama3\_70b $Diff_N$ & \textbf{\underline{459.45}}  & \textbf{\underline{17.78}} & \textbf{\underline{36.5}}   & \textbf{\underline{19.75}} & \textbf{\underline{0.12}} & \textbf{\underline{0.03}} & \textbf{\underline{1.43}} & \textbf{\underline{2.18}}  & 1.5                     & \textbf{\textit{-0.06}}   \\ \midrule
+            llama3\_70b $Diff_A$ & \textbf{\underline{1363.5}}  & \textbf{\underline{54.0}}  & \textbf{\underline{106.5}}  & \textbf{\underline{60.5}}  & \textbf{\underline{0.36}} & \textbf{\underline{0.09}} & \textbf{\underline{4.33}} & \textbf{\underline{6.38}}  & 7.46                    & \textbf{\textit{-0.11}}   \\ \bottomrule
         \end{tabular}%
@@
 
-\begin{figure*}[htp]
+\begin{figure*}[ht!]
     \centering
@@
 
-\begin{table*}[hpb]
+\begin{table*}[ht!]
     \centering
@@
             \toprule
-            claude3 $Diff\_N$     & \textbf{\underline{718.49}}  & \textbf{\underline{24.13}} & \textbf{\underline{44.43}}  & \textbf{\underline{23.43}} & \textbf{\underline{0.4}}  & 0.03                      & 0.94                      & \textbf{\underline{0.77}} & 1.2                        & 0.07                      \\ \midrule
-            claude3 $Diff\_A$     & \textbf{\underline{2345.54}} & \textbf{\underline{90.18}} & \textbf{\underline{148.36}} & \textbf{\underline{63.6}}  & \textbf{\underline{1.18}} & 0.07                      & 1.36                      & \textbf{\underline{2.55}} & 3.33                       & 0.52                      \\ \midrule
-            gemini $Diff\_N$      & \textbf{\underline{785.6}}   & \textbf{\underline{25.8}}  & \textbf{\underline{60.5}}   & \textbf{\underline{41.6}}  & \textbf{\underline{0.41}} & \textbf{\underline{0.05}} & \textbf{\underline{1.95}} & \textbf{\underline{2.67}} & -4.62                      & \textbf{\textit{-1.94}}   \\ \midrule
-            gemini $Diff\_A$      & \textbf{\underline{2512.0}}  & \textbf{\underline{81.5}}  & \textbf{\underline{210.34}} & \textbf{\underline{125.7}} & \textbf{\underline{1.25}} & \textbf{\underline{0.09}} & \textbf{\underline{5.84}} & \textbf{\underline{8.05}} & -9.46                      & \textbf{\textit{-5.82}}   \\ \midrule
-            gpt $Diff\_N$         & 47.8                         & 5.18                       & 6.13                        & -0.13                      & 0.02                      & 0.0                       & 0.05                      & 0.3                       & 3.06                       & -0.27                     \\ \midrule
-            gpt $Diff\_A$         & 77.0                         & 12.5                       & 13.0                        & -4.0                       & 0.04                      & 0.0                       & -0.17                     & 0.77                      & 8.45                       & -0.16                     \\ \midrule
-            command\_r $Diff\_N$  & \textbf{\underline{263.51}}  & 5.35                       & \textbf{\underline{25.08}}  & \textbf{\underline{17.37}} & \textbf{\textit{-0.09}}   & -0.01                     & \textbf{\underline{1.74}} & \textbf{\underline{1.42}} & \textbf{\underline{10.46}} & \textbf{\underline{1.56}} \\ \midrule
-            command\_r $Diff\_A$  & \textbf{\underline{896.0}}   & 20.5                       & \textbf{\underline{86.0}}   & \textbf{\underline{60.6}}  & \textbf{\textit{-0.21}}   & -0.03                     & \textbf{\underline{5.45}} & \textbf{\underline{4.95}} & \textbf{\underline{31.29}} & \textbf{\underline{5.67}} \\ \midrule
-            llama3\_70b $Diff\_N$ & \textbf{\underline{376.0}}   & \textbf{\underline{15.63}} & \textbf{\underline{38.63}}  & 15.74                      & \textbf{\underline{0.24}} & \textbf{\underline{0.06}} & 1.69                      & \textbf{\underline{2.91}} & 0.31                       & -0.3                      \\ \midrule
-            llama3\_70b $Diff\_A$ & \textbf{\underline{1088.5}}  & \textbf{\underline{45.0}}  & \textbf{\underline{113.0}}  & 41.0                       & \textbf{\underline{0.65}} & \textbf{\underline{0.18}} & 3.53                      & \textbf{\underline{8.56}} & 0.27                       & -0.68                     \\ \bottomrule
+            claude3 $Diff_N$     & \textbf{\underline{718.49}}  & \textbf{\underline{24.13}} & \textbf{\underline{44.43}}  & \textbf{\underline{23.43}} & \textbf{\underline{0.4}}  & 0.03                      & 0.94                      & \textbf{\underline{0.77}} & 1.2                        & 0.07                      \\ \midrule
+            claude3 $Diff_A$     & \textbf{\underline{2345.54}} & \textbf{\underline{90.18}} & \textbf{\underline{148.36}} & \textbf{\underline{63.6}}  & \textbf{\underline{1.18}} & 0.07                      & 1.36                      & \textbf{\underline{2.55}} & 3.33                       & 0.52                      \\ \midrule
+            gemini $Diff_N$      & \textbf{\underline{785.6}}   & \textbf{\underline{25.8}}  & \textbf{\underline{60.5}}   & \textbf{\underline{41.6}}  & \textbf{\underline{0.41}} & \textbf{\underline{0.05}} & \textbf{\underline{1.95}} & \textbf{\underline{2.67}} & -4.62                      & \textbf{\textit{-1.94}}   \\ \midrule
+            gemini $Diff_A$      & \textbf{\underline{2512.0}}  & \textbf{\underline{81.5}}  & \textbf{\underline{210.34}} & \textbf{\underline{125.7}} & \textbf{\underline{1.25}} & \textbf{\underline{0.09}} & \textbf{\underline{5.84}} & \textbf{\underline{8.05}} & -9.46                      & \textbf{\textit{-5.82}}   \\ \midrule
+            gpt $Diff_N$         & 47.8                         & 5.18                       & 6.13                        & -0.13                      & 0.02                      & 0.0                       & 0.05                      & 0.3                       & 3.06                       & -0.27                     \\ \midrule
+            gpt $Diff_A$         & 77.0                         & 12.5                       & 13.0                        & -4.0                       & 0.04                      & 0.0                       & -0.17                     & 0.77                      & 8.45                       & -0.16                     \\ \midrule
+            command\_r $Diff_N$  & \textbf{\underline{263.51}}  & 5.35                       & \textbf{\underline{25.08}}  & \textbf{\underline{17.37}} & \textbf{\textit{-0.09}}   & -0.01                     & \textbf{\underline{1.74}} & \textbf{\underline{1.42}} & \textbf{\underline{10.46}} & \textbf{\underline{1.56}} \\ \midrule
+            command\_r $Diff_A$  & \textbf{\underline{896.0}}   & 20.5                       & \textbf{\underline{86.0}}   & \textbf{\underline{60.6}}  & \textbf{\textit{-0.21}}   & -0.03                     & \textbf{\underline{5.45}} & \textbf{\underline{4.95}} & \textbf{\underline{31.29}} & \textbf{\underline{5.67}} \\ \midrule
+            llama3\_70b $Diff_N$ & \textbf{\underline{376.0}}   & \textbf{\underline{15.63}} & \textbf{\underline{38.63}}  & 15.74                      & \textbf{\underline{0.24}} & \textbf{\underline{0.06}} & 1.69                      & \textbf{\underline{2.91}} & 0.31                       & -0.3                      \\ \midrule
+            llama3\_70b $Diff_A$ & \textbf{\underline{1088.5}}  & \textbf{\underline{45.0}}  & \textbf{\underline{113.0}}  & 41.0                       & \textbf{\underline{0.65}} & \textbf{\underline{0.18}} & 3.53                      & \textbf{\underline{8.56}} & 0.27                       & -0.68                     \\ \bottomrule
         \end{tabular}%
@@
 
-\begin{figure*}[hpb]
+\begin{figure*}[ht!]
     \centering
@@
 We finally look at the second variant of the second scenario of our experiment.
-The hypothesis in this variant is that LLMs will generate text that exhibits
-HLFs much closer to the target corpus average if the input text is from that
+The hypothesis in this scenario variant was that the baseline would exhibit HLFs
+that are much closer to the target corpus average if the input text is from that
 corpus.
-The expectation is to see less significant results.
+The expectation is to see fewer significant HLF results and that, as a rule of
+thumb, HLF results are closer to the baseline in all cases.
 The results shown in Table~\ref{table-prompt-2-ifd-cnn-dailymail} contradict our
@@
             \toprule
-            claude3 $Diff\_N$     & \textbf{\underline{871.02}}  & \textbf{\underline{34.2}}   & \textbf{\underline{34.48}}  & \textbf{\textit{-7.38}}    & \textbf{\underline{0.27}} & 0.01                      & \textbf{\textit{-0.93}}    & \textbf{\textit{-0.72}}    & 1.16                       & -0.14                   \\ \midrule
-            claude3 $Diff\_A$     & \textbf{\underline{2756.12}} & \textbf{\underline{117.18}} & \textbf{\underline{126.36}} & \textbf{\textit{-11.2}}    & \textbf{\underline{0.81}} & -0.01                     & \textbf{\textit{-2.52}}    & \textbf{\textit{-0.74}}    & 3.69                       & -0.4                    \\ \midrule
-            gemini $Diff\_N$      & \textbf{\underline{1315.41}} & \textbf{\underline{46.19}}  & \textbf{\underline{114.47}} & \textbf{\underline{72.38}} & \textbf{\underline{1.17}} & 0.03                      & \textbf{\underline{5.25}}  & \textbf{\underline{6.64}}  & -1.6                       & \textbf{\textit{-2.1}}  \\ \midrule
-            gemini $Diff\_A$      & \textbf{\underline{3996.0}}  & \textbf{\underline{138.0}}  & \textbf{\underline{357.36}} & \textbf{\underline{212.7}} & \textbf{\underline{3.3}}  & 0.06                      & \textbf{\underline{15.49}} & \textbf{\underline{19.4}}  & -13.22                     & \textbf{\textit{-6.19}} \\ \midrule
-            gpt $Diff\_N$         & \textbf{\underline{521.29}}  & \textbf{\underline{23.3}}   & \textbf{\underline{43.13}}  & 19.85                      & \textbf{\underline{0.2}}  & 0.02                      & 1.08                       & \textbf{\underline{1.04}}  & 5.09                       & 0.67                    \\ \midrule
-            gpt $Diff\_A$         & \textbf{\underline{1464.0}}  & \textbf{\underline{65.5}}   & \textbf{\underline{121.84}} & 58.1                       & \textbf{\underline{0.58}} & 0.05                      & 2.93                       & \textbf{\underline{2.59}}  & 14.23                      & 1.89                    \\ \midrule
-            command\_r $Diff\_N$  & \textbf{\underline{371.08}}  & \textbf{\underline{13.31}}  & 22.06                       & \textbf{\underline{17.19}} & \textbf{\underline{0.17}} & \textbf{\underline{0.02}} & 0.88                       & 1.09                       & \textbf{\underline{4.05}}  & -0.15                   \\ \midrule
-            command\_r $Diff\_A$  & \textbf{\underline{1198.0}}  & \textbf{\underline{42.5}}   & 73.0                        & \textbf{\underline{56.5}}  & \textbf{\underline{0.69}} & \textbf{\underline{0.07}} & 2.85                       & 3.99                       & \textbf{\underline{13.89}} & -0.11                   \\ \midrule
-            llama3\_70b $Diff\_N$ & \textbf{\underline{1040.85}} & \textbf{\underline{41.99}}  & \textbf{\underline{80.58}}  & \textbf{\underline{49.12}} & \textbf{\underline{0.55}} & \textbf{\underline{0.04}} & \textbf{\underline{3.99}}  & \textbf{\underline{5.05}}  & 7.18                       & \textbf{\textit{-1.94}} \\ \midrule
-            llama3\_70b $Diff\_A$ & \textbf{\underline{2928.0}}  & \textbf{\underline{118.5}}  & \textbf{\underline{222.0}}  & \textbf{\underline{133.5}} & \textbf{\underline{1.43}} & \textbf{\underline{0.05}} & \textbf{\underline{10.77}} & \textbf{\underline{13.42}} & 17.21                      & \textbf{\textit{-6.2}}  \\ \bottomrule
+            claude3 $Diff_N$     & \textbf{\underline{871.02}}  & \textbf{\underline{34.2}}   & \textbf{\underline{34.48}}  & \textbf{\textit{-7.38}}    & \textbf{\underline{0.27}} & 0.01                      & \textbf{\textit{-0.93}}    & \textbf{\textit{-0.72}}    & 1.16                       & -0.14                   \\ \midrule
+            claude3 $Diff_A$     & \textbf{\underline{2756.12}} & \textbf{\underline{117.18}} & \textbf{\underline{126.36}} & \textbf{\textit{-11.2}}    & \textbf{\underline{0.81}} & -0.01                     & \textbf{\textit{-2.52}}    & \textbf{\textit{-0.74}}    & 3.69                       & -0.4                    \\ \midrule
+            gemini $Diff_N$      & \textbf{\underline{1315.41}} & \textbf{\underline{46.19}}  & \textbf{\underline{114.47}} & \textbf{\underline{72.38}} & \textbf{\underline{1.17}} & 0.03                      & \textbf{\underline{5.25}}  & \textbf{\underline{6.64}}  & -1.6                       & \textbf{\textit{-2.1}}  \\ \midrule
+            gemini $Diff_A$      & \textbf{\underline{3996.0}}  & \textbf{\underline{138.0}}  & \textbf{\underline{357.36}} & \textbf{\underline{212.7}} & \textbf{\underline{3.3}}  & 0.06                      & \textbf{\underline{15.49}} & \textbf{\underline{19.4}}  & -13.22                     & \textbf{\textit{-6.19}} \\ \midrule
+            gpt $Diff_N$         & \textbf{\underline{521.29}}  & \textbf{\underline{23.3}}   & \textbf{\underline{43.13}}  & 19.85                      & \textbf{\underline{0.2}}  & 0.02                      & 1.08                       & \textbf{\underline{1.04}}  & 5.09                       & 0.67                    \\ \midrule
+            gpt $Diff_A$         & \textbf{\underline{1464.0}}  & \textbf{\underline{65.5}}   & \textbf{\underline{121.84}} & 58.1                       & \textbf{\underline{0.58}} & 0.05                      & 2.93                       & \textbf{\underline{2.59}}  & 14.23                      & 1.89                    \\ \midrule
+            command\_r $Diff_N$  & \textbf{\underline{371.08}}  & \textbf{\underline{13.31}}  & 22.06                       & \textbf{\underline{17.19}} & \textbf{\underline{0.17}} & \textbf{\underline{0.02}} & 0.88                       & 1.09                       & \textbf{\underline{4.05}}  & -0.15                   \\ \midrule
+            command\_r $Diff_A$  & \textbf{\underline{1198.0}}  & \textbf{\underline{42.5}}   & 73.0                        & \textbf{\underline{56.5}}  & \textbf{\underline{0.69}} & \textbf{\underline{0.07}} & 2.85                       & 3.99                       & \textbf{\underline{13.89}} & -0.11                   \\ \midrule
+            llama3\_70b $Diff_N$ & \textbf{\underline{1040.85}} & \textbf{\underline{41.99}}  & \textbf{\underline{80.58}}  & \textbf{\underline{49.12}} & \textbf{\underline{0.55}} & \textbf{\underline{0.04}} & \textbf{\underline{3.99}}  & \textbf{\underline{5.05}}  & 7.18                       & \textbf{\textit{-1.94}} \\ \midrule
+            llama3\_70b $Diff_A$ & \textbf{\underline{2928.0}}  & \textbf{\underline{118.5}}  & \textbf{\underline{222.0}}  & \textbf{\underline{133.5}} & \textbf{\underline{1.43}} & \textbf{\underline{0.05}} & \textbf{\underline{10.77}} & \textbf{\underline{13.42}} & 17.21                      & \textbf{\textit{-6.2}}  \\ \bottomrule
         \end{tabular}%
@@
 
-\begin{figure*}[ht]
+\begin{figure*}[ht!]
     \centering
@@
 
-The outcomes from the latest experiment are ambiguous. The LLMs do not seem to
-utilize the examples given in the prompt to their advantage when they generate
-text that should mirror the linguistic features of those examples.
+Overall, the LLMs do not seem to utilize the examples given in the prompt to
+their advantage when generating text that should mirror the linguistic features
+of those examples.
 
@@
 LLMs ability to comply.
-This is highlighted in the experiment with age of acquisition versus the
-adjective variation.
+This is highlighted in the experiment with age of acquisition versus adjective
+variation.
 
@@
 exhibits certain linguistic features by instructing a large language model.
-To our surprise, it turns out state of the art large language models are
-receptive to instructions regarding the linguistic features of the output.
+It turns out state of the art large language models are receptive to
+instructions regarding the linguistic features of the output.
 This is especially true for concrete instructions.
@@
 features and input text seem like the way to obtain desirable results.
-Providing examples in the input prompt does not seem as beneficial.
+Providing examples in the input prompt does not seem to influence the HLFs of
+the LLM output in the expected manner.
+
+Finally, we've seen how we might use handcrafted linguistic features to assess
+LLM output.
+Setting up ``before and after'' scenarios to evaluate relative improvements of
+the LLM outcome in relation to the selected HLFs is one way to achieve this.
+Given enough measurements, the relative difference between the exhibited
+baseline and target HLFs can be quantified relatively easily using geometric or
+statistical means.
 
@@
             \toprule
-            claude3 $Diff\_N$     & \textbf{\textit{-281.9}}  & \textbf{\textit{-13.22}}   & \textbf{\textit{-25.79}} & -14.56                   & -0.06     & 0.0                     & -1.27                   & -1.14                   & \textbf{\underline{6.89}}  & \textbf{\underline{1.0}} \\ \midrule
-            claude3 $Diff\_A$     & \textbf{\textit{-774.0}}  & \textbf{\textit{-27.76}}   & \textbf{\textit{-70.0}}  & -38.0                    & -0.19     & 0.02                    & -3.25                   & -3.39                   & \textbf{\underline{21.68}} & \textbf{\underline{2.9}} \\ \midrule
-            gemini $Diff\_N$      & -109.27                   & 0.05                       & -7.25                    & -13.18                   & -0.14     & 0.01                    & -1.01                   & -0.4                    & 1.62                       & -0.4                     \\ \midrule
-            gemini $Diff\_A$      & -252.82                   & -0.78                      & -17.5                    & -23.14                   & -0.52     & 0.01                    & -1.85                   & -1.07                   & 7.14                       & -0.93                    \\ \midrule
-            gpt $Diff\_N$         & -21.62                    & 2.42                       & -3.8                     & -8.84                    & 0.0       & -0.01                   & -0.8                    & -0.53                   & 8.94                       & -0.37                    \\ \midrule
-            gpt $Diff\_A$         & -34.72                    & 9.68                       & -10.54                   & -24.57                   & -0.03     & -0.02                   & -2.37                   & -1.58                   & 19.26                      & -1.03                    \\ \midrule
-            command\_r $Diff\_N$  & -40.1                     & 1.22                       & 2.43                     & \textbf{\textit{-13.64}} & 0.05      & \textbf{\textit{-0.04}} & \textbf{\textit{-1.17}} & 0.59                    & \textbf{\textit{-7.25}}    & \textbf{\textit{-1.67}}  \\ \midrule
-            command\_r $Diff\_A$  & -75.5                     & 1.5                        & 12.94                    & \textbf{\textit{-34.0}}  & 0.19      & \textbf{\textit{-0.12}} & \textbf{\textit{-2.76}} & 2.83                    & \textbf{\textit{-20.49}}   & \textbf{\textit{-4.76}}  \\ \midrule
-            llama3\_70b $Diff\_N$ & \textbf{\textit{-210.45}} & \textbf{\underline{6.87}}  & \textbf{\textit{-13.46}} & \textbf{\textit{-15.17}} & -0.16     & \textbf{\textit{-0.02}} & \textbf{\textit{-1.78}} & \textbf{\textit{-1.31}} & 3.7                        & \textbf{\textit{-1.58}}  \\ \midrule
-            llama3\_70b $Diff\_A$ & \textbf{\textit{-607.52}} & \textbf{\underline{19.42}} & \textbf{\textit{-41.22}} & \textbf{\textit{-44.77}} & -0.42     & \textbf{\textit{-0.05}} & \textbf{\textit{-5.48}} & \textbf{\textit{-4.34}} & 9.79                       & \textbf{\textit{-4.98}}  \\ \bottomrule
+            claude3 $Diff_N$     & \textbf{\textit{-281.9}}  & \textbf{\textit{-13.22}}   & \textbf{\textit{-25.79}} & -14.56                   & -0.06     & 0.0                     & -1.27                   & -1.14                   & \textbf{\underline{6.89}}  & \textbf{\underline{1.0}} \\ \midrule
+            claude3 $Diff_A$     & \textbf{\textit{-774.0}}  & \textbf{\textit{-27.76}}   & \textbf{\textit{-70.0}}  & -38.0                    & -0.19     & 0.02                    & -3.25                   & -3.39                   & \textbf{\underline{21.68}} & \textbf{\underline{2.9}} \\ \midrule
+            gemini $Diff_N$      & -109.27                   & 0.05                       & -7.25                    & -13.18                   & -0.14     & 0.01                    & -1.01                   & -0.4                    & 1.62                       & -0.4                     \\ \midrule
+            gemini $Diff_A$      & -252.82                   & -0.78                      & -17.5                    & -23.14                   & -0.52     & 0.01                    & -1.85                   & -1.07                   & 7.14                       & -0.93                    \\ \midrule
+            gpt $Diff_N$         & -21.62                    & 2.42                       & -3.8                     & -8.84                    & 0.0       & -0.01                   & -0.8                    & -0.53                   & 8.94                       & -0.37                    \\ \midrule
+            gpt $Diff_A$         & -34.72                    & 9.68                       & -10.54                   & -24.57                   & -0.03     & -0.02                   & -2.37                   & -1.58                   & 19.26                      & -1.03                    \\ \midrule
+            command\_r $Diff_N$  & -40.1                     & 1.22                       & 2.43                     & \textbf{\textit{-13.64}} & 0.05      & \textbf{\textit{-0.04}} & \textbf{\textit{-1.17}} & 0.59                    & \textbf{\textit{-7.25}}    & \textbf{\textit{-1.67}}  \\ \midrule
+            command\_r $Diff_A$  & -75.5                     & 1.5                        & 12.94                    & \textbf{\textit{-34.0}}  & 0.19      & \textbf{\textit{-0.12}} & \textbf{\textit{-2.76}} & 2.83                    & \textbf{\textit{-20.49}}   & \textbf{\textit{-4.76}}  \\ \midrule
+            llama3\_70b $Diff_N$ & \textbf{\textit{-210.45}} & \textbf{\underline{6.87}}  & \textbf{\textit{-13.46}} & \textbf{\textit{-15.17}} & -0.16     & \textbf{\textit{-0.02}} & \textbf{\textit{-1.78}} & \textbf{\textit{-1.31}} & 3.7                        & \textbf{\textit{-1.58}}  \\ \midrule
+            llama3\_70b $Diff_A$ & \textbf{\textit{-607.52}} & \textbf{\underline{19.42}} & \textbf{\textit{-41.22}} & \textbf{\textit{-44.77}} & -0.42     & \textbf{\textit{-0.05}} & \textbf{\textit{-5.48}} & \textbf{\textit{-4.34}} & 9.79                       & \textbf{\textit{-4.98}}  \\ \bottomrule
         \end{tabular}%
@@
             \toprule
-            claude3 $Diff\_N$     & -121.77 & -4.61   & -9.9     & 18.41   & -0.11     & 0.03        & \textbf{\underline{1.4}}  & -0.33           & 1.81                       & 0.95       \\ \midrule
-            claude3 $Diff\_A$     & -394.5  & -11.5   & -37.5    & 60.0    & -0.37     & 0.09        & \textbf{\underline{4.66}} & -1.49           & 5.38                       & 2.62       \\ \midrule
-            gemini $Diff\_N$      & -23.49  & 1.52    & -0.55    & -1.29   & -0.02     & -0.01       & 0.09                      & -0.05           & \textbf{\underline{6.41}}  & 0.2        \\ \midrule
-            gemini $Diff\_A$      & -83.0   & 3.0     & -2.5     & -6.36   & -0.0      & -0.03       & 0.26                      & -0.08           & \textbf{\underline{19.63}} & 0.53       \\ \midrule
-            gpt $Diff\_N$         & 13.66   & 2.83    & -3.65    & -3.19   & 0.04      & -0.04       & -0.24                     & -0.22           & 7.5                        & 0.0        \\ \midrule
-            gpt $Diff\_A$         & 55.5    & 9.5     & -9.86    & -12.86  & 0.14      & -0.11       & -0.84                     & -0.22           & 18.67                      & -0.38      \\ \midrule
-            command\_r $Diff\_N$  & -188.71 & 0.69    & -10.73   & 1.17    & -0.23     & -0.01       & \textbf{\underline{0.74}} & -0.83           & \textbf{\textit{-5.02}}    & -0.08      \\ \midrule
-            command\_r $Diff\_A$  & -516.82 & 2.02    & -29.22   & 8.28    & -0.66     & -0.0        & \textbf{\underline{2.98}} & -2.48           & \textbf{\textit{-14.52}}   & -0.48      \\ \midrule
-            llama3\_70b $Diff\_N$ & -202.95 & -7.84   & -13.73   & 1.43    & -0.15     & -0.04       & 0.38                      & -1.39           & 10.67                      & 1.37       \\ \midrule
-            llama3\_70b $Diff\_A$ & -561.24 & -18.78  & -49.94   & 5.07    & -0.38     & -0.14       & 1.62                      & -4.81           & 29.26                      & 3.11       \\ \bottomrule
+            claude3 $Diff_N$     & -121.77 & -4.61   & -9.9     & 18.41   & -0.11     & 0.03        & \textbf{\underline{1.4}}  & -0.33           & 1.81                       & 0.95       \\ \midrule
+            claude3 $Diff_A$     & -394.5  & -11.5   & -37.5    & 60.0    & -0.37     & 0.09        & \textbf{\underline{4.66}} & -1.49           & 5.38                       & 2.62       \\ \midrule
+            gemini $Diff_N$      & -23.49  & 1.52    & -0.55    & -1.29   & -0.02     & -0.01       & 0.09                      & -0.05           & \textbf{\underline{6.41}}  & 0.2        \\ \midrule
+            gemini $Diff_A$      & -83.0   & 3.0     & -2.5     & -6.36   & -0.0      & -0.03       & 0.26                      & -0.08           & \textbf{\underline{19.63}} & 0.53       \\ \midrule
+            gpt $Diff_N$         & 13.66   & 2.83    & -3.65    & -3.19   & 0.04      & -0.04       & -0.24                     & -0.22           & 7.5                        & 0.0        \\ \midrule
+            gpt $Diff_A$         & 55.5    & 9.5     & -9.86    & -12.86  & 0.14      & -0.11       & -0.84                     & -0.22           & 18.67                      & -0.38      \\ \midrule
+            command\_r $Diff_N$  & -188.71 & 0.69    & -10.73   & 1.17    & -0.23     & -0.01       & \textbf{\underline{0.74}} & -0.83           & \textbf{\textit{-5.02}}    & -0.08      \\ \midrule
+            command\_r $Diff_A$  & -516.82 & 2.02    & -29.22   & 8.28    & -0.66     & -0.0        & \textbf{\underline{2.98}} & -2.48           & \textbf{\textit{-14.52}}   & -0.48      \\ \midrule
+            llama3\_70b $Diff_N$ & -202.95 & -7.84   & -13.73   & 1.43    & -0.15     & -0.04       & 0.38                      & -1.39           & 10.67                      & 1.37       \\ \midrule
+            llama3\_70b $Diff_A$ & -561.24 & -18.78  & -49.94   & 5.07    & -0.38     & -0.14       & 1.62                      & -4.81           & 29.26                      & 3.11       \\ \bottomrule
         \end{tabular}%
@@
             \toprule
-            claude3 $Diff\_N$     & -148.13                   & -1.63                   & -3.93                     & \textbf{\textit{-46.03}} & -0.11                   & -0.0        & \textbf{\textit{-3.08}} & -0.31                   & -4.3                    & \textbf{\textit{-1.53}} \\ \midrule
-            claude3 $Diff\_A$     & -526.14                   & -7.98                   & -23.86                    & \textbf{\textit{-133.5}} & -0.41                   & -0.0        & \textbf{\textit{-8.98}} & -1.75                   & -21.03                  & \textbf{\textit{-5.34}} \\ \midrule
-            gemini $Diff\_N$      & 74.27                     & 0.65                    & 4.67                      & 2.98                     & 0.04                    & -0.0        & -0.22                   & 1.05                    & 1.44                    & -0.65                   \\ \midrule
-            gemini $Diff\_A$      & 221.0                     & 2.0                     & 10.5                      & 7.44                     & 0.12                    & -0.03       & -0.85                   & 2.73                    & 3.28                    & -1.79                   \\ \midrule
-            gpt $Diff\_N$         & 0.65                      & 4.03                    & -1.19                     & -2.66                    & -0.03                   & -0.01       & -0.17                   & -0.13                   & 1.05                    & -0.67                   \\ \midrule
-            gpt $Diff\_A$         & 17.0                      & 11.68                   & -3.68                     & -7.0                     & -0.1                    & -0.04       & -0.43                   & -0.05                   & 0.57                    & -1.75                   \\ \midrule
-            command\_r $Diff\_N$  & \textbf{\textit{-236.11}} & \textbf{\textit{-3.6}}  & \textbf{\textit{-0.97}}   & \textbf{\textit{-27.04}} & \textbf{\textit{-0.23}} & 0.0         & -1.24                   & \textbf{\textit{-1.08}} & \textbf{\textit{-2.13}} & \textbf{\textit{-1.09}} \\ \midrule
-            command\_r $Diff\_A$  & \textbf{\textit{-727.0}}  & \textbf{\textit{-9.24}} & \textbf{\underline{0.76}} & \textbf{\textit{-68.5}}  & \textbf{\textit{-0.64}} & 0.02        & -2.89                   & \textbf{\textit{-3.08}} & \textbf{\textit{-5.5}}  & \textbf{\textit{-3.0}}  \\ \midrule
-            llama3\_70b $Diff\_N$ & \textbf{\textit{-333.03}} & \textbf{\textit{-5.19}} & 4.43                      & \textbf{\textit{-21.38}} & -0.16                   & -0.03       & \textbf{\textit{-1.77}} & -0.03                   & 1.87                    & -0.44                   \\ \midrule
-            llama3\_70b $Diff\_A$ & \textbf{\textit{-687.34}} & \textbf{\textit{-3.42}} & 10.7                      & \textbf{\textit{-61.64}} & -0.31                   & -0.04       & \textbf{\textit{-5.6}}  & -0.89                   & 3.0                     & -1.92                   \\ \bottomrule
+            claude3 $Diff_N$     & -148.13                   & -1.63                   & -3.93                     & \textbf{\textit{-46.03}} & -0.11                   & -0.0        & \textbf{\textit{-3.08}} & -0.31                   & -4.3                    & \textbf{\textit{-1.53}} \\ \midrule
+            claude3 $Diff_A$     & -526.14                   & -7.98                   & -23.86                    & \textbf{\textit{-133.5}} & -0.41                   & -0.0        & \textbf{\textit{-8.98}} & -1.75                   & -21.03                  & \textbf{\textit{-5.34}} \\ \midrule
+            gemini $Diff_N$      & 74.27                     & 0.65                    & 4.67                      & 2.98                     & 0.04                    & -0.0        & -0.22                   & 1.05                    & 1.44                    & -0.65                   \\ \midrule
+            gemini $Diff_A$      & 221.0                     & 2.0                     & 10.5                      & 7.44                     & 0.12                    & -0.03       & -0.85                   & 2.73                    & 3.28                    & -1.79                   \\ \midrule
+            gpt $Diff_N$         & 0.65                      & 4.03                    & -1.19                     & -2.66                    & -0.03                   & -0.01       & -0.17                   & -0.13                   & 1.05                    & -0.67                   \\ \midrule
+            gpt $Diff_A$         & 17.0                      & 11.68                   & -3.68                     & -7.0                     & -0.1                    & -0.04       & -0.43                   & -0.05                   & 0.57                    & -1.75                   \\ \midrule
+            command\_r $Diff_N$  & \textbf{\textit{-236.11}} & \textbf{\textit{-3.6}}  & \textbf{\textit{-0.97}}   & \textbf{\textit{-27.04}} & \textbf{\textit{-0.23}} & 0.0         & -1.24                   & \textbf{\textit{-1.08}} & \textbf{\textit{-2.13}} & \textbf{\textit{-1.09}} \\ \midrule
+            command\_r $Diff_A$  & \textbf{\textit{-727.0}}  & \textbf{\textit{-9.24}} & \textbf{\underline{0.76}} & \textbf{\textit{-68.5}}  & \textbf{\textit{-0.64}} & 0.02        & -2.89                   & \textbf{\textit{-3.08}} & \textbf{\textit{-5.5}}  & \textbf{\textit{-3.0}}  \\ \midrule
+            llama3\_70b $Diff_N$ & \textbf{\textit{-333.03}} & \textbf{\textit{-5.19}} & 4.43                      & \textbf{\textit{-21.38}} & -0.16                   & -0.03       & \textbf{\textit{-1.77}} & -0.03                   & 1.87                    & -0.44                   \\ \midrule
+            llama3\_70b $Diff_A$ & \textbf{\textit{-687.34}} & \textbf{\textit{-3.42}} & 10.7                      & \textbf{\textit{-61.64}} & -0.31                   & -0.04       & \textbf{\textit{-5.6}}  & -0.89                   & 3.0                     & -1.92                   \\ \bottomrule
         \end{tabular}%
@@
 \subsection{GPT}
-
-\Cref{fig:gpt-prompt1-cnn,fig:gpt-prompt1-yelp,fig:gpt-prompt2-cnn,fig:gpt-prompt2-yelp,fig:gpt-prompt2-cnn-ifd,fig:gpt-prompt2-yelp-ifd}
-show the experiment results for the Claude 3 Opus large language model.
 
@@
 
+\Cref{fig:gpt-prompt1-cnn,fig:gpt-prompt1-yelp,fig:gpt-prompt2-cnn,fig:gpt-prompt2-yelp,fig:gpt-prompt2-cnn-ifd,fig:gpt-prompt2-yelp-ifd}
+show the experiment results for the Claude 3 Opus large language model.
+
 \subsection{Gemini}
-
-\Cref{fig:gemini-prompt1-cnn,fig:gemini-prompt1-yelp,fig:gemini-prompt2-cnn,fig:gemini-prompt2-yelp,fig:gemini-prompt2-cnn-ifd,fig:gemini-prompt2-yelp-ifd}
-show the experiment results for the Claude 3 Opus large language model.
 
@@
 
+\Cref{fig:gemini-prompt1-cnn,fig:gemini-prompt1-yelp,fig:gemini-prompt2-cnn,fig:gemini-prompt2-yelp,fig:gemini-prompt2-cnn-ifd,fig:gemini-prompt2-yelp-ifd}
+show the experiment results for the Claude 3 Opus large language model.
+
 \subsection{Claude3}
-
-\Cref{fig:claude3-prompt1-cnn,fig:claude3-prompt1-yelp,fig:claude3-prompt2-cnn,fig:claude3-prompt2-yelp,fig:claude3-prompt2-cnn-ifd,fig:claude3-prompt2-yelp-ifd}
-show the experiment results for the Claude 3 Opus large language model.
 
@@
 
+\Cref{fig:claude3-prompt1-cnn,fig:claude3-prompt1-yelp,fig:claude3-prompt2-cnn,fig:claude3-prompt2-yelp,fig:claude3-prompt2-cnn-ifd,fig:claude3-prompt2-yelp-ifd}
+show the experiment results for the Claude 3 Opus large language model.
+
 \subsection{Llama3}
-
-\Cref{fig:llama3_70b-prompt1-cnn,fig:llama3_70b-prompt1-yelp,fig:llama3_70b-prompt2-cnn,fig:llama3_70b-prompt2-yelp,fig:llama3_70b-prompt2-cnn-ifd,fig:llama3_70b-prompt2-yelp-ifd}
-show the experiment results for the Claude 3 Opus large language model.
 
@@
 
+\Cref{fig:llama3_70b-prompt1-cnn,fig:llama3_70b-prompt1-yelp,fig:llama3_70b-prompt2-cnn,fig:llama3_70b-prompt2-yelp,fig:llama3_70b-prompt2-cnn-ifd,fig:llama3_70b-prompt2-yelp-ifd}
+show the experiment results for the Claude 3 Opus large language model.
+
 \subsection{Command R+}
-
-\Cref{fig:command_r-prompt1-cnn,fig:command_r-prompt1-yelp,fig:command_r-prompt2-cnn,fig:command_r-prompt2-yelp,fig:command_r-prompt2-cnn-ifd,fig:command_r-prompt2-yelp-ifd}
-show the experiment results for the Claude 3 Opus large language model.
 
@@
 
+\Cref{fig:command_r-prompt1-cnn,fig:command_r-prompt1-yelp,fig:command_r-prompt2-cnn,fig:command_r-prompt2-yelp,fig:command_r-prompt2-cnn-ifd,fig:command_r-prompt2-yelp-ifd}
+show the experiment results for the Claude 3 Opus large language model.
+
 \end{document}

### 2024-11-14 article/main.tex
refer to HLF results in a single way

@@
 the original sentence''~\cite{tst-review-2021}.
-This definition can be extended to entire articles or corpuses of text because
+This definition can be extended to entire articles or corpora of text because
 these are also the object of linguistic style and
@@
 With discussions on style transfer or author verification beyond this paper's
-scope, we present a method controlling LLM text generation through the use of
-HLFs.
+scope, we present a method for controlling LLM text generation through the use
+of HLFs.
 
@@
 To run the experiment we must select LLMs to evaluate and find datasets from
-which we can derive corpuses.
+which we can derive corpora.
 Then we must also select the HLFs for our experiment.
@@
 the LLM.\@
-Some context windows are smaller than others, settling the number of examples at
+Some context windows are smaller than others, setting the number of examples at
 three.
@@
 experiment.
-Following the process from the first scenario results in obtaining baseline HLF
-values for each variant of the second scenario.
+We repeat the same process from the first scenario to obtain baseline HLF values
+for each variant of the second scenario.
 
@@
 From here on we refer to the results obtained by using prompts with HLF
-instructions as `HLF-enhanced results', `HLF results' or even simply `results'.
+instructions as `HLF results'.
 
@@
             t\_sent         & total sentences                   & wordsent         & surface          & very concrete   & syntactic               \\ \bottomrule
-            n\_uverb        & total number of unique adjectives & partofspeech     & syntax           & concrete        & semantic                \\
+            n\_uverb        & total number of unique verbs      & partofspeech     & syntax           & concrete        & syntactic               \\
             n\_uadj         & total number of unique adjectives & partofspeech     & syntax           & concrete        & syntactic               \\ \bottomrule
@@
             fkgl            & Flesch-Kincaid grade level        & readformula      & surface          & highly abstract & pragmatic               \\
-            a\_kup\_pw      & Kuperman age of aquisition        & worddiff         & lexico-semantics & highly abstract & lexical                 \\ \bottomrule
+            a\_kup\_pw      & Kuperman age of acquisition       & worddiff         & lexico-semantics & highly abstract & lexical                 \\ \bottomrule
         \end{tabular}%
@@
 CNN/DailyMail contains stories and news with diverse styles.
-The style varies from one author to another and it also follows the publisher's
-editorial style.
+The style follows the publisher's guidelines, but, within those guidelines, it
+varies for each author.
 The articles are more formal, longer and more complex in structure than Yelp
@@
 
-Determining whether the baseline results differ from the HLF-enhanced results is
-a classic use case for the Kolmogorov-Smirnov
-test~\cite{kolmogorov1933,smirnov1939}, as implemented in the
-Scipy library~\cite{2020SciPy-NMeth}.
-We consider that the results are different for a resulting $p\_value \le 0.05$.
+We can determine whether the baseline results differ from the HLF results using
+the Kolmogorov-Smirnov test~\cite{kolmogorov1933,smirnov1939}, as implemented in
+the Scipy library~\cite{2020SciPy-NMeth}.
+We consider that the results are different for a $p\_value \le 0.05$.
 
@@
 
-The results on the CNN article corpus showcased in Table~\ref{table-prompt-1-cnn-dailymail}
-shows that LLMs take into account instructions concerning the HLFs of their
-output.
-
-The results on the review corpus generated from the Yelp dataset are similarly
-significant, but the HLF instructions determine outcomes that are further away
-from the target.
+Table~\ref{table-prompt-1-cnn-dailymail} shows that LLMs take into account
+HLF instructions when running the experiment.
+On the CNN article corpus, the HLF instructions lead to HLF results that are
+closer than the baseline to the target HLF.\@
+On the Yelp dataset, the HLF instructions generate results that are further away
+than the baseline from the target HLF.\@
 
@@
 
-There are two variants of this scenario: input text from within or from outside
-the corpus.
+There are two variants of this scenario depending on the provenance of the input
+text.
+The input text may be from outside the corpus or a text from within the corpus.
 Table~\ref{table-prompt-2-cnn-dailymail} contains results for the first variant:
-input from the corpus.
-We notice these results are similar to those in the first experiment scenario.
+input from outside the corpus.
+We notice the HLF results tend to be closer to the target than the baseline
+results.
+This picture is similar to the one we saw in the first scenario.
 
@@
 
-A possible explanation for having fewer significant differences in results might
-lie in the closeness to the target.
+On the other hand, there are fewer significant differences between baseline and
+HLF results.
+A possible explanation for this situation might lie in the closeness of the
+input prompt (which contains examples from the corpus)to the target.
 As we can see in Figure~\ref{fig-p2-gpt-twords}, this is not always the case.
@@
         \includegraphics[width=\linewidth]{plots/prompt_2/prompt_2-gemini-cnn_dailymail/prompt_2-gemini-cnn_dailymail_a_kup_pw.png}
-        \caption[center]{Gemini\\Kuperman Age of Aquisition}\label{fig-p2-gemini-a-kup-pw}
+        \caption[center]{Gemini\\Kuperman Age of Acquisition}\label{fig-p2-gemini-a-kup-pw}
     \end{minipage}
@@
         \includegraphics[width=\linewidth]{plots/prompt_2_ifd/prompt_2-llama3_70b-cnn_dailymail/prompt_2-llama3_70b-cnn_dailymail_a_kup_pw.png}
-        \caption[center]{Llama3\\Kuperman Age of Aquisition}\label{fig-p2-ifd-llama3-a-kup-pw}
+        \caption[center]{Llama3\\Kuperman Age of Acquisition}\label{fig-p2-ifd-llama3-a-kup-pw}
     \end{minipage}
@@
 the usage of a fixed external input text.
-This is relevant as evidentiated by the difference in results obtained using the
-Yelp reviews corpus when compared to the results obtained on the CNN article
-corpus.
+This is relevant as evidentiated by the difference in HLF results obtained using
+the Yelp reviews corpus when compared to the HLF results obtained on the CNN
+article corpus.
 While similarly significant, the HLF results on Yelp reviews are mostly worse
@@
 The importance of the choice of the external text is highlighted by the better
-results obtained using an input text from the corpus.
+HLF results obtained using an input text from the corpus.
 
-This leads us to our second question: how much do HLF instructions impact an
-LLM's writing?
+Let us move on to how much HLF instructions impact an LLM's writing.
 The metrics we compute dismiss the idea that HLF instructions have a consistent
 desired impact on the generated text.
-There isn't conclusive evidence that improving the LLM's odds of producing a
-better baseline results in a smaller difference between baseline and result
-relative to the target.
-More importantly, there is evidence that HLF instructions have an adverse effect
-when comparing the results on the Yelp reviews dataset.
+There isn't conclusive evidence that if we improve the LLM's chances of
+producing a better baseline, this will result in a smaller difference between
+baseline and HLF result relative to the target HLF.
+In fact, there is evidence that HLF instructions have an adverse effect when
+examining the HLF results obtained on the Yelp reviews dataset.
+This is especially true of highly abstract HLF instructions.
+LLMs do not seem to follow these instructions well in most scenarios.
+The impact of highly abstract HLF instructions tends to be adverse.
 
-Highly abstract HLF instructions are unlikely to have the desired effect.
-Furthermore, the impact tends to be adverse.
-This suggests that LLMs are more adept at following instructions regarding HLFs
-when those instructions are more concrete.
-
-The choice of input text in relation to the target HLF values also seems
-important.
+The choice of input text in relation to the target HLF values is consequential.
 LLMs don't yet seem able to cover the gulf between Oscar Wilde poems and the
@@
 target HLFs, there is a limit to this ability.
-
-Finally, selecting HLFs that are in tension with one another will affect the
-LLMs ability to comply.
-This is highlighted in the experiment with age of acquisition versus adjective
-variation.
 

### 2024-11-14 article/main.tex
capitalize Figure everywhere

@@
 \end{table*}
-\footnotetext{as described in the figure in~\cite{lugea2023stylistics}, chapter 1, page 7}
+\footnotetext{as described in the figure from~\cite{lugea2023stylistics}, chapter 1, page 7}
 
@@
 concrete generation instructions, such as the total number of words.
-This behaviour is highlighted in figure~\ref{fig-p2-ifd-gemini-twords}.
+This behaviour is highlighted in Figure~\ref{fig-p2-ifd-gemini-twords}.
 Additionally, we notice regressive behaviour for some models compared to the
-previously explored results in figure~\ref{fig-p2-ifd-claude3-nuadj}.
+previously explored results in Figure~\ref{fig-p2-ifd-claude3-nuadj}.
 Finally, the abstractness of the text generation instructions for certain HLFs
 leads to the generated text not being compliant as exemplified in
-figure~\ref{fig-p2-ifd-llama3-a-kup-pw}.
+Figure~\ref{fig-p2-ifd-llama3-a-kup-pw}.
 

### 2024-12-10 article/main.tex
reword section 3.1 and make other minor corrections

@@
 language~\cite{lftk-2023}.
-These values are usually easy to compute with methods that are idempotent.
+Examples of HLFs range from simple constructs such as counts or averages of
+words, sentences or specific parts of speech~(adjectives, verbs, a.s.o) to more
+complex statistics based on heuristics.
+All HLFs share the characteristics of computational ease and idempotence.
 This makes HLFs very attractive because of their dual potential.
-They can be used to instruct the generation of text (``write no more than 10
-sentences''), but they can also be used to analyse text.
-This allows us to verify if a text has the \textit{expected} value for an HLF.\@
+On the one hand, an LLM might be instructed using a prompt containing the value
+of an HLF.\@
+An example prompt instruction which asks the LLM to conform to a maximum
+\textit{sentence count} of 10 would read ``write no more than 10 sentences''.
+On the other hand, HLFs can be computed on the text output by an LLM, too.
+This dual nature of HLFs allows inspecting how closely does LLM generated text
+conform to the goal outlined in the input prompt HLF instruction.\@
 
-Taking Noam Chomsky's ideas that syntax and meaning are different
-things~\cite{chomsky2002syntactic}, we intuit that writing style is defined in
-terms of both.
-We shy away from discussing meaning, but syntax and similarly orderly constructs
-seem to be within our grasp.
-Traditionally, linguistic style revolved around morphological and syntactic
+If syntax differs from meaning~\cite{chomsky2002syntactic} the intuition is that
+writing style is defined in terms of both.
+The distinction is important in order to set boundaries.
+HLFs are largely focused on syntax and similarly orderly constructs.
+Therefore, we cover the influence of meaning on author style minimally and
+superficially at best.
+
+This limitation should not deter readers because, traditionally, linguistic
+style is centered around morphological and syntactic
 arrangements~\cite{lugea2023stylistics}.
-More recent descriptions find semantics and pragmatics to be inextricably linked
-to author style~\cite{verma2019lexical}, but argue that a great deal can still
-be inferred from lexicon, syntax and a `surface' level which contains basic HLFs
-such as the number of words in a sentence.
+Although recent descriptions find semantics and pragmatics to be inextricably
+linked to author style~\cite{verma2019lexical}, they too argue that a great deal
+can still be inferred from lexicon, syntax and a `surface' level which contains
+basic HLFs such as the number of words in a sentence.
 Some have observed that many characteristics of written text can be computed as
@@
 levels~\cite{dimarco1994model,lugea2023stylistics}.
-The influence of syntactic structure over text style is apparent also from a
-formalist perspective~\cite{chomsky2002syntactic}.
+The influence of syntactic structure over text style is apparent from a formal
+perspective~\cite{chomsky2002syntactic}.
 Our selection of HLFs is informed by the various aspects of linguistic style.
@@
 
-Our experiment analyses LLM output when asked to generate new text based on
-given input text and instructions derived from HLFs.
-We compute statistics of HLFs on a corpus to derive the text generation
-instructions.
-To run the experiment we must select LLMs to evaluate and find datasets from
-which we can derive corpora.
-Then we must also select the HLFs for our experiment.
-Before going over our selection for the experiment's parameters, let us go
-through the experimental process.
+Our experiment aims to find out how well LLMs can follow prompts which have
+instructions derived from HLFs.
+Two scenarios are investigated.
+
+The aim in the first scenario is to use an LLM to reword an input text so that
+its style resembles the text style of texts from a predetermined corpus of
+similarly styled texts.
+In this scenario the LLM rewords the input by following instructions which only
+contain HLFs.
+The input text used for this scenario is necessarily from outside the corpus.
+
+The second scenario is designed to show how much bearing do instructions
+containing HLFs have on the way LLMs generate text.
+The task of the LLM remains the same as in the first scenario.
+However, examples from the chosen corpus are added to the LLM prompt which means
+the LLM rewords text aided by the examples.
+The second scenario has two variants because of this particularity.
+In the first variant, similarly to the first scenario, the LLM rewords input
+from outside the corpus.
+In the second variant of the second scenario the LLM uses input text from within
+the corpus.
 
@@
 
-Assume we have a target LLM, a corpus containing texts, some examples extracted
-from the corpus, a selection of HLFs and input text both from within and from
-outside the corpus.
-This description of the experiment process assumes those parameters have stable
-values.
-Changing the parameters requires restarting the process.
+The experiment starts with choosing experiment parameter values.
+The following parameters have stable values throughout the experiment process: a
+selection of HLFs, a corpus containing texts, a target LLM, examples from the
+chosen corpus, one input text from the chosen corpus and one input text from
+outside the chosen corpus.
 
-The first activity involves establishing baselines for each experiment scenario.
-In the first scenario the target LLM is instructed to generate text retaining
-the meaning of the input text from outside the corpus without any further
-instruction related to style or the linguistic features of the generated text.
-This step is repeated 10 times, computing each HLF in our selection for the
-generated text.
-The computed values represent the baseline HLF values for the first scenario.
+Once the parameters have been determined, the HLF statistics are computed on the
+text corpus.
+Remembering that HLFs are just floating point numbers, the following statistics
+are computed for each HLF:~\textbf{min}~(the minimum value), \textbf{max}~(the
+maximum value) and \textbf{avg}~(the mean).
 
-The second scenario consists in generating text based on input text, but with
-the aid of some examples from the corpus.
-The examples are appended to the instruction prompt used in the first scenario.
-The number of examples is chosen so that we do not overrun the context window of
-the LLM.\@
-Some context windows are smaller than others, setting the number of examples at
-three.
-Adding examples to the LLM's instruction prompt also makes the provenance of the
-input text more important.
-Therefore we experiment with two variants of the second scenario: one with input
-text extraneous to the corpus and one with input text from the corpus.
-The text from outside the corpus is the same text we used in the first scenario.
-The text from within the corpus is chosen randomly once at the start of the
-experiment.
-We repeat the same process from the first scenario to obtain baseline HLF values
-for each variant of the second scenario.
+The subsequent step involves formulating prompt instructions that incorporate
+HLFs.
+For each HLF, an instruction is constructed in natural language, adhering to the
+previously calculated minimum, maximum, and mean values. For instance, an
+instruction corresponding to the total word count HLF might be expressed as
+follows:
 
-From here on we refer to `baseline HLF values' as `baseline'.
-We now have three baselines: one for the first scenario and one for each variant
-of the second scenario.
+\texttt{- Total Words: ensure the text contains 14 words (min=10, max=50).}
 
-The next step requires computing statistics on the corpus for each HLF:\@
-\textbf{min}~(the minimum value), \textbf{max}~(the maximum value) and
-\textbf{avg}~(the mean).
-These statistics allow writing an HLF prompt derived from each of our baseline
-prompts.
+The HLF instructions are used to build system prompts for the LLM.\@
+The prompts are assembled using Jinja2~\cite{jinja2} templates located in the
+\texttt{\textlinkanon{https://github.com/koosie0507/llm-tst-consistency/tree/main/src/llm_tst_consistency/templates}{templates}}
+directory of the experiment's source code.
 
-The HLF prompts instruct the LLM to generate text with HLFs that comply with the
-previously computed statistics.
-The individual (per-HLF) instructions are all part of the same HLF prompt.
-The prompts are constructed using the Jinja2~\cite{jinja2} templates which are
-located in the \texttt{\textlinkanon{https://github.com/koosie0507/llm-tst-consistency/tree/main/src/llm_tst_consistency/templates}{templates}}
-directory of the source code used to run the experiment.
-The instruction prompts corresponding to our two scenarios are
+The prompt templates correspond to each of the two experiment scenarios:
 `\texttt{prompt\_1}' and `\texttt{prompt\_2}', respectively.
-The rest of the templates in the directory are used to construct the individual
-HLF instructions.
+The static instruction in `\texttt{prompt\_1}' to ``\textit{Write text that
+conveys the meaning of the first user prompt}'' ensures that input text meaning
+is preserved.
+In addition to the primary directive from `\texttt{prompt\_1}',
+`\texttt{prompt\_2}' contains instructions concerning the interpretation of the
+provided examples, as well as logic to incorporate them.
+Both prompt templates contain logic to distinguish between the situation when
+HLF instructions were given versus not.
 
-In a similar manner to how we generated the baselines, we run 10 trials for each
-scenario and variant.
-We compare each baseline with the results obtained by using the corresponding
-HLF prompt.
-From here on we refer to the results obtained by using prompts with HLF
-instructions as `HLF results'.
+With the preliminary steps completed, baseline HLFs for the target LLM are
+computed.
+For the first scenario, the LLM is instructed to reword the input text 10 times
+by using `\texttt{prompt\_1}' without HLF instructions.
+The HLFs in our selection are computed for each generated text, resulting in
+a 10-dimensional vector for each HLF.\@
+These vectors represent the baseline HLFs for the LLM in the first scenario.
 
-The first scenario indicates whether LLMs acknowledge HLF instructions.
+The second scenario baseline values are computed similarly using the
+`\texttt{prompt\_2}' template and the examples provided as one of the experiment
+parameters.
+Remembering that the second scenario has two variants, we obtain two sets of
+baseline vectors: a set for input from the corpus and one for input outside the
+corpus.
+The input from outside the corpus is the same as the input for the first
+scenario.
+
+From here on we refer to the HLFs computed using the above process as
+\textit{baselines}.
+Note that we have 3 sets of baselines: 1 for the first scenario and 2 for the
+second scenario.
+
+For each type of baseline, we run 10 text generations using the corresponding
+prompt augmented with the HLF instructions computed earlier in the process.
+Doing this results in 3 sets of 10-dimensional vectors, each corresponding to
+the baselines.
+We refer to these vectors as \textit{HLF results}.
+
+The last step is to compare baselines with their corresponding HLF results for
+each HLF in our selection, for each scenario and variant.
+
+The comparison result for the first scenario indicates whether LLMs acknowledge
+HLF instructions at all.
 If they do, the HLF results should be significantly different from their
@@
 instructions.
-Working under the assumption that an LLM can generate text that is similar to an
-example from the corpus, the presence of HLF instructions in the prompt would
-remain among the few possible explanations were we to observe deviations from
-the baseline.
+This hypothesis emerges from the assumption that LLMs are able to mimic examples
+when they are provided in the system prompt.
+In such a case, the presence of HLF instructions would remain among the few
+possible explanations for deviations from the baseline.
 
@@
 
-We only use state of the art large language models for the experiment.
+The LLMs selected for the experiment must represent the state of the art.
 The Chatbot Arena~\cite{chiang2024chatbot} leaderboard is essential in our
-selection.
+selection based on this criterion.
 To cover more ground we find top performers from different vendors interesting.
@@
 instructions to conform to HLFs.
-LFTK~\cite{lftk-2023} is a framework built on top of spaCy~\cite{spacy}
-that provides implementations for computing many HLFs.
-We use these two libraries to perform our experiment.
-LFTK categorizes HLFs by domain and family and make available information on the
-suitability of each HLF for given tasks such as readability assessment or fake
-news detection.
+In this experiment we use LFTK~\cite{lftk-2023}, a framework built on top of
+spaCy~\cite{spacy} that provides implementations for computing many HLFs.
+LFTK categorizes HLFs by domain and family.
+We select HLFs from the catalog which is implemented by LFTK based on their
+domain and family, with the interest to make a diverse selection with respect to
+these attributes.
 Besides the domain and family, we also choose features based on their influence
@@
 
-The instructions for each HLF are specifically crafted using natural language.
+The instructions for each HLF are specifically crafted using natural language.\@
 The resulting instructions can be catalogued on the concrete-abstract spectrum.
@@
 and choose 2 features from each category.
-Table~\ref{table-hlf} showcases our selection of features.
+
+Table~\ref{table-hlf} showcases our HLF selection.
 The LFTK ID is used throughout tables and figures to refer to a specific HLF.\@
@@
 
-The experiment design rests upon measuring HLFs on a corpus of texts and
-computing statistics.
-LLMs also make use of examples extracted from a corpus to guide text generation.
-Our approach idempotently derives a corpus of texts from a well known dataset.
-Our selection of datasets is listed in Table~\ref{table-ds}.
+A corpus of texts is required to perform the experiment.
+We have conducted the experiment using corpora derived idempotently from the
+datasets listed in Table~\ref{table-ds}.
+
+The main criteria for choosing these datasets were the difference in style
+between the texts contained in one corpus versus the other and the similarity in
+style between the texts from the same corpus.
 
@@
 The style follows the publisher's guidelines, but, within those guidelines, it
-varies for each author.
+varies slightly for each author.
 The articles are more formal, longer and more complex in structure than Yelp
@@
 
-\subsection{External Input Texts}\label{input-text}
+\subsection{Input Text Choice}\label{input-text}
 
@@
 within it.
-This section describes how we use text external to the corpus, having already
-described our usage of text from within the corpus.
 
-We use a different external input text for each corpus.
-These texts are located in the
+An input text from within each corpus is chosen so that it exhibits the closest
+HLFs to the averages computed on the corpus.
+The choice is made when the text corpora are selected and is stable throughout
+all experimental evaluations.
+
+External input text is chosen in relation to each corpus.
+The chosen texts are located in the
 \texttt{\textlinkanon{https://github.com/koosie0507/llm-tst-consistency/tree/main/data}{data}}
@@
 Sterne~\cite{sterne2003life}.
-The text style is very dissimilar compared to news articles.
-Its linguistic features are also very unlike the average values computed for the
-CNN article corpus.
-A text that is similarly divorced from the style and linguistic features of most
-Yelp reviews is Oscar Wilde's \textit{Sonnet to Liberty}~\cite{wilde1909poems}.
+The text style is not similar to the style of news articles at all.
+Our selected HLFs are also very different from the average values computed for
+the CNN article corpus.
+A text that is similarly divorced from the style and HLFs of most Yelp reviews
+is Oscar Wilde's \textit{Sonnet to Liberty}~\cite{wilde1909poems}.
 Both texts are in the public domain and were obtained from Project
@@
 
-Handcrafted linguistic features are integer or floating point values that we
-can compute on text.
+HLFs are real numbers that we can derive from text idempotently.
 Plot figures show the HLF on the vertical axis and the trial number on the horizontal
@@
 
-We can determine whether the baseline results differ from the HLF results using
-the Kolmogorov-Smirnov test~\cite{kolmogorov1933,smirnov1939}, as implemented in
-the Scipy library~\cite{2020SciPy-NMeth}.
-We consider that the results are different for a $p\_value \le 0.05$.
+Because HLFs are just real numbers highlighting characteristics of text, they
+can be used for setting goals to aim for (`write a 100 word paragraph') and for
+describing text attributes (`this paragraph has 100 words').
+In our experiments we set goals for the chosen LLMs using instruction prompts
+and compute the corresponding HLFs on the generated text in order to measure
+LLM efficacy.
 
-To determine which of two results is closer to the target, we compute the area
-between each result and the target.
-Another way in which we measure the closeness of the result to the target is by
-computing the Euclidian norm of the vectorial difference between the result and
-target.
+As outlined in Subsection~\ref{subsec:experiment-process}, our experiment
+compares a baseline with the corresponding HLF results.
+Remember that in the context of a scenario variant, both the baseline and the
+HLF results for a specific HLF are vectors containing 10 real numbers.
+
+The Kolmogorov-Smirnov test~\cite{kolmogorov1933,smirnov1939} is useful for
+determining whether the baseline and HLF results vectors differ significantly.
+We use the Scipy~\cite{2020SciPy-NMeth} implementation of the test.
+In our experiment, the vectors are significantly different for a
+$p\_value \le 0.05$.
+
+For each HLF, the target HLF value is repeated 10 times to construct a vector
+which is compatible with the baselines and HLF results.
+This allows us to compute the Euclidian norm of the vectorial difference between
+the baselines and HLF results on one hand and the target vector, on the other.
+A similar computation can be performed based on the area between each of the
+vectors and the target.
+Because the area measurements are correlated with the Euclidian norm, we present
+results using only the Euclidian norm.
 
 Note that each HLF can have a different range of values than other HLFs.
-When we compute the area and the Euclidian norm using Numpy~\cite{harris2020array},
-we use the raw values of each HLF, without preprocessing them.
-Therefore the results for one HLF are not comparable with the results for
-another HLF.\@
-Since we don't quantify how different the baseline is from the result and we
-compute the relative closeness to a common target, comparing values for the same
-HLF but different LLMs is also ill-advised.
+While computing the Euclidian norm using Numpy~\cite{harris2020array}, we use
+the raw values of each HLF, without preprocessing.
+Therefore the norm computed for one HLF does not imply anything in relation to
+another HLF or LLM.\@
 
-Let us denote with $A_b$ and $A_h$ the areas between baseline and target and HLF
-result and target, respectively.
-Let us similarly denote with $N_b$ and $N_h$ the Euclidian norms.
-We define $Diff_A = A_b - A_h$ and $Diff_N = N_b - N_h$.
+Let us denote with $N_b$ and $N_h$ the Euclidian norms for the baseline and the
+HLF results, respectively.
+We denote the difference between the norms with $Diff_N = N_b - N_h$.
 
-The result tables show $Diff_A$ and $Diff_N$ for each LLM in the context of a
-prompt and a dataset.
-All result tables for each dataset and prompt variant are available in the
-appendix.
-
-Next, we present our perspective on the most relevant results of the experiment.
+The result tables show $Diff_N$ for each LLM in the context of a prompt and a
+text corpus.
+While the main body of the paper is focused on the CNN dataset and, an appendix
+to the paper documents the results collected by running the experiment for each
+dataset and prompt variant.
 
@@
             \toprule
-            claude3 $Diff_N$     & \textbf{\underline{623.16}}  & \textbf{\underline{24.74}} & \textbf{\underline{51.18}}  & \textbf{\underline{26.68}} & 0.06                      & -0.0                      & \textbf{\underline{0.46}} & \textbf{\underline{1.4}}   & 1.44                    & \textbf{\textit{-1.66}}   \\ \midrule
-            claude3 $Diff_A$     & \textbf{\underline{1774.0}}  & \textbf{\underline{74.0}}  & \textbf{\underline{158.84}} & \textbf{\underline{81.2}}  & 0.2                       & -0.01                     & \textbf{\underline{1.88}} & \textbf{\underline{4.49}}  & 5.01                    & \textbf{\textit{-4.09}}   \\ \midrule
-            gemini $Diff_N$      & \textbf{\underline{985.86}}  & \textbf{\underline{34.52}} & \textbf{\underline{100.36}} & \textbf{\underline{53.39}} & \textbf{\underline{0.52}} & 0.05                      & \textbf{\underline{2.62}} & \textbf{\underline{4.68}}  & -4.17                   & \textbf{\textit{-2.43}}   \\ \midrule
-            gemini $Diff_A$      & \textbf{\underline{2990.5}}  & \textbf{\underline{106.5}} & \textbf{\underline{320.34}} & \textbf{\underline{159.1}} & \textbf{\underline{1.5}}  & 0.09                      & \textbf{\underline{7.24}} & \textbf{\underline{13.15}} & -12.65                  & \textbf{\textit{-7.13}}   \\ \midrule
-            gpt $Diff_N$         & \textbf{\underline{125.87}}  & 1.46                       & 10.95                       & 10.77                      & 0.02                      & -0.01                     & 0.71                      & 0.81                       & \textbf{\textit{-1.72}} & 0.43                      \\ \midrule
-            gpt $Diff_A$         & \textbf{\underline{327.0}}   & 3.0                        & 29.0                        & 27.0                       & 0.08                      & -0.01                     & 1.82                      & 2.24                       & \textbf{\textit{-5.72}} & 0.44                      \\ \midrule
-            command\_r $Diff_N$  & \textbf{\underline{1024.78}} & \textbf{\underline{39.67}} & \textbf{\underline{81.58}}  & \textbf{\underline{41.46}} & \textbf{\underline{0.46}} & \textbf{\underline{0.03}} & \textbf{\underline{2.67}} & \textbf{\underline{3.78}}  & 1.4                     & \textbf{\underline{0.14}} \\ \midrule
-            command\_r $Diff_A$  & \textbf{\underline{3132.5}}  & \textbf{\underline{120.5}} & \textbf{\underline{246.84}} & \textbf{\underline{132.2}} & \textbf{\underline{1.36}} & \textbf{\underline{0.06}} & \textbf{\underline{7.88}} & \textbf{\underline{11.21}} & 3.9                     & \textbf{\underline{1.03}} \\ \midrule
-            llama3\_70b $Diff_N$ & \textbf{\underline{459.45}}  & \textbf{\underline{17.78}} & \textbf{\underline{36.5}}   & \textbf{\underline{19.75}} & \textbf{\underline{0.12}} & \textbf{\underline{0.03}} & \textbf{\underline{1.43}} & \textbf{\underline{2.18}}  & 1.5                     & \textbf{\textit{-0.06}}   \\ \midrule
-            llama3\_70b $Diff_A$ & \textbf{\underline{1363.5}}  & \textbf{\underline{54.0}}  & \textbf{\underline{106.5}}  & \textbf{\underline{60.5}}  & \textbf{\underline{0.36}} & \textbf{\underline{0.09}} & \textbf{\underline{4.33}} & \textbf{\underline{6.38}}  & 7.46                    & \textbf{\textit{-0.11}}   \\ \bottomrule
+            claude3     & \textbf{\underline{623.16}}  & \textbf{\underline{24.74}} & \textbf{\underline{51.18}}  & \textbf{\underline{26.68}} & 0.06                      & -0.0                      & \textbf{\underline{0.46}} & \textbf{\underline{1.4}}   & 1.44                    & \textbf{\textit{-1.66}}   \\ \midrule
+            gemini      & \textbf{\underline{985.86}}  & \textbf{\underline{34.52}} & \textbf{\underline{100.36}} & \textbf{\underline{53.39}} & \textbf{\underline{0.52}} & 0.05                      & \textbf{\underline{2.62}} & \textbf{\underline{4.68}}  & -4.17                   & \textbf{\textit{-2.43}}   \\ \midrule
+            gpt         & \textbf{\underline{125.87}}  & 1.46                       & 10.95                       & 10.77                      & 0.02                      & -0.01                     & 0.71                      & 0.81                       & \textbf{\textit{-1.72}} & 0.43                      \\ \midrule
+            command\_r  & \textbf{\underline{1024.78}} & \textbf{\underline{39.67}} & \textbf{\underline{81.58}}  & \textbf{\underline{41.46}} & \textbf{\underline{0.46}} & \textbf{\underline{0.03}} & \textbf{\underline{2.67}} & \textbf{\underline{3.78}}  & 1.4                     & \textbf{\underline{0.14}} \\ \midrule
+            llama3\_70b & \textbf{\underline{459.45}}  & \textbf{\underline{17.78}} & \textbf{\underline{36.5}}   & \textbf{\underline{19.75}} & \textbf{\underline{0.12}} & \textbf{\underline{0.03}} & \textbf{\underline{1.43}} & \textbf{\underline{2.18}}  & 1.5                     & \textbf{\textit{-0.06}}   \\ \midrule
         \end{tabular}%
-    }\caption{CNN Results}\label{table-prompt-1-cnn-dailymail}
+    }\caption{$Diff_N$ on the CNN-DailyMail corpus\\Scenario 1}\label{table-prompt-1-cnn-dailymail}
 \end{table*}
@@
 to the target.
-Figure~\ref{fig-llama-simp-ttr} shows an inconclusive result.
+Figure~\ref{fig-claude3-avpw} shows an inconclusive result.
 
@@
         \includegraphics[width=\linewidth]{plots/prompt_1/prompt_1-command_r-cnn_dailymail/prompt_1-command_r-cnn_dailymail_t_word.png}
-        \caption{Command-R+\\Total Words}\label{fig-command-r-t-word}
+        \caption{Command-R+~\texttt{t\_word}\\CNN-DailyMail,~Scenario 1}\label{fig-command-r-t-word}
     \end{minipage}
@@
         \includegraphics[width=\linewidth]{plots/prompt_1/prompt_1-gpt-cnn_dailymail/prompt_1-gpt-cnn_dailymail_fkgl.png}
-        \caption{GPT\\Flesch-Kincaid Grade Level}\label{fig-gpt-fkgl}
+        \caption{GPT~\texttt{fkgl}\\CNN-DailyMail,~Scenario 1}\label{fig-gpt-fkgl}
     \end{minipage}
@@
         \includegraphics[width=\linewidth]{plots/prompt_1/prompt_1-claude3-cnn_dailymail/prompt_1-claude3-cnn_dailymail_a_verb_pw.png}
-        \caption[center]{Claude3\\Average Verbs per Word}\label{fig-llama-simp-ttr}
+        \caption[center]{Claude3~\texttt{a\_verb\_pw}\\CNN-DailyMail,~Scenario 1}\label{fig-claude3-avpw}
     \end{minipage}
@@
 
-We attempt quantifying the impact of using HLF instructions on the output of the
-LLM by eliminating as many factors as possible that would lead to a different
-text being generated when compared to a baseline.
-In the second scenario of our experiment we use examples to allow the LLM to
-generate better baselines, thus hopefully narrowing the gap between baseline
-and HLF-enhanced result.
+The second scenario of our experiment uses examples to allow the LLM to
+generate better baselines.
+As pointed out at the end of Subsection~\ref{subsec:experiment-process}, we
+assume that the difference we measure between baselines and HLF results is due
+to the HLF instructions in the system prompt.
 
-There are two variants of this scenario depending on the provenance of the input
-text.
-The input text may be from outside the corpus or a text from within the corpus.
-Table~\ref{table-prompt-2-cnn-dailymail} contains results for the first variant:
-input from outside the corpus.
+The first variant of the second scenario involves using input text from outside
+the corpus.
+Table~\ref{table-prompt-2-cnn-dailymail} contains the measurements we obtained
+in this context.
 We notice the HLF results tend to be closer to the target than the baseline
 results.
-This picture is similar to the one we saw in the first scenario.
+In terms of significance and efficacy of the HLF instructions, the results are
+similar to the ones obtained in the first scenario.
 
@@
             \toprule
-            claude3 $Diff_N$     & \textbf{\underline{718.49}}  & \textbf{\underline{24.13}} & \textbf{\underline{44.43}}  & \textbf{\underline{23.43}} & \textbf{\underline{0.4}}  & 0.03                      & 0.94                      & \textbf{\underline{0.77}} & 1.2                        & 0.07                      \\ \midrule
-            claude3 $Diff_A$     & \textbf{\underline{2345.54}} & \textbf{\underline{90.18}} & \textbf{\underline{148.36}} & \textbf{\underline{63.6}}  & \textbf{\underline{1.18}} & 0.07                      & 1.36                      & \textbf{\underline{2.55}} & 3.33                       & 0.52                      \\ \midrule
-            gemini $Diff_N$      & \textbf{\underline{785.6}}   & \textbf{\underline{25.8}}  & \textbf{\underline{60.5}}   & \textbf{\underline{41.6}}  & \textbf{\underline{0.41}} & \textbf{\underline{0.05}} & \textbf{\underline{1.95}} & \textbf{\underline{2.67}} & -4.62                      & \textbf{\textit{-1.94}}   \\ \midrule
-            gemini $Diff_A$      & \textbf{\underline{2512.0}}  & \textbf{\underline{81.5}}  & \textbf{\underline{210.34}} & \textbf{\underline{125.7}} & \textbf{\underline{1.25}} & \textbf{\underline{0.09}} & \textbf{\underline{5.84}} & \textbf{\underline{8.05}} & -9.46                      & \textbf{\textit{-5.82}}   \\ \midrule
-            gpt $Diff_N$         & 47.8                         & 5.18                       & 6.13                        & -0.13                      & 0.02                      & 0.0                       & 0.05                      & 0.3                       & 3.06                       & -0.27                     \\ \midrule
-            gpt $Diff_A$         & 77.0                         & 12.5                       & 13.0                        & -4.0                       & 0.04                      & 0.0                       & -0.17                     & 0.77                      & 8.45                       & -0.16                     \\ \midrule
-            command\_r $Diff_N$  & \textbf{\underline{263.51}}  & 5.35                       & \textbf{\underline{25.08}}  & \textbf{\underline{17.37}} & \textbf{\textit{-0.09}}   & -0.01                     & \textbf{\underline{1.74}} & \textbf{\underline{1.42}} & \textbf{\underline{10.46}} & \textbf{\underline{1.56}} \\ \midrule
-            command\_r $Diff_A$  & \textbf{\underline{896.0}}   & 20.5                       & \textbf{\underline{86.0}}   & \textbf{\underline{60.6}}  & \textbf{\textit{-0.21}}   & -0.03                     & \textbf{\underline{5.45}} & \textbf{\underline{4.95}} & \textbf{\underline{31.29}} & \textbf{\underline{5.67}} \\ \midrule
-            llama3\_70b $Diff_N$ & \textbf{\underline{376.0}}   & \textbf{\underline{15.63}} & \textbf{\underline{38.63}}  & 15.74                      & \textbf{\underline{0.24}} & \textbf{\underline{0.06}} & 1.69                      & \textbf{\underline{2.91}} & 0.31                       & -0.3                      \\ \midrule
-            llama3\_70b $Diff_A$ & \textbf{\underline{1088.5}}  & \textbf{\underline{45.0}}  & \textbf{\underline{113.0}}  & 41.0                       & \textbf{\underline{0.65}} & \textbf{\underline{0.18}} & 3.53                      & \textbf{\underline{8.56}} & 0.27                       & -0.68                     \\ \bottomrule
+            claude3     & \textbf{\underline{718.49}}  & \textbf{\underline{24.13}} & \textbf{\underline{44.43}}  & \textbf{\underline{23.43}} & \textbf{\underline{0.4}}  & 0.03                      & 0.94                      & \textbf{\underline{0.77}} & 1.2                        & 0.07                      \\ \midrule
+            gemini      & \textbf{\underline{785.6}}   & \textbf{\underline{25.8}}  & \textbf{\underline{60.5}}   & \textbf{\underline{41.6}}  & \textbf{\underline{0.41}} & \textbf{\underline{0.05}} & \textbf{\underline{1.95}} & \textbf{\underline{2.67}} & -4.62                      & \textbf{\textit{-1.94}}   \\ \midrule
+            gpt         & 47.8                         & 5.18                       & 6.13                        & -0.13                      & 0.02                      & 0.0                       & 0.05                      & 0.3                       & 3.06                       & -0.27                     \\ \midrule
+            command\_r  & \textbf{\underline{263.51}}  & 5.35                       & \textbf{\underline{25.08}}  & \textbf{\underline{17.37}} & \textbf{\textit{-0.09}}   & -0.01                     & \textbf{\underline{1.74}} & \textbf{\underline{1.42}} & \textbf{\underline{10.46}} & \textbf{\underline{1.56}} \\ \midrule
+            llama3\_70b & \textbf{\underline{376.0}}   & \textbf{\underline{15.63}} & \textbf{\underline{38.63}}  & 15.74                      & \textbf{\underline{0.24}} & \textbf{\underline{0.06}} & 1.69                      & \textbf{\underline{2.91}} & 0.31                       & -0.3                      \\ \midrule
         \end{tabular}%
     }
-    \caption{prompt 2 - CNN Results\\Input Outside Corpus}\label{table-prompt-2-cnn-dailymail} % chktex-file 8
+    \caption{$Diff_N$ on the CNN-DailyMail corpus\\Scenario 2, Input Outside Corpus}\label{table-prompt-2-cnn-dailymail} % chktex-file 8
 \end{table*}
 
-On the other hand, there are fewer significant differences between baseline and
+We should note that there are fewer significant differences between baseline and
 HLF results.
-A possible explanation for this situation might lie in the closeness of the
-input prompt (which contains examples from the corpus)to the target.
-As we can see in Figure~\ref{fig-p2-gpt-twords}, this is not always the case.
-In~\cref{fig-p2-command-r-simpttr,fig-p2-gemini-a-kup-pw} we observe regressions
-in performance with significant differences between baseline and result.
-Even so, the results obtained in our first scenario are largely confirmed in the
-first variant of our second scenario.
+This seems to validate our assumption that using examples in the system prompt
+causes the baselines to be closer to the target HLF value.
 
@@
         \includegraphics[width=\linewidth]{plots/prompt_2/prompt_2-gpt-cnn_dailymail/prompt_2-gpt-cnn_dailymail_t_word.png}
-        \caption{GPT\\Total Words}\label{fig-p2-gpt-twords}
+        \caption{GPT~\texttt{t\_word}\\CNN-DailyMail, Scenario 2, Input Outside Corpus}\label{fig-p2-gpt-twords}
     \end{minipage}
@@
         \includegraphics[width=\linewidth]{plots/prompt_2/prompt_2-command_r-cnn_dailymail/prompt_2-command_r-cnn_dailymail_simp_ttr.png}
-        \caption{Command-R+\\Simple TTR}\label{fig-p2-command-r-simpttr}
+        \caption{Command-R+~\texttt{simp\_ttr}\\CNN-DailyMail, Scenario 2, Input Outside Corpus}\label{fig-p2-command-r-simpttr}
     \end{minipage}
@@
         \includegraphics[width=\linewidth]{plots/prompt_2/prompt_2-gemini-cnn_dailymail/prompt_2-gemini-cnn_dailymail_a_kup_pw.png}
-        \caption[center]{Gemini\\Kuperman Age of Acquisition}\label{fig-p2-gemini-a-kup-pw}
+        \caption[center]{Gemini~\texttt{a\_kup\_pw}\\CNN-DailyMail, Scenario 2, Input Outside Corpus}\label{fig-p2-gemini-a-kup-pw}
     \end{minipage}
@@
 
-We finally look at the second variant of the second scenario of our experiment.
-The hypothesis in this scenario variant was that the baseline would exhibit HLFs
-that are much closer to the target corpus average if the input text is from that
-corpus.
-The expectation is to see fewer significant HLF results and that, as a rule of
-thumb, HLF results are closer to the baseline in all cases.
-The results shown in Table~\ref{table-prompt-2-ifd-cnn-dailymail} contradict our
-expectations outright.
-The most surprising behaviour is exhibited for HLFs which are derived into more
-concrete generation instructions, such as the total number of words.
-This behaviour is highlighted in Figure~\ref{fig-p2-ifd-gemini-twords}.
-Additionally, we notice regressive behaviour for some models compared to the
-previously explored results in Figure~\ref{fig-p2-ifd-claude3-nuadj}.
-Finally, the abstractness of the text generation instructions for certain HLFs
-leads to the generated text not being compliant as exemplified in
-Figure~\ref{fig-p2-ifd-llama3-a-kup-pw}.
+There are exceptions to the general trend observed for the first scenario.
+One such exception is highlighted in Figure~\ref{fig-p2-gpt-twords}.
+In~\cref{fig-p2-command-r-simpttr,fig-p2-gemini-a-kup-pw} we even observe
+regressions in performance with significant differences between baseline and
+result.
+Even so, these are exceptions and the results obtained in our first scenario are
+largely confirmed.
+
+Finally, the second variant of the second scenario investigates the suggestion
+power of HLF instructions.
+By using input from the corpus, the LLM should produce baselines that comply
+even more to the target HLFs statistics computed on the corpus.
+We expect fewer significant differences between baselines and HLF results.
+Additionally, HLF results should be closer to the target than in our previous
+observations.
 
@@
             \toprule
-            claude3 $Diff_N$     & \textbf{\underline{871.02}}  & \textbf{\underline{34.2}}   & \textbf{\underline{34.48}}  & \textbf{\textit{-7.38}}    & \textbf{\underline{0.27}} & 0.01                      & \textbf{\textit{-0.93}}    & \textbf{\textit{-0.72}}    & 1.16                       & -0.14                   \\ \midrule
-            claude3 $Diff_A$     & \textbf{\underline{2756.12}} & \textbf{\underline{117.18}} & \textbf{\underline{126.36}} & \textbf{\textit{-11.2}}    & \textbf{\underline{0.81}} & -0.01                     & \textbf{\textit{-2.52}}    & \textbf{\textit{-0.74}}    & 3.69                       & -0.4                    \\ \midrule
-            gemini $Diff_N$      & \textbf{\underline{1315.41}} & \textbf{\underline{46.19}}  & \textbf{\underline{114.47}} & \textbf{\underline{72.38}} & \textbf{\underline{1.17}} & 0.03                      & \textbf{\underline{5.25}}  & \textbf{\underline{6.64}}  & -1.6                       & \textbf{\textit{-2.1}}  \\ \midrule
-            gemini $Diff_A$      & \textbf{\underline{3996.0}}  & \textbf{\underline{138.0}}  & \textbf{\underline{357.36}} & \textbf{\underline{212.7}} & \textbf{\underline{3.3}}  & 0.06                      & \textbf{\underline{15.49}} & \textbf{\underline{19.4}}  & -13.22                     & \textbf{\textit{-6.19}} \\ \midrule
-            gpt $Diff_N$         & \textbf{\underline{521.29}}  & \textbf{\underline{23.3}}   & \textbf{\underline{43.13}}  & 19.85                      & \textbf{\underline{0.2}}  & 0.02                      & 1.08                       & \textbf{\underline{1.04}}  & 5.09                       & 0.67                    \\ \midrule
-            gpt $Diff_A$         & \textbf{\underline{1464.0}}  & \textbf{\underline{65.5}}   & \textbf{\underline{121.84}} & 58.1                       & \textbf{\underline{0.58}} & 0.05                      & 2.93                       & \textbf{\underline{2.59}}  & 14.23                      & 1.89                    \\ \midrule
-            command\_r $Diff_N$  & \textbf{\underline{371.08}}  & \textbf{\underline{13.31}}  & 22.06                       & \textbf{\underline{17.19}} & \textbf{\underline{0.17}} & \textbf{\underline{0.02}} & 0.88                       & 1.09                       & \textbf{\underline{4.05}}  & -0.15                   \\ \midrule
-            command\_r $Diff_A$  & \textbf{\underline{1198.0}}  & \textbf{\underline{42.5}}   & 73.0                        & \textbf{\underline{56.5}}  & \textbf{\underline{0.69}} & \textbf{\underline{0.07}} & 2.85                       & 3.99                       & \textbf{\underline{13.89}} & -0.11                   \\ \midrule
-            llama3\_70b $Diff_N$ & \textbf{\underline{1040.85}} & \textbf{\underline{41.99}}  & \textbf{\underline{80.58}}  & \textbf{\underline{49.12}} & \textbf{\underline{0.55}} & \textbf{\underline{0.04}} & \textbf{\underline{3.99}}  & \textbf{\underline{5.05}}  & 7.18                       & \textbf{\textit{-1.94}} \\ \midrule
-            llama3\_70b $Diff_A$ & \textbf{\underline{2928.0}}  & \textbf{\underline{118.5}}  & \textbf{\underline{222.0}}  & \textbf{\underline{133.5}} & \textbf{\underline{1.43}} & \textbf{\underline{0.05}} & \textbf{\underline{10.77}} & \textbf{\underline{13.42}} & 17.21                      & \textbf{\textit{-6.2}}  \\ \bottomrule
+            claude3     & \textbf{\underline{871.02}}  & \textbf{\underline{34.2}}   & \textbf{\underline{34.48}}  & \textbf{\textit{-7.38}}    & \textbf{\underline{0.27}} & 0.01                      & \textbf{\textit{-0.93}}    & \textbf{\textit{-0.72}}    & 1.16                       & -0.14                   \\ \midrule
+            gemini      & \textbf{\underline{1315.41}} & \textbf{\underline{46.19}}  & \textbf{\underline{114.47}} & \textbf{\underline{72.38}} & \textbf{\underline{1.17}} & 0.03                      & \textbf{\underline{5.25}}  & \textbf{\underline{6.64}}  & -1.6                       & \textbf{\textit{-2.1}}  \\ \midrule
+            gpt         & \textbf{\underline{521.29}}  & \textbf{\underline{23.3}}   & \textbf{\underline{43.13}}  & 19.85                      & \textbf{\underline{0.2}}  & 0.02                      & 1.08                       & \textbf{\underline{1.04}}  & 5.09                       & 0.67                    \\ \midrule
+            command\_r  & \textbf{\underline{371.08}}  & \textbf{\underline{13.31}}  & 22.06                       & \textbf{\underline{17.19}} & \textbf{\underline{0.17}} & \textbf{\underline{0.02}} & 0.88                       & 1.09                       & \textbf{\underline{4.05}}  & -0.15                   \\ \midrule
+            llama3\_70b & \textbf{\underline{1040.85}} & \textbf{\underline{41.99}}  & \textbf{\underline{80.58}}  & \textbf{\underline{49.12}} & \textbf{\underline{0.55}} & \textbf{\underline{0.04}} & \textbf{\underline{3.99}}  & \textbf{\underline{5.05}}  & 7.18                       & \textbf{\textit{-1.94}} \\ \midrule
         \end{tabular}%
     }
-    \caption{CNN Results\\Input Inside Corpus}\label{table-prompt-2-ifd-cnn-dailymail}
+    \caption{$Diff_N$ on CNN-DailyMail\\Scenario 2, Input Inside Corpus}\label{table-prompt-2-ifd-cnn-dailymail}
 \end{table*}
+
+The results shown in Table~\ref{table-prompt-2-ifd-cnn-dailymail} contradict
+these expectations.
+The most surprising behaviour is exhibited for HLFs which are derived into more
+concrete generation instructions, such as the total number of
+words~(Figure~\ref{fig-p2-ifd-gemini-twords}).
+Additionally, we notice regressive behaviour for some models compared to the
+previously explored results in Figure~\ref{fig-p2-ifd-claude3-nuadj}.
+Finally, notice that the abstractness of the text generation instructions for
+certain HLFs leads to non-compliance in Figure~\ref{fig-p2-ifd-llama3-a-kup-pw}.
 
@@
         \includegraphics[width=\linewidth]{plots/prompt_2_ifd/prompt_2-claude3-cnn_dailymail/prompt_2-claude3-cnn_dailymail_n_uadj.png}
-        \caption{Claude3\\Number of Unique Adjectives}\label{fig-p2-ifd-claude3-nuadj}
+        \caption{Claude3\texttt{n\_uadj}\\CNN-DailyMail, Scenario 2, Input Inside Corpus}\label{fig-p2-ifd-claude3-nuadj}
     \end{minipage}
@@
         \includegraphics[width=\linewidth]{plots/prompt_2_ifd/prompt_2-gemini-cnn_dailymail/prompt_2-gemini-cnn_dailymail_t_word.png}
-        \caption{Gemini\\Total Words}\label{fig-p2-ifd-gemini-twords}
+        \caption{Gemini\texttt{t\_word}\\CNN-DailyMail, Scenario 2, Input Inside Corpus}\label{fig-p2-ifd-gemini-twords}
     \end{minipage}
@@
         \includegraphics[width=\linewidth]{plots/prompt_2_ifd/prompt_2-llama3_70b-cnn_dailymail/prompt_2-llama3_70b-cnn_dailymail_a_kup_pw.png}
-        \caption[center]{Llama3\\Kuperman Age of Acquisition}\label{fig-p2-ifd-llama3-a-kup-pw}
+        \caption[center]{Llama3\texttt{a\_kup\_pw}\\CNN-DailyMail, Scenario 2, Input Inside Corpus}\label{fig-p2-ifd-llama3-a-kup-pw}
     \end{minipage}
@@
 
-Overall, the LLMs do not seem to utilize the examples given in the prompt to
-their advantage when generating text that should mirror the linguistic features
-of those examples.
+The overall sentiment, taking into account the additional results obtained on
+the Yelp dataset, is that LLMs do \textit{not} take advantage of the examples
+from text corpora for the tasks performed in this experiment.
 
@@
 
-Let us move on to how much HLF instructions impact an LLM's writing.
+There is more doubt in response to how much HLF instructions impact an LLM's
+writing.
 The metrics we compute dismiss the idea that HLF instructions have a consistent
@@
 LLMs do not seem to follow these instructions well in most scenarios.
-The impact of highly abstract HLF instructions tends to be adverse.
 
@@
 The selection of language models, the reliance on a corpus for computing target
-linguistic features, the fixed choice of input text are all limitations of the
-current approach.
-Making different experiment design choices in all these respects might shed more
-light on this matter.
+linguistic features, the fixed choice of input text and not least of all the
+wording of the system prompts are all limitations of the current approach.
+
+Making different experiment design choices in all these respects might yield
+more positive and more powerful results.
 Using a selection of HLFs and not individually analysing the behaviour of the
@@
 
-We began by trying to understand whether it is possible to generate text that
-exhibits certain linguistic features by instructing a large language model.
+We designed an experiment that tries to understand whether it is possible to
+generate text that exhibits certain linguistic features by instructing a large
+language model.
 It turns out state of the art large language models are receptive to
@@
 the LLM output in the expected manner.
+Rewording text which already exhibits the desired linguistic features can have
+adverse effects, too.
 
@@
 the LLM outcome in relation to the selected HLFs is one way to achieve this.
-Given enough measurements, the relative difference between the exhibited
-baseline and target HLFs can be quantified relatively easily using geometric or
-statistical means.
+With enough measurements, the relative difference between baselines and target
+HLFs can be quantified using geometric or statistical means.
 

### 2024-12-10 article/main.tex
small rewording
remove redundant sentence
replace coling style with icaart style

@@
-\documentclass[11pt]{article}
+\documentclass[a4paper,twoside]{article}
 \pdfoutput=1
-
-\usepackage[review]{coling}
 
@@
 
-\usepackage{algorithmic}
-\usepackage{algorithm}
+\usepackage{algorithm2e}
 \usepackage{array}
@@
 \usepackage{amssymb}
+\usepackage{amstext}
 \usepackage{amsthm}
+\usepackage{apalike}
 \usepackage{balance}
 \usepackage{booktabs}
+\usepackage{calc}
 \usepackage[justification=centering]{caption}
@@
 \usepackage[T1]{fontenc}
+\usepackage[bottom]{footmisc}
 \usepackage{geometry}
@@
 \usepackage{multirow}
+\usepackage{pslatex}
 \usepackage{rotating}
 \usepackage{setspace}
-\usepackage{subfigure}
+\usepackage{subcaption}
 \usepackage{tabularx}
@@
 
-\title{LLM Output Compliance with Handcrafted Linguistic Features --- An Experiment}
-\authoranon{
-\author{Andrei Olar\\
-Universitatea Babeș-Bolyai, Cluj-Napoca, România\\
-\texttt{andrei.olar@ubbcluj.ro}\\}
-}
+\usepackage{SCITEPRESS}
 
 \begin{document}
-\maketitle
-\begin{abstract}
+
+\title{LLM Output Compliance with Handcrafted Linguistic Features --- An Experiment}
+
+\author{\authorname{Andrei Olar\sup{1}\orcidAuthor{0009-0006-7913-9276}}
+\affiliation{\sup{1}Mathematics and Computer Science Faculty, Babeș-Bolyai University, Cluj-Napoca, România}
+\email{andrei.olar@ubbcluj.ro}
+}
+\keywords{handcrafted linguistic features, large language models, controlled text generation, text style transfer}
+\abstract{
     Can we control the writing style of large language models (LLMs) by specifying
@@
     optimizing prompt design and feature selection.
-\end{abstract}
+}
+
+\onecolumn \maketitle \normalsize \setcounter{footnote}{0} \vfill
 
@@
 An example prompt instruction which asks the LLM to conform to a maximum
-\textit{sentence count} of 10 would read ``write no more than 10 sentences''.
+\textit{sentence count} of 10 could be ``write no more than 10 sentences''.
 On the other hand, HLFs can be computed on the text output by an LLM, too.
@@
 
-The experiment starts with choosing experiment parameter values.
+We start by choosing experiment parameter values.
 The following parameters have stable values throughout the experiment process: a
@@
 
-Once the parameters have been determined, the HLF statistics are computed on the
+Once the parameters have been determined, HLF statistics are computed on the
 text corpus.
@@
 
-The subsequent step involves formulating prompt instructions that incorporate
-HLFs.
 For each HLF, an instruction is constructed in natural language, adhering to the
@@
 provided examples, as well as logic to incorporate them.
-Both prompt templates contain logic to distinguish between the situation when
-HLF instructions were given versus not.
+Both prompt templates contain logic to distinguish between when HLF instructions
+were given or not.
 
-With the preliminary steps completed, baseline HLFs for the target LLM are
-computed.
+Baseline HLFs for the LLM output are computed after the preliminary steps are
+completed.
 For the first scenario, the LLM is instructed to reword the input text 10 times
 by using `\texttt{prompt\_1}' without HLF instructions.
-The HLFs in our selection are computed for each generated text, resulting in
-a 10-dimensional vector for each HLF.\@
-These vectors represent the baseline HLFs for the LLM in the first scenario.
+The HLFs in our selection are computed for each LLM output, resulting in a
+10-dimensional vector for each HLF.\@
+These vectors are the baseline HLFs for the LLM output in the first scenario.
 
-The second scenario baseline values are computed similarly using the
-`\texttt{prompt\_2}' template and the examples provided as one of the experiment
-parameters.
+The second scenario baseline HLFs are computed similarly to the first scenario
+baselines, using the `\texttt{prompt\_2}' template and the examples provided as
+an experiment parameter.
 Remembering that the second scenario has two variants, we obtain two sets of
-baseline vectors: a set for input from the corpus and one for input outside the
-corpus.
-The input from outside the corpus is the same as the input for the first
-scenario.
+baseline vectors, one for each variant.
+Note that the input from outside the corpus is the same as the input for the
+first scenario, while the input from inside the corpus is selected along with
+the corpus so that its HLFs are close to the average HLFs of the corpus.
 
@@
 \textit{baselines}.
-Note that we have 3 sets of baselines: 1 for the first scenario and 2 for the
-second scenario.
+We have 3 sets of baselines: 1 for the first scenario and 2 for the second
+scenario.
 
-For each type of baseline, we run 10 text generations using the corresponding
+For each set of baselines, we run 10 text generations using the corresponding
 prompt augmented with the HLF instructions computed earlier in the process.
@@
 
-The last step is to compare baselines with their corresponding HLF results for
-each HLF in our selection, for each scenario and variant.
+Lastly, we compare baselines with their corresponding HLF results for each HLF
+in our selection, for each scenario and variant.
 
@@
 baseline.
-
-Both variants of the second scenario quantify how similar the HLFs of the texts
-generated by the LLM are to the corpus average \textit{because of} the HLF
-instructions.
-This hypothesis emerges from the assumption that LLMs are able to mimic examples
-when they are provided in the system prompt.
+Both variants of the second scenario quantify how much HLF instructions
+influence text generation.
+The underlying assumption is that LLMs are able to mimic examples when they are
+provided in the system prompt.
 In such a case, the presence of HLF instructions would remain among the few
-possible explanations for deviations from the baseline.
+possible explanations for deviations from the baseline, especially when the
+baseline is already statistically close to the desired HLF values.
 
@@
 selection based on this criterion.
-To cover more ground we find top performers from different vendors interesting.
+To cover more ground we encourage vendor diversity in our selection.
 The final selection is available in Table~\ref{table-llm}.
@@
 In this experiment we use LFTK~\cite{lftk-2023}, a framework built on top of
-spaCy~\cite{spacy} that provides implementations for computing many HLFs.
+spaCy~\cite{spacy} which provides implementations for computing many HLFs.
 LFTK categorizes HLFs by domain and family.
-We select HLFs from the catalog which is implemented by LFTK based on their
-domain and family, with the interest to make a diverse selection with respect to
+We select HLFs based on their domain and family from the catalog implemented by
+LFTK, with the interest to make a diverse selection with respect to both
 these attributes.
@@
 HLFs to the averages computed on the corpus.
-The choice is made when the text corpora are selected and is stable throughout
-all experimental evaluations.
+The inputs are chosen from the corpora when the corpora are selected and are
+stable throughout all experimental evaluations.
 
@@
 
-HLFs are real numbers that we can derive from text idempotently.
-Plot figures show the HLF on the vertical axis and the trial number on the horizontal
-axis.
+HLFs are real numbers derived from text.
+Plot figures show the HLF on the vertical axis and the text generation trial
+number on the horizontal axis.
 We plot the baseline using a continuous grey line and the HLF result using a
@@
 Even so, these are exceptions and the results obtained in our first scenario are
-largely confirmed.
+largely confirmed on the CNN article corpus.
 
@@
 previously explored results in Figure~\ref{fig-p2-ifd-claude3-nuadj}.
-Finally, notice that the abstractness of the text generation instructions for
+Finally, we notice that the abstractness of the text generation instructions for
 certain HLFs leads to non-compliance in Figure~\ref{fig-p2-ifd-llama3-a-kup-pw}.
@@
 The overall sentiment, taking into account the additional results obtained on
-the Yelp dataset, is that LLMs do \textit{not} take advantage of the examples
-from text corpora for the tasks performed in this experiment.
+the Yelp dataset, is that LLMs do not take much advantage of the examples
+from text corpora, nor of the input from the corpora for the tasks performed in
+this experiment.
 
@@
 instructions.
-The importance of the choice of the external text is highlighted by the better
-HLF results obtained using an input text from the corpus.
 
-There is more doubt in response to how much HLF instructions impact an LLM's
-writing.
-The metrics we compute dismiss the idea that HLF instructions have a consistent
-desired impact on the generated text.
+The impact HLF instructions have on LLM output is limited.
+Our metrics do not show a consistent desired impact on the LLM output either.
 There isn't conclusive evidence that if we improve the LLM's chances of
-producing a better baseline, this will result in a smaller difference between
-baseline and HLF result relative to the target HLF.
-In fact, there is evidence that HLF instructions have an adverse effect when
-examining the HLF results obtained on the Yelp reviews dataset.
+producing a better baseline, this will result in closer HLF results to the 
+desired target values.\@
+In fact, there is some evidence that HLF instructions have an adverse effect
+when examining the HLF results obtained on the Yelp reviews dataset.
 This is especially true of highly abstract HLF instructions.
-LLMs do not seem to follow these instructions well in most scenarios.
 

### 2025-01-15 article/main.tex
prepare ICAART camera ready submission

@@
 \usepackage[justification=centering]{caption}
-\usepackage{cleveref}
 \usepackage{comment}
@@
 \usepackage{graphicx}
+\usepackage{hyperref}
+\usepackage{cleveref}
 \usepackage{inconsolata}
@@
 
-\usepackage[true]{anonymous-acm}
+\usepackage[false]{anonymous-acm}
 
@@
 \begin{document}
-
 \title{LLM Output Compliance with Handcrafted Linguistic Features --- An Experiment}
-
-\author{\authorname{Andrei Olar\sup{1}\orcidAuthor{0009-0006-7913-9276}}
+\author{
+    \authorname{Andrei Olar\sup{1}\orcidAuthor{0009-0006-7913-9276}}
     \affiliation{\sup{1}Mathematics and Computer Science Faculty, Babeș-Bolyai University, Cluj-Napoca, România}
@@
 }
-
 \onecolumn \maketitle \normalsize \setcounter{footnote}{0} \vfill
+\interfootnotelinepenalty=10000
 
@@
 
-If syntax differs from meaning~\cite{chomsky2002syntactic} the intuition is that
-writing style is defined in terms of both.
+Syntax differs from meaning~\cite{chomsky2002syntactic} and the intuition is
+that writing style is defined in terms of both.
 The distinction is important in order to set boundaries.
@@
 these attributes.
-Besides the domain and family, we also choose features based on their influence
-over text style as recognized in literature~\cite{verma2019lexical,lugea2023stylistics}.
+Besides the domain and family, we also choose features based on the
+level\footnotemark~at which they influence text style as recognized in
+literature~\cite{verma2019lexical,lugea2023stylistics}.
+\footnotetext{the notion of style level is described
+in~\cite{lugea2023stylistics}, chapter 1, page 7}
 
@@
 
-\begin{table*}[ht!]
+\begin{table*}[h!]
     \setlength\tabcolsep{6pt}
@@
             \toprule
-            LFTK ID         & Name                              & Family           & Domain           & Abstraction     & Style Level\footnotemark\\ \toprule
+            LFTK ID         & Name                              & Family           & Domain           & Abstraction     & Style Level \\ \toprule
             t\_word         & total words                       & wordsent         & surface          & very concrete   & lexical     \\
@@
 \end{table*}
-\footnotetext{as described in the figure from~\cite{lugea2023stylistics}, chapter 1, page 7}
 
@@
 
-The Huggingface ecosystem~\cite{lhoest-etal-2021-datasets} makes the selected
-datasets easily accessible.
+The datasets are available in the Huggingface
+ecosystem~\cite{lhoest-etal-2021-datasets}.
 Yelp reviews are loaded from \texttt{Yelp/yelp\_review\_full}, while CNN stories
@@
 text corpus.
-While the main body of the paper is focused on the CNN dataset and, an appendix
-to the paper documents the results collected by running the experiment for each
-dataset and prompt variant.
+The paper is focused on some key results observed on the CNN article corpus.
+Additional results are available
+\textlinkanon{https://github.com/koosie0507/llm-tst-consistency/tree/main/reports}{online}.
 
@@
         \includegraphics[width=\linewidth]{plots/prompt_2_ifd/prompt_2-claude3-cnn_dailymail/prompt_2-claude3-cnn_dailymail_n_uadj.png}
-        \caption{Claude3\texttt{n\_uadj}\\CNN-DailyMail, Scenario 2, Input Inside Corpus}\label{fig-p2-ifd-claude3-nuadj}
+        \caption{Claude3~\texttt{n\_uadj}\\CNN-DailyMail, Scenario 2, Input Inside Corpus}\label{fig-p2-ifd-claude3-nuadj}
     \end{minipage}
@@
         \includegraphics[width=\linewidth]{plots/prompt_2_ifd/prompt_2-gemini-cnn_dailymail/prompt_2-gemini-cnn_dailymail_t_word.png}
-        \caption{Gemini\texttt{t\_word}\\CNN-DailyMail, Scenario 2, Input Inside Corpus}\label{fig-p2-ifd-gemini-twords}
+        \caption{Gemini~\texttt{t\_word}\\CNN-DailyMail, Scenario 2, Input Inside Corpus}\label{fig-p2-ifd-gemini-twords}
     \end{minipage}
@@
         \includegraphics[width=\linewidth]{plots/prompt_2_ifd/prompt_2-llama3_70b-cnn_dailymail/prompt_2-llama3_70b-cnn_dailymail_a_kup_pw.png}
-        \caption[center]{Llama3\texttt{a\_kup\_pw}\\CNN-DailyMail, Scenario 2, Input Inside Corpus}\label{fig-p2-ifd-llama3-a-kup-pw}
+        \caption[center]{Llama3~\texttt{a\_kup\_pw}\\CNN-DailyMail, Scenario 2, Input Inside Corpus}\label{fig-p2-ifd-llama3-a-kup-pw}
     \end{minipage}
@@
 
-\bibliography{references}
-
-\appendix
-
-\section{Additional Result Tables}\label{sec:yelp-tables}
-\subsection{Yelp}
-\Cref{table-prompt-1-yelp,table-prompt-2-yelp,table-prompt-2-ifd-yelp} show the
-experiment results on the corpus of reviews extracted from the Yelp dataset.
-
-\begin{table*}[ht]
-    \centering
-    \resizebox{\textwidth}{!}{%
-        \begin{tabular}{lllllllllll}
-            model                 & t\_word                   & t\_sent                    & n\_uverb                 & n\_uadj                  & simp\_ttr & a\_verb\_pw             & corr\_adj\_var          & corr\_verb\_var         & fkgl                       & a\_kup\_pw               \\
-            \toprule
-            claude3 $Diff_N$     & \textbf{\textit{-281.9}}  & \textbf{\textit{-13.22}}   & \textbf{\textit{-25.79}} & -14.56                   & -0.06     & 0.0                     & -1.27                   & -1.14                   & \textbf{\underline{6.89}}  & \textbf{\underline{1.0}} \\ \midrule
-            claude3 $Diff_A$     & \textbf{\textit{-774.0}}  & \textbf{\textit{-27.76}}   & \textbf{\textit{-70.0}}  & -38.0                    & -0.19     & 0.02                    & -3.25                   & -3.39                   & \textbf{\underline{21.68}} & \textbf{\underline{2.9}} \\ \midrule
-            gemini $Diff_N$      & -109.27                   & 0.05                       & -7.25                    & -13.18                   & -0.14     & 0.01                    & -1.01                   & -0.4                    & 1.62                       & -0.4                     \\ \midrule
-            gemini $Diff_A$      & -252.82                   & -0.78                      & -17.5                    & -23.14                   & -0.52     & 0.01                    & -1.85                   & -1.07                   & 7.14                       & -0.93                    \\ \midrule
-            gpt $Diff_N$         & -21.62                    & 2.42                       & -3.8                     & -8.84                    & 0.0       & -0.01                   & -0.8                    & -0.53                   & 8.94                       & -0.37                    \\ \midrule
-            gpt $Diff_A$         & -34.72                    & 9.68                       & -10.54                   & -24.57                   & -0.03     & -0.02                   & -2.37                   & -1.58                   & 19.26                      & -1.03                    \\ \midrule
-            command\_r $Diff_N$  & -40.1                     & 1.22                       & 2.43                     & \textbf{\textit{-13.64}} & 0.05      & \textbf{\textit{-0.04}} & \textbf{\textit{-1.17}} & 0.59                    & \textbf{\textit{-7.25}}    & \textbf{\textit{-1.67}}  \\ \midrule
-            command\_r $Diff_A$  & -75.5                     & 1.5                        & 12.94                    & \textbf{\textit{-34.0}}  & 0.19      & \textbf{\textit{-0.12}} & \textbf{\textit{-2.76}} & 2.83                    & \textbf{\textit{-20.49}}   & \textbf{\textit{-4.76}}  \\ \midrule
-            llama3\_70b $Diff_N$ & \textbf{\textit{-210.45}} & \textbf{\underline{6.87}}  & \textbf{\textit{-13.46}} & \textbf{\textit{-15.17}} & -0.16     & \textbf{\textit{-0.02}} & \textbf{\textit{-1.78}} & \textbf{\textit{-1.31}} & 3.7                        & \textbf{\textit{-1.58}}  \\ \midrule
-            llama3\_70b $Diff_A$ & \textbf{\textit{-607.52}} & \textbf{\underline{19.42}} & \textbf{\textit{-41.22}} & \textbf{\textit{-44.77}} & -0.42     & \textbf{\textit{-0.05}} & \textbf{\textit{-5.48}} & \textbf{\textit{-4.34}} & 9.79                       & \textbf{\textit{-4.98}}  \\ \bottomrule
-        \end{tabular}%
-    }
-    \caption{Yelp Results - Prompt Without Examples}\label{table-prompt-1-yelp} % chktex-file 8
-\end{table*}
-
-\begin{table*}[!ht]
-    \centering
-    \resizebox{\textwidth}{!}{%
-        \begin{tabular}{lllllllllll}
-            model                 & t\_word & t\_sent & n\_uverb & n\_uadj & simp\_ttr & a\_verb\_pw & corr\_adj\_var            & corr\_verb\_var & fkgl                       & a\_kup\_pw \\
-            \toprule
-            claude3 $Diff_N$     & -121.77 & -4.61   & -9.9     & 18.41   & -0.11     & 0.03        & \textbf{\underline{1.4}}  & -0.33           & 1.81                       & 0.95       \\ \midrule
-            claude3 $Diff_A$     & -394.5  & -11.5   & -37.5    & 60.0    & -0.37     & 0.09        & \textbf{\underline{4.66}} & -1.49           & 5.38                       & 2.62       \\ \midrule
-            gemini $Diff_N$      & -23.49  & 1.52    & -0.55    & -1.29   & -0.02     & -0.01       & 0.09                      & -0.05           & \textbf{\underline{6.41}}  & 0.2        \\ \midrule
-            gemini $Diff_A$      & -83.0   & 3.0     & -2.5     & -6.36   & -0.0      & -0.03       & 0.26                      & -0.08           & \textbf{\underline{19.63}} & 0.53       \\ \midrule
-            gpt $Diff_N$         & 13.66   & 2.83    & -3.65    & -3.19   & 0.04      & -0.04       & -0.24                     & -0.22           & 7.5                        & 0.0        \\ \midrule
-            gpt $Diff_A$         & 55.5    & 9.5     & -9.86    & -12.86  & 0.14      & -0.11       & -0.84                     & -0.22           & 18.67                      & -0.38      \\ \midrule
-            command\_r $Diff_N$  & -188.71 & 0.69    & -10.73   & 1.17    & -0.23     & -0.01       & \textbf{\underline{0.74}} & -0.83           & \textbf{\textit{-5.02}}    & -0.08      \\ \midrule
-            command\_r $Diff_A$  & -516.82 & 2.02    & -29.22   & 8.28    & -0.66     & -0.0        & \textbf{\underline{2.98}} & -2.48           & \textbf{\textit{-14.52}}   & -0.48      \\ \midrule
-            llama3\_70b $Diff_N$ & -202.95 & -7.84   & -13.73   & 1.43    & -0.15     & -0.04       & 0.38                      & -1.39           & 10.67                      & 1.37       \\ \midrule
-            llama3\_70b $Diff_A$ & -561.24 & -18.78  & -49.94   & 5.07    & -0.38     & -0.14       & 1.62                      & -4.81           & 29.26                      & 3.11       \\ \bottomrule
-        \end{tabular}%
-    }
-    \caption{Yelp Results - Prompt With Examples\\Input External To Corpus}\label{table-prompt-2-yelp}
-\end{table*}
-\begin{table*}[!ht]
-    \centering
-    \resizebox{\textwidth}{!}{%
-        \begin{tabular}{lllllllllll}
-            model                 & t\_word                   & t\_sent                 & n\_uverb                  & n\_uadj                  & simp\_ttr               & a\_verb\_pw & corr\_adj\_var          & corr\_verb\_var         & fkgl                    & a\_kup\_pw              \\
-            \toprule
-            claude3 $Diff_N$     & -148.13                   & -1.63                   & -3.93                     & \textbf{\textit{-46.03}} & -0.11                   & -0.0        & \textbf{\textit{-3.08}} & -0.31                   & -4.3                    & \textbf{\textit{-1.53}} \\ \midrule
-            claude3 $Diff_A$     & -526.14                   & -7.98                   & -23.86                    & \textbf{\textit{-133.5}} & -0.41                   & -0.0        & \textbf{\textit{-8.98}} & -1.75                   & -21.03                  & \textbf{\textit{-5.34}} \\ \midrule
-            gemini $Diff_N$      & 74.27                     & 0.65                    & 4.67                      & 2.98                     & 0.04                    & -0.0        & -0.22                   & 1.05                    & 1.44                    & -0.65                   \\ \midrule
-            gemini $Diff_A$      & 221.0                     & 2.0                     & 10.5                      & 7.44                     & 0.12                    & -0.03       & -0.85                   & 2.73                    & 3.28                    & -1.79                   \\ \midrule
-            gpt $Diff_N$         & 0.65                      & 4.03                    & -1.19                     & -2.66                    & -0.03                   & -0.01       & -0.17                   & -0.13                   & 1.05                    & -0.67                   \\ \midrule
-            gpt $Diff_A$         & 17.0                      & 11.68                   & -3.68                     & -7.0                     & -0.1                    & -0.04       & -0.43                   & -0.05                   & 0.57                    & -1.75                   \\ \midrule
-            command\_r $Diff_N$  & \textbf{\textit{-236.11}} & \textbf{\textit{-3.6}}  & \textbf{\textit{-0.97}}   & \textbf{\textit{-27.04}} & \textbf{\textit{-0.23}} & 0.0         & -1.24                   & \textbf{\textit{-1.08}} & \textbf{\textit{-2.13}} & \textbf{\textit{-1.09}} \\ \midrule
-            command\_r $Diff_A$  & \textbf{\textit{-727.0}}  & \textbf{\textit{-9.24}} & \textbf{\underline{0.76}} & \textbf{\textit{-68.5}}  & \textbf{\textit{-0.64}} & 0.02        & -2.89                   & \textbf{\textit{-3.08}} & \textbf{\textit{-5.5}}  & \textbf{\textit{-3.0}}  \\ \midrule
-            llama3\_70b $Diff_N$ & \textbf{\textit{-333.03}} & \textbf{\textit{-5.19}} & 4.43                      & \textbf{\textit{-21.38}} & -0.16                   & -0.03       & \textbf{\textit{-1.77}} & -0.03                   & 1.87                    & -0.44                   \\ \midrule
-            llama3\_70b $Diff_A$ & \textbf{\textit{-687.34}} & \textbf{\textit{-3.42}} & 10.7                      & \textbf{\textit{-61.64}} & -0.31                   & -0.04       & \textbf{\textit{-5.6}}  & -0.89                   & 3.0                     & -1.92                   \\ \bottomrule
-        \end{tabular}%
-    }
-    \caption{Yelp Results - Prompt With Examples\\Input From Corpus}\label{table-prompt-2-ifd-yelp}
-\end{table*}
-
-\section{Additional Plots}
-
-\subsection{GPT}
-
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-gpt-cnn_dailymail/prompt_1-gpt-cnn_dailymail.png}
-    \caption{GPT on CNN Corpus\\Prompt Without Examples}\label{fig:gpt-prompt1-cnn}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-gpt-yelp/prompt_1-gpt-yelp.png}
-    \caption{GPT on Yelp Corpus\\Prompt Without Examples}\label{fig:gpt-prompt1-yelp}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-gpt-cnn_dailymail/prompt_2-gpt-cnn_dailymail.png}
-    \caption{GPT on CNN Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:gpt-prompt2-cnn}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-gpt-yelp/prompt_2-gpt-yelp.png}
-    \caption{GPT on Yelp Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:gpt-prompt2-yelp}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-gpt-cnn_dailymail/prompt_2-gpt-cnn_dailymail.png}
-    \caption{GPT on CNN Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:gpt-prompt2-cnn-ifd}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-gpt-yelp/prompt_2-gpt-yelp.png}
-    \caption{GPT on Yelp Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:gpt-prompt2-yelp-ifd}
-\end{figure*}
-
-\Cref{fig:gpt-prompt1-cnn,fig:gpt-prompt1-yelp,fig:gpt-prompt2-cnn,fig:gpt-prompt2-yelp,fig:gpt-prompt2-cnn-ifd,fig:gpt-prompt2-yelp-ifd}
-show the experiment results for the Claude 3 Opus large language model.
-
-\subsection{Gemini}
-
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-gemini-cnn_dailymail/prompt_1-gemini-cnn_dailymail.png}
-    \caption{Gemini on CNN Corpus\\Prompt Without Examples}\label{fig:gemini-prompt1-cnn}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-gemini-yelp/prompt_1-gemini-yelp.png}
-    \caption{Gemini on Yelp Corpus\\Prompt Without Examples}\label{fig:gemini-prompt1-yelp}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-gemini-cnn_dailymail/prompt_2-gemini-cnn_dailymail.png}
-    \caption{Gemini on CNN Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:gemini-prompt2-cnn}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-gemini-yelp/prompt_2-gemini-yelp.png}
-    \caption{Gemini on Yelp Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:gemini-prompt2-yelp}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-gemini-cnn_dailymail/prompt_2-gemini-cnn_dailymail.png}
-    \caption{Gemini on CNN Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:gemini-prompt2-cnn-ifd}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-gemini-yelp/prompt_2-gemini-yelp.png}
-    \caption{Gemini on Yelp Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:gemini-prompt2-yelp-ifd}
-\end{figure*}
-
-\Cref{fig:gemini-prompt1-cnn,fig:gemini-prompt1-yelp,fig:gemini-prompt2-cnn,fig:gemini-prompt2-yelp,fig:gemini-prompt2-cnn-ifd,fig:gemini-prompt2-yelp-ifd}
-show the experiment results for the Claude 3 Opus large language model.
-
-\subsection{Claude3}
-
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-claude3-cnn_dailymail/prompt_1-claude3-cnn_dailymail.png}
-    \caption{Claude3 on CNN Corpus\\Prompt Without Examples}\label{fig:claude3-prompt1-cnn}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-claude3-yelp/prompt_1-claude3-yelp.png}
-    \caption{Claude3 on Yelp Corpus\\Prompt Without Examples}\label{fig:claude3-prompt1-yelp}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-claude3-cnn_dailymail/prompt_2-claude3-cnn_dailymail.png}
-    \caption{Claude3 on CNN Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:claude3-prompt2-cnn}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-claude3-yelp/prompt_2-claude3-yelp.png}
-    \caption{Claude3 on Yelp Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:claude3-prompt2-yelp}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-claude3-cnn_dailymail/prompt_2-claude3-cnn_dailymail.png}
-    \caption{Claude3 on CNN Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:claude3-prompt2-cnn-ifd}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-claude3-yelp/prompt_2-claude3-yelp.png}
-    \caption{Claude3 on Yelp Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:claude3-prompt2-yelp-ifd}
-\end{figure*}
-
-\Cref{fig:claude3-prompt1-cnn,fig:claude3-prompt1-yelp,fig:claude3-prompt2-cnn,fig:claude3-prompt2-yelp,fig:claude3-prompt2-cnn-ifd,fig:claude3-prompt2-yelp-ifd}
-show the experiment results for the Claude 3 Opus large language model.
-
-\subsection{Llama3}
-
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-llama3_70b-cnn_dailymail/prompt_1-llama3_70b-cnn_dailymail.png}
-    \caption{Llama3 on CNN Corpus\\Prompt Without Examples}\label{fig:llama3_70b-prompt1-cnn}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-llama3_70b-yelp/prompt_1-llama3_70b-yelp.png}
-    \caption{Llama3 on Yelp Corpus\\Prompt Without Examples}\label{fig:llama3_70b-prompt1-yelp}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-llama3_70b-cnn_dailymail/prompt_2-llama3_70b-cnn_dailymail.png}
-    \caption{Llama3 on CNN Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:llama3_70b-prompt2-cnn}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-llama3_70b-yelp/prompt_2-llama3_70b-yelp.png}
-    \caption{Llama3 on Yelp Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:llama3_70b-prompt2-yelp}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-llama3_70b-cnn_dailymail/prompt_2-llama3_70b-cnn_dailymail.png}
-    \caption{Llama3 on CNN Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:llama3_70b-prompt2-cnn-ifd}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-llama3_70b-yelp/prompt_2-llama3_70b-yelp.png}
-    \caption{Llama3 on Yelp Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:llama3_70b-prompt2-yelp-ifd}
-\end{figure*}
-
-\Cref{fig:llama3_70b-prompt1-cnn,fig:llama3_70b-prompt1-yelp,fig:llama3_70b-prompt2-cnn,fig:llama3_70b-prompt2-yelp,fig:llama3_70b-prompt2-cnn-ifd,fig:llama3_70b-prompt2-yelp-ifd}
-show the experiment results for the Claude 3 Opus large language model.
-
-\subsection{Command R+}
-
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-command_r-cnn_dailymail/prompt_1-command_r-cnn_dailymail.png}
-    \caption{Command R on CNN Corpus\\Prompt Without Examples}\label{fig:command_r-prompt1-cnn}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-command_r-yelp/prompt_1-command_r-yelp.png}
-    \caption{Command R on Yelp Corpus\\Prompt Without Examples}\label{fig:command_r-prompt1-yelp}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-command_r-cnn_dailymail/prompt_2-command_r-cnn_dailymail.png}
-    \caption{Command R on CNN Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:command_r-prompt2-cnn}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-command_r-yelp/prompt_2-command_r-yelp.png}
-    \caption{Command R on Yelp Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:command_r-prompt2-yelp}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-command_r-cnn_dailymail/prompt_2-command_r-cnn_dailymail.png}
-    \caption{Command R on CNN Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:command_r-prompt2-cnn-ifd}
-\end{figure*}
-\begin{figure*}[ht]
-    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-command_r-yelp/prompt_2-command_r-yelp.png}
-    \caption{Command R on Yelp Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:command_r-prompt2-yelp-ifd}
-\end{figure*}
-
-\Cref{fig:command_r-prompt1-cnn,fig:command_r-prompt1-yelp,fig:command_r-prompt2-cnn,fig:command_r-prompt2-yelp,fig:command_r-prompt2-cnn-ifd,fig:command_r-prompt2-yelp-ifd}
-show the experiment results for the Claude 3 Opus large language model.
-
+\bibliographystyle{apalike}
+{\small
+\bibliography{references}}
+\balance
 \end{document}
